### 2024-08-30 paper.qmd
first commit.

---
title: "Tournaments"
author: "Greg"
format: html
editor: visual
---

## Tournaments

 - Single elimination (NCAA tournament, NFL playoffs)
  - Seeding issues (Tennis: https://www.liveabout.com/definition-of-seeding-3207821)
   Reseeding after each round.  Straight knockout.  
  - Second faces Third for silver medal.  Guarantees than sivler has 1 loss and bronze has 2 losses.  
  - With a 3rd place match (Badminton: https://www.nbcolympics.com/news/badminton-101-olympic-competition-format)
  - Single elimination with series (NBA, MLB, NHL)
  - Ladder tournaments (Korean baseball does best of 3 with one team starting up 1-0 in the opening round)
  - Seeding issues (NFL re-seeds, NCAA tournament straight bracket)
  - Group/Pool play followed by knockout stages (World Cup, Euro, Overwatch World CUp. )

 - Double elimination (Overwatch Champions League)
  - Repechage (Olympic Wrestling, Tae Kwon Do, Judo) (These also aware 2 bronze medals)
  - True Double elimination
  - Double elimination but a single loss before finals means you can onyl go for third place (College wrestling)

  - Triple Elimination (Curling! https://thegrandslamofcurling.com/bracketology-101-triple-knockout-explained-2021/)

 - Swiss Tournament (https://en.chessbase.com/post/125-years-swiss-system)

 - Round Robin
  - Single and Double.  
  - Baseball is an imbalanced multiple round robin qualifying round.

  - https://www.tandfonline.com/doi/abs/10.1080/01621459.1952.10501178

 - Heat based events (Track 5000m and below, swimming, )

 - Ranking Round followed by single elimination (Olympic Archery https://www.nbcolympics.com/news/archery-101-olympic-competition-format-and-scoring)

 - Elimination racing (Cycling, Bachelor/Bachelorette, Survivor)

 Tournament: a series of contests between a number of competitors, who compete for an overall prize. Oxford Dictionary
  Wiki: "A tournament is a competition involving at least three competitors, all participating in a sport or game"

 (https://www.espn.com/olympics/story/_/id/40536088/2024-paris-summer-olympics-track-field-medaling-format-rules)

Curling and DOE! https://khazna.ku.ac.ae/en/publications/applying-design-of-experiments-tooptimize-the-performance-level-o

Poker: https://arxiv.org/abs/physics/0703122

Graph Theory Stuff: 
Chickens: https://link.springer.com/article/10.1007/BF02476378

Are elections tournaments?

Tournament of Tournaments.

 - aggregation of rankings (golf fedex cup, formula one, tour de france)

https://liquipedia.net/overwatch/Overwatch_Champions_Series/2024/North_America/Stage_2
Swiss, Top Cut Swiss, Group play then a double elimiation bracket!

https://www.vgcguide.com/swiss-and-top-cut

```{r}
results <- data.frame(h = c("i","i","j","i","j","l"), a = c("j","l","l","k","k","k"), hpoints = c(3,3,3,1,1,1), apoints = c(0,0,0,1,1,1))

library(tidyverse)     
h <-results %>% group_by(h) %>% summarize(sum = sum(hpoints))  
a <- results %>% group_by(a) %>% summarize(sum = sum(apoints)) 
names(a) <- c("h","sum")
d <- rbind(h,a)
d %>% group_by(h) %>%  summarize(sum = sum(sum))  %>% arrange(-sum)

#Now reverse the outcomes
results <- data.frame(h = c("i","i","j","i","j","l"), a = c("j","l","l","k","k","k"), hpoints = c(0,0,0,1,1,1), apoints = c(3,3,3,1,1,1))

library(tidyverse)     
h <-results %>% group_by(h) %>% summarize(sum = sum(hpoints))  
a <- results %>% group_by(a) %>% summarize(sum = sum(apoints)) 
names(a) <- c("h","sum")
d <- rbind(h,a)
d %>% group_by(h) %>%  summarize(sum = sum(sum))  %>%arrange(-sum)

```

### 2024-09-13 paper.qmd
Update paper.qmd
Adding waterfall format, lines 35-41

@@
   - https://www.tandfonline.com/doi/abs/10.1080/01621459.1952.10501178
+
+  - Waterfall format
+   - Introduced in the Smash 64 community, [they claim](https://thesmashwriter.wordpress.com/2017/04/24/what-is-a-waterfall-bracket/)
+   - Recent example was [Super Smash Con 2023](https://www.start.gg/tournament/super-smash-con-2023/event/64-1v1-singles/overview)
+    - Players are seeded $1 \rightarrow n$ and grouped into $k$ divisions, 1 best, $n$ and $k$ worst. 
+    - The worst $\frac{n + \epsilon}{k}$ play in [relatively small round robin pools](https://www.start.gg/tournament/super-smash-con-2023/event/64-1v1-singles/brackets/1184919/1833431).
+Some fraction advance.
+In this linked example, There are 16 pools in division 4, where each pool is roughly 13 players, with 4 from each division advancing.
+    - The advancing players are then entered into pools with the slightly better-seeded players, and again complete round robins.
+This iterates over and over again.
+In this example, to make the numbers work, there are also [wildcard rounds](https://www.start.gg/tournament/super-smash-con-2023/event/64-1v1-singles/brackets/1184927/1833460) to advance an extra few players to further divisions.
+    - In Smash 64, this ultimately leads to a top 16 or 32, which then becomes a double-elimination tournament (all prior losses are ignored).
+The [top 32](https://www.start.gg/tournament/super-smash-con-2023/event/64-1v1-singles/brackets/1184923/1833456) is seeded by the [top division pool results](https://www.start.gg/tournament/super-smash-con-2023/event/64-1v1-singles/brackets/1184922/1833451).
 

### 2024-09-13 paper.qmd
Update paper.qmd
Adding blurb on graph theoretic tournaments.

@@
 
+ Tournament (graphy theory): [wikipedia definition](https://en.wikipedia.org/wiki/Tournament_(graph_theory)) and in [McShane (2019)](https://scholar.smu.edu/hum_sci_statisticalscience_etds/13/) -- essentially a "round robin tournament" -- a graph where every node (competitor: A, B, C) is connected by a directed edge (win A -> B or lose A <- B), where there are no loops (self-directed edges -- competitors cannot play against themselves).
+AKA a complete directed graph with no loops; has $n$ nodes and $\choose{n}{2}$ edges.
+John W.
+Moon's [Topic on Tournaments](https://www.gutenberg.org/ebooks/42833) is the classic graph theoretical tournaments text.
+This definition should be used to disambiguate _tournament_, especially since graph theory may be used.
+
  (https://www.espn.com/olympics/story/_/id/40536088/2024-paris-summer-olympics-track-field-medaling-format-rules)

### 2024-10-13 Tournament_Simulations.qmd
Add files via upload

---
title: "Tournaments"
author: "Zach Culp"
format: html
editor: visual
---

Creating the teams and the strengths

```{r}
# Define the number of teams 
num_teams <- 8
teams <- paste("Team", 1:num_teams)

# Assign random strengths to each team 
team_strengths <- data.frame(
  Team = teams,
  Strength = round(runif(num_teams, min = 0, max = 1),2)
)
#team_strengths
```

Simulating a match using the Bradley Terry model

```{r}
# Function to simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strengths) {
  strength1 <- strengths$Strength[strengths$Team == team1]
  strength2 <- strengths$Strength[strengths$Team == team2]

  prob_team1_wins <- strength1 / (strength1 + strength2)  # BT model probability

  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)  # Randomize outcome
  return(winner)
}
```

# Round Robin Tournaments

Example of a round robin structure (each team plays each other once)

```{r}
simulate_roundrobin <- function(teams, strengths){
  n <- length(teams)
  win_counts <- numeric(n)
  for (i in seq_along(teams)){
    for (j in seq_along(teams)){
      if (i != j){
        match_winner <- simulate_match(teams[i],teams[j],strengths)
        #cat(paste(teams[i], teams[j], match_winner), "\n")
        if (match_winner == teams[i]) {
          win_counts[i] <- win_counts[i] + 1
        } else {
          win_counts[j] <- win_counts[j] + 1
        }
      }
    }
  }
  results <- data.frame(
    Team = teams,
    Wins = win_counts
  )

  # Ranks teams based on number of wins (includes ties)
  results$Rank <- rank(-results$Wins, ties.method = "average")
  results <- results[order(results$Rank),]
  return(results)
}
round_robin <- simulate_roundrobin(teams, team_strengths)
round_robin
```

Simulation of many round robin tournaments.

```{r}
# Define the number of simulations
n <- 10000  # Set the number of simulations you want
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

# Run multiple simulations
for (i in 1:n) {
  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
  ranks <- simulation$Rank
  team_order <- simulation$Team
  total_ranks[i, team_order] <- ranks
}
# Calculate average ranks for each team
average_ranks <- colMeans(total_ranks, na.rm = TRUE)

# Print average ranks for each team
for (team in teams) {
  team_strength <- team_strengths$Strength[team_strengths$Team == team]
  cat(paste(team, "-> Avg. Rank:", average_ranks[team], "Strength:", team_strength), "\n")
}

```

This structure plays out exactly how it would be expected as the ranks follow the order of strengths.

# Single Elimination

Creates an example single elimination tournament with only one simulation.

```{r}
# defining the rank value for each finish
rank_values <- c(
  # Add more ranks as needed
  quarterfinalist = 6.5,
  semifinalist = 3.5,
  finalist = 2,
  champion = 1
)

# Simulate the single-elimination tournament
simulate_tournament <- function(teams, strengths) {
  ranks <- rep(NA, length(teams)) # Creates empty ranks to be filled later
  names(ranks) <- teams  # Assign team names to ranks for easier reference

  round_number <- 1  # To track the current round

  while (length(teams) > 1) {
    #cat("\n--- Round", round_number, "---\n")
    next_round <- c() # Makes sure only winners are included in next round

    for (i in seq(1, length(teams), by = 2)) { # Gets 2 teams for each match
      team1 <- teams[i]
      team2 <- teams[i+1]
      match_winner <- simulate_match(team1, team2, strengths)
      next_round <- c(next_round, match_winner)
      #cat(paste(team1, "vs", team2, "-> Winner:", match_winner, "\n"))

      # Assign ranks based on the current round
      if (round_number == 1) {  # Quarterfinals
        if (match_winner == team1) {
          if (is.na(ranks[team2])) {
            ranks[team2] <- rank_values["quarterfinalist"]
          }
        } else {
          if (is.na(ranks[team1])) {
            ranks[team1] <- rank_values["quarterfinalist"]
          }
        }
      } else if (round_number == 2) {  # Semifinals
        if (match_winner == team1) {
          if (is.na(ranks[team2])) {
            ranks[team2] <- rank_values["semifinalist"]
          }
        } else {
          if (is.na(ranks[team1])) {
            ranks[team1] <- rank_values["semifinalist"]
          }
        }
      } else if (round_number == 3) {
        if (match_winner == team1) {
          if(is.na(ranks[team2])) {
            ranks[team2] <- rank_values["finalist"]
          }
          if(is.na(ranks[team1])) {
            ranks[team1] <- rank_values["champion"]
          }
        } else {
          if(is.na(ranks[team1])) {
            ranks[team1] <- rank_values["finalist"]
          }
          if(is.na(ranks[team2])) {
            ranks[team2] <- rank_values["champion"]
          }
        }
      }
      # No rank assignment here for the final round; handled after loop
    }

    # If only two teams remain, handle the final match
    #if (length(next_round) == 2 && round_number == 3) {
    #  finalist <- setdiff(teams, next_round)
    #  # Assign ranks for the final match
    #  ranks[next_round[1]] <- rank_values["champion"]
    #  ranks[next_round[2]] <- rank_values["finalist"]
    #}

    teams <- next_round  # Winners move to the next round
    round_number <- round_number + 1
  }

  # Handle the case where only one team remains (champion)
  if (length(teams) == 1 && is.na(ranks[teams])) {
    ranks[teams] <- rank_values["champion"]
  }

  return(list(champion = teams[1], ranks = ranks))  # Return champion and ranks
}

# Define the teams for the tournament
teams <- c("Team 1", "Team 2", "Team 3", "Team 4", "Team 5", "Team 6", "Team 7", "Team 8")

# Run the tournament
cat("Starting tournament with teams:", paste(teams, collapse = ", "), "\n")
result <- simulate_tournament(teams,team_strengths)

# Print the results
cat("\nChampion of the tournament is:", result$champion, "\n")
cat("\nTeam Rankings:\n")
for (team in teams) {
  cat(paste(team, "-> Rank:", ifelse(is.na(result$ranks[team]), "Not Ranked", result$ranks[team]), "\n"))
}
team_strengths
```

Simulating games many times

```{r}
# Initialize the teams you want to simulate
team1 <- "Team 1"  # Replace with the actual team name
team2 <- "Team 2"  # Replace with the actual team name

# Initialize a vector to store match results (1 for team1 win, 0 for team2 win)
match_results <- numeric(1000)

# Run 1000 simulations of the match
for (i in 1:1000000) {
  winner <- simulate_match(team1, team2,team_strengths)
  match_results[i] <- ifelse(winner == team1, 1, 0)  # 1 if team1 wins, 0 if team2 wins
}

# Calculate the mean of the match results
mean_result <- mean(match_results)
cat("Mean result (fraction of wins for team1):", mean_result, "\n")
```

Simulating the tournament many times.

```{r}
# Define the number of simulations
n <- 10000  # Set the number of simulations you want

# Initialize a data frame to store total ranks for each team
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

# Run multiple simulations
for (i in 1:n) {
  simulation <- simulate_tournament(teams, team_strengths)  # Call the tournament simulation function
  total_ranks[i, ] <- simulation$ranks  # Store the ranks of the current simulation
}
total_ranks
# Calculate average ranks for each team
average_ranks <- colMeans(total_ranks, na.rm = TRUE)

# Print average ranks for each team
for (team in teams) {
  team_strength <- team_strengths$Strength[team_strengths$Team == team]
  cat(paste(team, "-> Avg. Rank:", average_ranks[team], "Strength:", team_strength), "\n")
}

```

As seen here, a completely randomized seeding can lead to a higher strength team losing more often than a lower strength team. Because of the high probability of winning the first game, Team 6 actually averages to the highest ranking.

Now, attempting it seeding the teams based on strength.

```{r}
# Sort the teams by their strengths in descending order (for seeding)
teams_seeded <- team_strengths[order(-team_strengths$Strength), ]  # Strongest at the top
seeded_strengths <- data.frame(
  Team = teams_seeded$Team,
  Strength = teams_seeded$Strength
)
# Display the seeding
#cat("Teams sorted by strength (seeds):\n")
#teams_seeded

simulate_seeded_tournament <- function(teams,strengths) {
  ranks <- rep(NA, length(teams))  # Create empty ranks to be filled later
  names(ranks) <- teams  # Set names for easier access
  round_number <- 1
  while (length(teams) > 1) {
    #cat("\n--- Round", round_number, "---\n")
    next_round <- c()  # Ensures only winners are included in the next round
    # Pair teams based on seeding (1 vs 8, 2 vs 7, etc.)
   for (i in 1:(length(teams) / 2)) {
      team1 <- teams[i]                  # Higher seed
      team2 <- teams[length(teams) - i + 1]  # Lower seed
      match_winner <- simulate_match(team1, team2, strengths)
      next_round <- c(next_round, match_winner)
      #cat(paste(team1, "vs", team2, "-> Winner:", match_winner, "\n\n"))

      # Assign ranks based on the current round
      if (round_number == 1) {  # Quarterfinals
        ranks[team1] <- rank_values["quarterfinalist"]
        ranks[team2] <- rank_values["quarterfinalist"]
      } else if (round_number == 2) {  # Semifinals
        if (match_winner == team1) {
          ranks[team2] <- rank_values["semifinalist"]
        } else {
          ranks[team1] <- rank_values["semifinalist"]
        }
      } else if (round_number == 3) {  # Finals
        if (match_winner == team1) {
          ranks[team1] <- rank_values["champion"]
          ranks[team2] <- rank_values["finalist"]
        } else {
          ranks[team2] <- rank_values["champion"]
          ranks[team1] <- rank_values["finalist"]
        }
      }
    }

    teams <- next_round
    round_number <- round_number + 1
  }

  return(list(champion = teams[1], ranks = ranks))  # Final winner
}

# Define the number of simulations
n <- 1000  # Set the number of simulations you want

# Initialize a data frame to store total ranks for each team
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

# Run multiple simulations
for (i in 1:n) {
  seeded_simulation <- simulate_seeded_tournament(teams_seeded$Team,seeded_strengths)  # Call the tournament simulation function
  seeded_simulation
  total_ranks[i, names(seeded_simulation$ranks)] <- seeded_simulation$ranks  # Store the ranks of the current simulation
}
# Calculate average ranks for each team
average_ranks <- colMeans(total_ranks, na.rm = TRUE)

# Print average ranks for each team
for (team in teams_seeded$Team) {
  team_strength <- teams_seeded$Strength[teams_seeded$Team == team]
  cat(paste(team, "-> Avg. Rank:", average_ranks[team], "Strength:", team_strength), "\n")
}

```

As seen in this example, the ranks are what you would expect given the strengths. This shows that a seeded single elimination tournament structure will likely produce more accurate depictions of strength than a randomized single elimination structure.

The next step will be to include potential byes to this single elimination tournament and then look into double elimination tournaments.

### 2024-10-18 all_possible_8team_brackets.R
all possible brackets.

@@ -1,0 +1,23 @@
+library(combinat)
+tourn <- permn(1:8)
+tourn <- do.call(rbind,tourn)
+
+dist <- apply(tourn, 1, function(x){
+  if (sum(x[c(5:8)] == 1)>0){x[c(1:8)] <- x[c(5:8,1:4)]}
+  
+  if (sum(x[c(3:4)] == min(x[1:4]))>0){x[c(1:4)] <- x[c(3:4,1:2)]}
+  if (sum(x[c(7:8)] == min(x[5:8]))>0){x[c(5:8)] <- x[c(7:8,5:6)]}
+  
+  for (i in c(1,3,5,7)){
+    x[c(i:(i+1))] <- sort(x[c(i:(i+1))])
+  }
+  return(x)
+})
+
+dist <- t(dist)
+
+test <- dist[!duplicated(dist),]
+dim(test)
+
+
+#as.data.frame(test) %>% arrange(V1, V2)

### 2024-10-18 paper.qmd
Added to library

--- 2024-08-30 paper.qmd
@@
 title: "Tournaments"
-author: "Greg"
+author: "Greg, Josie, Zach, Ryan"
 format: html
@@
 ---
+
+Llull (1283)
+
+Kendall and Smith (1940): Judges makes consistent preferences.
+
+Slater's I (1961): Measure of inconsistency
+
+Thurstone Mosteller model (1928): Bradley Terry Model with a normal CDF transformation.
+
+Kendall-Wei (1952)
+Elo (1978)
+Glicko (1998) and Glicko2 (2006?): Essentially Bayesian Elo.  
+Trueskill (2006): Used in halo 2 or 3.
+
+Weak stochastic probablity. $P(A \rightarrow B) > 0.5$  and $P(B \rightarrow C) > 0.5$ then $P(A \rightarrow C) > 0.5$.
+
+CRSP: Chain Rock Scissor Paper.  Non transitive model.
+
+Ratings vs rankings: Ratings are numeric estimates of strengths.  Rankings are a function of ratings.
+
+Let's start with what a BAD tournament would be.
+
+Quarto manuscript template:https://github.com/mikemahoney218/quarto-arxiv
+
+##Papers
+
+https://www.sciencedirect.com/science/article/pii/S0378375803003215
+"Maximum weight perfect match problem"
+
+## Deterministic Binary Outcome Tournaments
+
+ - Knockout (KO) lKO
+  - 1KO
+  - 2KO
+      - True 3rd, college wrestling, repechage
+  - 3KO
+  - Seeding issues,.
+
+ - nRR 
+  - 1RR
+  - 2RR
+  - partial RR
+
+## Dynamic Binary Outcome Tournaments
+
+  - Swiss
+  - Dynamic Knockout (reseeding after every round)
 
@@
 
- - Heat based events (Track 5000m and below, swimming, )
+ - Heat based events (Track 5000m and below, swimming)
 
@@
 ```
+
+Note for Greg: 
+Can we build a simulator that computes the probability of a team winning in two steps: 1) the team strength comes from a distribution itself and then the actual outcome come from flipping a coin based on the particular draw of the distribtuion.
+
+Can we model a sport and how muhc variability it has?
+So like estimate the probability of rteam A beating team B as a function of relative strengths but also use that other distribution to model the variability of the sport.
+Does this exist?

### 2024-11-07 round_robin.qmd
round robin simulations

---
title: "round robin simulations"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Bradley Terry

teams and strengths

```{r}
set.seed(4329)
library(tidyverse)

#set up
num_teams <- 8
teams <- paste("Team", 1:num_teams)

team_strengths <- data.frame(Team = teams,
  Strength = round(runif(num_teams, min = 0, max = 1),2))

# add column for true ranks of team_strengths
(team_strengths <- team_strengths %>% arrange(-Strength) %>% mutate(Rank = row_number()) %>% arrange(Team))
```

simulate one match using Bradley Terry model
```{r}
# simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strengths) {
  strength1 <- strengths$Strength[strengths$Team == team1]
  strength2 <- strengths$Strength[strengths$Team == team2]

  prob_team1_wins <- strength1 / (strength1 + strength2)  # BT model probability

  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)  # Randomize outcome
  return(winner)
}
```

simulate single round robin tournament

```{r}
simulate_roundrobin <- function(teams, strengths){
  n <- length(teams)
  win_counts <- numeric(n)
  for (i in seq_along(teams)){
    for (j in seq_along(teams)){
      if (i != j){
        match_winner <- simulate_match(teams[i],teams[j],strengths)
        if (match_winner == teams[i]) {
          win_counts[i] <- win_counts[i] + 1
        } else {
          win_counts[j] <- win_counts[j] + 1
        }
      }
    }
  }
  results <- data.frame(
    Team = teams,
    Wins = win_counts
  )

  # Ranks teams based on number of wins (includes ties)
  results$RankHat <- rank(-results$Wins, ties.method = "average")
  results <- results[order(results$RankHat),]

  results <- results %>% inner_join(team_strengths, by = 'Team') %>% select(Team, Strength, Rank, RankHat, Wins)

  return(results)
}

rr <- simulate_roundrobin(teams, team_strengths)

# kendall correlation
kendallcor <- cor.test(rr$RankHat,rr$Rank, method="kendall")
```

simulate many round robin tournaments using the same set of 8 teams

```{r}
# Define the number of simulations
n <- 1000  # Set the number of simulations you want
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

kendallcor <- c()

# Run multiple simulations
for (i in 1:n) {
  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
  ranks <- simulation$Rank
  team_order <- simulation$Team
  total_ranks[i, team_order] <- ranks

  k <- cor.test(simulation$RankHat,simulation$Rank, method="kendall")$estimate
  kendallcor[i] <- k
}

# Calculate average ranks for each team
average_ranks <- data.frame(AvgRankHat = colMeans(total_ranks, na.rm = TRUE), Strength = team_strengths$Strength, Rank = team_strengths$Rank)
average_ranks

# distribution of kendall coefficients
hist(kendallcor)
mean(kendallcor)
median(kendallcor) # rank hat matches true rank about 70% of the time
```

# Finalizing Round Robin Function

### generate winning probabilities with$\exp(i-j)/(1+\exp(i-j))$

round robin function inputs: teams, strengths, schedule, (seed?)

## create teams and strengths - strengths are just 1-8

```{r}
library(tidyverse)

# Define the number of teams 
num_teams <- 8
teams <- paste("Team", 1:num_teams)

# Assign random strengths to each team 
team_strengths <- data.frame(
  Team = teams,
  Strength = (1:num_teams)
)

# add column for true ranks of team_strengths
team_strengths <- team_strengths %>% arrange(-Strength) %>% mutate(Rank = row_number()) %>% arrange(Team)
```

## simulate a single match

```{r}
simulate_match <- function(team1, team2, strengths) {
  strength1 <- strengths$Strength[strengths$Team == team1]
  strength2 <- strengths$Strength[strengths$Team == team2]

  prob_team1_wins <- exp(strength1-strength2) / (1 + exp(strength1-strength2))

  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)
  return(winner)
}
```

## simulate a single round robin tournament

```{r}
simulate_roundrobin <- function(teams, strengths){
  n <- length(teams)
  win_counts <- numeric(n)
  for (i in seq_along(teams)){
    for (j in seq_along(teams)){
      if (i != j){
        match_winner <- simulate_match(teams[i],teams[j],strengths)
        if (match_winner == teams[i]) {
          win_counts[i] <- win_counts[i] + 1
        } else {
          win_counts[j] <- win_counts[j] + 1
        }
      }
    }
  }
  results <- data.frame(
    Team = teams,
    Wins = win_counts
  )

  # Ranks teams based on number of wins (includes ties)
  results$RankHat <- rank(-results$Wins, ties.method = "average")
  results <- results[order(results$RankHat),]

  results <- results %>% inner_join(team_strengths, by = 'Team') %>% select(Team, Strength, Rank, RankHat, Wins)

  return(results)
}

rr <- simulate_roundrobin(teams, team_strengths)
# kendall correlation
kendallcor <- cor(rr$RankHat, rr$Rank, method="kendall")
# spearman correlation
spearmancor <- cor(rr$RankHat, rr$Rank, method="spearman")
```

Kendall Coefficient

Kendall Coefficient is a statistic used to measure the ordinal association between two measured quantities.

-   If the agreement between the two rankings is perfect (i.e., the two rankings are the same) the coefficient has value 1.
-   If the disagreement between the two rankings is perfect (i.e., one ranking is the reverse of the other) the coefficient has value −1
    -   We do not have any values of the coefficient that are less than 0

Spearman Coefficent

Spearman Correlation is a non-parametric correlation also known as rank-based correlation coefficient. It measures the strength and direction of the monotonic relationship between two ranked variables. It ranges from -1 to +1:

-   **+1**: A perfect positive monotonic relationship.

-   **0**: No monotonic relationship.

-   **-1**: A perfect negative monotonic relationship.

### simulate many round robin tournaments using the same set of 8 teams, get kendall and spearman coefficients

```{r}
# Define the number of simulations
n <- 1000  # Set the number of simulations
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

kendallcor <- c()
spearmancor <- c()

# Run multiple simulations
for (i in 1:n) {
  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
  ranks <- simulation$Rank
  team_order <- simulation$Team
  total_ranks[i, team_order] <- ranks

  kendallcor[i] <- cor(simulation$RankHat, simulation$Rank, method="kendall")
  spearmancor[i] <- cor(simulation$RankHat, simulation$Rank, method="spearman")
}

# Calculate average ranks for each team
average_ranks <- data.frame(AvgRankHat = colMeans(total_ranks, na.rm = TRUE), Strength = team_strengths$Strength, Rank = team_strengths$Rank)
average_ranks

# distribution of kendall correlation
print('kendall')
hist(kendallcor)
mean(kendallcor)
median(kendallcor)

# distribution of spearman correlation
print('spearman')
hist(spearmancor)
mean(spearmancor)
median(spearmancor)
```

# ROUND ROBIN CODE

## simulate a single match

```{r}
simulate_match <- function(team1, team2, strengths) {
  strength1 <- strengths$Strength[strengths$Team == team1]
  strength2 <- strengths$Strength[strengths$Team == team2]

  prob_team1_wins <- exp(strength1-strength2) / (1 + exp(strength1-strength2))

  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)
  return(winner)
}
```

## simulate a single round robin tournament

```{r}
simulate_roundrobin <- function(teams, strengths){
  n <- length(teams)
  win_counts <- numeric(n)
  for (i in seq_along(teams)){
    for (j in seq_along(teams)){
      if (i != j){
        match_winner <- simulate_match(teams[i],teams[j],strengths)
        if (match_winner == teams[i]) {
          win_counts[i] <- win_counts[i] + 1
        } else {
          win_counts[j] <- win_counts[j] + 1
        }
      }
    }
  }
  results <- data.frame(
    Team = teams,
    Wins = win_counts
  )

  # Ranks teams based on number of wins (includes ties)
  results$RankHat <- rank(-results$Wins, ties.method = "average")
  results <- results[order(results$RankHat),]

  results <- results %>% inner_join(team_strengths, by = 'Team') %>% select(Team, Strength, Rank, RankHat, Wins)

  return(results)
}
```

# code that is changed for different team strengths/number sims, etc.
## generate team strengths using normal distribution
```{r}
# Define the number of teams 
num_teams <- 8
teams <- paste("Team", 1:num_teams)

# Assign random strengths to each team 
team_strengths <- data.frame(
  Team = teams,
  Strength = rep(0,8))

# add column for true ranks of team_strengths
team_strengths <- team_strengths %>% arrange(-Strength) %>% mutate(Rank = row_number()) %>% arrange(Team)
team_strengths
```

run round robin simulation
```{r}
# Define the number of simulations
n <- 1000  # Set the number of simulations
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

kendallcor <- c()
spearmancor <- c()

# Run multiple simulations
for (i in 1:n) {
  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
  ranks <- simulation$Rank
  team_order <- simulation$Team
  total_ranks[i, team_order] <- ranks

  kendallcor[i] <- cor(simulation$RankHat, simulation$Rank, method="kendall")
  spearmancor[i] <- cor(simulation$RankHat, simulation$Rank, method="spearman")
}

# Calculate average ranks for each team
average_ranks <- data.frame(AvgRankHat = colMeans(total_ranks, na.rm = TRUE), Strength = team_strengths$Strength, Rank = team_strengths$Rank)
average_ranks

# distribution of kendall correlation
print('kendall')
hist(kendallcor)
mean(kendallcor)
median(kendallcor)

# distribution of spearman correlation
print('spearman')
hist(spearmancor)
mean(spearmancor)
median(spearmancor)
```

# controlling the strengths but still have normal distribution
```{r}
thetas <- qnorm((seq(0,1,length=9) + 1/16)[-9])
thetas

# qq plot way
qnorm(((1:8) - 0.5) / 8, 0, 0.5)

# plot with variability on x and average of correlation coefficient on y

# qunif
```

### 2024-11-22 single_elimination.qmd

---
title: "single_elimination"
format: html
editor: visual
---

# Simulate a Single Match

```{r}
# Function to simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strength1, strength2) {
  odds <- exp(strength1 - strength2)/(1+exp(strength1 - strength2))

  winner <- ifelse(runif(1) < odds, team1, team2)  # Randomize outcome
  return(winner)
}
```

# Single Elimination Tournament Structures

```{r}
num_teams <- 8
teams <- paste("Team", 1:num_teams)
team_map <- setNames(1:8, teams)
tourn <- do.call(rbind, args = combinat::permn(1:8))
tourn
dist <- apply(X = tourn, MARGIN = 1, FUN = function(x){
  if (sum(x[5:8] == 1) > 0) {
    x[1:8] <- x[c(5:8, 1:4)]
  }
  if (sum(x[3:4] == min(x[1:4])) > 0) {
    x[1:4] <- x[c(3:4, 1:2)]
  }
  if (sum(x[7:8] == min(x[5:8])) > 0) {
    x[5:8] <- x[c(7:8, 5:6)]
  }
  for (i in c(1, 3, 5, 7)) {
    x[i:(i + 1)] <- sort(x[i:(i + 1)])
  }
return(x)
})

dist <- t(dist)
seeds <- dist[!duplicated(dist), ]

# How to check for a specific seed
which(apply(seeds, 1, function(x) all(x == c(1, 8, 2, 7, 3, 6, 4, 5))))

## Important Structures Rows ##
# 1-8: row 1
# (1,8), (2,7), (3,6), (4,5): row 23
```

# Setting Up the Teams and Strengths

```{r}
# Assign random strengths to each team 
team_strengths <- data.frame(
  Team = teams,
  Strength = c(1,2,3,4,5,6,7,8),
  Wins = rep(0,8),
  Ranks = rep(NA,8),
  Exp_Ranks = c(8,7,6,5,4,3,2,1)
)

normal_strengths <- data.frame(
  Team = teams,
  Strength = round(rnorm(8,0,1),2),
  Wins = rep(0,8),
  Ranks = rep(NA,8)
)
normal_strengths$Exp_Ranks <- rank(-normal_strengths$Strength, ties.method = "average")

unif_strength <- data.frame(
  Team = teams,
  Strength = round(runif(8,0,8),2),
  Wins = rep(0,8),
  Ranks = rep(NA,8)
)
unif_strength$Exp_Ranks <- rank(-unif_strength$Strength, ties.method = "average")

# Store strength data frames in a list
strengths_list <- list(
  "Default" = team_strengths,
  "Normal" = normal_strengths,
  "Uniform" = unif_strength
)

```

# Simulating a Tournament

```{r}
# defining the rank value for each finish
rank_values <- c(
  # Add more ranks as needed
  quarterfinalist = 6.5,
  semifinalist = 3.5,
  finalist = 2,
  champion = 1
)

# Simulate the single-elimination tournament
simulate_tournament <- function(df,seeds = NULL) {

  # Sets seed to that of common best vs worst structure
  if (is.null(seeds)){
    seeds <- matrix(c(1,8,2,7,3,6,4,5), nrow = 1)
  }

  results <- list()  # Initialize the results list

  for (i in 1:nrow(seeds)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeds[i, ]
    permuted_df <- df
    permuted_df$Team <- df$Team[permutation]  # Update team order 
    permuted_df$Strength <- df$Strength[permutation]
    permuted_df$Ranks <- NA  # Initialize ranks column
    permuted_df$Wins <- 0  # Initialize wins column
    permuted_df$Exp_Rank <- df$Exp_Rank[permutation]

    teams <- permuted_df$Team
    strengths <- permuted_df$Strength

    round_number <- log2(length(teams))

  while (length(teams) > 1) {
    #cat("\n--- Round", round_number, "---\n")
    next_round <- c() # Makes sure only winners are included in next round

    for (j in seq(1, length(teams), by = 2)) { # Gets 2 teams for each match
      team1 <- teams[j]
      team2 <- teams[j+1]

      strength1 <- strengths[permuted_df$Team == team1]
      strength2 <- strengths[permuted_df$Team == team2]

      match_winner <- simulate_match(team1, team2, strength1, strength2)
      next_round <- c(next_round, match_winner)
      #cat(paste(team1, "vs", team2, "-> Winner:", match_winner, "\n"))

      # Can add more rounds by the same code with rounnd_number increasing by 1 and rank_values[round name]

      # Assign ranks based on the current round
      if (round_number == 3) {  # Quarterfinals
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <- 
          rank_values["quarterfinalist"]
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <-   
          rank_values["quarterfinalist"]
        }
      } else if (round_number == 2) {  # Semifinals
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <-
          rank_values["semifinalist"]
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <-
          rank_values["semifinalist"]
        }
      } else if (round_number == 1) { # Finals
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <- rank_values["finalist"]
        permuted_df$Ranks[permuted_df$Team == team1] <- rank_values["champion"]
        permuted_df$Wins[permuted_df$Team == team1] <-
        permuted_df$Wins[permuted_df$Team == team1] + 1
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <- rank_values["finalist"]
          permuted_df$Ranks[permuted_df$Team == team2] <- rank_values["champion"]
          permuted_df$Wins[permuted_df$Team == team2] <-
          permuted_df$Wins[permuted_df$Team == team2] + 1
        }
      }
    }

    teams <- next_round  # Winners move to the next round
    round_number <- round_number - 1
  }

  results[[i]] <- permuted_df # Store the results of this permutation
  }
  return(do.call(rbind, results))
}

```

# Single Tournament Simulation Results

```{r}
result <- simulate_tournament(team_strengths,seeds)
result[order(-result$Strength),]

norm_result <- simulate_tournament(normal_strengths,seeds)
norm_result[order(-norm_result$Strength),]

unif_result <- simulate_tournament(unif_strength,seeds)
unif_result[order(-unif_result$Strength),]
```

# Multiple Simulations

```{r}
tourn_sims <- function(df, seeds = NULL, replicates = 100) {
  tournament_replicates <- replicate(replicates, {
    # Reset wins and ranks for each tournament
    df$Wins <- rep(0, nrow(df))
    df$Ranks <- rep(NA, nrow(df))

    if (is.null(seeds)) {
      seeds <- matrix(c(1, 8, 2, 7, 3, 6, 4, 5), nrow = 1)
    }

    # Run the tournament
    result <- simulate_tournament(df, seeds)

    return(result[, c("Team", "Strength", "Wins", "Ranks", "Exp_Rank")])  
  }, simplify = FALSE)  # Prevents simplification to atomic vector

  combined_results <- do.call(rbind, tournament_replicates)

  # Calculate average ranks and total wins using dplyr
  results_summary <- combined_results %>%
    group_by(Team, Strength, Exp_Rank) %>%  # Track SeedNum
    summarise(
      Average_Rank = mean(Ranks, na.rm = TRUE),
      Total_Wins = sum(Wins, na.rm = TRUE)
    ) %>%
    arrange(desc(Strength))

  return(results_summary)
}

# Run the simulation for each strength configuration
results_sim <- lapply(strengths_list, tourn_sims)

# Print the results for each configuration
for (name in names(results_sim)) {
  cat("\nResults for strength configuration:", name, "\n")
  print(results_sim[[name]])

  # Calculate Kendall tau correlation
  kendall_cor <- cor(results_sim[[name]]$Average_Rank, results_sim[[name]]$Exp_Rank, method = "kendall", use = "complete.obs")
  cat("Kendall Tau Correlation:", kendall_cor, "\n")

  # Calculate Spearman correlation
  spearman_cor <- cor(results_sim[[name]]$Average_Rank, results_sim[[name]]$Exp_Rank, method = "spearman", use = "complete.obs")
  cat("Spearman Correlation:", spearman_cor, "\n")

  # Create a barplot for total wins
  barplot(
    results_sim[[name]]$Total_Wins,
    names.arg = results_sim[[name]]$Strength,
    main = paste("Total Wins for", name),
    xlab = "Strength",
    col = "blue",  # Added color for better visualization
    border = "black"  # Added border color for bars
  )
}

```

# Spearman Correlation Best and Worst Cases

```{r}
all_results <- function (strength){
    all_seeds <- list()
    total_seeds <- nrow(seeds)  # Store the total number of seeds
    for (i in 1:total_seeds) {
        current_seed <- seeds[i, , drop = FALSE]
        result <- suppressMessages(tourn_sims(team_strengths, current_seed))
        result$SeedNum <- i  # Add seed number to the result
        all_seeds[[paste0("Seed_", i)]] <- result
    }
    return(all_seeds)
}

# Store strength data frames in a list
sims_list <- list(
    "Default" = all_results(team_strengths),
    "Normal" = all_results(normal_strengths),
    "Uniform" = all_results(unif_strength)
)

# Initialize a list to store Spearman correlation results separately
spearman_results_list <- list()

# Loop through each simulation and calculate Spearman correlation
for (name in names(sims_list)) {
    sim_results <- do.call(rbind, sims_list[[name]])

    # Calculate the Spearman correlation only once for the unique ranks
    spearman_cor <- cor(sim_results$Average_Rank, sim_results$Exp_Rank, method = "spearman", use = "complete.obs")

    # Store the overall Spearman correlation result
    spearman_results_list[[name]] <- data.frame(Configuration = name, 
                                                  Spearman_Correlation = spearman_cor)

    # Keep only relevant columns and ensure distinct SeedNum
    distinct_results <- sim_results[, c("SeedNum", "Average_Rank", "Exp_Rank")]

    # Calculate the Spearman correlation for each distinct SeedNum
    distinct_spearman_results <- distinct_results %>%
        group_by(SeedNum) %>%
        summarize(Spearman_Correlation = cor(Average_Rank, Exp_Rank, method = "spearman", use = "complete.obs")) %>%
        ungroup()

    # Get top 5 highest distinct Spearman correlations for this configuration
    top_5 <- distinct_spearman_results %>%
        arrange(desc(Spearman_Correlation)) %>%
        slice_head(n = 5)  # Select top 5

    # Get top 5 lowest distinct Spearman correlations for this configuration
    lowest_5 <- distinct_spearman_results %>%
        arrange(Spearman_Correlation) %>%
        slice_head(n = 5)  # Select lowest 5

    # Print results with seed numbers
    cat(paste("\nTop 5 Distinct Spearman Correlations for configuration:", name, "\n"))
    print(top_5)  # Only SeedNum and Spearman_Correlation

    cat(paste("\nLowest 5 Distinct Spearman Correlations for configuration:", name, "\n"))
    print(lowest_5)  # Only SeedNum and Spearman_Correlation
}

# Print overall Spearman results for all configurations
 for (name in names(spearman_results_list)) {
     cat(paste("\nSpearman Correlation for configuration:", name, "\n"))
     print(spearman_results_list[[name]])
 }

```

### 2024-12-13 paper.qmd
update.

@@
 
+@Glickman2025: Review of methods for estimating team and player strength.
+So this paper doesn't directly relate to our tournaments paper.
+We are interested in structure and this paper is interested in estimating strength.
+However, we are interested in knockout tournements (and other tournament structures) that require seeding, which is estimating player/team strength.
+So estimating team player/team could be considered part of the tournament structure.
+
 Llull (1283)
 
-Kendall and Smith (1940): Judges makes consistent preferences.
+@KendallSmith1940: Judges makes consistent preferences.
+All about paired comparisons.  Doesn't seem to be about tournaments.
 
-Slater's I (1961): Measure of inconsistency
+@Slater1961 Slater's I (1961): Measure of inconsistency
 
@@
 
-Kendall-Wei (1952)
-Elo (1978)
-Glicko (1998) and Glicko2 (2006?): Essentially Bayesian Elo.  
+Kendall-Wei (1952) Elo (1978) Glicko (1998) and Glicko2 (2006?): Essentially Bayesian Elo.\
 Trueskill (2006): Used in halo 2 or 3.
@@
 
-https://www.sciencedirect.com/science/article/pii/S0378375803003215
-"Maximum weight perfect match problem"
+https://www.sciencedirect.com/science/article/pii/S0378375803003215 "Maximum weight perfect match problem"
 
@@
 -   Knockout (KO) lKO
+
 -   1KO
+
 -   2KO
+
     -   True 3rd, college wrestling, repechage
+
 -   3KO
+
 -   Seeding issues,.
@@
 -   nRR
+
 -   1RR
+
 -   2RR
+
 -   partial RR
@@
 -   Single elimination (NCAA tournament, NFL playoffs)
-  - Seeding issues (Tennis: https://www.liveabout.com/definition-of-seeding-3207821)
-   Reseeding after each round.  Straight knockout.  
-  - Second faces Third for silver medal.  Guarantees than sivler has 1 loss and bronze has 2 losses.  
+
+-   Seeding issues (Tennis: https://www.liveabout.com/definition-of-seeding-3207821) Reseeding after each round. Straight knockout.\
+
+-   Second faces Third for silver medal. Guarantees than sivler has 1 loss and bronze has 2 losses.\
+
 -   With a 3rd place match (Badminton: https://www.nbcolympics.com/news/badminton-101-olympic-competition-format)
+
 -   Single elimination with series (NBA, MLB, NHL)
+
 -   Ladder tournaments (Korean baseball does best of 3 with one team starting up 1-0 in the opening round)
+
 -   Seeding issues (NFL re-seeds, NCAA tournament straight bracket)
+
 -   Group/Pool play followed by knockout stages (World Cup, Euro, Overwatch World CUp. )
@@
 -   Double elimination (Overwatch Champions League)
+
 -   Repechage (Olympic Wrestling, Tae Kwon Do, Judo) (These also aware 2 bronze medals)
+
 -   True Double elimination
+
 -   Double elimination but a single loss before finals means you can onyl go for third place (College wrestling)
@@
 -   Round Robin
-  - Single and Double.  
+
+-   Single and Double.\
+
 -   Baseball is an imbalanced multiple round robin qualifying round.
@@
 -   Waterfall format
+
     -   Introduced in the Smash 64 community, [they claim](https://thesmashwriter.wordpress.com/2017/04/24/what-is-a-waterfall-bracket/)
@@
 
- Tournament: a series of contests between a number of competitors, who compete for an overall prize. Oxford Dictionary
-  Wiki: "A tournament is a competition involving at least three competitors, all participating in a sport or game"
+Tournament: a series of contests between a number of competitors, who compete for an overall prize.
+Oxford Dictionary Wiki: "A tournament is a competition involving at least three competitors, all participating in a sport or game"
 
-Tournament (graphy theory): [wikipedia definition](https://en.wikipedia.org/wiki/Tournament_(graph_theory)) and in [McShane (2019)](https://scholar.smu.edu/hum_sci_statisticalscience_etds/13/) -- essentially a "round robin tournament" -- a graph where every node (competitor: A, B, C) is connected by a directed edge (win A -> B or lose A <- B), where there are no loops (self-directed edges -- competitors cannot play against themselves).
+Tournament (graphy theory): [wikipedia definition](https://en.wikipedia.org/wiki/Tournament_(graph_theory)) and in [McShane (2019)](https://scholar.smu.edu/hum_sci_statisticalscience_etds/13/) -- essentially a "round robin tournament" -- a graph where every node (competitor: A, B, C) is connected by a directed edge (win A -\> B or lose A \<- B), where there are no loops (self-directed edges -- competitors cannot play against themselves).
 AKA a complete directed graph with no loops; has $n$ nodes and $\choose{n}{2}$ edges.
@@
 Moon's [Topic on Tournaments](https://www.gutenberg.org/ebooks/42833) is the classic graph theoretical tournaments text.
-This definition should be used to disambiguate _tournament_, especially since graph theory may be used.
+This definition should be used to disambiguate *tournament*, especially since graph theory may be used.
 
- (https://www.espn.com/olympics/story/_/id/40536088/2024-paris-summer-olympics-track-field-medaling-format-rules)
+(https://www.espn.com/olympics/story/\_/id/40536088/2024-paris-summer-olympics-track-field-medaling-format-rules)
 
@@
 
-Graph Theory Stuff: 
-Chickens: https://link.springer.com/article/10.1007/BF02476378
+Graph Theory Stuff: Chickens: https://link.springer.com/article/10.1007/BF02476378
 
@@
 
-https://liquipedia.net/overwatch/Overwatch_Champions_Series/2024/North_America/Stage_2
-Swiss, Top Cut Swiss, Group play then a double elimiation bracket!
+https://liquipedia.net/overwatch/Overwatch_Champions_Series/2024/North_America/Stage_2 Swiss, Top Cut Swiss, Group play then a double elimiation bracket!
 
@@
 
-Note for Greg: 
-Can we build a simulator that computes the probability of a team winning in two steps: 1) the team strength comes from a distribution itself and then the actual outcome come from flipping a coin based on the particular draw of the distribtuion.
+Note for Greg: Can we build a simulator that computes the probability of a team winning in two steps: 1) the team strength comes from a distribution itself and then the actual outcome come from flipping a coin based on the particular draw of the distribtuion.
 

### 2024-12-18 single_elimination_shiny.qmd
Updated Code 12/18/2024

---
title: "single_elimination"
format: 
  html:
    embed-resources: true
editor: visual
---

```{r}
library(tidyverse)
library(dplyr)
library(shiny)
```

# Simulate a Single Match

```{r}
# Function to simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strength1, strength2) {
  if (is.na(team1)) {
    return(team2)
  } else if (is.na(team2)) {
    return(team1)
  } else {
    p <- exp(strength1 - strength2) / (1 + exp(strength1 - strength2))
    return(ifelse(runif(1) < p, team1, team2))  # Randomize outcome
  }
}

```

# Single Elimination Tournament Structures

## Creating the traditional tournament structure for n teams

```{r}
generate_tournament_structure <- function(n) {

  is_power_of_two <- function(x) {
    x > 0 && (x & (x - 1)) == 0
  }

  # If n is not a power of 2, round up to the next power of 2 and add placeholders
  if (!is_power_of_two(n)) {
    next_power_of_two <- 2^ceiling(log2(n))
    extra_teams <- next_power_of_two - n
    #message(sprintf("Rounding up to %d teams. Adding %d NA placeholders.", next_power_of_two, extra_teams))
    n <- next_power_of_two
  } else {
    extra_teams <- 0
  }

  rounds <- log(n, base = 2) - 1
  teams <- c(1, 2)  # Initial list of teams

  for (i in 1:rounds) {
    teams <- nextLayer(teams)
  }

  while (extra_teams > 0){
      teams[which.max(teams)] <- NA
      extra_teams <- extra_teams - 1
  }

  return(teams)
}

nextLayer <- function(teams) {
  out <- c()  # Initialize an empty vector
  length <- length(teams) * 2 + 1  # Calculate the length for pairing

  # Generate the next layer by pairing teams
  for (i in teams) {
    out <- c(out, i, length - i)  # Push the current team and its pair
  }

  return(out)
}

```

# Simulating a Tournament

```{r}
# Simulate the single-elimination tournament
simulate_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1, third_place = T) {

if (num_teams <= 3) {
  stop("Number of Teams must be greater than 3.")
}
teams <- paste("Team",num_teams:1)
if (log2(num_teams) %% 1 != 0) {
  power_of_2 <- 2^ceiling(log2(num_teams))
  teams <- c(teams, rep(NA, power_of_2 - num_teams))
}
if (distribution == "Normal"){
  strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
  normal_strengths <- data.frame(
  Team = teams,
  Strength = c(strengths, rep(NA, length(teams) - num_teams)),
  Wins = rep(0,length(teams)),
  Ranks = rep(NA,length(teams))
)
normal_strengths$True_Rank <- rank(-normal_strengths$Strength, ties.method = "average")
df <- arrange(normal_strengths, True_Rank)
}
else if (distribution == "Uniform"){
  strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
  unif_strength <- data.frame(
  Team = teams,
  Strength = c(strengths, rep(NA, length(teams) - num_teams)),
  Wins = rep(0,length(teams)),
  Ranks = rep(NA,length(teams))
)
unif_strength$True_Rank <- rank(-unif_strength$Strength, ties.method = "average")
df <- arrange(unif_strength, True_Rank)
}
else {
  stop("Distribution must be Normal or Uniform")
}

  # Sets seed to that of common best vs worst structure
  if (is.null(seeding_structure)){
    seeding_structure <- matrix(generate_tournament_structure(num_teams), nrow = 1)
  } else{
    seeding_structure <- matrix(seeding_structure, nrow = 1)
  }
  if(series %% 2 == 0) {
    stop("Series must be an odd number to avoid potential ties.")
  }

  results <- list()  # Initialize the results list

  for (i in 1:nrow(seeding_structure)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeding_structure[i, ]
    permuted_df <- df
    permuted_df$Team <- df$Team[permutation]  # Update team order 
    permuted_df$Strength <- df$Strength[permutation]
    permuted_df$Ranks <- NA  # Initialize ranks column
    permuted_df$Wins <- 0  # Initialize wins column
    permuted_df$True_Rank <- df$True_Rank[permutation]

    teams <- permuted_df$Team
    strengths <- permuted_df$Strength
    round_number <- ceiling(log2(length(teams)))

    # Track the losers of the semifinals for the third-place match
    losers_semis <- c()

  while (length(teams) > 1) {
    #cat("\n--- New Round ---\n")
    next_round <- c() # Makes sure only winners are included in next round

    for (j in seq(1, length(teams), by = 2)) { # Gets 2 teams for each match
      team1 <- teams[j]
      team2 <- teams[j+1]

      strength1 <- strengths[which(permuted_df$Team == team1)]
      strength2 <- strengths[which(permuted_df$Team == team2)]

      wins_team1 <- 0
      wins_team2 <- 0
      for (game in 1:series){
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      #cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", 
      #      team1, team2, strength1, strength2, match_winner))
      if (match_winner == team1) {
            wins_team1 <- wins_team1 + 1
          } else {
            wins_team2 <- wins_team2 + 1
          }

          # Determine the overall winner if a majority is reached
          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
            break
          }
        }

        # Determine the match winner based on series results
        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
      next_round <- c(next_round, series_winner)

      # Function to calculate the rank values for a given round
    get_rank_values <- function(n) {
      # Calculate the rank values as the average of the last n/2 teams
      suppressWarnings({
        if (n / 2 + 1 == num_teams) {
          rank_values <- num_teams
        } else {
          rank_values <- (n / 2 + 1):min(num_teams, n)
        }
      })
      return(rank_values)
    }
      if (round_number > 2){
      # Assign ranks based on the current round
        n <- 2^(round_number)
        rank_values <- suppressWarnings(get_rank_values(n))
        #print(match_winner)
        if (match_winner == team1) {
          if (ties == TRUE){
          permuted_df$Ranks[permuted_df$Team == team2] <- suppressWarnings(mean(rank_values))
          }
          else if (ties == FALSE){
          permuted_df$Ranks[permuted_df$Team == team2] <- n
          }
        } else {
          if (ties == TRUE){
          permuted_df$Ranks[permuted_df$Team == team1] <-  suppressWarnings(mean(rank_values))
          }
          else if (ties == FALSE){
          permuted_df$Ranks[permuted_df$Team == team1] <- n
          }
        }
      } else if (round_number == 2) {  # Semifinals
        n <- 2^(round_number)
        rank_values <- get_rank_values(n)
        if (match_winner == team1) {
          if (ties == TRUE){
          permuted_df$Ranks[permuted_df$Team == team2] <-
          mean(rank_values)
          }
          else if (ties == FALSE & third_place == FALSE){
          permuted_df$Ranks[permuted_df$Team == team2] <- n
          }
          losers_semis <- c(losers_semis, team2)
        } else {
          if (ties == TRUE){
          permuted_df$Ranks[permuted_df$Team == team1] <-
          mean(rank_values)
          }
          else if (ties == FALSE & third_place == FALSE){
          permuted_df$Ranks[permuted_df$Team == team1] <- n
          }
          losers_semis <- c(losers_semis, team1)

          }
        }else if (round_number == 1) { # Finals
          permuted_df <- permuted_df %>%
            filter(!is.na(Team))
           n <- 2^(round_number)
        rank_values <- get_rank_values(n)
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <- 2
        permuted_df$Ranks[permuted_df$Team == team1] <- 1
        permuted_df$Wins[permuted_df$Team == team1] <-
        permuted_df$Wins[permuted_df$Team == team1] + 1
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <- 2
          permuted_df$Ranks[permuted_df$Team == team2] <- 1
          permuted_df$Wins[permuted_df$Team == team2] <-
          permuted_df$Wins[permuted_df$Team == team2] + 1
        }
      }
    }

    teams <- next_round  # Winners move to the next round
    round_number <- round_number - 1
  }

  # Third place match between the losers of the semifinals
  if (third_place == TRUE){
    team1 <- losers_semis[1]
    team2 <- losers_semis[2]
    strength1 <- permuted_df$Strength[permuted_df$Team == team1]
    strength2 <- permuted_df$Strength[permuted_df$Team == team2]
    wins_team1 <- 0
    wins_team2 <- 0
    for (game in 1:series) {
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      if (match_winner == team1) {
        wins_team1 <- wins_team1 + 1
      } else {
        wins_team2 <- wins_team2 + 1
      }
     # Determine the overall winner if a majority is reached
      if (wins_team1 > series / 2 || wins_team2 > series / 2) {
      break
      }
    }
   # Assign third place and fourth place ranks
      if (wins_team1 > wins_team2) {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team1))
        permuted_df$Ranks[permuted_df$Team == team1] <- 3
        permuted_df$Ranks[permuted_df$Team == team2] <- 4
      } else {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team2))
        permuted_df$Ranks[permuted_df$Team == team2] <- 3
        permuted_df$Ranks[permuted_df$Team == team1] <- 4
      }
    }

  results[[i]] <- permuted_df # Store the results of this permutation
  }

  final_results <- do.call(rbind, results)

  # Generalized function to remove ties by assigning unique ranks
remove_ties <- function(df) {
  suppressWarnings({
  df <- df %>%
    group_by(Ranks) %>%
    mutate(
      Ranks = ifelse(
        Ranks == 2,  # Do not change if rank is 2
        Ranks, 
        ifelse(
          log2(Ranks) %% 1 == 0,  # Check if Ranks is a power of 2
          ifelse(
            Ranks / 2 + 1 == num_teams,  
            num_teams,  
            sample((Ranks / 2 + 1):(min(Ranks, num_teams)), n(), replace = FALSE)  # Regular case
          ),
          Ranks  # Keep Ranks as is if it's not a power of 2
        )
      )
    ) %>%
    ungroup()
  })
  return(df)
}

  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- remove_ties(final_results)
  }

  final_results <- final_results[order(final_results$Ranks),]
  return(final_results)
}

test <- simulate_tournament(num_teams = 10, distribution = "Normal", ties=F, series=3, third_place=T, seeding_structure = c(1,NA,2,NA,3,NA,4,5,6,7,8,NA,9,NA,10,NA))
test

```

## Shiny User Interface Tournament Simulation

```{r}
# Define the Shiny app UI
ui <- fluidPage(
  titlePanel("Single-Elimination Tournament Simulator"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 4, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Tournament Results"),
      tableOutput("ranked_results")
    )
  )
)

# Define the server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_tournament_structure(input$num_teams))
    }
  })

  tournament_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs

    seed_structure <- custom_seeding()

    simulate_tournament(
      num_teams = input$num_teams,
      distribution = input$distribution,
      ties = input$ties,
      series = input$series,
      third_place = input$third_place,
      seeding_structure = seed_structure
    )
  })

  output$ranked_results <- renderTable({
    results <- tournament_results()
    results[order(-results$Strength), ]  # Order by Strength descending
  })
}

shinyApp(ui = ui, server = server)
```

## Shiny Number of Wins

```{r}
# Shiny app UI
ui <- fluidPage(
  titlePanel("Tournament Simulation with Total Wins Visualization"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Total Wins Summary"),
      plotOutput("wins_plot")
    )
  )
)

# Shiny server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_tournament_structure(input$num_teams))
    }
  })

  simulation_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
    seed_structure <- custom_seeding()
    # Run simulations
    replicate_results <- replicate(
      input$num_simulations,
      simulate_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        third_place = input$third_place,
        seeding_structure = seed_structure
      ),
      simplify = FALSE
    )

    # Combine all results and summarize total wins
    all_results <- do.call(rbind, replicate_results)
    total_wins_summary <- all_results %>%
      group_by(Team) %>%
      summarise(
        Total_Wins = sum(Wins, na.rm = TRUE),
        .groups = 'drop'
      ) %>%
      arrange(desc(Total_Wins))
    return(total_wins_summary)
  })

  output$wins_plot <- renderPlot({
    req(simulation_results())
    results <- simulation_results()

    ggplot(results, aes(x = reorder(Team, -Total_Wins), y = Total_Wins, fill = Total_Wins)) +
      geom_text(aes(label = Total_Wins), vjust = -0.5, size = 4, color = "black") +
      geom_bar(stat = "identity") +
      labs(
        title = "Total Wins Across Simulations",
        x = "Team",
        y = "Total Wins"
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_fill_gradient(low = "skyblue", high = "blue")
  })
}

# Run the Shiny app
shinyApp(ui = ui, server = server)
```

# Multiple Simulations for 1 Seeding Structure

```{r}
library(shiny)

ui <- fluidPage(
  titlePanel("Tournament Simulation"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", 8, min = 4, max = 32, step = 1),
      selectInput("distribution", "Distribution Type:", choices = c("Normal", "Uniform")),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
      numericInput("replicates", "Number of Replicates:", 100, min = 1, max = 1000, step = 1),
      numericInput("series", "Number of Games in a Series (odd number):", 5, min = 3, max = 11, step = 2),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Tournament Simulation")
    ),

    mainPanel(
      tableOutput("summary_results"),
      verbatimTextOutput("kendall_correlation"),
      verbatimTextOutput("spearman_correlation")
    )
  )
)

server <- function(input, output, session) {
 custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_tournament_structure(input$num_teams))
    }
  })

  # Function to run tournament simulations
  observeEvent(input$run_sim, {
    req(input$num_teams, input$distribution, input$replicates, input$series)

    # Initialize lists to store correlation results
    kendall_correlations <- numeric(input$replicates)
    spearman_correlations <- numeric(input$replicates)

    seed_structure <- custom_seeding()

    # Run tournament simulations for the specified number of replicates
    simulation_results <- list()
    for (replicate in 1:input$replicates) {
      simulation_results[[replicate]] <- simulate_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        third_place = input$third_place,
        seeding_structure = seed_structure
      )

      # Combine all results into one data frame
      all_results <- do.call(rbind, simulation_results)

      # Compute average ranks across all replicates
      average_ranks <- all_results %>%
        group_by(Team, True_Rank) %>%
        summarise(Average_Rank = mean(Ranks, na.rm = TRUE), .groups = 'drop') %>%
        arrange(Team)

      # Calculate Kendall Tau correlation for the current replicate
      kendall_correlations[replicate] <- cor(average_ranks$Average_Rank, average_ranks$True_Rank, method = "kendall", use = "complete.obs")

      # Calculate Spearman correlation for the current replicate
      spearman_correlations[replicate] <- cor(average_ranks$Average_Rank, average_ranks$True_Rank, method = "spearman", use = "complete.obs")
    }

    # Compute average Kendall Tau and Spearman correlations across all replicates
    avg_kendall_cor <- mean(kendall_correlations, na.rm = TRUE)
    avg_spearman_cor <- mean(spearman_correlations, na.rm = TRUE)

    # Show summary results in table
    output$summary_results <- renderTable({
      average_ranks
    })

    # Display average Kendall Tau correlation
    output$kendall_correlation <- renderPrint({
      cat("Average Kendall Tau Correlation:", avg_kendall_cor, "\n")
    })

    # Display average Spearman correlation
    output$spearman_correlation <- renderPrint({
      cat("Average Spearman Correlation:", avg_spearman_cor, "\n")
    })
  })
}

shinyApp(ui, server)

```

## Shiny Probability of Top 3 Correct

```{r}
ui <- fluidPage(
  titlePanel("Tournament Simulation with Total Wins Visualization"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Top 3 Correct"),
      verbatimTextOutput("top3_output")
    )
  )
)

# Shiny server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_tournament_structure(input$num_teams))
    }
  })

  simulation_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
    seed_structure <- custom_seeding()

    # Run simulations
    replicate_results <- replicate(
      input$num_simulations,
      simulate_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        third_place = input$third_place,
        seeding_structure = seed_structure
      ),
      simplify = FALSE
    )

    # Combine all results into one dataframe and add a Simulation column
    all_results <- do.call(rbind, replicate_results)
    all_results$Simulation <- rep(1:input$num_simulations, each = nrow(all_results) / input$num_simulations)

    return(all_results)
  })

  # Identify top-3 ranks for each simulation and calculate the probability
  top_three_probability <- reactive({
    all_results <- simulation_results()

    # Group by Simulation and filter for top 3 True_Rank and Ranks
    top_three_results <- all_results %>%
      group_by(Simulation) %>%
      filter(True_Rank %in% 1:3 & Ranks %in% 1:3) %>%
      arrange(Simulation, Ranks) %>%
      summarize(top_3_true_rank = list(True_Rank), top_3_rank = list(Ranks), .groups = 'drop') %>%
      mutate(match = mapply(function(true, rank) setequal(true, rank), top_3_true_rank, top_3_rank))

    # Calculate probability of matching top 3 True_Rank with Ranks
    match_count <- sum(top_three_results$match)
    total_simulations <- n_distinct(all_results$Simulation)

    probability <- match_count / total_simulations
    return(probability)
  })

  # Output the probability to the UI
  output$top3_output <- renderText({
    probability <- top_three_probability()

    paste("Probability of Top 3 Seeds with Ranks 1-3: ", round(probability,4))
  })
}

shinyApp(ui=ui,server=server)
```

## Shiny Conditional Probability Curve

```{r}
ui <- fluidPage(
  titlePanel("Conditional Probability of Team Ranks in Tournament"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", 
                placeholder = "e.g. 1 8 4 5 3 6 7 2"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      plotOutput("probability_plot")
    )
  )
)

# Server
server <- function(input, output) {

  # Helper Function: Parse Custom Seeding Structure
  parse_seeding <- function(seeding_input, num_teams) {
    if (seeding_input != "") {
      seeding_vector <- unlist(strsplit(seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)

      num_teams_next_power_of_two <- 2^ceiling(log2(num_teams))
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(
          paste("Seeding structure length must match", num_teams_next_power_of_two), 
          type = "error"
        )
        return(NULL)
      }
    } else {
      return(NULL)
    }
  }

  # Curve Function to Calculate Conditional Probabilities
  curve <- function(sim, num_teams) {
    probability <- numeric(num_teams)
    for (i in 1:num_teams) {
      if (i == 1) {
        probability[i] <- nrow(sim[sim$Ranks == 1 & sim$True_Rank == 1, ]) / length(unique(sim$Simulation))
      } else {
        top_ranks <- sim[sim$Ranks <= (i - 1) & sim$True_Rank <= (i - 1), ]
        better_teams <- 1:(i - 1)
        numerator <- sum(sapply(unique(top_ranks$Simulation), function(sim_id) {
          sim_subset <- sim[sim$Simulation == sim_id, ]
          if (all(better_teams %in% sim_subset$True_Rank[sim_subset$Ranks <= (i - 1)])) {
            return(sum(sim_subset$Ranks == i & sim_subset$True_Rank == i))
          } else {
            return(0)
          }
        }))
        denominator <- sum(sapply(unique(top_ranks$Simulation), function(sim_id) {
          sim_subset <- top_ranks[top_ranks$Simulation == sim_id, ]
          all(better_teams %in% sim_subset$True_Rank)
        }))
        probability[i] <- numerator / denominator
      }
    }
    return(probability)
  }

  # Reactive Event: Run Simulations
  simulation_data <- eventReactive(input$run_sim, {
    num_teams <- input$num_teams
    distribution <- input$distribution
    series <- input$series
    num_simulations <- input$num_simulations
    third_place <- input$third_place

    # Parse custom seeding structure
    seeding_structure <- parse_seeding(input$seeding_input, num_teams)

    # Run simulations
    replicate_results <- replicate(
      num_simulations,
      simulate_tournament(
        num_teams = num_teams,
        distribution = distribution,
        series = series,
        ties = FALSE,
        third_place = third_place,
        seeding_structure = seeding_structure
      ),
      simplify = FALSE
    )

    # Combine all results into one dataframe
    all_results <- do.call(rbind, replicate_results)
    all_results$Simulation <- rep(1:num_simulations, each = nrow(all_results) / num_simulations)

    # Calculate conditional probabilities
    sim_probs <- curve(all_results, num_teams)
    probabilities <- data.frame(x = 1:num_teams, y = sim_probs)
    probabilities$Line <- "Simulated"

    # Baseline probabilities
    baseline_prob <- 1 / (num_teams:1)
    baseline <- data.frame(x = 1:num_teams, y = baseline_prob)
    baseline$Line <- "Baseline"

    # Combine for plotting
    rbind(baseline, probabilities)
  })

  # Plot Output
  output$probability_plot <- renderPlot({
    req(simulation_data())
    ggplot(data = simulation_data(), aes(x = x, y = y, color = Line, group = Line)) +
      geom_point() +
      geom_line() +
      ylim(0, 1) +
      labs(
        x = "Rank",
        y = "Conditional Probability",
        title = "Simulated vs Baseline Conditional Probabilities"
      ) +
      theme_minimal()
  })
}

# Run the App
shinyApp(ui = ui, server = server)
```

### 2025-01-04 double_elimination.qmd
Updated code 1/4/2025

---
title: "Double Elimination Tournament"
author: "Zach Culp"
format: html
editor: visual
---

# Double Elimination Tournament

```{r}
library(tidyverse)
library(dplyr)
library(shiny)
```

## Simulate a Single Match

```{r}
# Function to simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strength1, strength2) {
  if(is.na(team1) && is.na(team2)){
    return(NA)
  }
  if (is.na(team1)) {
    return(team2)
  } else if (is.na(team2)) {
    return(team1)
  } else {
    p <- exp(strength1 - strength2) / (1 + exp(strength1 - strength2))
    return(ifelse(runif(1) < p, team1, team2))  # Randomize outcome
  }
}
```

## Creating the traditional tournament structure for n teams

```{r}
generate_elimination_tournament_structure <- function(n) {

  is_power_of_two <- function(x) {
    x > 0 && (x & (x - 1)) == 0
  }

  # If n is not a power of 2, round up to the next power of 2 and add placeholders
  if (!is_power_of_two(n)) {
    next_power_of_two <- 2^ceiling(log2(n))
    extra_teams <- next_power_of_two - n
    #message(sprintf("Rounding up to %d teams. Adding %d NA placeholders.", next_power_of_two, extra_teams))
    n <- next_power_of_two
  } else {
    extra_teams <- 0
  }

  rounds <- log(n, base = 2) - 1
  teams <- c(1, 2)  # Initial list of teams

  for (i in 1:rounds) {
    teams <- nextLayer(teams)
  }

  left_half <- teams[1:(length(teams)/2)]
  right_half <- teams[(length(teams)/2+1):length(teams)]
  right_half <- rev(right_half)
  teams <- c(left_half,right_half)

  # Ensure that the lower number team is on the left side of each matchup
  teams <- unlist(lapply(seq(1, length(teams), by = 2), function(i) {
    if (teams[i] > teams[i + 1]) {
      c(teams[i + 1], teams[i])  # Swap if needed
    } else {
      c(teams[i], teams[i + 1])
    }
  })
  )

  while (extra_teams > 0){
      teams[which.max(teams)] <- NA
      extra_teams <- extra_teams - 1
  }

  return(teams)
}

nextLayer <- function(teams) {
  out <- c()  # Initialize an empty vector
  length <- length(teams) * 2 + 1  # Calculate the length for pairing

  # Generate the next layer by pairing teams
  for (i in teams) {
    out <- c(out, i, length - i)  # Push the current team and its pair
  }

  return(out)
}

```

## Creating a Single Simulation of a Double Elimination Tournament

```{r}
simulate_double_elimination_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1) {

  if (num_teams <= 3) {
    stop("Number of Teams must be greater than 3.")
  }
  teams <- paste(num_teams:1)
  if (log2(num_teams) %% 1 != 0) {
    power_of_2 <- 2^ceiling(log2(num_teams))
    teams <- c(teams, rep(NA, power_of_2 - num_teams))
  }
  if (distribution == "Normal"){
    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Uniform"){
    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
    unif_strength <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams)),
      rank_hat = rep(NA,length(teams))
    )
    df <- arrange(unif_strength, true_rank)
  }
     else if (distribution == "Manual") {
    cat("Please enter the strengths for each team in descending rank order:\n")
    strengths <- numeric(num_teams)
    for (i in num_teams:1) {
      repeat {
        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
        if (!is.na(input)) {
          strengths[i] <- input
          break
        } else {
          cat("Invalid input. Please enter a numeric value.\n")
        }
      }
    }
    manual_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(manual_strengths, true_rank)
  }
  else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

  # Sets seed to that of common best vs worst structure
  if (is.null(seeding_structure)){
    seeding_structure <- matrix(generate_elimination_tournament_structure(num_teams), nrow = 1)
  } else{
    seeding_structure <- matrix(seeding_structure, nrow = 1)
  }
  if(series %% 2 == 0) {
    stop("Series must be an odd number to avoid potential ties.")
  }

  results <- list()  # Initialize the results list

  for (i in 1:nrow(seeding_structure)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeding_structure[i, ]
    permuted_df <- df
    permuted_df$true_rank <- df$true_rank[permutation]  # Update team order 
    teams <- permuted_df$true_rank
    permuted_df$true_strength <- df$true_strength[permutation]
    permuted_df$rank_hat <- NA  # Initialize ranks column
    permuted_df$game_wins <- 0  # Initialize wins column
    permuted_df$game_losses <- 0

    team_rank <- num_teams
    winners_bracket <- teams
    strengths <- permuted_df$true_strength

    losers_bracket <- c()

    # gets results of the winners bracket
    while (length(winners_bracket) > 1 || length(losers_bracket > 0)) {
      eliminated <- c()
      if (length(winners_bracket) > 1) {
      #cat("\n--- Winner's Bracket --- \n")
      next_winners <- c() 
      round_losers <- c()

      for (j in seq(1, length(winners_bracket), by = 2)) { # Gets 2 teams for each match
        if (j + 1 > length(winners_bracket)) break
        team1 <- winners_bracket[j]
        team2 <- winners_bracket[j+1]

        strength1 <- strengths[which(permuted_df$true_rank == team1)]
        strength2 <- strengths[which(permuted_df$true_rank == team2)]

        wins_team1 <- 0
        wins_team2 <- 0
        losses_team1 <- 0
        losses_team2 <- 0
        for (game in 1:series){
          match_winner <- simulate_match(team1, team2, strength1, strength2)
          #cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
          if(!is.na(team1)) {
          if (match_winner == team1) {
            wins_team1 <- wins_team1 + 1
            losses_team2 <- losses_team2 + 1
          } else {
            wins_team2 <- wins_team2 + 1
            losses_team1 <- losses_team1 + 1
          }
          }
          else {wins_team2 <- (series+1) / 2}

          # Determine the overall winner if a majority is reached
          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
            break
          }
        }
        # Update the game_wins and game_losses outside of the loop
        if (!any(is.na(c(team1, team2)))) {
          #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
        permuted_df <- permuted_df %>%
          mutate(game_wins = case_when(
            true_rank == team1 ~ game_wins + wins_team1,
            true_rank == team2 ~ game_wins + wins_team2,
            TRUE ~ game_wins
          )) %>%
          mutate(game_losses = case_when(
            true_rank == team1 ~ game_losses + losses_team1,
            true_rank == team2 ~ game_losses + losses_team2,
            TRUE ~ game_losses
          ))
        }
        # Determine the match winner based on series results
        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
        series_loser <- ifelse(series_winner == team1, team2, team1)
        #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n: | Loser: %s\n", team1, team2, series_winner, series_loser))
        next_winners <- c(next_winners, series_winner)
        round_losers <- c(round_losers, series_loser)
    }
     winners_bracket <- next_winners  # Winners move to the next round
     losers_bracket <- c(losers_bracket, round_losers)
     if(length(losers_bracket) %% 2 != 0){losers_bracket <- c(losers_bracket, NA)}

      }

    # Losers' bracket
    if (length(losers_bracket) > 0) {
      if(length(losers_bracket) > 1){
      #cat("\n--- Loser's Bracket --- \n")
      }
      next_losers <- c()

      # alternates games for losers bracket
      if(length(losers_bracket) > 2){
        if(length(losers_bracket) == 4 && is.na(losers_bracket[4])){
          losers_bracket <- losers_bracket
        } else{
          top_bracket <- losers_bracket[seq(1,length(losers_bracket), by = 2)]
          bottom_bracket <- losers_bracket[seq(2,length(losers_bracket), by = 2)]
          losers_bracket <- c(top_bracket,bottom_bracket)
        }
      }

      for (j in seq(1, length(losers_bracket), by = 2)) {
        if (j + 1 > length(losers_bracket)) break
        team1 <- losers_bracket[j]
        team2 <- losers_bracket[j + 1]
        strength1 <- strengths[which(permuted_df$true_rank == team1)]
        strength2 <- strengths[which(permuted_df$true_rank == team2)]

        wins_team1 <- 0
        wins_team2 <- 0
        losses_team1 <- 0
        losses_team2 <- 0

        # Play series
        for (game in 1:series) {
          match_winner <- simulate_match(team1, team2, strength1, strength2)
          if(is.na(match_winner)){
            wins_team1 == wins_team1 + 1
          }
          if(!is.na(team1)) {
            if (match_winner == team1) {
              wins_team1 <- wins_team1 + 1
              losses_team2 <- losses_team2 + 1
            } else {
              wins_team2 <- wins_team2 + 1
              losses_team1 <- losses_team1 + 1
            }
          }
          else {wins_team2 <- (series+1) / 2}

          # Determine overall series winner
          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
            break
          }
        }
        if (!any(is.na(c(team1, team2)))) {
          #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
          permuted_df <- permuted_df %>%
          mutate(game_wins = case_when(
            true_rank == team1 ~ game_wins + wins_team1,
            true_rank == team2 ~ game_wins + wins_team2,
            TRUE ~ game_wins
          )) %>%
          mutate(game_losses = case_when(
            true_rank == team1 ~ game_losses + losses_team1,
            true_rank == team2 ~ game_losses + losses_team2,
            TRUE ~ game_losses
          ))
        }

        if(is.na(team1) && is.na(team2)){
          series_winner <- NA
          next_losers <- c(next_losers, NA)
        } else{
        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
        series_loser <- ifelse(wins_team1 > wins_team2, team2, team1)
        next_losers <- c(next_losers, series_winner)
        eliminated <- c(eliminated, series_loser)
        #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
        }
      }
        team_rank <- team_rank - 1
        permuted_df$rank_hat[permuted_df$true_rank %in% eliminated] <- team_rank
        losers_bracket <- next_losers

    }

      # Final match for loser's bracket (aka third place game)
if (length(winners_bracket) == 1 && length(losers_bracket) == 1) {

# Final matchup between winners and losers
    #cat("\n--- Final Match ---\n")
    final_match <- c(winners_bracket, losers_bracket)
    if(length(final_match) == 2){
    team1 <- final_match[1]
    team2 <- final_match[2]
    strength1 <- strengths[which(permuted_df$true_rank == team1)]
    strength2 <- strengths[which(permuted_df$true_rank == team2)]

    wins_team1 <- 0
    wins_team2 <- 0
    losses_team1 <- 0
    losses_team2 <- 0

    for (game in 1:series) {
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      if (match_winner == team1) {
        wins_team1 <- wins_team1 + 1
        losses_team2 <- losses_team2 + 1
      } else {
        wins_team2 <- wins_team2 + 1
        losses_team1 <- losses_team1 + 1
      }

      if (wins_team1 > series / 2 || wins_team2 > series / 2) {
        break
      }
    }
    if (!any(is.na(c(team1, team2)))) {
      #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
      permuted_df <- permuted_df %>%
        mutate(game_wins = case_when(
          true_rank == team1 ~ game_wins + wins_team1,
          true_rank == team2 ~ game_wins + wins_team2,
          TRUE ~ game_wins
        )) %>%
        mutate(game_losses = case_when(
          true_rank == team1 ~ game_losses + losses_team1,
          true_rank == team2 ~ game_losses + losses_team2,
          TRUE ~ game_losses
          ))
    }

    if (wins_team1 > wins_team2) {
      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
      permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
      permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
    } else {
      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
      # second game if the loser's bracket team wins
      for (game in 1:series) {
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      if (match_winner == team1) {
        wins_team1 <- wins_team1 + 1
        losses_team2 <- losses_team2 + 1
      } else {
        wins_team2 <- wins_team2 + 1
        losses_team1 <- losses_team1 + 1
      }

      if (wins_team1 > series / 2 || wins_team2 > series / 2) {
        break
      }
    }
      if (!any(is.na(c(team1, team2)))) {
        #cat(sprintf("Wins - %s: %d | %s: %d\n", team1, wins_team1, team2, wins_team2))
        permuted_df <- permuted_df %>%
          mutate(game_wins = case_when(
            true_rank == team1 ~ game_wins + wins_team1,
            true_rank == team2 ~ game_wins + wins_team2 - 1, # wins_team2 = 2 in loop but only need to add one game
            TRUE ~ game_wins
          )) %>%
          mutate(game_losses = case_when(
            true_rank == team1 ~ game_losses + losses_team1 - 1,
            true_rank == team2 ~ game_losses + losses_team2,
            TRUE ~ game_losses
          ))
      }
      if (wins_team1 > wins_team2) {
      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
      permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
      permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
    } else {
      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
      permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
      permuted_df$rank_hat[permuted_df$true_rank == team1] <- 2
        }
      }
    }
  }
}
    permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
    results[[i]] <- permuted_df
}
  final_results <- do.call(rbind, results)

  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(rank_hat) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
  }
  else{
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(rank_hat) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
  }

  final_results <- final_results %>%
    mutate(distribution = distribution) %>%
    mutate(true_rank = as.numeric(true_rank)) %>%
    arrange(true_rank)
  return(final_results)
}

test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1)
test
#manual_test <- simulate_double_elimination_tournament(num_teams = 4, distribution = "Manual", ties=F, series=1)
#manual_test

```

## Shiny User Interface Tournament Simulation

```{r}
ui <- fluidPage(
  titlePanel("Double-Elimination Tournament Simulator"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 4, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Tournament Results"),
      tableOutput("ranked_results")
    )
  )
)

# Define the server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  tournament_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs

    seed_structure <- custom_seeding()

    simulate_double_elimination_tournament(
      num_teams = input$num_teams,
      distribution = input$distribution,
      ties = input$ties,
      series = input$series,
      seeding_structure = seed_structure
    )
  })

  output$ranked_results <- renderTable({
    results <- tournament_results()
    results[order(results$true_rank), ]  # Order by Strength descending
  })
}

shinyApp(ui = ui, server = server)
```

## Shiny Number of Wins

```{r}
ui <- fluidPage(
  titlePanel("Tournament Simulation with Total Wins Visualization"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Total Wins Summary"),
      plotOutput("wins_plot")
    )
  )
)

# Shiny server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  simulation_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
    seed_structure <- custom_seeding()
    # Run simulations
    replicate_results <- replicate(
      input$num_simulations,
      simulate_double_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        seeding_structure = seed_structure
      ),
      simplify = FALSE
    )

    # Combine all results and summarize total wins
    all_results <- do.call(rbind, replicate_results)
    total_wins_summary <- all_results %>%
      group_by(true_rank) %>%
      summarise(
        total_wins = sum(rank_hat==1, na.rm = TRUE),
        .groups = 'drop'
      ) %>%
      arrange(desc(total_wins))
    return(total_wins_summary)
  })

  output$wins_plot <- renderPlot({
    req(simulation_results())
    results <- simulation_results()

    ggplot(results, aes(x = reorder(true_rank, -total_wins), y = total_wins, fill = total_wins)) +
      geom_text(aes(label = total_wins), vjust = -0.5, size = 4, color = "black") +
      geom_bar(stat = "identity") +
      labs(
        title = "Total Wins Across Simulations",
        x = "True Rank",
        y = "Total Wins"
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_fill_gradient(low = "skyblue", high = "blue")
  })
}

# Run the Shiny app
shinyApp(ui = ui, server = server)
```

# Multiple Simulations for 1 Seeding Structure

```{r}
library(shiny)

ui <- fluidPage(
  titlePanel("Tournament Simulation"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", 8, min = 4, step = 1),
      selectInput("distribution", "Distribution Type:", choices = c("Normal", "Uniform")),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      numericInput("replicates", "Number of Replicates:", 100, min = 1, max = 1000, step = 1),
      numericInput("series", "Number of Games in a Series (odd number):", 1, min = 1, max = 11, step = 2),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      actionButton("run_sim", "Run Tournament Simulation")
    ),

    mainPanel(
      tableOutput("summary_results"),
      verbatimTextOutput("kendall_correlation"),
      verbatimTextOutput("spearman_correlation")
    )
  )
)

server <- function(input, output, session) {
 custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  # Function to run tournament simulations
  observeEvent(input$run_sim, {
    req(input$num_teams, input$distribution, input$replicates, input$series)

    # Initialize lists to store correlation results
    kendall_correlations <- numeric(input$replicates)
    spearman_correlations <- numeric(input$replicates)

    seed_structure <- custom_seeding()

    # Run tournament simulations for the specified number of replicates
    simulation_results <- list()
    for (replicate in 1:input$replicates) {
      simulation_results[[replicate]] <- simulate_double_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        seeding_structure = seed_structure
      )

      # Combine all results into one data frame
      all_results <- do.call(rbind, simulation_results)

      # Compute average ranks across all replicates
      average_ranks <- all_results %>%
        group_by(true_rank) %>%
        summarise(avg_rank = mean(rank_hat, na.rm = TRUE), .groups = 'drop') %>%
        arrange(true_rank)

      # Calculate Kendall Tau correlation for the current replicate
      kendall_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "kendall", use = "complete.obs")

      # Calculate Spearman correlation for the current replicate
      spearman_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "spearman", use = "complete.obs")
    }

    # Compute average Kendall Tau and Spearman correlations across all replicates
    avg_kendall_cor <- mean(kendall_correlations, na.rm = TRUE)
    avg_spearman_cor <- mean(spearman_correlations, na.rm = TRUE)

    # Show summary results in table
    output$summary_results <- renderTable({
      average_ranks
    })

    # Display average Kendall Tau correlation
    output$kendall_correlation <- renderPrint({
      cat("Average Kendall Tau Correlation:", avg_kendall_cor, "\n")
    })

    # Display average Spearman correlation
    output$spearman_correlation <- renderPrint({
      cat("Average Spearman Correlation:", avg_spearman_cor, "\n")
    })
  })
}

shinyApp(ui, server)
```

## Shiny Probability of Top 3 Correct

```{r}
ui <- fluidPage(
  titlePanel("Probability of Top 3 Teams in the Correct Order"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Top 3 Correct"),
      verbatimTextOutput("top3_output")
    )
  )
)

# Shiny server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  simulation_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
    seed_structure <- custom_seeding()

    # Run simulations
    replicate_results <- replicate(
      input$num_simulations,
      simulate_double_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        seeding_structure = seed_structure
      ),
      simplify = FALSE
    )

    # Combine all results into one dataframe and add a Simulation column
    all_results <- do.call(rbind, replicate_results)
    all_results$simulation <- rep(1:input$num_simulations, each = nrow(all_results) / input$num_simulations)

    return(all_results)
  })

  # Identify top-3 ranks for each simulation and calculate the probability
  top_three_probability <- reactive({
    all_results <- simulation_results()

    # Group by simulation and filter for top 3 true_rank and rank_hat
    top_three_results <- all_results %>%
      group_by(simulation) %>%
      filter(true_rank %in% 1:3 & rank_hat %in% 1:3) %>%
      arrange(simulation, rank_hat) %>%
      summarize(top_3_true_rank = list(true_rank), top_3_rank = list(rank_hat), .groups = 'drop') %>%
      mutate(match = mapply(function(true, rank) setequal(true, rank), top_3_true_rank, top_3_rank))

    # Calculate probability of matching top 3 true_rank with rank_hat
    match_count <- sum(top_three_results$match)
    total_simulations <- n_distinct(all_results$simulation)

    probability <- match_count / total_simulations
    return(probability)
  })

  # Output the probability to the UI
  output$top3_output <- renderText({
    probability <- top_three_probability()

    paste("Probability of Top 3 Seeds with Ranks 1-3: ", round(probability,4))
  })
}

shinyApp(ui=ui,server=server)
```

## Shiny Conditional Probability Curve

```{r}
ui <- fluidPage(
  titlePanel("Conditional Probability of Team Ranks in Tournament"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", 
                placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      plotOutput("probability_plot")
    )
  )
)

# Server
server <- function(input, output) {

  # Helper Function: Parse Custom Seeding Structure
  parse_seeding <- function(seeding_input, num_teams) {
    if (seeding_input != "") {
      seeding_vector <- unlist(strsplit(seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)

      num_teams_next_power_of_two <- 2^ceiling(log2(num_teams))
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(
          paste("Seeding structure length must match", num_teams_next_power_of_two), 
          type = "error"
        )
        return(NULL)
      }
    } else {
      return(NULL)
    }
  }

  # Curve Function to Calculate Conditional Probabilities
  curve <- function(sim, num_teams) {
    probability <- numeric(num_teams)
    for (i in 1:num_teams) {
      if (i == 1) {
        probability[i] <- nrow(sim[sim$rank_hat == 1 & sim$true_rank == 1, ]) / length(unique(sim$simulation))
      } else {
        top_ranks <- sim[sim$rank_hat <= (i - 1) & sim$true_rank <= (i - 1), ]
        better_teams <- 1:(i - 1)
        numerator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
          sim_subset <- sim[sim$simulation == sim_id, ]
          if (all(better_teams %in% sim_subset$true_rank[sim_subset$rank_hat <= (i - 1)])) {
            return(sum(sim_subset$rank_hat == i & sim_subset$true_rank == i))
          } else {
            return(0)
          }
        }))
        denominator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
          sim_subset <- top_ranks[top_ranks$simulation == sim_id, ]
          all(better_teams %in% sim_subset$true_rank)
        }))
        probability[i] <- numerator / denominator
      }
    }
    return(probability)
  }

  # Reactive Event: Run Simulations
  simulation_data <- eventReactive(input$run_sim, {
    num_teams <- input$num_teams
    distribution <- input$distribution
    series <- input$series
    num_simulations <- input$num_simulations

    # Parse custom seeding structure
    seeding_structure <- parse_seeding(input$seeding_input, num_teams)

    # Run simulations
    replicate_results <- replicate(
      num_simulations,
      simulate_double_elimination_tournament(
        num_teams = num_teams,
        distribution = distribution,
        series = series,
        ties = FALSE,
        seeding_structure = seeding_structure
      ),
      simplify = FALSE
    )

    # Combine all results into one dataframe
    #all_results <- do.call(rbind, replicate_results)
    #all_results$simulation <- rep(1:num_simulations, each = nrow(all_results) / num_simulations)
    all_results <- list_rbind(x=replicate_results, names_to = "simulation")

    # Calculate conditional probabilities
    sim_probs <- curve(all_results, num_teams)
    probabilities <- data.frame(x = 1:num_teams, y = sim_probs)
    probabilities$line <- "Simulated"

    # Baseline probabilities
    baseline_prob <- 1 / (num_teams:1)
    baseline <- data.frame(x = 1:num_teams, y = baseline_prob)
    baseline$line <- "Baseline"

    # Combine for plotting
    rbind(baseline, probabilities)
  })

  # Plot Output
  output$probability_plot <- renderPlot({
    req(simulation_data())
    ggplot(data = simulation_data(), aes(x = x, y = y, color = line, group = line)) +
      geom_point() +
      geom_line() +
      ylim(0, 1) +
      labs(
        x = "Rank",
        y = "Conditional Probability",
        title = "Simulated vs Baseline Conditional Probabilities"
      ) +
      theme_minimal()
  })
}

# Run the App
shinyApp(ui = ui, server = server)
```

### 2025-01-24 round_robin.qmd
Update round_robin.qmd

@@
 ---
-title: "round robin simulations"
+title: "fix rr"
 format: html
@@
 
-# Bradley Terry
+# Fix Round Robin Code
 
-teams and strengths
+### reference: strength weights
 
 ```{r}
-set.seed(4329)
+num <- 4
+
+# uniform 
+qunif(1:num/(num+1), 0, num+1)
+
+# normal
+qnorm(((1:num) - 0.5) / num, 0, 1)
+```
+
+# Set Up
+
+simulate one match
+
+```{r}
 library(tidyverse)
 
-#set up
-num_teams <- 8
-teams <- paste("Team", 1:num_teams)
+simulate_match <- function(team1, team2, strengths) {
+  strength1 <- strengths$True_Strength[strengths$Team == team1]
+  strength2 <- strengths$True_Strength[strengths$Team == team2]
 
-team_strengths <- data.frame(Team = teams,
-  Strength = round(runif(num_teams, min = 0, max = 1),2))
+  prob_team1_wins <- exp(strength1-strength2) / (1 + exp(strength1 - strength2))
 
-# add column for true ranks of team_strengths
-(team_strengths <- team_strengths %>% arrange(-Strength) %>% mutate(Rank = row_number()) %>% arrange(Team))
-```
-
-simulate one match using Bradley Terry model
-```{r}
-# simulate a match using Bradley-Terry model
-simulate_match <- function(team1, team2, strengths) {
-  strength1 <- strengths$Strength[strengths$Team == team1]
-  strength2 <- strengths$Strength[strengths$Team == team2]
-
-  prob_team1_wins <- strength1 / (strength1 + strength2)  # BT model probability
-
-  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)  # Randomize outcome
+  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)
   return(winner)
@@
 
-simulate single round robin tournament
+round robin tournament function for n rounds
 
 ```{r}
-simulate_roundrobin <- function(teams, strengths){
+simulate_roundrobin_r <- function(teams, strengths, r){
   n <- length(teams)
   win_counts <- numeric(n)
+  for (i in 1:r){
     for (i in seq_along(teams)){
@@
     }
+  }
   results <- data.frame(
     Team = teams,
-    Wins = win_counts
+    Game_Wins = win_counts
   )
@@
   # Ranks teams based on number of wins (includes ties)
-  results$RankHat <- rank(-results$Wins, ties.method = "average")
-  results <- results[order(results$RankHat),]
+  results$Rank_Hat <- rank(-results$Game_Wins, ties.method = "average")
+  results <- results[order(results$Rank_Hat),]
 
-  results <- results %>% inner_join(team_strengths, by = 'Team') %>% select(Team, Strength, Rank, RankHat, Wins)
+  results <- results %>% inner_join(strengths, by = 'Team') %>% select(Team, True_Rank, True_Strength, Rank_Hat, Game_Wins)
 
@@
 }
-
-rr <- simulate_roundrobin(teams, team_strengths)
-
-# kendall correlation
-kendallcor <- cor.test(rr$RankHat,rr$Rank, method="kendall")
 ```
 
-simulate many round robin tournaments using the same set of 8 teams
+## Uniform Distribution of Strengths
+
+run tournament for n rounds
+
+# Saving Output in Different Way
+
+- have code to simulate one match, round robin tournament for n rounds
+
+- just need to fix run_tournament function to save output differently?
+
+## Uniform Dist, Save Each Sim
 
 ```{r}
-# Define the number of simulations
-n <- 1000  # Set the number of simulations you want
-total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
-colnames(total_ranks) <- teams
-
-kendallcor <- c()
-
-# Run multiple simulations
-for (i in 1:n) {
-  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
-  ranks <- simulation$Rank
-  team_order <- simulation$Team
-  total_ranks[i, team_order] <- ranks
-
-  k <- cor.test(simulation$RankHat,simulation$Rank, method="kendall")$estimate
-  kendallcor[i] <- k
-}
-
-# Calculate average ranks for each team
-average_ranks <- data.frame(AvgRankHat = colMeans(total_ranks, na.rm = TRUE), Strength = team_strengths$Strength, Rank = team_strengths$Rank)
-average_ranks
-
-# distribution of kendall coefficients
-hist(kendallcor)
-mean(kendallcor)
-median(kendallcor) # rank hat matches true rank about 70% of the time
-```
-
-# Finalizing Round Robin Function
-
-### generate winning probabilities with$\exp(i-j)/(1+\exp(i-j))$
-
-round robin function inputs: teams, strengths, schedule, (seed?)
-
-## create teams and strengths - strengths are just 1-8
-
-```{r}
-library(tidyverse)
-
-# Define the number of teams 
-num_teams <- 8
+run_tournament <- function(num_teams, num_rounds){
+  # define teams
   teams <- paste("Team", 1:num_teams)
-
-# Assign random strengths to each team 
+  # put Teams and True_Strength into data frame
   team_strengths <- data.frame(
     Team = teams,
-  Strength = (1:num_teams)
-)
+    True_Strength = qunif(1:num_teams/(num_teams+1), 0, sqrt(12)))
 
-# add column for true ranks of team_strengths
-team_strengths <- team_strengths %>% arrange(-Strength) %>% mutate(Rank = row_number()) %>% arrange(Team)
-```
+  # add True_Rank and Seed, order columns
+  team_strengths <- team_strengths %>% arrange(-True_Strength) %>% mutate(True_Rank = row_number()) %>% arrange(Team) %>% mutate(Seed = NA) %>% select(Team, True_Rank, True_Strength, Seed)
 
-## simulate a single match
-
-```{r}
-simulate_match <- function(team1, team2, strengths) {
-  strength1 <- strengths$Strength[strengths$Team == team1]
-  strength2 <- strengths$Strength[strengths$Team == team2]
-
-  prob_team1_wins <- exp(strength1-strength2) / (1 + exp(strength1-strength2))
-
-  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)
-  return(winner)
-}
-```
-
-## simulate a single round robin tournament
-
-```{r}
-simulate_roundrobin <- function(teams, strengths){
-  n <- length(teams)
-  win_counts <- numeric(n)
-  for (i in seq_along(teams)){
-    for (j in seq_along(teams)){
-      if (i != j){
-        match_winner <- simulate_match(teams[i],teams[j],strengths)
-        if (match_winner == teams[i]) {
-          win_counts[i] <- win_counts[i] + 1
-        } else {
-          win_counts[j] <- win_counts[j] + 1
-        }
-      }
-    }
-  }
-  results <- data.frame(
-    Team = teams,
-    Wins = win_counts
-  )
-
-  # Ranks teams based on number of wins (includes ties)
-  results$RankHat <- rank(-results$Wins, ties.method = "average")
-  results <- results[order(results$RankHat),]
-
-  results <- results %>% inner_join(team_strengths, by = 'Team') %>% select(Team, Strength, Rank, RankHat, Wins)
-
-  return(results)
-}
-
-rr <- simulate_roundrobin(teams, team_strengths)
-# kendall correlation
-kendallcor <- cor(rr$RankHat, rr$Rank, method="kendall")
-# spearman correlation
-spearmancor <- cor(rr$RankHat, rr$Rank, method="spearman")
-```
-
-Kendall Coefficient
-
-Kendall Coefficient is a statistic used to measure the ordinal association between two measured quantities.
-
--   If the agreement between the two rankings is perfect (i.e., the two rankings are the same) the coefficient has value 1.
--   If the disagreement between the two rankings is perfect (i.e., one ranking is the reverse of the other) the coefficient has value −1
-    -   We do not have any values of the coefficient that are less than 0
-
-Spearman Coefficent
-
-Spearman Correlation is a non-parametric correlation also known as rank-based correlation coefficient.
-It measures the strength and direction of the monotonic relationship between two ranked variables.
-It ranges from -1 to +1:
-
--   **+1**: A perfect positive monotonic relationship.
-
--   **0**: No monotonic relationship.
-
--   **-1**: A perfect negative monotonic relationship.
-
-### simulate many round robin tournaments using the same set of 8 teams, get kendall and spearman coefficients
-
-```{r}
-# Define the number of simulations
-n <- 1000  # Set the number of simulations
+  # simulation set up
+  n <- 1000
   total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
@@
 
-# Run multiple simulations
+  ####### new object to save all of the simulations
+  simulation_list <- list()
+
+  # run multiple simulations
   for (i in 1:n) {
-  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
-  ranks <- simulation$Rank
-  team_order <- simulation$Team
+    simulation_list[[i]] <- simulate_roundrobin_r(teams, team_strengths, num_rounds)
+
+    ranks <- simulation_list[[i]]$Rank_Hat
+    team_order <- simulation_list[[i]]$Team
     total_ranks[i, team_order] <- ranks
 
-  kendallcor[i] <- cor(simulation$RankHat, simulation$Rank, method="kendall")
-  spearmancor[i] <- cor(simulation$RankHat, simulation$Rank, method="spearman")
+    kendallcor[i] <- cor(simulation_list[[i]]$Rank_Hat, simulation_list[[i]]$True_Rank, method="kendall")
+    spearmancor[i] <- cor(simulation_list[[i]]$Rank_Hat, simulation_list[[i]]$True_Rank, method="spearman")
   }
 
-# Calculate average ranks for each team
-average_ranks <- data.frame(AvgRankHat = colMeans(total_ranks, na.rm = TRUE), Strength = team_strengths$Strength, Rank = team_strengths$Rank)
-average_ranks
+  # SUMMARY - Calculate average ranks for each team
+  average_ranks <- data.frame(AvgRankHat = colMeans(total_ranks, na.rm = TRUE), Strength = team_strengths$True_Strength, Rank = team_strengths$True_Rank)
 
-# distribution of kendall correlation
-print('kendall')
-hist(kendallcor)
-mean(kendallcor)
-median(kendallcor)
+  # remove NA correlation values
+  kendallcor <- na.omit(kendallcor)
+  spearmancor <- na.omit(spearmancor)
 
-# distribution of spearman correlation
-print('spearman')
-hist(spearmancor)
-mean(spearmancor)
-median(spearmancor)
-```
+  # kendall and spearman correlations
+  k_mean <- mean(kendallcor)
+  s_mean <- mean(spearmancor)
 
-# ROUND ROBIN CODE
-
-## simulate a single match
-
-```{r}
-simulate_match <- function(team1, team2, strengths) {
-  strength1 <- strengths$Strength[strengths$Team == team1]
-  strength2 <- strengths$Strength[strengths$Team == team2]
-
-  prob_team1_wins <- exp(strength1-strength2) / (1 + exp(strength1-strength2))
-
-  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)
-  return(winner)
-}
-```
-
-## simulate a single round robin tournament
-
-```{r}
-simulate_roundrobin <- function(teams, strengths){
-  n <- length(teams)
-  win_counts <- numeric(n)
-  for (i in seq_along(teams)){
-    for (j in seq_along(teams)){
-      if (i != j){
-        match_winner <- simulate_match(teams[i],teams[j],strengths)
-        if (match_winner == teams[i]) {
-          win_counts[i] <- win_counts[i] + 1
-        } else {
-          win_counts[j] <- win_counts[j] + 1
-        }
-      }
-    }
-  }
-  results <- data.frame(
-    Team = teams,
-    Wins = win_counts
-  )
-
-  # Ranks teams based on number of wins (includes ties)
-  results$RankHat <- rank(-results$Wins, ties.method = "average")
-  results <- results[order(results$RankHat),]
-
-  results <- results %>% inner_join(team_strengths, by = 'Team') %>% select(Team, Strength, Rank, RankHat, Wins)
+  results <- list(average_ranks = average_ranks, k_mean = k_mean, s_mean = s_mean, simulation_list = simulation_list)
 
@@
 
-# code that is changed for different team strengths/number sims, etc.
-## generate team strengths using normal distribution
+### Run the simulation
+
 ```{r}
-# Define the number of teams 
-num_teams <- 8
-teams <- paste("Team", 1:num_teams)
+# run simulation using run_tournament() 
+  # inputs are number of teams and number of rounds
+results <- run_tournament(8, 1)
+sims <- results$simulation_list
 
-# Assign random strengths to each team 
-team_strengths <- data.frame(
-  Team = teams,
-  Strength = rep(0,8))
-
-# add column for true ranks of team_strengths
-team_strengths <- team_strengths %>% arrange(-Strength) %>% mutate(Rank = row_number()) %>% arrange(Team)
-team_strengths
+results
 ```
 
-run round robin simulation
-```{r}
-# Define the number of simulations
-n <- 1000  # Set the number of simulations
-total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
-colnames(total_ranks) <- teams
+- run 1000 simulations for different numbers of teams
+  - uniform and normal distributions for strengths
+  - 3-8 teams, 16 teams
 
-kendallcor <- c()
-spearmancor <- c()
-
-# Run multiple simulations
-for (i in 1:n) {
-  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
-  ranks <- simulation$Rank
-  team_order <- simulation$Team
-  total_ranks[i, team_order] <- ranks
-
-  kendallcor[i] <- cor(simulation$RankHat, simulation$Rank, method="kendall")
-  spearmancor[i] <- cor(simulation$RankHat, simulation$Rank, method="spearman")
-}
-
-# Calculate average ranks for each team
-average_ranks <- data.frame(AvgRankHat = colMeans(total_ranks, na.rm = TRUE), Strength = team_strengths$Strength, Rank = team_strengths$Rank)
-average_ranks
-
-# distribution of kendall correlation
-print('kendall')
-hist(kendallcor)
-mean(kendallcor)
-median(kendallcor)
-
-# distribution of spearman correlation
-print('spearman')
-hist(spearmancor)
-mean(spearmancor)
-median(spearmancor)
-```
-
-# controlling the strengths but still have normal distribution
-```{r}
-thetas <- qnorm((seq(0,1,length=9) + 1/16)[-9])
-thetas
-
-# qq plot way
-qnorm(((1:8) - 0.5) / 8, 0, 0.5)
-
-# plot with variability on x and average of correlation coefficient on y
-
-# qunif
-```
+- save them as list objects

### 2025-02-06 simulation_curves.qmd
update 2/6/2025

---
title: "Simulation Curves"
format: html
editor: visual
---

## Simulation Curves

The curves show the conditional probabilities of the true ranking being equal to the observed ranking in the simulations given that the teams with a better true rank (1 denotes the best) are ranked better in the observed rankings. For example, in a 4 team tournament the probabilities would look like the following:

$$
P(R_1 = 1)
$$

$$
P(R_2 = 2 | R_1 \leq 1) = \frac{P(R_1 \leq 1 \cap R_2 = 2)}{P(R_1 \leq 1)}
$$

$$
P(R_3 = 3 | R_1 \leq 2 \cap R_2 \leq2) = \frac{P(R_1 \leq 2 \cap R_2 \leq2 \cap R_3 = 3)}{P(R_1 \leq 2 \cap R_2 \leq2)}
$$

$$
P(R_4 = 4 | R_1 \leq 3 \cap R_2 \leq 3 \cap R_3 \leq 3) = \frac{P(R_1 \leq 3 \cap R_2 \leq 3 \cap R_3 \leq 3 \cap R_4 = 4) }{P(R_1 \leq 3 \cap R_2 \leq 3 \cap R_3 \leq 3)} = 1
$$

Each curve is compared to the baseline curve, which can be found by the following formula:

$$
P(R_i = i) = \frac{1}{n - i}, \text{ where n is the number of teams and i is the true rank}
$$

The baseline curve is the same conditional probabilities solved if each team had the same true rank (each team would be equally likely to win every game). Points above the baseline curve suggest that the tournament structure is more accurate to the true ranks than random chance.

## Calculating the Curve

```{r}
curve <- function(sim, num_teams) {
  probability <- numeric(num_teams)
  for (i in 1:num_teams) {
    if (i == 1) {
      probability[i] <- nrow(sim[sim$rank_hat == 1 & sim$true_rank == 1, ]) / length(unique(sim$simulation))
    } else {
      top_ranks <- sim[sim$rank_hat <= (i - 1) & sim$true_rank <= (i - 1), ]
      better_teams <- 1:(i - 1)
      numerator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
        sim_subset <- sim[sim$simulation == sim_id, ]
        if (all(better_teams %in% sim_subset$true_rank[sim_subset$rank_hat <= (i - 1)])) {
           return(sum(sim_subset$rank_hat == i & sim_subset$true_rank == i))
        } else {
          return(0)
        }
      }))
      denominator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
        sim_subset <- top_ranks[top_ranks$simulation == sim_id, ]
        all(better_teams %in% sim_subset$true_rank)
      }))
      probability[i] <- numerator / denominator
    }
  }

  # Baseline probabilities
  baseline_prob <- 1 / (num_teams:1)
  baseline <- data.frame(x = 1:num_teams, y = baseline_prob)
  baseline$line <- "Baseline"

  # Simulated probabilities
  probabilities <- data.frame(x = 1:num_teams, y = probability)
  probabilities$line <- "Simulated"

  # Combine for plotting
  test_plot <- rbind(baseline, probabilities)

  # Generate ggplot
  p <- ggplot(data = test_plot, aes(x = x, y = y, color = line, group = line)) +
    geom_point() +
    geom_line() +
    ylim(0, 1) +
    labs(
      x = "Rank",
      y = "Conditional Probability",
      title = "Simulated vs Baseline Conditional Probabilities"
    ) +
    theme_minimal()

  return(p)
}

```

### Round Robin Tournament Structure

```{r}
rr_probs_plot <- curve(test_rr, 8)
print(rr_probs_plot)

# with randomly assigning ties
test_rr_noTies <- test_rr %>%
  group_by(simulation) %>%
  mutate(rank_hat = rank(rank_hat,ties.method="random"))
rr_noTies_probs_plot <- curve(test_rr_noTies, 8)
print(rr_noTies_probs_plot)

```

### Single Elimination Tournament Structure

```{r}
se_norm_results$simulation <- rep(1:10000, each = nrow(se_norm_results) / 10000)
se_probs_plot <- curve(se_norm_results, 8)
print(se_probs_plot)

```

### Double Elimination Tournament Structure

```{r}
de_norm_results_eight$simulation <- rep(1:10000, each = nrow(de_norm_results_eight) / 10000)
de_probs_plot <- curve(de_norm_results_eight, 8)
print(de_probs_plot)
```

### 2025-02-06 single_elimination_shiny.qmd
update 2/6/2025

@@
 ```
+
+### Things To Do
+
+-   Add a re-seeded option
+
+-   Custom strengths (likely as an input to the function)
 
@@
 ```{r}
-generate_tournament_structure <- function(n) {
+generate_elimination_tournament_structure <- function(n) {
 
@@
 
+  left_half <- teams[1:(length(teams)/2)]
+  right_half <- teams[(length(teams)/2+1):length(teams)]
+  right_half <- rev(right_half)
+  teams <- c(left_half,right_half)
+
+  # Ensure that the lower number team is on the left side of each matchup
+  teams <- unlist(lapply(seq(1, length(teams), by = 2), function(i) {
+    if (teams[i] > teams[i + 1]) {
+      c(teams[i + 1], teams[i])  # Swap if needed
+    } else {
+      c(teams[i], teams[i + 1])
+    }
+  })
+  )
+
   while (extra_teams > 0){
@@
 }
+generate_elimination_tournament_structure(8)
+generate_elimination_tournament_structure(16)
 
@@
 # Simulate the single-elimination tournament
-simulate_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1, third_place = T) {
+simulate_single_elimination_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1, third_place = T) {
 
@@
   }
-teams <- paste("Team",num_teams:1)
+  teams <- paste(num_teams:1)
   if (log2(num_teams) %% 1 != 0) {
@@
     normal_strengths <- data.frame(
-  Team = teams,
-  Strength = c(strengths, rep(NA, length(teams) - num_teams)),
-  Wins = rep(0,length(teams)),
-  Ranks = rep(NA,length(teams))
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      rank_hat = rep(NA,length(teams)),
+      game_wins = rep(0,length(teams)),
+      game_losses = rep(0,length(teams))
     )
-normal_strengths$True_Rank <- rank(-normal_strengths$Strength, ties.method = "average")
-df <- arrange(normal_strengths, True_Rank)
+    df <- arrange(normal_strengths, true_rank)
   }
@@
     unif_strength <- data.frame(
-  Team = teams,
-  Strength = c(strengths, rep(NA, length(teams) - num_teams)),
-  Wins = rep(0,length(teams)),
-  Ranks = rep(NA,length(teams))
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      game_wins = rep(0,length(teams)),
+      game_losses = rep(0,length(teams)),
+      rank_hat = rep(NA,length(teams))
     )
-unif_strength$True_Rank <- rank(-unif_strength$Strength, ties.method = "average")
-df <- arrange(unif_strength, True_Rank)
+    df <- arrange(unif_strength, true_rank)
+  }
+  else if (distribution == "Manual") {
+    cat("Please enter the strengths for each team in descending rank order:\n")
+    strengths <- numeric(num_teams)
+    for (i in num_teams:1) {
+      repeat {
+        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
+        if (!is.na(input)) {
+          strengths[i] <- input
+          break
+        } else {
+          cat("Invalid input. Please enter a numeric value.\n")
+        }
+      }
+    }
+    manual_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- arrange(manual_strengths, true_rank)
   }
   else {
-  stop("Distribution must be Normal or Uniform")
+    stop("Distribution not found: Enter Manual to input your own strengths")
   }
@@
   if (is.null(seeding_structure)){
-    seeding_structure <- matrix(generate_tournament_structure(num_teams), nrow = 1)
+    seeding_structure <- matrix(generate_elimination_tournament_structure(num_teams), nrow = 1)
   } else{
@@
     permuted_df <- df
-    permuted_df$Team <- df$Team[permutation]  # Update team order 
-    permuted_df$Strength <- df$Strength[permutation]
-    permuted_df$Ranks <- NA  # Initialize ranks column
-    permuted_df$Wins <- 0  # Initialize wins column
-    permuted_df$True_Rank <- df$True_Rank[permutation]
+    permuted_df$true_rank <- df$true_rank[permutation]  # Update team order 
+    teams <- permuted_df$true_rank
+    permuted_df$true_strength <- df$true_strength[permutation]
+    permuted_df$rank_hat <- NA  # Initialize ranks column
+    permuted_df$game_wins <- 0  # Initialize wins column
+    permuted_df$game_losses <- 0
 
-    teams <- permuted_df$Team
-    strengths <- permuted_df$Strength
+    strengths <- permuted_df$true_strength
     round_number <- ceiling(log2(length(teams)))
@@
 
-      strength1 <- strengths[which(permuted_df$Team == team1)]
-      strength2 <- strengths[which(permuted_df$Team == team2)]
+        strength1 <- strengths[which(permuted_df$true_rank == team1)]
+        strength2 <- strengths[which(permuted_df$true_rank == team2)]
 
@@
         wins_team2 <- 0
+        losses_team1 <- 0
+        losses_team2 <- 0
         for (game in 1:series){
           match_winner <- simulate_match(team1, team2, strength1, strength2)
-      #cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", 
-      #      team1, team2, strength1, strength2, match_winner))
+          #cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
+          if(!is.na(team1)) {
           if (match_winner == team1) {
             wins_team1 <- wins_team1 + 1
+            losses_team2 <- losses_team2 + 1
           } else {
             wins_team2 <- wins_team2 + 1
+            losses_team1 <- losses_team1 + 1
           }
+          }
+          else {wins_team2 <- series / 2 + 1}
 
@@
 
+        # adds wins and losses of the series to the teams unless they have a bye
+        if (!any(is.na(c(team1, team2)))) {
+          permuted_df <- permuted_df %>%
+            mutate(game_wins = ifelse(true_rank == team1, game_wins + wins_team1, game_wins)) %>%
+            mutate(game_wins = ifelse(true_rank == team2, game_wins + wins_team2, game_wins)) %>%
+            mutate(game_losses = ifelse(true_rank == team1, game_losses + losses_team1, game_losses)) %>%
+            mutate(game_losses = ifelse(true_rank == team2, game_losses + losses_team2, game_losses)) 
+           }
+
         # Determine the match winner based on series results
@@
 
-      # Function to calculate the rank values for a given round
-    get_rank_values <- function(n) {
-      # Calculate the rank values as the average of the last n/2 teams
-      suppressWarnings({
-        if (n / 2 + 1 == num_teams) {
-          rank_values <- num_teams
-        } else {
-          rank_values <- (n / 2 + 1):min(num_teams, n)
-        }
-      })
-      return(rank_values)
-    }
         if (round_number > 2){
@@
           n <- 2^(round_number)
-        rank_values <- suppressWarnings(get_rank_values(n))
-        #print(match_winner)
-        if (match_winner == team1) {
-          if (ties == TRUE){
-          permuted_df$Ranks[permuted_df$Team == team2] <- suppressWarnings(mean(rank_values))
-          }
-          else if (ties == FALSE){
-          permuted_df$Ranks[permuted_df$Team == team2] <- n
-          }
+          #print(series_winner)
+            if (!is.na(team1) & series_winner == team1) {
+                permuted_df$rank_hat[permuted_df$true_rank == team2] <- n
             } else {
-          if (ties == TRUE){
-          permuted_df$Ranks[permuted_df$Team == team1] <-  suppressWarnings(mean(rank_values))
-          }
-          else if (ties == FALSE){
-          permuted_df$Ranks[permuted_df$Team == team1] <- n
-          }
+                permuted_df$rank_hat[permuted_df$true_rank == team1] <- n
             }
@@
           n <- 2^(round_number)
-        rank_values <- get_rank_values(n)
           if (match_winner == team1) {
-          if (ties == TRUE){
-          permuted_df$Ranks[permuted_df$Team == team2] <-
-          mean(rank_values)
-          }
-          else if (ties == FALSE & third_place == FALSE){
-          permuted_df$Ranks[permuted_df$Team == team2] <- n
-          }
+              permuted_df$rank_hat[permuted_df$true_rank == team2] <- n
             losers_semis <- c(losers_semis, team2)
           } else {
-          if (ties == TRUE){
-          permuted_df$Ranks[permuted_df$Team == team1] <-
-          mean(rank_values)
-          }
-          else if (ties == FALSE & third_place == FALSE){
-          permuted_df$Ranks[permuted_df$Team == team1] <- n
-          }
+              permuted_df$rank_hat[permuted_df$true_rank == team1] <- n
             losers_semis <- c(losers_semis, team1)
-
           }
@@
           permuted_df <- permuted_df %>%
-            filter(!is.na(Team))
+            filter(!is.na(true_rank))
           n <- 2^(round_number)
-        rank_values <- get_rank_values(n)
-        if (match_winner == team1) {
-          permuted_df$Ranks[permuted_df$Team == team2] <- 2
-        permuted_df$Ranks[permuted_df$Team == team1] <- 1
-        permuted_df$Wins[permuted_df$Team == team1] <-
-        permuted_df$Wins[permuted_df$Team == team1] + 1
+          if (series_winner == team1) {
+            permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
+            permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
           } else {
-          permuted_df$Ranks[permuted_df$Team == team1] <- 2
-          permuted_df$Ranks[permuted_df$Team == team2] <- 1
-          permuted_df$Wins[permuted_df$Team == team2] <-
-          permuted_df$Wins[permuted_df$Team == team2] + 1
+            permuted_df$rank_hat[permuted_df$true_rank == team1] <- 2
+            permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
           }
@@
       team2 <- losers_semis[2]
-    strength1 <- permuted_df$Strength[permuted_df$Team == team1]
-    strength2 <- permuted_df$Strength[permuted_df$Team == team2]
+      strength1 <- permuted_df$true_strength[permuted_df$true_rank == team1]
+      strength2 <- permuted_df$true_strength[permuted_df$true_rank == team2]
       wins_team1 <- 0
@@
         #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team1))
-        permuted_df$Ranks[permuted_df$Team == team1] <- 3
-        permuted_df$Ranks[permuted_df$Team == team2] <- 4
+        permuted_df$rank_hat[permuted_df$true_rank == team1] <- 3
+        permuted_df$rank_hat[permuted_df$true_rank == team2] <- 4
       } else {
         #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team2))
-        permuted_df$Ranks[permuted_df$Team == team2] <- 3
-        permuted_df$Ranks[permuted_df$Team == team1] <- 4
+        permuted_df$rank_hat[permuted_df$true_rank == team2] <- 3
+        permuted_df$rank_hat[permuted_df$true_rank == team1] <- 4
       }
@@
 
+    permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
     results[[i]] <- permuted_df # Store the results of this permutation
@@
 
-  # Generalized function to remove ties by assigning unique ranks
-remove_ties <- function(df) {
-  suppressWarnings({
-  df <- df %>%
-    group_by(Ranks) %>%
-    mutate(
-      Ranks = ifelse(
-        Ranks == 2,  # Do not change if rank is 2
-        Ranks, 
-        ifelse(
-          log2(Ranks) %% 1 == 0,  # Check if Ranks is a power of 2
-          ifelse(
-            Ranks / 2 + 1 == num_teams,  
-            num_teams,  
-            sample((Ranks / 2 + 1):(min(Ranks, num_teams)), n(), replace = FALSE)  # Regular case
-          ),
-          Ranks  # Keep Ranks as is if it's not a power of 2
-        )
-      )
-    ) %>%
-    ungroup()
-  })
-  return(df)
+  # Remove ties if ties = FALSE
+  if (!ties) {
+    final_results <- final_results %>%
+      filter(!is.na(true_rank)) %>%
+      arrange(true_rank) %>%
+      mutate(rank_hat = rank(rank_hat, ties.method = "first"))
+  } else{
+    final_results <- final_results %>%
+      filter(!is.na(true_rank)) %>%
+      arrange(rank_hat) %>%
+      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
   }
 
-  # Remove ties if ties = FALSE
-  if (!ties) {
-    final_results <- remove_ties(final_results)
-  }
-
-  final_results <- final_results[order(final_results$Ranks),]
+  final_results <- final_results %>%
+    mutate(distribution = distribution) %>%
+    mutate(true_rank = as.numeric(true_rank)) %>%
+    arrange(true_rank)
   return(final_results)
@@
 
-test <- simulate_tournament(num_teams = 10, distribution = "Normal", ties=F, series=3, third_place=T, seeding_structure = c(1,NA,2,NA,3,NA,4,5,6,7,8,NA,9,NA,10,NA))
+test <- simulate_single_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, third_place=T)
 test
+
+#manual_test <- simulate_single_elimination_tournament(num_teams = 4, distribution = "Manual", ties=T, series=1, third_place=T)
+#manual_test
+
+# Gets all results of a normal distribution num_teams between 4 and 64 single-game elimination
+all_normal_results <- list()
+for (teams in c(4,8,16,32,64)) {
+  all_normal_results[[as.character(teams)]] <- list()
+  for (replicate in 1:10000) {
+    # Store each simulation result in the list with unique indexing
+    all_normal_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
+      num_teams = teams,
+      distribution = "Normal",
+      ties = FALSE,
+      series = 1,
+      third_place = TRUE
+    )
+  }
+}
+
+all_results_norm <- do.call(rbind, lapply(names(all_normal_results), function(teams) {
+  team_results <- do.call(rbind, all_normal_results[[teams]])
+  team_results$teams <- as.numeric(teams)
+  return(team_results)
+}))
+
+all_results_norm
+
+# Gets all results of a uniform distribution num_teams between 4 and 64 single-game elimination
+all_unif_results <- list()
+for (teams in c(4,8,16,32,64)) {
+  all_unif_results[[as.character(teams)]] <- list()
+  for (replicate in 1:100) {
+    # Store each simulation result in the list with unique indexing
+    all_unif_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
+      num_teams = teams,
+      distribution = "Uniform",
+      ties = FALSE,
+      series = 1,
+      third_place = TRUE
+    )
+  }
+}
+
+all_results_unif <- do.call(rbind, lapply(names(all_unif_results), function(teams) {
+  team_results <- do.call(rbind, all_unif_results[[teams]])
+  team_results$teams <- as.numeric(teams)
+  return(team_results)
+}))
+
+all_results_unif
 
@@
       numericInput("num_teams", "Number of Teams:", value = 8, min = 4, step = 1),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
+      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
       selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
+      uiOutput("manual_strengths_ui"), # Dynamic UI for Manual strengths
       checkboxInput("ties", "Allow Ties?", value = TRUE),
@@
     } else {
-      return(generate_tournament_structure(input$num_teams))
+      return(generate_elimination_tournament_structure(input$num_teams))
     }
@@
 
-    simulate_tournament(
+    simulate_single_elimination_tournament(
       num_teams = input$num_teams,
@@
     results <- tournament_results()
-    results[order(-results$Strength), ]  # Order by Strength descending
+    results[order(results$true_rank), ]  # Order by Strength descending
   })
@@
       numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
+      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
       selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
@@
       input$num_simulations,
-      simulate_tournament(
+      simulate_single_elimination_tournament(
         num_teams = input$num_teams,
@@
     total_wins_summary <- all_results %>%
-      group_by(Team) %>%
+      group_by(true_rank) %>%
       summarise(
-        Total_Wins = sum(Wins, na.rm = TRUE),
+        total_wins = sum(rank_hat==1, na.rm = TRUE),
         .groups = 'drop'
       ) %>%
-      arrange(desc(Total_Wins))
+      arrange(desc(total_wins))
     return(total_wins_summary)
@@
 
-    ggplot(results, aes(x = reorder(Team, -Total_Wins), y = Total_Wins, fill = Total_Wins)) +
-      geom_text(aes(label = Total_Wins), vjust = -0.5, size = 4, color = "black") +
+    ggplot(results, aes(x = reorder(true_rank, -total_wins), y = total_wins, fill = total_wins)) +
+      geom_text(aes(label = total_wins), vjust = -0.5, size = 4, color = "black") +
       geom_bar(stat = "identity") +
@@
         title = "Total Wins Across Simulations",
-        x = "Team",
+        x = "True Rank",
         y = "Total Wins"
@@
 ```{r}
-library(shiny)
-
 ui <- fluidPage(
@@
     sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", 8, min = 4, max = 32, step = 1),
+      numericInput("num_teams", "Number of Teams:", 8, min = 4, step = 1),
       selectInput("distribution", "Distribution Type:", choices = c("Normal", "Uniform")),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
+      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
       numericInput("replicates", "Number of Replicates:", 100, min = 1, max = 1000, step = 1),
-      numericInput("series", "Number of Games in a Series (odd number):", 5, min = 3, max = 11, step = 2),
+      numericInput("series", "Number of Games in a Series (odd number):", 1, min = 1, max = 11, step = 2),
       checkboxInput("ties", "Allow Ties?", value = TRUE),
@@
     for (replicate in 1:input$replicates) {
-      simulation_results[[replicate]] <- simulate_tournament(
+      simulation_results[[replicate]] <- simulate_single_elimination_tournament(
         num_teams = input$num_teams,
@@
       average_ranks <- all_results %>%
-        group_by(Team, True_Rank) %>%
-        summarise(Average_Rank = mean(Ranks, na.rm = TRUE), .groups = 'drop') %>%
-        arrange(Team)
+        group_by(true_rank) %>%
+        summarise(avg_rank = mean(rank_hat, na.rm = TRUE), .groups = 'drop') %>%
+        arrange(true_rank)
 
       # Calculate Kendall Tau correlation for the current replicate
-      kendall_correlations[replicate] <- cor(average_ranks$Average_Rank, average_ranks$True_Rank, method = "kendall", use = "complete.obs")
+      kendall_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "kendall", use = "complete.obs")
 
       # Calculate Spearman correlation for the current replicate
-      spearman_correlations[replicate] <- cor(average_ranks$Average_Rank, average_ranks$True_Rank, method = "spearman", use = "complete.obs")
+      spearman_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "spearman", use = "complete.obs")
     }
@@
 ui <- fluidPage(
-  titlePanel("Tournament Simulation with Total Wins Visualization"),
+  titlePanel("Probability of Top 3 Teams in the Correct Order"),
 
@@
     all_results <- do.call(rbind, replicate_results)
-    all_results$Simulation <- rep(1:input$num_simulations, each = nrow(all_results) / input$num_simulations)
+    all_results$simulation <- rep(1:input$num_simulations, each = nrow(all_results) / input$num_simulations)
 
@@
 
-    # Group by Simulation and filter for top 3 True_Rank and Ranks
+    # Group by simulation and filter for top 3 true_rank and rank_hat
     top_three_results <- all_results %>%
-      group_by(Simulation) %>%
-      filter(True_Rank %in% 1:3 & Ranks %in% 1:3) %>%
-      arrange(Simulation, Ranks) %>%
-      summarize(top_3_true_rank = list(True_Rank), top_3_rank = list(Ranks), .groups = 'drop') %>%
+      group_by(simulation) %>%
+      filter(true_rank %in% 1:3 & rank_hat %in% 1:3) %>%
+      arrange(simulation, rank_hat) %>%
+      summarize(top_3_true_rank = list(true_rank), top_3_rank = list(rank_hat), .groups = 'drop') %>%
       mutate(match = mapply(function(true, rank) setequal(true, rank), top_3_true_rank, top_3_rank))
 
-    # Calculate probability of matching top 3 True_Rank with Ranks
+    # Calculate probability of matching top 3 true_rank with rank_hat
     match_count <- sum(top_three_results$match)
-    total_simulations <- n_distinct(all_results$Simulation)
+    total_simulations <- n_distinct(all_results$simulation)
 
@@
       textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", 
-                placeholder = "e.g. 1 8 4 5 3 6 7 2"),
+                placeholder = "e.g. 1 8 4 5 3 6 2 7"),
       selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
@@
       if (i == 1) {
-        probability[i] <- nrow(sim[sim$Ranks == 1 & sim$True_Rank == 1, ]) / length(unique(sim$Simulation))
+        probability[i] <- nrow(sim[sim$rank_hat == 1 & sim$true_rank == 1, ]) / length(unique(sim$simulation))
       } else {
-        top_ranks <- sim[sim$Ranks <= (i - 1) & sim$True_Rank <= (i - 1), ]
+        top_ranks <- sim[sim$rank_hat <= (i - 1) & sim$true_rank <= (i - 1), ]
         better_teams <- 1:(i - 1)
-        numerator <- sum(sapply(unique(top_ranks$Simulation), function(sim_id) {
-          sim_subset <- sim[sim$Simulation == sim_id, ]
-          if (all(better_teams %in% sim_subset$True_Rank[sim_subset$Ranks <= (i - 1)])) {
-            return(sum(sim_subset$Ranks == i & sim_subset$True_Rank == i))
+        numerator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
+          sim_subset <- sim[sim$simulation == sim_id, ]
+          if (all(better_teams %in% sim_subset$true_rank[sim_subset$rank_hat <= (i - 1)])) {
+            return(sum(sim_subset$rank_hat == i & sim_subset$true_rank == i))
           } else {
@@
         }))
-        denominator <- sum(sapply(unique(top_ranks$Simulation), function(sim_id) {
-          sim_subset <- top_ranks[top_ranks$Simulation == sim_id, ]
-          all(better_teams %in% sim_subset$True_Rank)
+        denominator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
+          sim_subset <- top_ranks[top_ranks$simulation == sim_id, ]
+          all(better_teams %in% sim_subset$true_rank)
         }))
@@
       num_simulations,
-      simulate_tournament(
+      simulate_single_elimination_tournament(
         num_teams = num_teams,
@@
     all_results <- do.call(rbind, replicate_results)
-    all_results$Simulation <- rep(1:num_simulations, each = nrow(all_results) / num_simulations)
+    all_results$simulation <- rep(1:num_simulations, each = nrow(all_results) / num_simulations)
+    # list_rbind(x=all_results, names_to = "simulation")
 
@@
     probabilities <- data.frame(x = 1:num_teams, y = sim_probs)
-    probabilities$Line <- "Simulated"
+    probabilities$line <- "Simulated"
 
@@
     baseline <- data.frame(x = 1:num_teams, y = baseline_prob)
-    baseline$Line <- "Baseline"
+    baseline$line <- "Baseline"
 
@@
     req(simulation_data())
-    ggplot(data = simulation_data(), aes(x = x, y = y, color = Line, group = Line)) +
+    ggplot(data = simulation_data(), aes(x = x, y = y, color = line, group = line)) +
       geom_point() +

### 2025-02-07 round_robin.qmd
Update round_robin.qmd

@@
 ---
-title: "fix rr"
+title: "rr"
 format: html
@@
 
-# Fix Round Robin Code
+# Round Robin Code
 
@@
 
-  results <- results %>% inner_join(strengths, by = 'Team') %>% select(Team, True_Rank, True_Strength, Rank_Hat, Game_Wins)
+  results <- results %>% 
+    inner_join(strengths, by = 'Team') %>% 
+    select(Team, True_Rank, True_Strength, Rank_Hat, Game_Wins)
 
@@
 ```{r}
-run_tournament <- function(num_teams, num_rounds){
+run_tournament <- function(num_teams=4, num_rounds=1, dist='uniform'){
   # define teams
   teams <- paste("Team", 1:num_teams)
+
+  if (dist == "uniform"){
+    stren <- qunif(1:num_teams/(num_teams+1), 0, num_teams+1)
+  }
+  else if (dist == "normal"){
+    stren = qnorm(((1:num_teams) - 0.5) / num_teams, 0, 1)
+  }
+
   # put Teams and True_Strength into data frame
@@
     Team = teams,
-    True_Strength = qunif(1:num_teams/(num_teams+1), 0, sqrt(12)))
+    True_Strength = stren)
 
   # add True_Rank and Seed, order columns
-  team_strengths <- team_strengths %>% arrange(-True_Strength) %>% mutate(True_Rank = row_number()) %>% arrange(Team) %>% mutate(Seed = NA) %>% select(Team, True_Rank, True_Strength, Seed)
+  team_strengths <- team_strengths %>% 
+    arrange(-True_Strength) %>% 
+    mutate(True_Rank = row_number()) %>% 
+    arrange(Team) %>% 
+    mutate(Seed = NA) %>% 
+    select(Team, True_Rank, True_Strength, Seed)
 
@@
 
-  results <- list(average_ranks = average_ranks, k_mean = k_mean, s_mean = s_mean, simulation_list = simulation_list)
+  results <- list(average_ranks = average_ranks, k_mean = k_mean, s_mean = s_mean, simulation_list = simulation_list, all_k = kendallcor, all_s = spearmancor)
 
@@
   # inputs are number of teams and number of rounds
-results <- run_tournament(8, 1)
+results <- run_tournament(8, 1, "uniform")
 sims <- results$simulation_list
 
-results
+# results
 ```
@@
     -   3-8 teams, 16 teams
+-   save them as list objects
 
-- save them as list objects
+## Correlation Coefficients Histograms
+
+```{r}
+hist(results$all_k, main = "Kendall Cor Dist", xlab = "Kendall Correlation")
+hist(results$all_s, main = "Spearman Cor Dist", xlab = "Spearman Correlation")
+```

### 2025-02-20 mutual_information.R
mutual information

@@ -1,0 +1,154 @@
+dat <- expand.grid(x = c("12","21"),
+                   y = c("12","21"))
+p <- 0.9
+
+f <- function(p){
+dat$prob <- c(p,1-p,1-p,p)/2
+sum(dat$prob*log(dat$prob/(0.5*0.5),2))
+}
+f <- Vectorize(f)
+plot(0:999/1000,f(0:999/1000),xlab = "Prob Team 1 beats Team 2", ylab = "Bits", type = "l")
+
+
+
+
+dat <- expand.grid(x = c("12"),
+                   y = c("12","21"))
+p <- 0.9
+
+f <- function(p){
+  dat$prob <- c(p,1-p)/2
+  2*sum(dat$prob*log(dat$prob/(0.5*0.5),2))
+}
+f <- Vectorize(f)
+plot(0:999/1000,f(0:999/1000),xlab = "Prob Team 1 beats Team 2", ylab = "Bits", type = "l")
+
+
+dat <- expand.grid(x = c("123"),
+                   y = c("123","132","213","231","312","321"))
+
+p <- 0.9
+dat$prob <- c(p,rep((1-p)/5,5))/factorial(3)
+factorial(3)*sum(dat$prob*log(dat$prob/(1/factorial(3)*1/factorial(3)),2))
+}
+
+p <- 1/6 
+f <- function(p){
+  dat$prob <- c(rep((1-p)/5,5),p)/factorial(3)
+  factorial(3)*sum(dat$prob*log(dat$prob/(1/factorial(3)*1/factorial(3)),2))
+}
+f <- Vectorize(f)
+plot(0:999/1000,f(0:999/1000),xlab = "Prob Team 1 beats Team 2", ylab = "Bits", type = "l")
+abline(v = 1/6)
+
+f <- function(p){
+  dat$prob <- c(0.01,0.95,0.01,0.01,0.01,0.01)/factorial(3)
+  factorial(3)*sum(dat$prob*log(dat$prob/(1/factorial(3)*1/factorial(3)),2))
+}
+f <- Vectorize(f)
+points(0:999/1000,f(0:999/1000),xlab = "Prob Team 1 beats Team 2", ylab = "Bits", type = "l")
+abline(v = 1/6)
+
+
+
+joint * log(joint/(marginx*marginy))
+
+
+dat <- expand.grid(x = c("123","132","213","231","312","321"),
+           y = c("123","132","213","231","312","321"))
[…]

### 2025-02-20 double_elimination.qmd

@@
       filter(!is.na(true_rank)) %>%
-      arrange(rank_hat) %>%
-      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
+      arrange(true_rank) %>%
+      mutate(rank_hat = rank(rank_hat, ties.method = "first"))
   }
@@
 
-test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1)
-test
+#test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1)
+#test
 #manual_test <- simulate_double_elimination_tournament(num_teams = 4, distribution = "Manual", ties=F, series=1)
 #manual_test
+
+# Gets all results of a normal distribution num_teams between 4 and 64 single-game elimination
+de_norm_results <- list()
+for (teams in c(4,8,16,32,64)) {
+  de_norm_results[[as.character(teams)]] <- list()
+  for (replicate in 1:10000) {
+    # Store each simulation result in the list with unique indexing
+    de_norm_results[[as.character(teams)]][[replicate]] <- simulate_double_elimination_tournament(
+      num_teams = teams,
+      distribution = "Normal",
+      ties = FALSE,
+      series = 1
+    )
+  }
+}
+
+de_norm_results <- do.call(rbind, lapply(names(all_normal_results), function(teams) {
+  team_results <- do.call(rbind, all_normal_results[[teams]])
+  team_results$teams <- as.numeric(teams)
+  return(team_results)
+}))
+de_norm_results
 

### 2025-02-20 single_elimination.qmd

@@
 title: "single_elimination"
-format: html
+format: 
+  html:
+    embed-resources: true
 editor: visual
 ---
+
+```{r}
+library(tidyverse)
+library(dplyr)
+library(shiny)
+```
 
@@
 simulate_match <- function(team1, team2, strength1, strength2) {
-  odds <- exp(strength1 - strength2)/(1+exp(strength1 - strength2))
-
-  winner <- ifelse(runif(1) < odds, team1, team2)  # Randomize outcome
+  if(is.na(team1)){
+    winner <- team2
+  }
+  else if(is.na(team2)){
+    winner <- team1
+  }
+  else {
+  p <- exp(strength1 - strength2)/(1+exp(strength1 - strength2))
+  winner <- ifelse(runif(1) < p, team1, team2)  # Randomize outcom
+  }
   return(winner)
 }
+# sample("A","B"),p=(p,1-p)
+
 ```
@@
 
+## 8 Teams (all combinations)
+
 ```{r}
 num_teams <- 8
-teams <- paste("Team", 1:num_teams)
+teams <- paste("Team",num_teams:1)
 team_map <- setNames(1:8, teams)
 tourn <- do.call(rbind, args = combinat::permn(1:8))
-tourn
 dist <- apply(X = tourn, MARGIN = 1, FUN = function(x){
@@
 dist <- t(dist)
-seeds <- dist[!duplicated(dist), ]
-
+seeding_structure <- dist[!duplicated(dist), ]
+seeding_structure_char <- apply(seeding_structure,1,function(x) {paste0(x,collapse="")})
 # How to check for a specific seed
-which(apply(seeds, 1, function(x) all(x == c(1, 8, 2, 7, 3, 6, 4, 5))))
+which(apply(seeding_structure, 1, function(x) all(x == c(1, 8, 4, 5, 2, 7, 3, 6))))
 
@@
 
+## 4 Teams (all combinations)
+
+```{r}
+num_teams <- 4
+teams <- paste("Team",num_teams:1)
+team_map <- setNames(1:num_teams, teams)
+tourn <- do.call(rbind, args = combinat::permn(1:num_teams))
+dist <- apply(X = tourn, MARGIN = 1, FUN = function(x){
+  if (sum(x[3:4] == min(x[1:4])) > 0) {
+    x[1:4] <- x[c(3:4, 1:2)]
+  }
+  for (i in c(1, 3)) {
+    x[i:(i + 1)] <- sort(x[i:(i + 1)])
+  }
+return(x)
+})
+dist <- t(dist)
+seeding_structure <- dist[!duplicated(dist), ]
+seeding_structure_char <- apply(seeding_structure,1,function(x) {paste0(x,collapse#="")})
+
+#num_teams <- 5
+#teams <- paste("Team",num_teams:1)
+#team_map <- setNames(1:num_teams, teams)
+#tourn <- do.call(rbind, args = combinat::permn(1:num_teams))
+#tourn[tourn[,1] == 1,]
+```
+
+## Creating the traditional tournament structure for n teams
+
+```{r}
+generate_tournament_structure <- function(n) {
+
+  is_power_of_two <- function(x) {
+    x > 0 && (x & (x - 1)) == 0
+  }
+
+  # If n is not a power of 2, round up to the next power of 2 and add placeholders
+  if (!is_power_of_two(n)) {
+    next_power_of_two <- 2^ceiling(log2(n))
+    extra_teams <- next_power_of_two - n
+    message(sprintf("Rounding up to %d teams. Adding %d NA placeholders.", next_power_of_two, extra_teams))
+    n <- next_power_of_two
+  } else {
+    extra_teams <- 0
+  }
+
+  rounds <- log(n, base = 2) - 1
+  teams <- c(1, 2)  # Initial list of teams
+
+  for (i in 1:rounds) {
+    teams <- nextLayer(teams)
+  }
+
+  while (extra_teams > 0){
+      teams[which.max(teams)] <- NA
+      extra_teams <- extra_teams - 1
+  }
+
+  return(teams)
+}
+
+nextLayer <- function(teams) {
+  out <- c()  # Initialize an empty vector
+  length <- length(teams) * 2 + 1  # Calculate the length for pairing
+
+  # Generate the next layer by pairing teams
+  for (i in teams) {
+    out <- c(out, i, length - i)  # Push the current team and its pair
+  }
+
+  return(out)
+}
+
+```
+
 # Setting Up the Teams and Strengths
@@
 ```{r}
-# Assign random strengths to each team 
-team_strengths <- data.frame(
-  Team = teams,
-  Strength = c(1,2,3,4,5,6,7,8),
-  Wins = rep(0,8),
-  Ranks = rep(NA,8),
-  Exp_Ranks = c(8,7,6,5,4,3,2,1)
-)
-
+# Assign random strengths to each team coming from normal and uniform distributions
 normal_strengths <- data.frame(
   Team = teams,
-  Strength = round(rnorm(8,0,1),2),
-  Wins = rep(0,8),
-  Ranks = rep(NA,8)
+  Strength = sapply(num_teams, function(n) { qnorm(1:n/(n+1)) }),
+  Wins = rep(0,num_teams),
+  Ranks = rep(NA,num_teams)
 )
-normal_strengths$Exp_Ranks <- rank(-normal_strengths$Strength, ties.method = "average")
+normal_strengths$True_Rank <- rank(-normal_strengths$Strength, ties.method = "average")
+normal_strengths <- arrange(normal_strengths, True_Rank)
 
@@
   Team = teams,
-  Strength = round(runif(8,0,8),2),
-  Wins = rep(0,8),
-  Ranks = rep(NA,8)
+  Strength = sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) }),
+  Wins = rep(0,num_teams),
+  Ranks = rep(NA,num_teams)
 )
-unif_strength$Exp_Ranks <- rank(-unif_strength$Strength, ties.method = "average")
+unif_strength$True_Rank <- rank(-unif_strength$Strength, ties.method = "average")
+unif_strength <- arrange(unif_strength, True_Rank)
 
@@
 strengths_list <- list(
-  "Default" = team_strengths,
   "Normal" = normal_strengths,
@@
   # Add more ranks as needed
+  round_of_sixteen = 12.5,
   quarterfinalist = 6.5,
   semifinalist = 3.5,
+  fourth_place = 4,
+  third_place = 3,
   finalist = 2,
@@
 
+# Remove ties in rankings
+noTies <- function(df) {
+  quarterfinals <- df %>%
+    filter(Ranks == rank_values["quarterfinalist"]) %>%
+    mutate(Ranks = sample(5:8, n(), replace = FALSE))  # Assigns unique ranks
+  semifinals <- df %>%
+    filter(Ranks == rank_values["semifinalist"]) %>%
+    mutate(Ranks = sample(3:min(n(),4), n(), replace = FALSE))
+
+  df <- df %>%
+    filter(Ranks != rank_values["quarterfinalist"], Ranks != rank_values["semifinalist"]) %>%
+    bind_rows(quarterfinals,semifinals)  # Replace old quarterfinalists with unique ranks
+
+  return(df)
+}
+
 # Simulate the single-elimination tournament
-simulate_tournament <- function(df,seeds = NULL) {
+simulate_tournament <- function(df,seeding_structure = NULL, ties=T, series=1, third_place = T) {
 
   # Sets seed to that of common best vs worst structure
-  if (is.null(seeds)){
-    seeds <- matrix(c(1,8,2,7,3,6,4,5), nrow = 1)
+  n <- length(unique(df$Team))
+  if (is.null(seeding_structure)){
+    seeding_structure <- matrix(generate_tournament_structure(n), nrow = 1)
+  } 
+  if(series %% 2 == 0) {
+    stop("Series must be an odd number to avoid potential ties.")
   }
@@
 
-  for (i in 1:nrow(seeds)) {
+  for (i in 1:nrow(seeding_structure)) {
     # Create a new data frame based on the current permutation in 'test'
-    permutation <- seeds[i, ]
+    permutation <- seeding_structure[i, ]
     permuted_df <- df
@@
     permuted_df$Wins <- 0  # Initialize wins column
-    permuted_df$Exp_Rank <- df$Exp_Rank[permutation]
+    permuted_df$True_Rank <- df$True_Rank[permutation]
 
@@
 
+    # Track the losers of the semifinals for the third-place match
+    losers_semis <- c()
+
   while (length(teams) > 1) {
-    #cat("\n--- Round", round_number, "---\n")
+    #cat("\n--- New Round ---\n")
     next_round <- c() # Makes sure only winners are included in next round
@@
 
+      wins_team1 <- 0
+      wins_team2 <- 0
+      for (game in 1:series){
       match_winner <- simulate_match(team1, team2, strength1, strength2)
-      next_round <- c(next_round, match_winner)
-      #cat(paste(team1, "vs", team2, "-> Winner:", match_winner, "\n"))
+      if (match_winner == team1) {
+            wins_team1 <- wins_team1 + 1
+          } else {
+            wins_team2 <- wins_team2 + 1
+          }
 
-      # Can add more rounds by the same code with rounnd_number increasing by 1 and rank_values[round name]
+          # Determine the overall winner if a majority is reached
+          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
+            break
+          }
+        }
+
+        # Determine the match winner based on series results
+        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
+      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
+      next_round <- c(next_round, series_winner)
+
+      # Can add more rounds by the same code with round_number increasing by 1 and rank_values[round name]
 
@@
           rank_values["semifinalist"]
+          losers_semis <- c(losers_semis, team2)
         } else {
@@
           rank_values["semifinalist"]
+          losers_semis <- c(losers_semis, team1)
         }
+
       } else if (round_number == 1) { # Finals
@@
 
+  # Third place match between the losers of the semifinals
+  if (third_place == TRUE){
+    team1 <- losers_semis[1]
+    team2 <- losers_semis[2]
+    strength1 <- strengths[permuted_df$Team == team1]
+    strength2 <- strengths[permuted_df$Team == team2]
+    wins_team1 <- 0
+    wins_team2 <- 0
+    for (game in 1:series) {
+      match_winner <- simulate_match(team1, team2, strength1, strength2)
+      if (match_winner == team1) {
+        wins_team1 <- wins_team1 + 1
+      } else {
+        wins_team2 <- wins_team2 + 1
+      }
+     # Determine the overall winner if a majority is reached
+      if (wins_team1 > series / 2 || wins_team2 > series / 2) {
+      break
+    }
+   # Assign third place and fourth place ranks
+      if (wins_team1 > wins_team2) {
+        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team1))
+        permuted_df$Ranks[permuted_df$Team == team1] <- rank_values["third_place"]
+        permuted_df$Ranks[permuted_df$Team == team2] <- rank_values["fourth_place"]
+      } else {
+        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team2))
+        permuted_df$Ranks[permuted_df$Team == team2] <- rank_values["third_place"]
+        permuted_df$Ranks[permuted_df$Team == team1] <- rank_values["fourth_place"]
+      }
+    }
+  }
+
   results[[i]] <- permuted_df # Store the results of this permutation
   }
-  return(do.call(rbind, results))
+
+  final_results <- do.call(rbind, results)
+  # Remove ties if ties = FALSE
+  if (!ties) {
+    final_results <- noTies(final_results)
   }
+
+  return(final_results)
+}
+
+test <- simulate_tournament(normal_strengths,ties=T, series=3)
+test[order(-test$Strength),]
+
+# Run the tournament simulation
+simulation_results <- replicate(1000, simulate_tournament(unif_strength), simplify = FALSE)
+
+# Combine all replicate results
+all_results <- do.call(rbind, simulation_results)
+
+# Summarize total wins
+total_wins_summary <- all_results %>%
+  group_by(Team) %>%
+  summarise(
+    Total_Wins = sum(Wins, na.rm = TRUE),
+    .groups = 'drop'
+  ) %>%
+  arrange(desc(Total_Wins))
+
+# Print the total wins summary
+print(total_wins_summary)
+
+# Run the tournament simulation
+simulation_series <- replicate(1000, simulate_tournament(unif_strength, series=7), simplify = FALSE)
+
+# Combine all replicate results
+series_results <- do.call(rbind, simulation_series)
+
+# Summarize total wins
+total_wins_series <- series_results %>%
+  group_by(Team) %>%
+  summarise(
+    Total_Wins = sum(Wins, na.rm = TRUE),
+    .groups = 'drop'
+  ) %>%
+  arrange(desc(Total_Wins))
+
+# Print the total wins summary
+print(total_wins_series)
 
@@
 ```{r}
-result <- simulate_tournament(team_strengths,seeds)
-result[order(-result$Strength),]
-
-norm_result <- simulate_tournament(normal_strengths,seeds)
+norm_result <- simulate_tournament(normal_strengths,seeding_structure)
 norm_result[order(-norm_result$Strength),]
 
-unif_result <- simulate_tournament(unif_strength,seeds)
+unif_result <- simulate_tournament(unif_strength,seeding_structure)
 unif_result[order(-unif_result$Strength),]
@@
 
-# Multiple Simulations
+# Multiple Simulations for 1 Seeding Structure
 
 ```{r}
-tourn_sims <- function(df, seeds = NULL, replicates = 100) {
-  tournament_replicates <- replicate(replicates, {
+tourn_sims <- function(df, seeding_structure = NULL, seed_char = NULL, replicates = 100, series=1) {
+  # Default seeding structure if none provided
+  if (is.null(seeding_structure)) {
+    seeding_structure <- matrix(c(1,8,4,5,2,7,3,6), nrow = 1)
+    seed_char <- apply(seeding_structure, 1, function(x) paste0(x, collapse = ""))
+  }
+
+  # List to store results for each seed configuration
+  tournament_replicates <- lapply(1:nrow(seeding_structure), function(seed_num) {
+    replicate(replicates, {
       # Reset wins and ranks for each tournament
@@
 
-    if (is.null(seeds)) {
-      seeds <- matrix(c(1, 8, 2, 7, 3, 6, 4, 5), nrow = 1)
-    }
+      # Run the tournament with the current seed configuration
+      result <- simulate_tournament(df, seeding_structure[seed_num, , drop = FALSE], series = series)
 
-    # Run the tournament
-    result <- simulate_tournament(df, seeds)
+      # Add Seed_Num column to keep track of the current seed configuration
+      result$Seed_Num <- seed_num
+      result$Seeding_Structure <- seed_char[seed_num]
 
-    return(result[, c("Team", "Strength", "Wins", "Ranks", "Exp_Rank")])  
-  }, simplify = FALSE)  # Prevents simplification to atomic vector
+      return(result[, c("Team", "Strength", "Wins", "Ranks", "True_Rank", "Seed_Num", "Seeding_Structure")])  
+    }, simplify = FALSE)
+  })
 
-  combined_results <- do.call(rbind, tournament_replicates)
+  # Combine results for all seed configurations
+  combined_results <- do.call(rbind, unlist(tournament_replicates, recursive = FALSE))
 
-  # Calculate average ranks and total wins using dplyr
+  # Summarize results
   results_summary <- combined_results %>%
-    group_by(Team, Strength, Exp_Rank) %>%  # Track SeedNum
+    group_by(Seed_Num, Seeding_Structure, Team, Strength, True_Rank) %>%
     summarise(
       Average_Rank = mean(Ranks, na.rm = TRUE),
-      Total_Wins = sum(Wins, na.rm = TRUE)
+      Total_Wins = sum(Wins, na.rm = TRUE),
+      .groups = 'drop'
     ) %>%
@@
   # Calculate Kendall tau correlation
-  kendall_cor <- cor(results_sim[[name]]$Average_Rank, results_sim[[name]]$Exp_Rank, method = "kendall", use = "complete.obs")
+  kendall_cor <- cor(results_sim[[name]]$Average_Rank, results_sim[[name]]$True_Rank, method = "kendall", use = "complete.obs")
   cat("Kendall Tau Correlation:", kendall_cor, "\n")
@@
   # Calculate Spearman correlation
-  spearman_cor <- cor(results_sim[[name]]$Average_Rank, results_sim[[name]]$Exp_Rank, method = "spearman", use = "complete.obs")
+  spearman_cor <- cor(results_sim[[name]]$Average_Rank, results_sim[[name]]$True_Rank, method = "spearman", use = "complete.obs")
   cat("Spearman Correlation:", spearman_cor, "\n")
@@
   # Create a barplot for total wins
-  barplot(
+  bar <- barplot(
     results_sim[[name]]$Total_Wins,
-    names.arg = results_sim[[name]]$Strength,
+    names.arg = results_sim[[name]]$True_Rank,
     main = paste("Total Wins for", name),
-    xlab = "Strength",
+    xlab = "True Rank",
+    ylim = c(0,100),
     col = "blue",  # Added color for better visualization
@@
   )
+  text(
+    x = bar,
+    y = results_sim[[name]]$Total_Wins + 50,
+    labels = results_sim[[name]]$Total_Wins,
+    cex = 0.8,
+    col = "black"
+  )
 }
 
+num_games <- 7
+series_sim <- lapply(strengths_list, function(strength_config) {
+  tourn_sims(strength_config, series = num_games)
+})
+# Print the results for each configuration
+for (name in names(series_sim)) {
+  cat("\nResults for strength configuration:", name, "\n")
+  print(series_sim[[name]])
+
+  # Calculate Kendall tau correlation
+  kendall_cor <- cor(series_sim[[name]]$Average_Rank, series_sim[[name]]$True_Rank, method = "kendall", use = "complete.obs")
+  cat("Kendall Tau Correlation:", kendall_cor, "\n")
+
+  # Calculate Spearman correlation
+  spearman_cor <- cor(series_sim[[name]]$Average_Rank, series_sim[[name]]$True_Rank, method = "spearman", use = "complete.obs")
+  cat("Spearman Correlation:", spearman_cor, "\n")
+
+  # Create a barplot for total wins
+  bar <- barplot(
+    series_sim[[name]]$Total_Wins,
+    names.arg = series_sim[[name]]$True_Rank,
+    main = paste("Total Wins for", name),
+    xlab = "True Rank",
+    ylim = c(0,100),
+    col = "blue",  # Added color for better visualization
+    border = "black"  # Added border color for bars
+  )
+  text(
+    x = bar,
+    y = series_sim[[name]]$Total_Wins + 50,
+    labels = series_sim[[name]]$Total_Wins,
+    cex = 0.8,
+    col = "black"
+  )
+}
+```
+
+# Probability of Each Seed Winning With Each Structure
+
+```{r}
+library(dplyr)
+library(ggplot2)
+
+sims_norm <- tourn_sims(normal_strengths,seeding_structure, seeding_structure_char, 100)
+sims_unif <- tourn_sims(unif_strength,seeding_structure, seeding_structure_char, 100)
+
+graphing <- function(df, distribution_name){
+  # Loop over each unique True_Rank
+  for (rank in unique(df$True_Rank)){
+    # Filter data for the current True_Rank
+    team_data <- df %>% filter(True_Rank == rank)
+
+    # Rank the teams based on Total_Wins in descending order, ensuring ties are handled correctly
+    best <- team_data %>%
+      arrange(desc(Total_Wins)) %>%
+      head(25)
+
+    worst <- team_data %>%
+      arrange(Total_Wins) %>%
+      head(25) %>%
+      arrange(desc(Total_Wins))
+
+    # Set Seeding_Structure as a factor to ensure the x-axis ordering by Seeding_Structure
+    best$Seeding_Structure <- factor(best$Seeding_Structure, levels = unique(best$Seeding_Structure))
+    worst$Seeding_Structure <- factor(worst$Seeding_Structure, levels = unique(worst$Seeding_Structure))
+
+    # Generate the plot for the current True_Rank - Best Plot
+    best_plot <- ggplot(best, aes(x = Seeding_Structure, y = Total_Wins, fill = as.factor(True_Rank))) +
+      geom_bar(stat = "identity", position = "dodge", width = 0.7) +  
+      geom_text(aes(label = Total_Wins), size = 3, vjust = -0.3, color = "black") +
+      labs(
+        title = paste("Most Total Wins by Seed:", rank, "in Distribution:", distribution_name),
+        x = "Seed Number",
+        y = "Total Wins",
+        fill = "True Rank"
+      ) +
+      theme_minimal() +
+      theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability
+
+    # Print the plot for each True_Rank
+    print(best_plot)
+
+    # Generate the plot for the current True_Rank - Worst Plot
+    worst_plot <- ggplot(worst, aes(x = Seeding_Structure, y = Total_Wins, fill = as.factor(True_Rank))) +
+      geom_bar(stat = "identity", position = "dodge", width = 0.7) + 
+      geom_text(aes(label = Total_Wins), size = 3, vjust = -0.3, color = "black") +
+      facet_wrap(~ True_Rank, scales = "free_x") +  
+      labs(
+        title = paste("Least Total Wins by Seed:", rank, "in Distribution:", distribution_name),
+        x = "Seed Number",
+        y = "Total Wins",
+        fill = "True Rank"
+      ) +
+      theme_minimal() +
+      theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability
+
+    # Print the plot for each True_Rank
+    print(worst_plot)
+  }
+}
+
+# Example: Call graphing function for both distributions
+graphing(sims_norm, "Normal")
+graphing(sims_unif, "Uniform")
+
+```
+
+```{r}
+seeding_structure[c(42,86,104,74,41,139,75,117,141,140),]
+```
+
+# Individual Spearman Correlations by Seeding Structure
+
+```{r}
+spearman_results <- function(df, seeding_structure, replicates = 100) {
+  tournament_replicates <- replicate(replicates, {
+    # Reset wins and ranks for each tournament
+    df$Wins <- rep(0, nrow(df))
+    df$Ranks <- rep(NA, nrow(df))  # Ensure Ranks is initialized
+    df$True_Rank <- rank(-df$Strength)
+
+    # Run the tournament
+    result <- simulate_tournament(df, seeding_structure, ties=FALSE)
+
+    # Calculate Spearman correlation for this replicate
+    spearman_cor <- cor(result$Ranks, result$True_Rank, method = "spearman", use = "complete.obs")
+
+    # Return the average Spearman correlation (not the full simulation details)
+    return(spearman_cor)
+  }, simplify = TRUE)  # Keeps the result as a vector of Spearman correlations
+
+  # Calculate average Spearman correlation for the entire set of replicates
+  average_spearman <- mean(tournament_replicates, na.rm = TRUE)
+
+  return(average_spearman)
+}
+
+output_spearman <- function(strengths_list, seeding_structure, seeding_structure_char = NULL) {
+  # Preallocate results list
+  results <- vector("list", length = nrow(seeding_structure) * length(strengths_list))
+  counter <- 1  # Index for results
+
+  # Loop over each seed configuration
+  for (seed_index in 1:nrow(seeding_structure)) {
+    current_seed <- matrix(seeding_structure[seed_index, ], nrow = 1)
+    current_structure <- if (!is.null(seeding_structure_char)) {
+      seeding_structure_char[seed_index]
+    } else {
+      paste(seeding_structure[seed_index, ], collapse = "-")
+    }
+
+    # Loop over each distribution type
+    for (dist_name in names(strengths_list)) {
+      df <- strengths_list[[dist_name]]
+
+      # Calculate average Spearman correlation
+      avg_spearman <- spearman_results(df, current_seed)
+
+      # Store results
+      results[[counter]] <- data.frame(
+        Seed = seed_index,
+        Seeding_Structure = current_structure,
+        Distribution = dist_name,
+        Average_Spearman = avg_spearman
+      )
+      counter <- counter + 1
+    }
+  }
+
+  # Combine all results into a single data frame
+  final_results <- do.call(rbind, results)
+  return(final_results)
+}
+
+# Example usage
+#output_results(strengths_list, seeds = seeds)
+
+```
+
+# Average Spearman Correlation by Seeding Structure
+
+```{r}
+spearman_by_seed <- function(df, seeding_structure, seeding_structure_char, replicates = 100) {
+  tournament_replicates <- replicate(replicates, {
+    # Reset wins and ranks for each tournament
+    df$Wins <- rep(0, nrow(df))
+    df$Ranks <- rep(NA, nrow(df))  # Ensure Ranks is initialized
+    df$True_Rank <- rank(-df$Strength)
+
+    # Run the tournament
+    result <- simulate_tournament(df, seeding_structure,ties=FALSE)
+
+    # Calculate Spearman correlation for this replicate
+    spearman_cor <- cor(result$Ranks, result$True_Rank, method = "spearman", use = "complete.obs")
+
+    # Return the average Spearman correlation (not the full simulation details)
+    return(spearman_cor)
+  }, simplify = TRUE)  # Keeps the result as a vector of Spearman correlations
+
+  return(tournament_replicates)
+}
+
+output_results <- function(strengths_list, seeding_structure, seeding_structure_char = NULL) {
+  # Preallocate results list
+  results <- vector("list", length = nrow(seeding_structure) * length(strengths_list))
+  counter <- 1  # Index for results
+
+  # Loop over each seed configuration
+  for (seed_index in 1:nrow(seeding_structure)) {
+    current_seed <- matrix(seeding_structure[seed_index, ], nrow = 1)
+    current_structure <- if (!is.null(seeding_structure_char)) {
+      seeding_structure_char[seed_index]
+    } else {
+      paste(seeding_structure[seed_index, ], collapse = "-")
+    }
+
+    # Loop over each distribution type
+    for (dist_name in names(strengths_list)) {
+      df <- strengths_list[[dist_name]]
+
+      # Calculate average Spearman correlation
+      avg_spearman <- spearman_by_seed(df, current_seed)
+
+      # Store results
+      results[[counter]] <- data.frame(
+        Seed = seed_index,
+        Seeding_Structure = current_structure,
+        Distribution = dist_name,
+        Spearman = avg_spearman
+      )
+      counter <- counter + 1
+    }
+  }
+
+  # Combine all results into a single data frame
+  final_results <- do.call(rbind, results)
+  return(final_results)
+}
+
+# Example usage
+#output_results(strengths_list, seeds = seeds)
 ```
@@
 ```{r}
-all_results <- function (strength){
-    all_seeds <- list()
-    total_seeds <- nrow(seeds)  # Store the total number of seeds
-    for (i in 1:total_seeds) {
-        current_seed <- seeds[i, , drop = FALSE]
-        result <- suppressMessages(tourn_sims(team_strengths, current_seed))
-        result$SeedNum <- i  # Add seed number to the result
-        all_seeds[[paste0("Seed_", i)]] <- result
-    }
-    return(all_seeds)
+graph_results <- output_results(strengths_list, seeding_structure = seeding_structure, seeding_structure_char) 
+library(ggplot2)
+#ggplot(data=graph_results, aes(x=as.factor(Seeding_Structure),y=Spearman)) + geom_boxplot() + facet_grid(~Distribution)
+
+#ggplot(data=graph_results, aes(x=Spearman,fill=as.factor(Seeding_Structure))) + geom_density(alpha=0.5) + facet_grid(~Distribution)
+
+sim_results <- output_spearman(strengths_list, seeding_structure = seeding_structure, seeding_structure_char) 
+#ggplot(data=sim_results, aes(x=as.factor(Seeding_Structure),y=Average_Spearman)) + geom_bar(stat="identity", position = "dodge") + facet_grid(~Distribution)
+# Sort and extract top 5 highest Spearman correlations
+norm_top_5_seeds <- sim_results %>%
+  filter(Distribution == "Normal") %>%
+  arrange(desc(Average_Spearman)) %>%
+  slice_head(n = 5)  # Top 5 highest Spearman correlations
+
+unif_top_5_seeds <- sim_results %>%
+  filter(Distribution == "Uniform") %>%
+  arrange(desc(Average_Spearman)) %>%
+  slice_head(n = 5)  # Top 5 highest Spearman correlations
+
+# Sort and extract top 5 lowest Spearman correlations
+norm_worst_5_seeds <- sim_results %>%
+  filter(Distribution == "Normal") %>%
+  arrange(Average_Spearman) %>%
+  slice_head(n = 5)  # Top 5 lowest Spearman correlations
+
+unif_worst_5_seeds <- sim_results %>%
+  filter(Distribution == "Uniform") %>%
+  arrange(Average_Spearman) %>%
+  slice_head(n = 5)  # Top 5 highest Spearman correlations
+
+print(norm_top_5_seeds)
+print(norm_worst_5_seeds)
+print(unif_top_5_seeds)
+print(unif_worst_5_seeds)
+```
+
+# Probability of Top 3 Correct
+
+```{r}
+# Function to run multiple simulations
+run_simulations <- function(df, seeding_structure = NULL, seed_char = NULL, replicates = 1000, series = 1,ties=FALSE) {
+  # Default seeding structure
+  if (is.null(seeding_structure)) {
+    seeding_structure <- matrix(c(1, 8, 4, 5, 2, 7, 3, 6), nrow = 1)
   }
 
-# Store strength data frames in a list
-sims_list <- list(
-    "Default" = all_results(team_strengths),
-    "Normal" = all_results(normal_strengths),
-    "Uniform" = all_results(unif_strength)
+  # Store results for all simulations
+  all_simulations <- lapply(1:replicates, function(rep) {
+    # Simulate tournament for each seeding structure
+    lapply(1:nrow(seeding_structure), function(seed_num) {
+      simulate_tournament(df,seeding_structure,series = series,third_place = TRUE
+      ,ties=FALSE) %>%
+        mutate(Simulation = rep, Seed_Num = seed_num, Seeding_Structure = seed_char[seed_num])
+    }) %>%
+      do.call(rbind, .)
+  }) %>%
+    do.call(rbind, .)
+
+  return(all_simulations)
+}
+
+# Analyze Top-Three Ranks
+analyze_top_three <- function(simulations) {
+
+  # Identify top-3 ranks for each simulation
+  top_three_results <- simulations %>%
+    group_by(Simulation) %>%
+    filter(Ranks %in% 1:3 & True_Rank %in% 1:3) %>%
+    summarize(match_count = n_distinct(True_Rank), .groups = 'drop') %>%
+    filter(match_count == 3)
+
+  top_three_prob = nrow(top_three_results) / n_distinct(simulations$Simulation)
+  return(top_three_prob)
+}
+
+simulations <- run_simulations(normal_strengths, replicates = 1000, series = 3, ties=FALSE)
+
+# Analyze top-three occurrences
+top_three_summary <- analyze_top_three(simulations)
+
+# Print results
+print(top_three_summary)
+
+```
+
+-   Probability that the top 3 are correct (1,2,3) in any order
+
+-   16 team seeding
+
+-   byes and third team game
+
+-   curve showing top i seeds correct (for 2 is 1 and 2 in the final); overlap a curve showing the correct order; probability of 1, 2\|1, 3\|(1,2) or (2,1) etc; use %in%
+
+-   series best of n
+
+    ```{r}
+    curve <- function(sim){
+      probability <- numeric(num_teams)
+      for (i in 1:num_teams){
+        if (i==1){
+          probability[i] <- nrow(sim[sim$Ranks == 1 & sim$True_Rank == 1, ]) / length(unique(sim$Simulation))
+        }
+        else{
+        top_ranks <- sim[sim$Ranks <= (i-1) & sim$True_Rank <= (i-1), ]
+        better_teams <- 1:(i-1)
+        numerator <- sum(sapply(unique(top_ranks$Simulation), function(sim_id){
+          sim_subset <- sim[sim$Simulation == sim_id, ]
+
+          if(all(better_teams %in% sim_subset$True_Rank[sim_subset$Ranks <= (i - 1)])){
+          return(sum(sim_subset$Ranks == i & sim_subset$True_Rank == i))
+          }
+          else{
+            return(0)
+          }
+          }))
+
+        denominator <- sum(sapply(unique(top_ranks$Simulation), function(sim_id){
+            sim_subset <- top_ranks[top_ranks$Simulation == sim_id, ]
+            all(better_teams %in% sim_subset$True_Rank)
+          }))
+        #print(i)
+        #print(numerator)
+        #print(denominator)
+        probability[i] <- numerator / denominator
+        }
+      }
+      return(probability)
+    }
+    simulations <- run_simulations(normal_strengths, replicates = 2000, series = 7, ties=FALSE)
+    sim_probs <- curve(simulations)
+    probabilities <- data.frame(x=c(1:num_teams), y=sim_probs)
+    probabilities$Line <- "Simulated"
+
+    baseline_prob <- 1/(num_teams:1)
+    baseline <- data.frame(x=c(1:num_teams), y=baseline_prob)
+    baseline$Line <- "Baseline"
+
+    prob_plot <- rbind(baseline,probabilities)
+
+    ggplot(data=prob_plot, aes(x=x,y=y, color=Line, group=Line)) + geom_point() + geom_line() + ylim(0,1) + labs(
+      x = "Rank",
+      y = "Conditional Probability"
     )
 
-# Initialize a list to store Spearman correlation results separately
-spearman_results_list <- list()
-
-# Loop through each simulation and calculate Spearman correlation
-for (name in names(sims_list)) {
-    sim_results <- do.call(rbind, sims_list[[name]])
-
-    # Calculate the Spearman correlation only once for the unique ranks
-    spearman_cor <- cor(sim_results$Average_Rank, sim_results$Exp_Rank, method = "spearman", use = "complete.obs")
-
-    # Store the overall Spearman correlation result
-    spearman_results_list[[name]] <- data.frame(Configuration = name, 
-                                                  Spearman_Correlation = spearman_cor)
-
-    # Keep only relevant columns and ensure distinct SeedNum
-    distinct_results <- sim_results[, c("SeedNum", "Average_Rank", "Exp_Rank")]
-
-    # Calculate the Spearman correlation for each distinct SeedNum
-    distinct_spearman_results <- distinct_results %>%
-        group_by(SeedNum) %>%
-        summarize(Spearman_Correlation = cor(Average_Rank, Exp_Rank, method = "spearman", use = "complete.obs")) %>%
-        ungroup()
-
-    # Get top 5 highest distinct Spearman correlations for this configuration
-    top_5 <- distinct_spearman_results %>%
-        arrange(desc(Spearman_Correlation)) %>%
-        slice_head(n = 5)  # Select top 5
-
-    # Get top 5 lowest distinct Spearman correlations for this configuration
-    lowest_5 <- distinct_spearman_results %>%
-        arrange(Spearman_Correlation) %>%
-        slice_head(n = 5)  # Select lowest 5
-
-    # Print results with seed numbers
-    cat(paste("\nTop 5 Distinct Spearman Correlations for configuration:", name, "\n"))
-    print(top_5)  # Only SeedNum and Spearman_Correlation
-
-    cat(paste("\nLowest 5 Distinct Spearman Correlations for configuration:", name, "\n"))
-    print(lowest_5)  # Only SeedNum and Spearman_Correlation
-}
-
-# Print overall Spearman results for all configurations
- for (name in names(spearman_results_list)) {
-     cat(paste("\nSpearman Correlation for configuration:", name, "\n"))
-     print(spearman_results_list[[name]])
- }
-
+    # Interesting to note that 3 is almost always greater than 2
     ```

### 2025-02-20 swaps.qmd

---
title: "Swaps"
author: "Zach Culp"
format: html
editor: visual
---

## Number of Swaps to Get Correct Seeding

Notation: PS = number of positions away from the true rank that the predicted rank is

For example, if $R_1 = 3$, then $PS_1 = -2$. PS follows a discrete probability distribution with range of $-n+1 < PS < n-1$, where $n$ is the number of teams in the tournament. My prediction is that E(PS) should be 0 in a fair tournament. Then, if we see multiple tournaments with a predicted PS around 0, the variance can be a determining factor to how well the tournament is structures.

```{r}
library(tidyverse)
se_norm_results <- all_results_norm %>% filter(teams == 8)

se_norm_results$simulation <- rep(1:10000, each = nrow(curves_test) / 10000)

create_ps <- function(df){
  df <- df %>%
  mutate(ps = true_rank - rank_hat) %>%
  group_by(true_rank) %>%
  summarize(

    mean_ps = mean(ps),
    sd_ps = sd(ps)
  )
  return(df)
}
se_test <- create_ps(se_norm_results)
de_test <- create_ps(de_norm_results %>% filter(teams == 8))
rr_test <- create_ps(test_rr_noTies)

library(combinat)
perms_list <- permn(8)
all_data <- list()
# Loop through each permutation and create the data frame
for (i in 1:length(perms_list)) {
  # Create a temporary data frame for the current permutation
  temp_df <- data.frame(
    true_rank = rep(1:8, each = 1),  # Each row gets a true_rank from 1 to 8
    rank_hat = perms_list[[i]],      # The current permutation's ranks
    simulation = rep(i, 8)           # Assign the simulation number to each row
  )
  # Add the temporary data frame to the list
  all_data[[i]] <- temp_df
}
# Combine all the data frames into one
perms_df <- do.call(rbind, all_data)
perms_test <- create_ps(perms_df)
```

### Graph

```{r}
# Combine the two datasets into one for easier plotting
se_test <- se_test %>%
  mutate(type = "Single Elimination")  # Add a label for the type of data

perms_test <- perms_test %>%
  mutate(type = "Baseline")   # Add a label for the type of data

rr_test <- rr_test %>%
  mutate(type = "Round Robin")

de_test <- de_test %>%
  mutate(type = "Double Elimination")

# Combine both data frames into one
combined_swaps_data <- bind_rows(se_test, perms_test, rr_test, de_test)

# Plotting using ggplot2
ggplot(combined_swaps_data, aes(x = true_rank, y = mean_ps, color = type, group = type)) +
  geom_point(size = 3) +  # Plot the points
  geom_line() +  # Connect the points with lines
  geom_errorbar(aes(ymin = mean_ps - sd_ps, ymax = mean_ps + sd_ps), width = 0.2) +  # Add error bars (standard deviation)
  scale_color_manual(values = c("Single Elimination" = "blue", "Baseline" = "red", "Round Robin" = "green", "Double Elimination" = "yellow")) +  # Set colors for the groups
  labs(
    title = "Comparison of Mean PS with Standard Deviation for Simulated vs Baseline",
    x = "True Rank",
    y = "Mean PS",
    color = "Simulation Type"
  ) +
  theme_minimal() + # Apply a clean theme
  theme(
      panel.background = element_rect(fill = "gray75"),  # Set background color to light gray
      plot.background = element_rect(fill = "gray80"),  # Set the plot background color to a lighter gray
      legend.background = element_rect(fill = "gray80")  # Set the legend background color
    )
```

### Baseline Probabilities of Under/Over Performance

An overall graph of the probabilities of correctly predicting the rankings versus overpredicting and underpredicting. If the tournament results were randomly decided the probabilities would be the following:

$P(R_i = i) = \frac{1}{n}$

$P(R_i < i) = \frac{i-1}{n}$

$P(R_i > i) = \frac{n-i}{n}$

```{r}
# Function to calculate probabilities of ps conditions
calculate_probs <- function(df) {
  df %>%
    mutate(ps = true_rank - rank_hat) %>%
    group_by(true_rank) %>%
    summarize(
      ps_equal = sum(ps == 0) / n(),
      ps_less = sum(ps < 0) / n(),
      ps_more = sum(ps > 0) / n()
    )
}

# Apply the function to each data frame
probs_rr_test <- calculate_probs(test_rr_noTies)
probs_se_test <- calculate_probs(se_norm_results)
probs_perms_test <- calculate_probs(perms_df)
probs_de_test <- calculate_probs(de_norm_results_eight)

# Combine the two datasets into one for easier plotting
probs_se_test <- probs_se_test %>%
  mutate(type = "Single Elimination")  # Add a label for the type of data

probs_perms_test <- probs_perms_test %>%
  mutate(type = "Baseline")   # Add a label for the type of data

probs_rr_test <- probs_rr_test %>%
  mutate(type = "Round Robin")

probs_de_test <- probs_de_test %>%
  mutate(type = "Double Elimination")

# Combining datasets
combined_probs_swaps_data <- bind_rows(probs_rr_test, probs_se_test, probs_perms_test, probs_de_test)

# R_i < i plot
ggplot(combined_probs_swaps_data, aes(x = true_rank, y = ps_less, color = type, group = type)) +
  geom_point(size = 3) +  
  geom_line() +
  scale_color_manual(values = c("Single Elimination" = "blue", 
                                "Baseline" = "red", 
                                "Round Robin" = "green",
                                "Double Elimination" = "yellow")) +  
  labs(
    title = "Probability of Simulation Under Predicting Ranks for Each Tournament",
    x = "True Rank",
    y = expression(P(R[i] < i)),
    color = "Simulation Type"
  ) +
  theme_minimal() +  # Apply a clean theme
  theme(
    panel.background = element_rect(fill = "gray75"),  
    plot.background = element_rect(fill = "gray80"), 
    legend.background = element_rect(fill = "gray80")  
  ) +
  scale_y_continuous(limits = c(0, 1))

# R_i = i plot
ggplot(combined_probs_swaps_data, aes(x = true_rank, y = ps_equal, color = type, group = type)) +
  geom_point(size = 3) +  
  geom_line() +
  scale_color_manual(values = c("Single Elimination" = "blue", 
                                "Baseline" = "red", 
                                "Round Robin" = "green",
                                "Double Elimination" = "yellow")) +  
  labs(
    title = "Probability of Simulation Correctly Predicting Ranks for Each Tournament",
    x = "True Rank",
    y = expression(P(R[i] == i)),
    color = "Simulation Type"
  ) +
  theme_minimal() +  # Apply a clean theme
  theme(
    panel.background = element_rect(fill = "gray75"),  
    plot.background = element_rect(fill = "gray80"), 
    legend.background = element_rect(fill = "gray80")  
  ) +
  scale_y_continuous(limits = c(0, 1))

# R_i > i plot
ggplot(combined_probs_swaps_data, aes(x = true_rank, y = ps_more, color = type, group = type)) +
  geom_point(size = 3) +  
  geom_line() +
  scale_color_manual(values = c("Single Elimination" = "blue", 
                                "Baseline" = "red", 
                                "Round Robin" = "green",
                                "Double Elimination" = "yellow")) +  
  labs(
    title = "Probability of Simulation Over Predicting Ranks for Each Tournament",
    x = "True Rank",
    y = expression(P(R[i] > i)),
    color = "Simulation Type"
  ) +
  theme_minimal() +  # Apply a clean theme
  theme(
    panel.background = element_rect(fill = "gray75"),  
    plot.background = element_rect(fill = "gray80"), 
    legend.background = element_rect(fill = "gray80")  
  ) +
  scale_y_continuous(limits = c(0, 1))

```

\

\

### 2025-02-20 sporttau.R
updated.

@@ -1,0 +1,88 @@
+nfl <- read.csv("./nfl2023.csv")
+
+library(rjags)
+library(tidyverse)
+
+nfl <- nfl %>% filter(Winner.tie != "", Loser.tie!="") %>% select(Winner.tie, Loser.tie)
+
+teams <- sort(unique(c(nfl$Winner.tie,nfl$Loser.tie)))
+
+x <- matrix(0, nrow = nrow(nfl), ncol = length(teams))
+
+
+for (i in 1:50) {
+  x[i, which(teams == nfl$Winner.tie[i])] <- -1
+  x[i, which(teams == nfl$Loser.tie[i])] <- 1
+}
+
+for (i in 51:nrow(nfl)) {
+  x[i, which(teams == nfl$Winner.tie[i])] <- 1
+  x[i, which(teams == nfl$Loser.tie[i])] <- -1
+}
+
+y <- c(rep(0,50),rep(1,nrow(nfl)-50))
+a <- glm(y ~ 0 + x, family = "binomial")
+a$coefficients[32] <- 0
+data.frame(teams,beta = c(a$coefficients)) %>% arrange(beta)
+
+
+data <- list(
+  N = nrow(x), # Number of observations
+  J = ncol(x),        # Number of classes
+  Y=y,     # Response variable
+  X=x      #Predictor matrix Fixed effects
+  
+)
+
+#With predictors
+model_string <- "model{
+
+for (i in 1:N){
+  Y[i] ~ dbinom(p[i], 1)
+  logit(p[i]) = inprod(X[i,], theta[]) + eps
+}  
+
+
+for (j in 1:(J-1)){
+  theta[j] ~ dnorm(0, 0.001)
+}
+
+theta[J] <- 0
+
+eps ~ dnorm(0,tau)
+tau <- 1/sigma
+sigma ~ dnorm(0,0.001)T(0,10000)
+
+}"
+
+setwd("/Users/gregorymatthews/")
+write(model_string,"sporttau.bug")
+
+# Run JAGS
+jags <- jags.model("/Users/gregorymatthews/sporttau.bug",
+                   data = data,
+                   n.chains = 3, 
+                   n.adapt = 100
+)
+
+update(jags, 1000)
+
+# ztest <- jags.samples(jags,c('beta','p','b',"sigma2b"),1000, thin = 1)
+# str(ztest)
+
[…]

### 2025-02-27 mutual_information_curves.qmd
mutual information for tournaments

---
title: "Mutual Information Plots"
format: html
editor: visual
---

## Mutual Information Plots

From Wikipedia, weight mutual information formula:

$I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}$

```{r}
library(combinat)
library(tidyverse)

mut_info <- function(sims, num_teams, rank_weights = NULL){
  # Handle missing weights (default to equal weights if not provided)
  if (is.null(rank_weights)){
    rank_weights <- rep(1,num_teams)
  }
  # Extract y values from sims, assuming `rank_hat` contains a list of rankings per simulation
  y_df <- sims %>%
    group_by(simulation) %>%
    summarise(rank_hat = list(rank_hat)) %>%  # Ensure rank_hat is a list column
    mutate(rank_hat = map(rank_hat, unlist)) %>%  # Unlist rank_hat properly
    unnest_wider(rank_hat, names_sep = "_")  # Expand into separate columns

  # Rename y columns
  colnames(y_df)[-1] <- paste0("y", 1:num_teams)

  # Create x matrix, representing positions 1:num_teams
  x_matrix <- matrix(rep(1:num_teams, times = nrow(y_df)), ncol = num_teams, byrow = TRUE)
  x_df <- as.data.frame(x_matrix)
  colnames(x_df) <- paste0("x", 1:num_teams)

  # Combine x and y data frames
  perms_df <- cbind(x_df, y_df %>% select(-simulation))
  perms_df <- perms_df %>%
    mutate(across(starts_with("x"), as.numeric)) %>%
    mutate(across(starts_with("y"), as.numeric))

  # Generate all possible permutations of rankings
  all_perms <- as.data.frame(do.call(rbind, permn(1:num_teams)))
  colnames(all_perms) <- paste0("y", 1:num_teams)

  # Count occurrences of each observed ranking
  prob_df <- perms_df %>%
    group_by(across(starts_with("x")),across(starts_with("y")))  %>%
    summarise(count = n(), .groups = "drop")

  # Merge with full permutation set to ensure all permutations exist
  prob_df <- right_join(all_perms, prob_df, by = colnames(all_perms)) %>%
    mutate(count = replace_na(count, 0)) %>%
    mutate(probability = count / sum(count))  # Normalize probabilities

    w <- numeric(nrow(prob_df))  # Initialize weight vector for the number of rows in prob_df

  for (i in 1:num_teams) {
    x_i <- paste0("x", i)
    y_i <- paste0("y", i)

    # Loop through each row of prob_df
    for (j in 1:nrow(prob_df)) {
      # Calculate the squared difference between x_i and y_i for each row j
      diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
    w[j] <- w[j] + rank_weights[i]* diff^2  # Sum squared differences for each row
    }
  }
  w <- ifelse(w == 0.0000, 1, 1 / w)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
  w <- w/sum(w)
  # Compute weights dynamically for each row
  weighted_df <- prob_df %>%
    mutate(w = w)

  mutual_info <- factorial(num_teams) * sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/nrow(weighted_df) * weighted_df$probability)))

  # Return the weighted data frame with mutual information weights
  return(mutual_info)
}
mut_info(all_results_norm,8, c(10,5,1,0,0,0,0,0))
mut_info(de_norm_results_eight,8, c(10,5,1,0,0,0,0,0))
mut_info(test_rr_noTies,8, c(10,5,1,0,0,0,0,0))
mut_info(seven_games_se_results,8, c(10,5,1,0,0,0,0,0))

de_norm_results_eight <- de_norm_results %>% filter(teams == 8)
de_norm_results_eight$simulation <- rep(1:10000, each = nrow(de_norm_results_eight) / 10000)
all_results_norm$simulation <- rep(1:10000, each = nrow(all_results_norm) / 10000)
seven_games_se_results$simulation <- rep(1:10000, each = nrow(seven_games_se_results) / 10000)
# with randomly assigning ties
test_rr_noTies <- test_rr %>%
  group_by(simulation) %>%
  mutate(rank_hat = rank(rank_hat,ties.method="random"))

# Initialize variables
num_vars <- 8  # Total length of the vector
indices <- 1:num_vars  # Store the number of 1's in each step

# Function to compute mutual information for a given dataset
compute_mut_plots <- function(dataset) {
  results <- numeric(num_vars)
  for (i in 1:num_vars) {
    binary_vector <- c(rep(1, i), rep(0, num_vars - i))  # Construct vector
    results[i] <- mut_info(dataset, num_vars, binary_vector)  # Compute MI
  }
  return(results)
}

# Compute mutual information for each dataset
mi_de_norm <- compute_mut_plots(de_norm_results_eight)
mi_se_norm <- compute_mut_plots(all_results_norm)
mi_test_rr <- compute_mut_plots(test_rr_noTies)
mi_seven_se <- compute_mut_plots(seven_games_se_results)

# Plot results
plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm, mi_test_rr, mi_seven_se)),
     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Mutual Information vs. Number of 1's")

# Add other datasets
lines(indices, mi_se_norm, type="b", pch=19, col="red")
lines(indices, mi_test_rr, type="b", pch=19, col="green")
lines(indices, mi_seven_se, type = "b", pch=19, col="black")

# Add legend
legend("topleft", legend=c("Double Elimination", "Single Elimination", "Round Robin", "7 Game Single Elimination"), 
       col=c("blue", "red", "green", "black"), pch=19, lty=1)

### Single Elimination vs Double Elimination ###

# Plot results
plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm)),
     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Single Elimination vs Double Elimination")

# Add other datasets
lines(indices, mi_se_norm, type="b", pch=19, col="red")

# Add legend
legend("bottomright", legend=c("Double Elimination", "Single Elimination"), 
       col=c("blue", "red"), pch=19, lty=1)

```

### 2025-04-21 mutual_information_curves.qmd

@@
   }
-  w <- ifelse(w == 0.0000, 1, 1 / w)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
+  w <- 1 / (w + 1)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
   w <- w/sum(w)
@@
 
-  mutual_info <- factorial(num_teams) * sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/nrow(weighted_df) * weighted_df$probability)))
+  mutual_info <- factorial(num_teams) * sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/factorial(num_teams) * weighted_df$probability)))
 
@@
 mut_info(seven_games_se_results,8, c(10,5,1,0,0,0,0,0))
+mut_info(perms_df,8,c(1,0,0,0,0,0,0,0))
 
@@
 de_norm_results_eight$simulation <- rep(1:10000, each = nrow(de_norm_results_eight) / 10000)
+test_unif$simulation <- rep(1:10000, each = nrow(test_unif) / 10000)
+equal_strengths$simulation <- rep(1:10000, each = nrow(equal_strengths) / 10000)
 all_results_norm$simulation <- rep(1:10000, each = nrow(all_results_norm) / 10000)
+bad_seeding_structure$simulation <- rep(1:10000, each = nrow(bad_seeding_structure) / 10000)
 seven_games_se_results$simulation <- rep(1:10000, each = nrow(seven_games_se_results) / 10000)
@@
 
+perms_list <- permn(8)
+all_data <- list()
+# Loop through each permutation and create the data frame
+for (i in 1:length(perms_list)) {
+  # Create a temporary data frame for the current permutation
+  temp_df <- data.frame(
+    true_rank = rep(1:8, each = 1),  # Each row gets a true_rank from 1 to 8
+    rank_hat = perms_list[[i]],      # The current permutation's ranks
+    simulation = rep(i, 8)           # Assign the simulation number to each row
+  )
+  # Add the temporary data frame to the list
+  all_data[[i]] <- temp_df
+}
+# Combine all the data frames into one
+perms_df <- do.call(rbind, all_data)
+
 # Compute mutual information for each dataset
@@
 mi_seven_se <- compute_mut_plots(seven_games_se_results)
+mi_random <- compute_mut_plots(equal_strengths)
+mi_se_unif <- compute_mut_plots(test_unif)
+mi_bad <- compute_mut_plots(bad_seeding_structure)
+mi_perms <- compute_mut_plots(perms_df)
 
 # Plot results
-plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm, mi_test_rr, mi_seven_se)),
+plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm, mi_test_rr, mi_seven_se, mi_random, mi_se_unif)),
      xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Mutual Information vs. Number of 1's")
@@
 lines(indices, mi_seven_se, type = "b", pch=19, col="black")
+lines(indices, mi_random, type="b", pch=19, col="purple")
+lines(indices, mi_se_unif, type="b", pch=19, col="orange")
 
 # Add legend
-legend("topleft", legend=c("Double Elimination", "Single Elimination", "Round Robin", "7 Game Single Elimination"), 
-       col=c("blue", "red", "green", "black"), pch=19, lty=1)
+legend("topleft", legend=c("Double Elimination", "Single Elimination", "Round Robin", "7 Game Single Elimination", "Random Single Elimination", "Uniform Single Elimination"), 
+       col=c("blue", "red", "green", "black", "purple", "orange"), pch=19, lty=1)
 
-### Single Elimination vs Double Elimination ###
-
-# Plot results
-plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm)),
-     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Single Elimination vs Double Elimination")
-
-# Add other datasets
-lines(indices, mi_se_norm, type="b", pch=19, col="red")
-
-# Add legend
-legend("bottomright", legend=c("Double Elimination", "Single Elimination"), 
-       col=c("blue", "red"), pch=19, lty=1)
-
+mi_plots_df <- data.frame(
+  indices = indices,
+  DE = mi_de_norm,
+  Norm_SE = mi_se_norm,
+  Random_SE = mi_random,
+  Unif_SE = mi_se_unif,
+  Bad_Seeds_SE = mi_bad,
+  Perms = mi_perms
+)
+# Reshape data into long format for ggplot
+df_long <- pivot_longer(mi_plots_df, cols = -indices, names_to = "Method", values_to = "MI")
+ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
+  geom_line() +           # Add lines
+  geom_point() +          # Add points
+  labs(
+    title = "Single Elimination vs Double Elimination",
+    x = "Number of 1's in the vector",
+    y = "Mutual Information Bits"
+  ) +
+  theme_minimal() +       # Use a clean theme
+  theme(legend.position = "right")
 ```

### 2025-04-21 tournament analysis.qmd

---
title: "Statistical Analysis of Tournament Structures"
author: "Greg, Josie, Zach, Ryan"
format: pdf
editor: visual
---

### Introduction

  There are hundreds of different tournament structures that dictate how a set of teams compete against each other to determine an overall ranking. Different tournaments have unique strengths and weaknesses, affecting how well they reflect the true rankings of the teams. The goal of a tournament is to find the best teams, but how can we quantify how well a tournament performs? In some instances, the tournament organizers only award the overall winner, but other times, the top three or ten teams are awarded. Because of this, organizers may prefer one structure over another based on their needs. Tournament organizers must also take into account factors like cost, timeliness, entertainment value, and fairness when choosing a tournament structure.

	To evaluate a tournament’s effectiveness, we propose a numerical metric that quantifies how accurately it orders teams based on their true rankings. Tournament results can be viewed as a “message” attempting to convey the true team rankings. However, various factors introduce noise, leading to information loss. We use principles from information theory to measure this information loss across multiple tournament simulations to assess the reliability of different formats.

### 2025-04-25 manuscript/paper.Rmd
paper.

---
title: |
  |  \large Tournaments 
author: | 
  | \large Zach Culp \vspace{-1.1mm}
  | \normalsize  \vspace{-1mm}
  | \normalsize Loyola Chicago \vspace{-1mm}
  | \normalsize Chicago \vspace{-1mm}
  | \normalsize [`email`](mailto:ypu@something.edu) \vspace{-1mm}
  |
  | \large Josie Peterburgs \vspace{-1.1mm}
  | \normalsize College of Education \vspace{-1mm}
  | \normalsize University of Texas at Austin \vspace{-1mm}
  | \normalsize Austin, TX 78712 \vspace{-1mm}
  | \normalsize [`brian.mills@austin.utexas.edu`](mailto:brian.mills@austin.utexas.edu) \vspace{-1mm}
  |
  | \large Ryan McShane \vspace{-1.1mm}
  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
  | \normalsize Loyola University Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  | \normalsize [`lderango@luc.edu`](mailto:lderango@luc.edu) \vspace{-1mm}
  |
  | \large Gregory J. Matthews \vspace{-1.1mm}
  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
  | \normalsize Center for Data Science and Consulting \vspace{-1mm}
  | \normalsize Loyola University Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm}
abstract: | 
 | Here is the abstract.
 | 
  \vspace{2mm}
  | Keywords: survival analysis, hockey
bibliography: references.bib
fontsize: 12pt
link-citations: true
linkcolor: cyan
urlcolor: cyan
output:
  pdf_document:
    df_print: kable
    number_sections: true
    keep_tex: true
header-includes:
 \usepackage{setspace}
 \setstretch{1.15}
 \usepackage{float}
 \floatplacement{figure}{t}
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)

gjm <- function(x, color = "red") {
  sprintf("\\textcolor{%s}{%s}", color, x)}
```

```{r pkgs}
library(tidyverse)
library(dplyr)
library(stringr)
theme_set(theme_minimal())
```

# Introduction
There are hundreds of different tournament structures that dictate how a set of teams compete against each other to determine an overall ranking. Different tournament structures have unique strengths and weaknesses, affecting how well they reflect the true rankings of the teams. The goal of a tournament is to find the best teams, but how can we quantify how well a tournament performs? In some instances, the tournament organizers only award the overall winner, but other times, the top three or ten teams are awarded. Because of this, organizers may prefer one structure over another based on their needs. Tournament organizers must also take into account factors like cost, timeliness, entertainment value, and fairness when choosing a tournament structure.

To evaluate a tournament’s effectiveness, we propose a numerical metric that quantifies how accurately it orders teams based on their true rankings. Tournament results can be viewed as a "message" attempting to convey the true team rankings. However, various factors introduce noise, leading to information loss. We use principles from information theory to measure this information loss across multiple tournament simulations to assess the reliability of different formats.

# Methods
Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 > \theta_2 > \ldots > \theta_n)$.  Next, let $r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when theta is sorted from largest to smallest so that $r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}$.  Next define $T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure $s$.  $\phi$ contains all the information about the scheduling of teams, which could be fully known prior to the tournament (i.e. round robin) or determined as the tournament progresses (i.e. single elimination tournament).  We then let  $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an $n$- dimensional vector valued random variable that gives the results of a tournament as a vector of the indexes of the vector $\boldsymbol{\theta}$.  The index in the first position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished first, The index in the second position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished second, and so on.

Examples: 
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (1, 2, 3, 4) This means that the tournament got the order correct.

$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) This means that the third best team according to thetas won the tournament, second best team finished second, the best team finished third and the worst team finished 4th.

# ryan way 
Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 < \theta_2 < \ldots < \theta_n)$.  Next, let $r(\boldsymbol{\theta})$ be the rank of the vector $\boldsymbol{\theta}$ from largest to smallest so that $r(\boldsymbol{\theta}) = \{n, n-1, \ldots, 2, 1\}$.  Next define $T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure $s$.  $\phi$ contains all the information about the scheduling of teams, which could be fully known prior to the tournament (i.e. round robin) or determined as the tournament progresses (i.e. single elimination tournament).  We then let  $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be a vector valued random variable that gives the results of a tournament as a vector of ranks.

Figure out what $\hat{r}$ should look like.

$r = \{1,2,3,4\}$ which is the

Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n) $ be the permutation of indexes of $\boldsymbol{\theta}$ such that $x_1 = {i | \theta_i = max(\boldstyle{\theta})}$

Let $x$ be

x is the true ranking and y is the output

$$
I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
$$

Rajski's distance: 
$$
1 - \frac{I(X,Y)}{H(X,Y)}
$$

# Results

Zach, put your stuff here.

# Conclusion {#sec:conclusion}

# Acknowledgements {-}

# Supplementary Material {-}

# References

### 2025-04-25 manuscript/paper.tex
paper.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace} \setstretch{1.15} \usepackage{float} \floatplacement{figure}{t}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  colorlinks=true,
  linkcolor={cyan},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={cyan},
  pdfcreator={LaTeX via pandoc}}

\title{~\large Tournaments}
\author{\large Zach Culp \vspace{-1.1mm}\\
\normalsize Epi \vspace{-1mm}\\
\normalsize Oxford something at Emory \vspace{-1mm}\\
\normalsize Georgia? \vspace{-1mm}\\
\normalsize \href{mailto:ypu@something.edu}{\texttt{email}}
\vspace{-1mm}\\
\strut \\
\large Josie Peterburgs \vspace{-1.1mm}\\
\normalsize College of Education \vspace{-1mm}\\
\normalsize University of Texas at Austin \vspace{-1mm}\\
\normalsize Austin, TX 78712 \vspace{-1mm}\\
\normalsize \href{mailto:brian.mills@austin.utexas.edu}{\texttt{brian.mills@austin.utexas.edu}}
\vspace{-1mm}\\
\strut \\
\large Ryan McShane \vspace{-1.1mm}\\
\normalsize Department of Mathematics and Statistics \vspace{-1mm}\\
\normalsize Loyola University Chicago \vspace{-1mm}\\
\normalsize Chicago, IL 60660 \vspace{-1mm}\\
\normalsize \href{mailto:lderango@luc.edu}{\texttt{lderango@luc.edu}}
\vspace{-1mm}\\
\strut \\
\large Gregory J. Matthews \vspace{-1.1mm}\\
\normalsize Department of Mathematics and Statistics \vspace{-1mm}\\
\normalsize Center for Data Science and Consulting \vspace{-1mm}\\
\normalsize Loyola University Chicago \vspace{-1mm}\\
\normalsize Chicago, IL 60660 \vspace{-1mm}\\
\normalsize \href{mailto:gmatthews1@luc.edu}{\texttt{gmatthews1@luc.edu}}
\vspace{-1mm}}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Here is the abstract.\\

\vspace{2mm} \textbar{} Keywords: survival analysis, hockey
\end{abstract}

\section{Introduction}\label{introduction}

Popkin et al. (\citeproc{ref-Popkin2023}{2023})

There are hundreds of different tournament structures that dictate how a
set of teams compete against each other to determine an overall ranking.
Different tournaments have unique strengths and weaknesses, affecting
how well they reflect the true rankings of the teams. The goal of a
tournament is to find the best teams, but how can we quantify how well a
tournament performs? In some instances, the tournament organizers only
award the overall winner, but other times, the top three or ten teams
are awarded. Because of this, organizers may prefer one structure over
another based on their needs. Tournament organizers must also take into
account factors like cost, timeliness, entertainment value, and fairness
when choosing a tournament structure.

\begin{verbatim}
To evaluate a tournament’s effectiveness, we propose a numerical metric that quantifies how accurately it orders teams based on their true rankings. Tournament results can be viewed as a “message” attempting to convey the true team rankings. However, various factors introduce noise, leading to information loss. We use principles from information theory to measure this information loss across multiple tournament simulations to assess the reliability of different formats.
\end{verbatim}

\section{Methods}\label{methods}

\section{Conclusion}\label{sec:conclusion}

\section*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

\section*{Supplementary Material}\label{supplementary-material}
\addcontentsline{toc}{section}{Supplementary Material}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Popkin2023}
Popkin, Charles A., Cole R. Morrissette, Thomas A. Fortney, Kyle L.
McCormick, Prakash Gorroochurn, and Michael J. Stuart. 2023. {``Fighting
and Penalty Minutes Associated with Long-Term Mortality Among National
Hockey League Players, 1967 to 2022.''} \emph{JAMA Network Open} 6 (5):
e2311308--8.

\end{CSLReferences}

\end{document}

### 2025-04-25 tournament analysis.qmd
paper.

@@
 
+```         
 To evaluate a tournament’s effectiveness, we propose a numerical metric that quantifies how accurately it orders teams based on their true rankings.
@@
 We use principles from information theory to measure this information loss across multiple tournament simulations to assess the reliability of different formats.
+```

### 2025-04-29 manuscript/paper.Rmd
updated manuscript

@@
 We then let  $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an $n$- dimensional vector valued random variable that gives the results of a tournament as a vector of the indexes of the vector $\boldsymbol{\theta}$.
-The index in the first position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished first, The index in the second position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished second, and so on.
+The index in the first position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished first, the index in the second position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished second, and so on.
 
-Examples: 
-$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (1, 2, 3, 4) This means that the tournament got the order correct.
+For example, in a 4 team tournament, $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that the team with index 3 (i.e.
+the true third best team) won the tournament, the true second best team finished second, the true best team finished third and the true 4th team finished 4th in that particular tournament.
+If $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (1, 2, 3, 4), this means that the random outcome of the tournament was the same as the true ordering of the teams.
 
-$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) This means that the third best team according to thetas won the tournament, second best team finished second, the best team finished third and the worst team finished 4th.
+If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the "goodness" of a tournament in terms of its ability to accurately transmit the true ranking.
+We can then use concepts from information theory to assess the ability of a tournament to correctly rank teams.
+Specifically, we start with the concept of mutual information (CITE).
+For two random variables $X$ and $Y$ mutual information is defined as:
+
+$$
+I(X;Y)=\sum_{y\in Y}\sum_{x\in X}p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
+$$.
+
+In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and $\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate $I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$.
+We note here that $r(\boldsymbol{\theta})$ is not a random variable, however, since the indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary (we use 1, 2, 3, etc.
+for convenience, but any indexing is valid), we can view this a random variable where all permutation so of the $n$ teams are equally likely.
+Therefore when calculating  the mutual information in this setting, we set $p(r(\boldsymbol{\theta})) = \frac{1}{n!}$.
+By a similar argument we set $\hat{r}(T(\phi, s),\boldsymbol{\theta}) = \frac{1}{n!}$.
+This
+
+Note that one is 1/N!
+
+And this does work because consistently getting the order incorrect is perfect information.
+
+$$
+I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
+$$
+
+Let $x$ be
+
+x is the true ranking and y is the output
+
+Rajski's distance: 
+$$
+1 - \frac{I(X,Y)}{H(X,Y)}
+$$
 

### 2025-04-29 manuscript/paper.Rmd
paper updated.

@@
 In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and $\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate $I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$.
-We note here that $r(\boldsymbol{\theta})$ is not a random variable, however, since the indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary (we use 1, 2, 3, etc.
+For simplicity, we drop the arguments from the functions $r$ and $\hat{r}$ for ease of exposition.
+
+We note here that $r$ is not a random variable, however, since the indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary (we use 1, 2, 3, etc.
 for convenience, but any indexing is valid), we can view this a random variable where all permutation so of the $n$ teams are equally likely.
-Therefore when calculating  the mutual information in this setting, we set $p(r(\boldsymbol{\theta})) = \frac{1}{n!}$.
-By a similar argument we set $\hat{r}(T(\phi, s),\boldsymbol{\theta}) = \frac{1}{n!}$.
-This
+Therefore when calculating  the mutual information in this setting, we set $p(r) = \frac{1}{n!}$.
+By a similar argument we set $p(\hat{r}) = \frac{1}{n!}$.
+
+We define the mutual information of $r$ and $\hat{r}$ to be:
+
+$$
+I(r,\hat{r})=\sum_{r}\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})} = n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{(\frac{1}{n!})^2}
+$$.
+
+While there are $n!$ different permutations for the result of $r$, we don't need to sum across these as any specific choice of $r$ is just an arbitrary labeling of the true strength parameters vector $\boldsymbol{\theta}$.
+So given $r$, the distribution of $\hat{r}$ is the same up to the labeling.
+Therefore, we only need to consider a single permutation of $r$, we compute the the quantity inside the summation, and then multiple by $n!$ (effectively summing across $r$).
+
+In order to compute this quantity, we need to compute $p(r,\hat{r})$.
+This probability is found by calculating the probability of a given permutation of $\hat{r}$ and $r$ is assumed to be the permutation from $\{1, 2, \ldots, n\}$.
+We estimate $p(r,\hat{r})$ empirically through simulation.
+
+However, using mutual information in this form does not suit our needs in this setting.
+The problem is that mutual information in this form will yield high values of mutual information when the output from the tournament $\hat{r}$ is highly consistent *even if the ranking from the tournament is incorrect*.
+As an example, if $n=4$ and the true order of \boldsymbol{\theta} is $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in both of these cases the mutual information will be maximized at:
+
+$$
+I(r,\hat{r})= n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{\left(\frac{1}{n!}\right)^2} = n! log((n!)^2)
+$$
+and for the specific case when $n=4$ would be: 
+$$
+4! log_2(4!^2) = 24*log_2(24^2) = 220.0782
+$$.
+
+In order to alleviate this problem, we instead consider *weighted* mutual information.
+We want to give more weight to permutations from $\hat{r}$ that are "closer" to $r$.
+There are any number of reasonable choices for this weighting function, and we choose to use squared error.
+Specifically,
+$$
+w(r, \hat{r}) =
+
+\left\{  
+\begin{array}{cc}
+    \begin{array}{cc}
+      \frac{1}{\hat{r}'r} & r \ne \hat{r} \\
+      1 & r = \hat{r} \\
+    \end{array}
+\end{array}
+\right. 
+$$
+In general, though any loss function $l(r, \hat{r})$ can be used:
+$$
+w(r, \hat{r}) =
+
+\left\{  
+\begin{array}{cc}
+    \begin{array}{cc}
+      \frac{1}{l(r, \hat{r})} & r \ne \hat{r} \\
+      1 & r = \hat{r} \\
+    \end{array}
+\end{array}
+\right. 
+$$
 
@@
 
-# ryan way 
-Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 < \theta_2 < \ldots < \theta_n)$.
+<!-- # ryan way  -->
+<!-- Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 < \theta_2 < \ldots < \theta_n)$.
 Next, let $r(\boldsymbol{\theta})$ be the rank of the vector $\boldsymbol{\theta}$ from largest to smallest so that $r(\boldsymbol{\theta}) = \{n, n-1, \ldots, 2, 1\}$.
@@
 We then let  $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be a vector valued random variable that gives the results of a tournament as a vector of ranks.
+-->
 
-Figure out what $\hat{r}$ should look like.
+<!-- Figure out what $\hat{r}$ should look like.   -->
 
@@
 
-Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n) $ be the permutation of indexes of $\boldsymbol{\theta}$ such that $x_1 = {i | \theta_i = max(\boldstyle{\theta})}$
+Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n)$ be the permutation of indexes of $\boldsymbol{\theta}$ such that $x_1 = {i | \theta_i = max(\boldsymbol{\theta})}$
 

### 2025-05-01 double_elimination.qmd
update 5/1/25

@@
           match_winner <- simulate_match(team1, team2, strength1, strength2)
-          #cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
+          cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
           if(!is.na(team1)) {
@@
         if (!any(is.na(c(team1, team2)))) {
-          #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
+          cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
         permuted_df <- permuted_df %>%
@@
         series_loser <- ifelse(series_winner == team1, team2, team1)
-        #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n: | Loser: %s\n", team1, team2, series_winner, series_loser))
+        cat(sprintf("Matchup:  %s vs %s | Winner: %s\n: | Loser: %s\n", team1, team2, series_winner, series_loser))
         next_winners <- c(next_winners, series_winner)
@@
       if(length(losers_bracket) > 1){
-      #cat("\n--- Loser's Bracket --- \n")
+      cat("\n--- Loser's Bracket --- \n")
       }
@@
         if (!any(is.na(c(team1, team2)))) {
-          #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
+          cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
           permuted_df <- permuted_df %>%
@@
         eliminated <- c(eliminated, series_loser)
-        #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
+        cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
         }
@@
 # Final matchup between winners and losers
-    #cat("\n--- Final Match ---\n")
+    cat("\n--- Final Match ---\n")
     final_match <- c(winners_bracket, losers_bracket)
@@
     if (!any(is.na(c(team1, team2)))) {
-      #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
+      cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
       permuted_df <- permuted_df %>%
@@
     if (wins_team1 > wins_team2) {
-      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
+      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
       permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
@@
     } else {
-      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
+      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
       # second game if the loser's bracket team wins
@@
       if (!any(is.na(c(team1, team2)))) {
-        #cat(sprintf("Wins - %s: %d | %s: %d\n", team1, wins_team1, team2, wins_team2))
+        cat(sprintf("Wins - %s: %d | %s: %d\n", team1, wins_team1, team2, wins_team2))
         permuted_df <- permuted_df %>%
@@
       if (wins_team1 > wins_team2) {
-      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
+      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
       permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
@@
     } else {
-      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
+      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
       permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
@@
       arrange(true_rank) %>%
-      mutate(rank_hat = rank(rank_hat, ties.method = "first"))
+      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
   }
@@
 
-#test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1)
+test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1)
 #test

### 2025-05-01 group_stage.qmd
update 5/1/25

--- 2025-02-06 single_elimination_shiny.qmd
@@
 ---
-title: "single_elimination"
-format: 
-  html:
-    embed-resources: true
+title: "Group Stage Play"
+author: "Zach Culp"
+format: html
 editor: visual
@@
 
-```{r}
-library(tidyverse)
-library(dplyr)
-library(shiny)
-```
+## Group Stage
 
-### Things To Do
-
--   Add a re-seeded option
-
--   Custom strengths (likely as an input to the function)
-
-# Simulate a Single Match
+### Simulate Game
 
 ```{r}
-# Function to simulate a match using Bradley-Terry model
 simulate_match <- function(team1, team2, strength1, strength2) {
+  if(is.na(team1) && is.na(team2)){
+    return(NA)
+  }
   if (is.na(team1)) {
@@
 }
-
 ```
 
-# Single Elimination Tournament Structures
-
-## Creating the traditional tournament structure for n teams
+### Simulate Each Group
 
 ```{r}
-generate_elimination_tournament_structure <- function(n) {
+simulate_group_play <- function(strengths, rounds=1){
+  teams <- strengths$true_rank
+  n <- length(teams)
+  win_counts <- numeric(n)
+  loss_counts <- numeric(n)
+  for (i in 1:rounds){
+    for (i in seq_along(teams)){
+      for (j in seq_along(teams)){
+        if (i != j){
+          team1 <- teams[i]
+          team2 <- teams[j]
+          strength1 <- strengths$true_strength[strengths$true_rank == team1]
+          strength2 <- strengths$true_strength[strengths$true_rank == team2]
+          match_winner <- simulate_match(teams[i],teams[j],strength1, strength2)
+          if (match_winner == teams[i]) {
+            win_counts[i] <- win_counts[i] + 1
+            loss_counts[j] <- loss_counts[j] + 1
+          } else {
+            win_counts[j] <- win_counts[j] + 1
+            loss_counts[i] <- loss_counts[i] + 1
+          }
+        }
+      }
+    }
+  }
+  results <- data.frame(
+    true_rank = teams,
+    game_wins = win_counts,
+    game_loses = loss_counts,
+    true_strength = strengths$true_strength,
+    groups = strengths$groups
+  )
 
-  is_power_of_two <- function(x) {
-    x > 0 && (x & (x - 1)) == 0
+  # Ranks teams based on number of wins (includes ties)
+  results$rank_hat <- rank(-results$game_wins, ties.method = "random")
+  results <- results[order(results$rank_hat),]
+
+  return(results)
 }
 
-  # If n is not a power of 2, round up to the next power of 2 and add placeholders
-  if (!is_power_of_two(n)) {
-    next_power_of_two <- 2^ceiling(log2(n))
-    extra_teams <- next_power_of_two - n
-    #message(sprintf("Rounding up to %d teams. Adding %d NA placeholders.", next_power_of_two, extra_teams))
-    n <- next_power_of_two
-  } else {
-    extra_teams <- 0
-  }
-
-  rounds <- log(n, base = 2) - 1
-  teams <- c(1, 2)  # Initial list of teams
-
-  for (i in 1:rounds) {
-    teams <- nextLayer(teams)
-  }
-
-  left_half <- teams[1:(length(teams)/2)]
-  right_half <- teams[(length(teams)/2+1):length(teams)]
-  right_half <- rev(right_half)
-  teams <- c(left_half,right_half)
-
-  # Ensure that the lower number team is on the left side of each matchup
-  teams <- unlist(lapply(seq(1, length(teams), by = 2), function(i) {
-    if (teams[i] > teams[i + 1]) {
-      c(teams[i + 1], teams[i])  # Swap if needed
-    } else {
-      c(teams[i], teams[i + 1])
-    }
-  })
-  )
-
-  while (extra_teams > 0){
-      teams[which.max(teams)] <- NA
-      extra_teams <- extra_teams - 1
-  }
-
-  return(teams)
-}
-
-nextLayer <- function(teams) {
-  out <- c()  # Initialize an empty vector
-  length <- length(teams) * 2 + 1  # Calculate the length for pairing
-
-  # Generate the next layer by pairing teams
-  for (i in teams) {
-    out <- c(out, i, length - i)  # Push the current team and its pair
-  }
-
-  return(out)
-}
-generate_elimination_tournament_structure(8)
-generate_elimination_tournament_structure(16)
-
-```
-
-# Simulating a Tournament
-
-```{r}
-# Simulate the single-elimination tournament
-simulate_single_elimination_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1, third_place = T) {
-
-  if (num_teams <= 3) {
-    stop("Number of Teams must be greater than 3.")
-  }
+simulate_groups <- function(group_length, num_teams, distribution, num_advance){
   teams <- paste(num_teams:1)
-  if (log2(num_teams) %% 1 != 0) {
-    power_of_2 <- 2^ceiling(log2(num_teams))
-    teams <- c(teams, rep(NA, power_of_2 - num_teams))
-  }
   if (distribution == "Normal"){
     strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
+    normal_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      rank_hat = rep(NA,length(teams)),
+      game_wins = rep(0,length(teams)),
+      game_losses = rep(0,length(teams))
+    )
+    df <- arrange(normal_strengths, true_rank)
+  }
+  else if (distribution == "Same"){
+    strengths <- sapply(num_teams, function(n) {rep(0,n)})
     normal_strengths <- data.frame(
@@
   }
-  else if (distribution == "Manual") {
-    cat("Please enter the strengths for each team in descending rank order:\n")
-    strengths <- numeric(num_teams)
-    for (i in num_teams:1) {
-      repeat {
-        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
-        if (!is.na(input)) {
-          strengths[i] <- input
-          break
-        } else {
-          cat("Invalid input. Please enter a numeric value.\n")
-        }
-      }
-    }
-    manual_strengths <- data.frame(
-      true_rank = as.numeric(teams),
-      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
-      rank_hat = rep(NA, length(teams)),
-      game_wins = rep(0, length(teams)),
-      game_losses = rep(0, length(teams))
-    )
-    df <- arrange(manual_strengths, true_rank)
-  }
-  else {
-    stop("Distribution not found: Enter Manual to input your own strengths")
+
+  num_groups <- num_teams / group_length
+  # Create tiers based on sorted true_strength
+  df <- df %>% arrange(desc(true_strength))
+  tiers <- split(df, ceiling(seq_along(df$true_rank) / num_groups))
+
+  # Initialize group vector
+  df$groups <- NA
+
+  # Assign one team from each tier to each group randomly
+  for (tier in tiers) {
+    group_ids <- sample(1:num_groups)  # random group order for this tier
+    df$groups[match(tier$true_rank, df$true_rank)] <- group_ids
   }
 
-  # Sets seed to that of common best vs worst structure
-  if (is.null(seeding_structure)){
-    seeding_structure <- matrix(generate_elimination_tournament_structure(num_teams), nrow = 1)
-  } else{
-    seeding_structure <- matrix(seeding_structure, nrow = 1)
+  all_results <- list()
+  for (g in unique(df$groups)){
+    group_df <- df[df$groups == g, ]
+    group_results <- simulate_group_play(strengths = group_df)
+    group_results$groups <- g
+    all_results[[g]] <- group_results
   }
-  if(series %% 2 == 0) {
-    stop("Series must be an odd number to avoid potential ties.")
+  print(all_results)
+  final_df <- do.call(rbind, all_results)
+  teams_in_knockout <- final_df %>%
+    arrange(rank_hat, game_wins) %>%
+    head(num_advance * num_groups)
+  return(teams_in_knockout)
+}
+create_knockout_matches <- function(df) {
+  group_list <- sort(unique(df$groups))
+  matches <- list()
+
+  for (i in seq(1, length(group_list), by = 2)) {
+    g1 <- group_list[i]
+    g2 <- group_list[i + 1]
+
+    # Get the winner from group g1 and runner-up from group g2
+    team1 <- df[df$groups == g1 & df$rank_hat == 1, ]
+    team2 <- df[df$groups == g2 & df$rank_hat == 2, ]
+
+    # Get the winner from group g2 and runner-up from group g1
+    team3 <- df[df$groups == g2 & df$rank_hat == 1, ]
+    team4 <- df[df$groups == g1 & df$rank_hat == 2, ]
+
+    matches[[length(matches) + 1]] <- data.frame(
+      Match = paste("Match", length(matches) + 1),
+      TeamA = team1$true_rank,
+      TeamB = team2$true_rank
+    )
+
+    matches[[length(matches) + 1]] <- data.frame(
+      Match = paste("Match", length(matches) + 1),
+      TeamA = team3$true_rank,
+      TeamB = team4$true_rank
+    )
+  }
+
+  do.call(rbind, matches)
+}
+simulate_knockout <- function(df, series=1, third_place = T){
+
+  seeding_structure <- create_knockout_matches(df)
+
+  for (i in nrow(df)){
+    for (g in unique(df$groups)){
+
+    }
   }
@@
       arrange(true_rank) %>%
-      mutate(rank_hat = rank(rank_hat, ties.method = "first"))
+      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
   } else{
@@
 
-test <- simulate_single_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, third_place=T)
-test
-
-#manual_test <- simulate_single_elimination_tournament(num_teams = 4, distribution = "Manual", ties=T, series=1, third_place=T)
-#manual_test
-
-# Gets all results of a normal distribution num_teams between 4 and 64 single-game elimination
-all_normal_results <- list()
-for (teams in c(4,8,16,32,64)) {
-  all_normal_results[[as.character(teams)]] <- list()
-  for (replicate in 1:10000) {
-    # Store each simulation result in the list with unique indexing
-    all_normal_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
-      num_teams = teams,
-      distribution = "Normal",
-      ties = FALSE,
-      series = 1,
-      third_place = TRUE
-    )
-  }
-}
-
-all_results_norm <- do.call(rbind, lapply(names(all_normal_results), function(teams) {
-  team_results <- do.call(rbind, all_normal_results[[teams]])
-  team_results$teams <- as.numeric(teams)
-  return(team_results)
-}))
-
-all_results_norm
-
-# Gets all results of a uniform distribution num_teams between 4 and 64 single-game elimination
-all_unif_results <- list()
-for (teams in c(4,8,16,32,64)) {
-  all_unif_results[[as.character(teams)]] <- list()
-  for (replicate in 1:100) {
-    # Store each simulation result in the list with unique indexing
-    all_unif_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
-      num_teams = teams,
-      distribution = "Uniform",
-      ties = FALSE,
-      series = 1,
-      third_place = TRUE
-    )
-  }
-}
-
-all_results_unif <- do.call(rbind, lapply(names(all_unif_results), function(teams) {
-  team_results <- do.call(rbind, all_unif_results[[teams]])
-  team_results$teams <- as.numeric(teams)
-  return(team_results)
-}))
-
-all_results_unif
-
+knockout_teams <- simulate_groups(4,8,"Normal",2)
+create_knockout_matches(knockout_teams)
 ```
-
-## Shiny User Interface Tournament Simulation
-
-```{r}
-# Define the Shiny app UI
-ui <- fluidPage(
-  titlePanel("Single-Elimination Tournament Simulator"),
-
-  sidebarLayout(
-    sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", value = 8, min = 4, step = 1),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
-      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
-      uiOutput("manual_strengths_ui"), # Dynamic UI for Manual strengths
-      checkboxInput("ties", "Allow Ties?", value = TRUE),
-      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
-      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
-      actionButton("run_sim", "Run Simulation")
-    ),
-
-    mainPanel(
-      h4("Tournament Results"),
-      tableOutput("ranked_results")
-    )
-  )
-)
-
-# Define the server logic
-server <- function(input, output) {
-  custom_seeding <- reactive({
-    if (input$seeding_input != "") {
-      # Convert the comma-separated string into a numeric vector
-      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
-      seeding_vector[seeding_vector == "NA"] <- NA
-      seeding_vector <- as.numeric(seeding_vector)
-      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
-      # Validate that it matches the number of teams
-      if (length(seeding_vector) == num_teams_next_power_of_two) {
-        return(seeding_vector)
-      } else {
-        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
-        return(NULL)
-      }
-    } else {
-      return(generate_elimination_tournament_structure(input$num_teams))
-    }
-  })
-
-  tournament_results <- eventReactive(input$run_sim, {
-    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
-
-    seed_structure <- custom_seeding()
-
-    simulate_single_elimination_tournament(
-      num_teams = input$num_teams,
-      distribution = input$distribution,
-      ties = input$ties,
-      series = input$series,
-      third_place = input$third_place,
-      seeding_structure = seed_structure
-    )
-  })
-
-  output$ranked_results <- renderTable({
-    results <- tournament_results()
-    results[order(results$true_rank), ]  # Order by Strength descending
-  })
-}
-
-shinyApp(ui = ui, server = server)
-```
-
-## Shiny Number of Wins
-
-```{r}
-# Shiny app UI
-ui <- fluidPage(
-  titlePanel("Tournament Simulation with Total Wins Visualization"),
-
-  sidebarLayout(
-    sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
-      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
-      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
-      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
-      checkboxInput("ties", "Allow Ties?", value = TRUE),
-      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
-      actionButton("run_sim", "Run Simulation")
-    ),
-
-    mainPanel(
-      h4("Total Wins Summary"),
-      plotOutput("wins_plot")
-    )
-  )
-)
-
-# Shiny server logic
-server <- function(input, output) {
-  custom_seeding <- reactive({
-    if (input$seeding_input != "") {
-      # Convert the comma-separated string into a numeric vector
-      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
-      seeding_vector[seeding_vector == "NA"] <- NA
-      seeding_vector <- as.numeric(seeding_vector)
-      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
-      # Validate that it matches the number of teams
-      if (length(seeding_vector) == num_teams_next_power_of_two) {
-        return(seeding_vector)
-      } else {
-        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
-        return(NULL)
-      }
-    } else {
-      return(generate_elimination_tournament_structure(input$num_teams))
-    }
-  })
-
-  simulation_results <- eventReactive(input$run_sim, {
-    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
-    seed_structure <- custom_seeding()
-    # Run simulations
-    replicate_results <- replicate(
-      input$num_simulations,
-      simulate_single_elimination_tournament(
-        num_teams = input$num_teams,
-        distribution = input$distribution,
-        ties = input$ties,
-        series = input$series,
-        third_place = input$third_place,
-        seeding_structure = seed_structure
-      ),
-      simplify = FALSE
-    )
-
-    # Combine all results and summarize total wins
-    all_results <- do.call(rbind, replicate_results)
-    total_wins_summary <- all_results %>%
-      group_by(true_rank) %>%
-      summarise(
-        total_wins = sum(rank_hat==1, na.rm = TRUE),
-        .groups = 'drop'
-      ) %>%
-      arrange(desc(total_wins))
-    return(total_wins_summary)
-  })
-
-  output$wins_plot <- renderPlot({
-    req(simulation_results())
-    results <- simulation_results()
-
-    ggplot(results, aes(x = reorder(true_rank, -total_wins), y = total_wins, fill = total_wins)) +
-      geom_text(aes(label = total_wins), vjust = -0.5, size = 4, color = "black") +
-      geom_bar(stat = "identity") +
-      labs(
-        title = "Total Wins Across Simulations",
-        x = "True Rank",
-        y = "Total Wins"
-      ) +
-      theme_minimal() +
-      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-      scale_fill_gradient(low = "skyblue", high = "blue")
-  })
-}
-
-# Run the Shiny app
-shinyApp(ui = ui, server = server)
-```
-
-# Multiple Simulations for 1 Seeding Structure
-
-```{r}
-ui <- fluidPage(
-  titlePanel("Tournament Simulation"),
-
-  sidebarLayout(
-    sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", 8, min = 4, step = 1),
-      selectInput("distribution", "Distribution Type:", choices = c("Normal", "Uniform")),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
-      numericInput("replicates", "Number of Replicates:", 100, min = 1, max = 1000, step = 1),
-      numericInput("series", "Number of Games in a Series (odd number):", 1, min = 1, max = 11, step = 2),
-      checkboxInput("ties", "Allow Ties?", value = TRUE),
-      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
-      actionButton("run_sim", "Run Tournament Simulation")
-    ),
-
-    mainPanel(
-      tableOutput("summary_results"),
-      verbatimTextOutput("kendall_correlation"),
-      verbatimTextOutput("spearman_correlation")
-    )
-  )
-)
-
-server <- function(input, output, session) {
- custom_seeding <- reactive({
-    if (input$seeding_input != "") {
-      # Convert the comma-separated string into a numeric vector
-      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
-      seeding_vector[seeding_vector == "NA"] <- NA
-      seeding_vector <- as.numeric(seeding_vector)
-      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
-      # Validate that it matches the number of teams
-      if (length(seeding_vector) == num_teams_next_power_of_two) {
-        return(seeding_vector)
-      } else {
-        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
-        return(NULL)
-      }
-    } else {
-      return(generate_elimination_tournament_structure(input$num_teams))
-    }
-  })
-
-  # Function to run tournament simulations
-  observeEvent(input$run_sim, {
-    req(input$num_teams, input$distribution, input$replicates, input$series)
-
-    # Initialize lists to store correlation results
-    kendall_correlations <- numeric(input$replicates)
-    spearman_correlations <- numeric(input$replicates)
-
-    seed_structure <- custom_seeding()
-
-    # Run tournament simulations for the specified number of replicates
-    simulation_results <- list()
-    for (replicate in 1:input$replicates) {
-      simulation_results[[replicate]] <- simulate_single_elimination_tournament(
-        num_teams = input$num_teams,
-        distribution = input$distribution,
-        ties = input$ties,
-        series = input$series,
-        third_place = input$third_place,
-        seeding_structure = seed_structure
-      )
-
-      # Combine all results into one data frame
-      all_results <- do.call(rbind, simulation_results)
-
-      # Compute average ranks across all replicates
-      average_ranks <- all_results %>%
-        group_by(true_rank) %>%
-        summarise(avg_rank = mean(rank_hat, na.rm = TRUE), .groups = 'drop') %>%
-        arrange(true_rank)
-
-      # Calculate Kendall Tau correlation for the current replicate
-      kendall_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "kendall", use = "complete.obs")
-
-      # Calculate Spearman correlation for the current replicate
-      spearman_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "spearman", use = "complete.obs")
-    }
-
-    # Compute average Kendall Tau and Spearman correlations across all replicates
-    avg_kendall_cor <- mean(kendall_correlations, na.rm = TRUE)
-    avg_spearman_cor <- mean(spearman_correlations, na.rm = TRUE)
-
-    # Show summary results in table
-    output$summary_results <- renderTable({
-      average_ranks
-    })
-
-    # Display average Kendall Tau correlation
-    output$kendall_correlation <- renderPrint({
-      cat("Average Kendall Tau Correlation:", avg_kendall_cor, "\n")
-    })
-
-    # Display average Spearman correlation
-    output$spearman_correlation <- renderPrint({
-      cat("Average Spearman Correlation:", avg_spearman_cor, "\n")
-    })
-  })
-}
-
-shinyApp(ui, server)
-
-```
-
-## Shiny Probability of Top 3 Correct
-
-```{r}
-ui <- fluidPage(
-  titlePanel("Probability of Top 3 Teams in the Correct Order"),
-
-  sidebarLayout(
-    sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
-      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
-      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
-      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
-      checkboxInput("ties", "Allow Ties?", value = TRUE),
-      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
-      actionButton("run_sim", "Run Simulation")
-    ),
-
-    mainPanel(
-      h4("Top 3 Correct"),
-      verbatimTextOutput("top3_output")
-    )
-  )
-)
-
-# Shiny server logic
-server <- function(input, output) {
-  custom_seeding <- reactive({
-    if (input$seeding_input != "") {
-      # Convert the comma-separated string into a numeric vector
-      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
-      seeding_vector[seeding_vector == "NA"] <- NA
-      seeding_vector <- as.numeric(seeding_vector)
-      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
-      # Validate that it matches the number of teams
-      if (length(seeding_vector) == num_teams_next_power_of_two) {
-        return(seeding_vector)
-      } else {
-        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
-        return(NULL)
-      }
-    } else {
-      return(generate_elimination_tournament_structure(input$num_teams))
-    }
-  })
-
-  simulation_results <- eventReactive(input$run_sim, {
-    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
-    seed_structure <- custom_seeding()
-
-    # Run simulations
-    replicate_results <- replicate(
-      input$num_simulations,
-      simulate_single_elimination_tournament(
-        num_teams = input$num_teams,
-        distribution = input$distribution,
-        ties = input$ties,
-        series = input$series,
-        third_place = input$third_place,
-        seeding_structure = seed_structure
-      ),
-      simplify = FALSE
-    )
-
-    # Combine all results into one dataframe and add a Simulation column
-    all_results <- do.call(rbind, replicate_results)
-    all_results$simulation <- rep(1:input$num_simulations, each = nrow(all_results) / input$num_simulations)
-
-    return(all_results)
-  })
-
-  # Identify top-3 ranks for each simulation and calculate the probability
-  top_three_probability <- reactive({
-    all_results <- simulation_results()
-
-    # Group by simulation and filter for top 3 true_rank and rank_hat
-    top_three_results <- all_results %>%
-      group_by(simulation) %>%
-      filter(true_rank %in% 1:3 & rank_hat %in% 1:3) %>%
-      arrange(simulation, rank_hat) %>%
-      summarize(top_3_true_rank = list(true_rank), top_3_rank = list(rank_hat), .groups = 'drop') %>%
-      mutate(match = mapply(function(true, rank) setequal(true, rank), top_3_true_rank, top_3_rank))
-
-    # Calculate probability of matching top 3 true_rank with rank_hat
-    match_count <- sum(top_three_results$match)
-    total_simulations <- n_distinct(all_results$simulation)
-
-    probability <- match_count / total_simulations
-    return(probability)
-  })
-
-  # Output the probability to the UI
-  output$top3_output <- renderText({
-    probability <- top_three_probability()
-
-    paste("Probability of Top 3 Seeds with Ranks 1-3: ", round(probability,4))
-  })
-}
-
-shinyApp(ui=ui,server=server)
-```
-
-## Shiny Conditional Probability Curve
-
-```{r}
-ui <- fluidPage(
-  titlePanel("Conditional Probability of Team Ranks in Tournament"),
-
-  sidebarLayout(
-    sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", 
-                placeholder = "e.g. 1 8 4 5 3 6 2 7"),
-      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
-      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
-      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
-      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
-      actionButton("run_sim", "Run Simulation")
-    ),
-
-    mainPanel(
-      plotOutput("probability_plot")
-    )
-  )
-)
-
-# Server
-server <- function(input, output) {
-
-  # Helper Function: Parse Custom Seeding Structure
-  parse_seeding <- function(seeding_input, num_teams) {
-    if (seeding_input != "") {
-      seeding_vector <- unlist(strsplit(seeding_input, " "))
-      seeding_vector[seeding_vector == "NA"] <- NA
-      seeding_vector <- as.numeric(seeding_vector)
-
-      num_teams_next_power_of_two <- 2^ceiling(log2(num_teams))
-      if (length(seeding_vector) == num_teams_next_power_of_two) {
-        return(seeding_vector)
-      } else {
-        showNotification(
-          paste("Seeding structure length must match", num_teams_next_power_of_two), 
-          type = "error"
-        )
-        return(NULL)
-      }
-    } else {
-      return(NULL)
-    }
-  }
-
-  # Curve Function to Calculate Conditional Probabilities
-  curve <- function(sim, num_teams) {
-    probability <- numeric(num_teams)
-    for (i in 1:num_teams) {
-      if (i == 1) {
-        probability[i] <- nrow(sim[sim$rank_hat == 1 & sim$true_rank == 1, ]) / length(unique(sim$simulation))
-      } else {
-        top_ranks <- sim[sim$rank_hat <= (i - 1) & sim$true_rank <= (i - 1), ]
-        better_teams <- 1:(i - 1)
-        numerator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
-          sim_subset <- sim[sim$simulation == sim_id, ]
-          if (all(better_teams %in% sim_subset$true_rank[sim_subset$rank_hat <= (i - 1)])) {
-            return(sum(sim_subset$rank_hat == i & sim_subset$true_rank == i))
-          } else {
-            return(0)
-          }
-        }))
-        denominator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
-          sim_subset <- top_ranks[top_ranks$simulation == sim_id, ]
-          all(better_teams %in% sim_subset$true_rank)
-        }))
-        probability[i] <- numerator / denominator
-      }
-    }
-    return(probability)
-  }
-
-  # Reactive Event: Run Simulations
-  simulation_data <- eventReactive(input$run_sim, {
-    num_teams <- input$num_teams
-    distribution <- input$distribution
-    series <- input$series
-    num_simulations <- input$num_simulations
-    third_place <- input$third_place
-
-    # Parse custom seeding structure
-    seeding_structure <- parse_seeding(input$seeding_input, num_teams)
-
-    # Run simulations
-    replicate_results <- replicate(
-      num_simulations,
-      simulate_single_elimination_tournament(
-        num_teams = num_teams,
-        distribution = distribution,
-        series = series,
-        ties = FALSE,
-        third_place = third_place,
-        seeding_structure = seeding_structure
-      ),
-      simplify = FALSE
-    )
-
-    # Combine all results into one dataframe
-    all_results <- do.call(rbind, replicate_results)
-    all_results$simulation <- rep(1:num_simulations, each = nrow(all_results) / num_simulations)
-    # list_rbind(x=all_results, names_to = "simulation")
-
-    # Calculate conditional probabilities
-    sim_probs <- curve(all_results, num_teams)
-    probabilities <- data.frame(x = 1:num_teams, y = sim_probs)
-    probabilities$line <- "Simulated"
-
-    # Baseline probabilities
-    baseline_prob <- 1 / (num_teams:1)
-    baseline <- data.frame(x = 1:num_teams, y = baseline_prob)
-    baseline$line <- "Baseline"
-
-    # Combine for plotting
-    rbind(baseline, probabilities)
-  })
-
-  # Plot Output
-  output$probability_plot <- renderPlot({
-    req(simulation_data())
-    ggplot(data = simulation_data(), aes(x = x, y = y, color = line, group = line)) +
-      geom_point() +
-      geom_line() +
-      ylim(0, 1) +
-      labs(
-        x = "Rank",
-        y = "Conditional Probability",
-        title = "Simulated vs Baseline Conditional Probabilities"
-      ) +
-      theme_minimal()
-  })
-}
-
-# Run the App
-shinyApp(ui = ui, server = server)
-```

### 2025-05-01 mutual_information_curves.qmd
update 5/1/25

--- 2025-02-27 mutual_information_curves.qmd
@@
 
-From Wikipedia, weight mutual information formula:
+From Wikipedia, weighted mutual information formula:
 
 $I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}$
+
+Rajski's distance: $1 - \frac{I(X,Y)}{H(X,Y)}$
+
+Interpretation: the fraction of uncertainty not explained by mutual information
+
+<https://en.wikipedia.org/wiki/Inverse_distance_weighting>
+
+Every permutation is not possible to elimination tournaments, so should we divide by every possible permutation for 8 teams, or just every possible permutation in the tournament structure?
 
@@
     for (j in 1:nrow(prob_df)) {
+      #x_vals <- as.numeric(prob_df[j, paste0("x", 1:num_teams)])
+      #y_vals <- as.numeric(prob_df[j, paste0("y", 1:num_teams)])
       # Calculate the squared difference between x_i and y_i for each row j
       diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
-    w[j] <- w[j] + rank_weights[i]* diff^2  # Sum squared differences for each row
+      #rank_diff <- abs(rank(x_vals) - rank(y_vals))
+      w[j] <- w[j] + rank_weights[i] * diff^2  # Sum squared differences for each row
+      # Count inversions (pairs where order is flipped)
+      #inversions <- sum(outer(x_vals, x_vals, ">") & outer(y_vals, y_vals, "<"))
+      #w[j] <- sum(exp(rank_diff) - 1)
+      #w[j] <- w[j] + rank_weights[i] * (exp(inversions) - 1)
+      #w[j] <- cor(x_vals, y_vals, method = "spearman", use = "complete.obs")
+      # Kendal's Tau as weights?
     }
   }
-  w <- ifelse(w == 0.0000, 1, 1 / w)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
-  w <- w/sum(w)
+  #w <- (w + 1) / 2
+  #w[w == Inf] <- 0
+  w <- 1 / (w+1)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
+  #w <- w/sum(w)
   # Compute weights dynamically for each row
@@
     mutate(w = w)
+  #print(weighted_df)
 
-  mutual_info <- factorial(num_teams) * sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/nrow(weighted_df) * weighted_df$probability)))
+  numerator <- sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/factorial(num_teams) * (1/factorial(num_teams)))))
+  denominator <- log2(factorial(num_teams))
+
+  Raj_dist <- 1 - numerator/denominator
 
   # Return the weighted data frame with mutual information weights
-  return(mutual_info)
+  return(Raj_dist)
 }
-mut_info(all_results_norm,8, c(10,5,1,0,0,0,0,0))
-mut_info(de_norm_results_eight,8, c(10,5,1,0,0,0,0,0))
-mut_info(test_rr_noTies,8, c(10,5,1,0,0,0,0,0))
-mut_info(seven_games_se_results,8, c(10,5,1,0,0,0,0,0))
 
@@
 de_norm_results_eight$simulation <- rep(1:10000, each = nrow(de_norm_results_eight) / 10000)
+
 all_results_norm$simulation <- rep(1:10000, each = nrow(all_results_norm) / 10000)
@@
   mutate(rank_hat = rank(rank_hat,ties.method="random"))
+perms_list <- permn(8)
+all_data <- list()
+# Loop through each permutation and create the data frame
+for (i in 1:length(perms_list)) {
+# Create a temporary data frame for the current permutation
+temp_df <- data.frame(
+true_rank = rep(1:8, each = 1),  # Each row gets a true_rank from 1 to 8
+rank_hat = perms_list[[i]],      # The current permutation's ranks
+simulation = rep(i, 8)           # Assign the simulation number to each row
+)
+# Add the temporary data frame to the list
+all_data[[i]] <- temp_df
+}
+# Combine all the data frames into one
+perms_df <- do.call(rbind, all_data)
 
-# Initialize variables
-num_vars <- 8  # Total length of the vector
-indices <- 1:num_vars  # Store the number of 1's in each step
+best <- data.frame(
+  true_rank = rep(1:8, times = 10000),  # Repeat 1-8 for 10,000 simulations
+  rank_hat = rep(1:8, times = 10000),   # Ensure rank_hat matches true_rank
+  simulation = rep(1:10000, each = 8)   # Assign each set to a simulation
+)
+worst <- data.frame(
+  true_rank = rep(1:8, times = 10000),  # Repeat 1-8 for 10,000 simulations
+  rank_hat = rep(8:1, times = 10000),   # Ensure rank_hat matches true_rank
+  simulation = rep(1:10000, each = 8)   # Assign each set to a simulation
+)
 
@@
 compute_mut_plots <- function(dataset) {
-  results <- numeric(num_vars)
-  for (i in 1:num_vars) {
-    binary_vector <- c(rep(1, i), rep(0, num_vars - i))  # Construct vector
-    results[i] <- mut_info(dataset, num_vars, binary_vector)  # Compute MI
+  num_teams <- length(unique(dataset$true_rank))
+  results <- numeric(num_teams)
+  for (i in 1:num_teams) {
+    binary_vector <- c(rep(1, i), rep(0, num_teams - i))  # Construct vector
+    results[i] <- mut_info(dataset, num_teams, binary_vector)  # Compute MI
   }
@@
 mi_seven_se <- compute_mut_plots(seven_games_se_results)
+mi_random <- compute_mut_plots(equal_strengths)
+mi_se_unif <- compute_mut_plots(test_unif)
+mi_bad <- compute_mut_plots(bad_seeding_structure)
+mi_perms <- compute_mut_plots(perms_df)
+mi_best <- compute_mut_plots(best)
+mi_worst <- compute_mut_plots(worst)
 
-# Plot results
-plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm, mi_test_rr, mi_seven_se)),
-     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Mutual Information vs. Number of 1's")
+mi_plots_df <- data.frame(
+indices = 1:8,
+"Double Elimination" = (1-mi_de_norm),
+"Single Elimination" = (1-mi_se_norm),
+"Equal Strengths Single Elimination" = (1-mi_random),
+#"Unif_SE" = (1-mi_se_unif),
+"Bad Seeding Structure Single Elimination" = (1-mi_bad),
+"Every Possible Permutation" = (1-mi_perms),
+"Round Robin" = (1-mi_test_rr),
+"Worst Possible" = (1-mi_worst),
+"Best Possible" = (1-mi_best)
+)
 
-# Add other datasets
-lines(indices, mi_se_norm, type="b", pch=19, col="red")
-lines(indices, mi_test_rr, type="b", pch=19, col="green")
-lines(indices, mi_seven_se, type = "b", pch=19, col="black")
+# Reshape data into long format for ggplot
+df_long <- pivot_longer(mi_plots_df, cols = -indices, names_to = "Method", values_to = "MI")
+ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
+geom_line() +           # Add lines
+geom_point() +          # Add points
+labs(
+title = "Comparison of Tournament Structures with Normal Strengths",
+x = "Top i Number of Teams",
+y = "Mutual Information \\ Entropy (log10 scale)"
+) +
+theme_minimal() +       # Use a clean theme
+theme(legend.position = "right") +
+  scale_y_log10(
+    labels = scales::label_number() 
+  )
 
-# Add legend
-legend("topleft", legend=c("Double Elimination", "Single Elimination", "Round Robin", "7 Game Single Elimination"), 
-       col=c("blue", "red", "green", "black"), pch=19, lty=1)
+normalize <- function(x) {
+  return((x - min(x)) / (max(x) - min(x)))
+}
 
-### Single Elimination vs Double Elimination ###
+# Apply normalization to all columns except "indices"
+mi_norm_plot <- mi_plots_df
+mi_norm_plot[ , -1] <- t(apply(mi_plots_df[ , -1], 1, normalize))
+# Reshape data into long format for ggplot
+df_long_norm <- pivot_longer(mi_norm_plot, cols = -indices, names_to = "Method", values_to = "MI")
+windowsFonts(Arial = windowsFont("Arial"))
+ggplot(df_long_norm, aes(x = indices, y = MI, color = Method, group = Method)) +
+geom_line(linewidth = 2.5) +           # Add lines
+geom_point(size=3) +          # Add points
+labs(
+title = "Comparison of Tournament Structures with Normal Strengths",
+x = "Top i Number of Teams",
+y = "Normalized Mutual Information \\ Entropy (log10 scale)"
+) +
+theme_minimal() +       # Use a clean theme
+theme(text = element_text(family = "Arial", size = 32),
+  legend.position = "bottom",
+  legend.box = "vertical",
+  plot.title = element_text(hjust = 0.5, size = 36),
+  axis.title = element_text(size = 30),
+  axis.text = element_text(size=30)) +
+  guides(color = guide_legend(nrow = 4)) +
+  scale_y_log10(
+    labels = scales::label_number() 
+  ) +
+  scale_color_viridis_d(option = "H")
+ggsave(
+  filename = "tournament_comparison_plot.png",  # Change file type if needed
+  plot = last_plot(),  # Saves the last ggplot created
+  dpi = 600,        
+  bg = "white",
+  units = "in",
+  width = 15.28,
+  height = 13.43
+)
+```
 
-# Plot results
-plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm)),
-     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Single Elimination vs Double Elimination")
+include number of games (might be a range)
 
-# Add other datasets
-lines(indices, mi_se_norm, type="b", pch=19, col="red")
+Single Elimination - third place and without third place; series = 3,5,7
 
-# Add legend
-legend("bottomright", legend=c("Double Elimination", "Single Elimination"), 
-       col=c("blue", "red"), pch=19, lty=1)
+Double Elimination
 
-```
+-   Repechage
+
+-   True Double Elimination
+
+-   Consolation Bracket (loser's bracket best place is third)
+
+Triple Elimination
+
+Deterministic Outcomes
+
+Round Robin - with k rounds (1,2,3,5,7)
+
+Stepladder - 8 vs 7, winner plays 6, etc.
+
+Group Stage - winners of group advance, top 2 advance
+
+Staged Round Robin/Jacobian Ladder - top 2 in each group form new group, bottom 2 new group (each new group is a new round robin)
+
+High School Tennis Ladder - 1 vs 2, 3 vs 4, etc., Re-rank 2 vs 3, 4 vs 5, 6 vs 7, number of teams - 1 iterations

### 2025-05-01 single_elimination_shiny.qmd
update 5/1/25

@@
     strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
+    normal_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      rank_hat = rep(NA,length(teams)),
+      game_wins = rep(0,length(teams)),
+      game_losses = rep(0,length(teams))
+    )
+    df <- arrange(normal_strengths, true_rank)
+  }
+  else if (distribution == "Same"){
+    strengths <- sapply(num_teams, function(n) {rep(0,n)})
     normal_strengths <- data.frame(
@@
 
-test <- simulate_single_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, third_place=T)
+  test <- list()
+for (teams in c(8)) {
+  test[[as.character(teams)]] <- list()
+  for (replicate in 1:10000) {
+    # Store each simulation result in the list with unique indexing
+    test[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
+      num_teams = teams,
+      distribution = "Same",
+      ties = FALSE,
+      series = 1,
+      third_place = TRUE
+    )
+  }
+}
+}
+test <- simulate_single_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, third_place=T, seeding_structure = c(1,2,3,4,5,6,7,8))
 test
@@
 all_normal_results <- list()
-for (teams in c(4,8,16,32,64)) {
+for (teams in c(8)) {
   all_normal_results[[as.character(teams)]] <- list()
@@
       series = 1,
-      third_place = TRUE
+      third_place = FALSE
     )
@@
 all_results_unif
+
+seven_games_se_results <- list()
+for (teams in c(8)) {
+  seven_games_se_results[[as.character(teams)]] <- list()
+  for (replicate in 1:10000) {
+    # Store each simulation result in the list with unique indexing
+    seven_games_se_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
+      num_teams = teams,
+      distribution = "Normal",
+      ties = FALSE,
+      series = 7,
+      third_place = TRUE
+    )
+  }
+}
+
+seven_games_se_results <- do.call(rbind, lapply(names(seven_games_se_results), function(teams) {
+  team_results <- do.call(rbind, seven_games_se_results[[teams]])
+  team_results$teams <- as.numeric(teams)
+  return(team_results)
+}))
 

### 2025-05-01 all_possible_8team_brackets.R

@@ -1,5 +1,5 @@
 tourn <- do.call(rbind, args = combinat::permn(1:8))
-
+tourn
 dist <- apply(X = tourn, MARGIN = 1, FUN = function(x){
   if (sum(x[5:8] == 1) > 0) {
     x[1:8] <- x[c(5:8, 1:4)]

### 2025-05-01 manuscript/paper.Rmd
paper update with Results section

@@
   | \normalsize Chicago \vspace{-1mm}
-  | \normalsize [`email`](mailto:ypu@something.edu) \vspace{-1mm}
   |
@@
   | \normalsize Austin, TX 78712 \vspace{-1mm}
-  | \normalsize [`brian.mills@austin.utexas.edu`](mailto:brian.mills@austin.utexas.edu) \vspace{-1mm}
   |
@@
   | \normalsize Chicago, IL 60660 \vspace{-1mm}
-  | \normalsize [`lderango@luc.edu`](mailto:lderango@luc.edu) \vspace{-1mm}
   |
@@
   | \normalsize Chicago, IL 60660 \vspace{-1mm}
+  | \normalsize [`email`](mailto:ypu@something.edu) \vspace{-1mm}
+  | \normalsize [`brian.mills@austin.utexas.edu`](mailto:brian.mills@austin.utexas.edu) \vspace{-1mm}
+  | \normalsize [`lderango@luc.edu`](mailto:lderango@luc.edu) \vspace{-1mm}
   | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm}
@@
  \floatplacement{figure}{t}
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
@@
 # Introduction
-There are hundreds of different tournament structures that dictate how a set of teams compete against each other to determine an overall ranking.
-Different tournament structures have unique strengths and weaknesses, affecting how well they reflect the true rankings of the teams.
-The goal of a tournament is to find the best teams, but how can we quantify how well a tournament performs?
-In some instances, the tournament organizers only award the overall winner, but other times, the top three or ten teams are awarded.
-Because of this, organizers may prefer one structure over another based on their needs.
-Tournament organizers must also take into account factors like cost, timeliness, entertainment value, and fairness when choosing a tournament structure.
 
-To evaluate a tournament’s effectiveness, we propose a numerical metric that quantifies how accurately it orders teams based on their true rankings.
-Tournament results can be viewed as a "message" attempting to convey the true team rankings.
-However, various factors introduce noise, leading to information loss.
-We use principles from information theory to measure this information loss across multiple tournament simulations to assess the reliability of different formats.
+There are hundreds of different tournament structures that dictate how a
+set of teams compete against each other to determine an overall ranking.
+Different tournament structures have unique strengths and weaknesses,
+affecting how well they reflect the true rankings of the teams. The goal
+of a tournament is to find the best teams, but how can we quantify how
+well a tournament performs? In some instances, the tournament organizers
+only award the overall winner, but other times, the top three or ten
+teams are awarded. Because of this, organizers may prefer one structure
+over another based on their needs. Tournament organizers must also take
+into account factors like cost, timeliness, entertainment value, and
+fairness when choosing a tournament structure.
+
+To evaluate a tournament’s effectiveness, we propose a numerical metric
+that quantifies how accurately it orders teams based on their true
+rankings. Tournament results can be viewed as a "message" attempting to
+convey the true team rankings. However, various factors introduce noise,
+leading to information loss. We use principles from information theory
+to measure this information loss across multiple tournament simulations
+to assess the reliability of different formats.
 
 # Methods
-Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 > \theta_2 > \ldots > \theta_n)$.
-Next, let $r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when theta is sorted from largest to smallest so that $r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}$.
-Next define $T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure $s$.
-$\phi$ contains all the information about the scheduling of teams, which could be fully known prior to the tournament (i.e.
-round robin) or determined as the tournament progresses (i.e.
-single elimination tournament).
-We then let  $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an $n$- dimensional vector valued random variable that gives the results of a tournament as a vector of the indexes of the vector $\boldsymbol{\theta}$.
-The index in the first position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished first, the index in the second position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished second, and so on.
 
-For example, in a 4 team tournament, $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that the team with index 3 (i.e.
-the true third best team) won the tournament, the true second best team finished second, the true best team finished third and the true 4th team finished 4th in that particular tournament.
-If $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (1, 2, 3, 4), this means that the random outcome of the tournament was the same as the true ordering of the teams.
+Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with
+an associated true strength parameter
+$\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that
+$(\theta_1 > \theta_2 > \ldots > \theta_n)$. Next, let
+$r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when
+theta is sorted from largest to smallest so that
+$r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}$. Next define
+$T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure
+$s$. $\phi$ contains all the information about the scheduling of teams,
+which could be fully known prior to the tournament (i.e. round robin) or
+determined as the tournament progresses (i.e. single elimination
+tournament). We then let $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an
+$n$- dimensional vector valued random variable that gives the results of
+a tournament as a vector of the indexes of the vector
+$\boldsymbol{\theta}$. The index in the first position of the vector
+$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
+team that finished first, the index in the second position of the vector
+$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
+team that finished second, and so on.
 
-If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the "goodness" of a tournament in terms of its ability to accurately transmit the true ranking.
-We can then use concepts from information theory to assess the ability of a tournament to correctly rank teams.
-Specifically, we start with the concept of mutual information (CITE).
-For two random variables $X$ and $Y$ mutual information is defined as:
+For example, in a 4 team tournament,
+$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that
+the team with index 3 (i.e. the true third best team) won the
+tournament, the true second best team finished second, the true best
+team finished third and the true 4th team finished 4th in that
+particular tournament. If $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ =
+(1, 2, 3, 4), this means that the random outcome of the tournament was
+the same as the true ordering of the teams.
+
+If one views the true ranking of the teams in a tournament as a message
+to be sent to a receiver and the outcome of the tournament as a message
+that is received, we can measure the "goodness" of a tournament in terms
+of its ability to accurately transmit the true ranking. We can then use
+concepts from information theory to assess the ability of a tournament
+to correctly rank teams. Specifically, we start with the concept of
+mutual information (CITE). For two random variables $X$ and $Y$ mutual
+information is defined as:
 
@@
 
-In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and $\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate $I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$.
-For simplicity, we drop the arguments from the functions $r$ and $\hat{r}$ for ease of exposition.
+In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and
+$\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate
+$I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$. For
+simplicity, we drop the arguments from the functions $r$ and $\hat{r}$
+for ease of exposition.
 
-We note here that $r$ is not a random variable, however, since the indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary (we use 1, 2, 3, etc.
-for convenience, but any indexing is valid), we can view this a random variable where all permutation so of the $n$ teams are equally likely.
-Therefore when calculating  the mutual information in this setting, we set $p(r) = \frac{1}{n!}$.
-By a similar argument we set $p(\hat{r}) = \frac{1}{n!}$.
+We note here that $r$ is not a random variable, however, since the
+indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary
+(we use 1, 2, 3, etc. for convenience, but any indexing is valid), we
+can view this a random variable where all permutation so of the $n$
+teams are equally likely. Therefore when calculating the mutual
+information in this setting, we set $p(r) = \frac{1}{n!}$. By a similar
+argument we set $p(\hat{r}) = \frac{1}{n!}$.
 
@@
 
-While there are $n!$ different permutations for the result of $r$, we don't need to sum across these as any specific choice of $r$ is just an arbitrary labeling of the true strength parameters vector $\boldsymbol{\theta}$.
-So given $r$, the distribution of $\hat{r}$ is the same up to the labeling.
-Therefore, we only need to consider a single permutation of $r$, we compute the the quantity inside the summation, and then multiple by $n!$ (effectively summing across $r$).
+While there are $n!$ different permutations for the result of $r$, we
+don't need to sum across these as any specific choice of $r$ is just an
+arbitrary labeling of the true strength parameters vector
+$\boldsymbol{\theta}$. So given $r$, the distribution of $\hat{r}$ is
+the same up to the labeling. Therefore, we only need to consider a
+single permutation of $r$, we compute the the quantity inside the
+summation, and then multiple by $n!$ (effectively summing across $r$).
 
 In order to compute this quantity, we need to compute $p(r,\hat{r})$.
-This probability is found by calculating the probability of a given permutation of $\hat{r}$ and $r$ is assumed to be the permutation from $\{1, 2, \ldots, n\}$.
-We estimate $p(r,\hat{r})$ empirically through simulation.
+This probability is found by calculating the probability of a given
+permutation of $\hat{r}$ and $r$ is assumed to be the permutation from
+$\{1, 2, \ldots, n\}$. We estimate $p(r,\hat{r})$ empirically through
+simulation.
 
-However, using mutual information in this form does not suit our needs in this setting.
-The problem is that mutual information in this form will yield high values of mutual information when the output from the tournament $\hat{r}$ is highly consistent *even if the ranking from the tournament is incorrect*.
-As an example, if $n=4$ and the true order of \boldsymbol{\theta} is $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in both of these cases the mutual information will be maximized at:
+However, using mutual information in this form does not suit our needs
+in this setting. The problem is that mutual information in this form
+will yield high values of mutual information when the output from the
+tournament $\hat{r}$ is highly consistent *even if the ranking from the
+tournament is incorrect*. As an example, if $n=4$ and the true order of
+\boldsymbol{\theta} is $r = \{1, 2, 3, 4\}$ and
+$p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as
+when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in
+both of these cases the mutual information will be maximized at:
 
@@
 I(r,\hat{r})= n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{\left(\frac{1}{n!}\right)^2} = n! log((n!)^2)
-$$
-and for the specific case when $n=4$ would be: 
-$$
+$$ and for the specific case when $n=4$ would be: $$
 4! log_2(4!^2) = 24*log_2(24^2) = 220.0782
@@
 
-In order to alleviate this problem, we instead consider *weighted* mutual information.
-We want to give more weight to permutations from $\hat{r}$ that are "closer" to $r$.
-There are any number of reasonable choices for this weighting function, and we choose to use squared error.
-Specifically,
-$$
-w(r, \hat{r}) =
+In order to alleviate this problem, we instead consider *weighted*
+mutual information. We want to give more weight to permutations from
+$\hat{r}$ that are "closer" to $r$. There are any number of reasonable
+choices for this weighting function, and we choose to use squared error.
+Specifically, \$\$ w(r, \hat{r}) =
 
-\left\{  
+\\left{\
+
 \begin{array}{cc}
@@
 \end{array}
-\right. 
-$$
+
+\right. $$
 In general, though any loss function $l(r, \hat{r})$ can be used:
-$$
-w(r, \hat{r}) =
+$$ w(r, \hat{r}) =
 
-\left\{  
+\\left{\
+
 \begin{array}{cc}
@@
 \end{array}
-\right. 
-$$
+
+\right. \$\$
 
@@
 
-And this does work because consistently getting the order incorrect is perfect information.
+And this does work because consistently getting the order incorrect is
+perfect information.
 
@@
 
-Rajski's distance: 
-$$
+Rajski's distance: $$
 1 - \frac{I(X,Y)}{H(X,Y)}
@@
 <!-- # ryan way  -->
+
 <!-- Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 < \theta_2 < \ldots < \theta_n)$.
@@
 
-Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n)$ be the permutation of indexes of $\boldsymbol{\theta}$ such that $x_1 = {i | \theta_i = max(\boldsymbol{\theta})}$
+Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n)$ be the permutation of
+indexes of $\boldsymbol{\theta}$ such that
+$x_1 = {i | \theta_i = max(\boldsymbol{\theta})}$
 
@@
 
-Zach, put your stuff here.
+In order to fairly compare tournament structures, we assume the true
+strengths of the teams follow a normal distribution, where each team is
+assigned an equally spaced quantile and each team's strength is the
+z-score of its respective quantile. Using this assumption, 10,000
+simulations, all with 8 total teams, were ran for a single game round
+robin (each team plays each other once), a single elimination structure
+with the usual seeding structure (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7), a
+single elimination structure with a poor seeding structure (1 vs 2, 3 vs
+4, 5 vs 6, 7 vs 8). For a better understanding of the results, the best
+possible results (every structure has the best in rank 1, the 2nd best
+in rank 2, etc), the worst possible results (every structure has the
+worst in rank 1, the 2nd worst in rank 2, etc), every possible
+permutation of the ranks once, and a single elimination tournament with
+the usual seeding structure where every team is of equal strengths.
+Because the metric found is unitless, we were able to normalize the
+values using the best and worst results to make the range from 0 to 1,
+with 0 as the worst and 1 as the best. After normalizing, we get the
+following graph (note the y-axis is on a log10 scale):
+
+![](images/Normal_Information_Curve_Graph.png){width="408"}
+
+Based on the graph, the intuitive best tournament, round robin, does
+show to be the best. Double elimination is slightly better than single
+elimination, with the gap growing wider as the top number of teams
+looked at increases. However, those are under the assumption that the
+true strengths were ordered correctly in the seeding. If there is a poor
+seeding structure, as seen in the graph, it can drastically impact the
+effectiveness of the tournament. In the extreme case seen, the
+tournament can actually become worse than deciding games by random
+chance (like flipping a coin). Another important discovery is that the
+single elimination structure with equal strengths is better than every
+possible permutation. The line for every possible permutation would
+likely be similar to a round robin structure with all equal strengths,
+so perhaps the closer the teams are in true strength, the closer round
+robin and single elimination become.
 
@@
 
-# Acknowledgements {-}
+# Acknowledgements {.unnumbered}
 
-# Supplementary Material {-}
+# Supplementary Material {.unnumbered}
 

### 2025-08-29 manuscript/paper.Rmd
First draft paper

@@
   | \normalsize  \vspace{-1mm}
-  | \normalsize Loyola Chicago \vspace{-1mm}
-  | \normalsize Chicago \vspace{-1mm}
+  | \normalsize Loyola University Chicago \vspace{-1mm}
+  | \normalsize Chicago, IL 60660 \vspace{-1mm}
   |
-  | \large Josie Peterburgs \vspace{-1.1mm}
-  | \normalsize College of Education \vspace{-1mm}
-  | \normalsize University of Texas at Austin \vspace{-1mm}
-  | \normalsize Austin, TX 78712 \vspace{-1mm}
+  | \large Josie Peterburs \vspace{-1.1mm}
+  | \normalsize \vspace{-1mm}
+  | \normalsize Loyola University Chicago \vspace{-1mm}
+  | \normalsize Chicago, IL 60660 \vspace{-1mm}
   |
@@
   | \normalsize [`email`](mailto:ypu@something.edu) \vspace{-1mm}
-  | \normalsize [`brian.mills@austin.utexas.edu`](mailto:brian.mills@austin.utexas.edu) \vspace{-1mm}
-  | \normalsize [`lderango@luc.edu`](mailto:lderango@luc.edu) \vspace{-1mm}
+  | \normalsize [`zculp@luc.edu`](mailto:zculp@luc.edu) \vspace{-1mm}
+  | \normalsize [`jpeterburs@luc.edu`](mailto:jpeterburs@luc.edu) \vspace{-1mm}
+  | \normalsize [`rmcshane@uchicago.edu`](mailto:rmcshane@uchicago.edu) \vspace{-1mm}
   | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm}
 abstract: | 
- | Here is the abstract.
+ | We evaluate multiple tournament formats to assess their effectiveness in accurately reproducing the true underlying ranking of participating teams.
+Using repeated simulations for each structure, we compare teams’ true performance ranks to their simulated outcomes.
+To quantify the accuracy of each format, we calculate weighted mutual information between true and simulated rankings across a large number of replicates.
+"Remarkable" weights are applied to look at only the top n teams/competitors in a tournament.
+This information-theoretic approach allows us to objectively compare formats and identify which structures most effectively preserve rank order.
+Results are visualized through comparative plots, providing clear insights into the trade-offs and strengths of each tournament design.
  | 
   \vspace{2mm}
-  | Keywords: survival analysis, hockey
+  | Keywords: tournament structures, mutual information
 bibliography: references.bib
@@
 I(X;Y)=\sum_{y\in Y}\sum_{x\in X}p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
-$$.
+$$
 
@@
 I(r,\hat{r})=\sum_{r}\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})} = n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{(\frac{1}{n!})^2}
-$$.
+$$
 
@@
 tournament is incorrect*. As an example, if $n=4$ and the true order of
-\boldsymbol{\theta} is $r = \{1, 2, 3, 4\}$ and
+$\boldsymbol{\theta}$ is $r = \{1, 2, 3, 4\}$ and
 $p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as
@@
 4! log_2(4!^2) = 24*log_2(24^2) = 220.0782
-$$.
+$$
 
@@
 choices for this weighting function, and we choose to use squared error.
-Specifically, \$\$ w(r, \hat{r}) =
+Specifically,
 
-\\left{\
+$$
+w(r, \hat{r}) =
+\begin{cases}
+\frac{1}{\hat{r}'r}, & r \ne \hat{r} \\
+1, & r = \hat{r}
+\end{cases}
+$$
 
-\begin{array}{cc}
-    \begin{array}{cc}
-      \frac{1}{\hat{r}'r} & r \ne \hat{r} \\
-      1 & r = \hat{r} \\
-    \end{array}
-\end{array}
+In general, though any loss function $l(r, \hat{r})$ can be used:
 
-\right. $$
-In general, though any loss function $l(r, \hat{r})$ can be used:
-$$ w(r, \hat{r}) =
-
-\\left{\
-
-\begin{array}{cc}
-    \begin{array}{cc}
-      \frac{1}{l(r, \hat{r})} & r \ne \hat{r} \\
-      1 & r = \hat{r} \\
-    \end{array}
-\end{array}
-
-\right. \$\$
+$$
+w(r, \hat{r}) =
+\begin{cases}
+\frac{1}{l(r, \hat{r})}, & r \ne \hat{r} \\
+1, & r = \hat{r}
+\end{cases}
+$$
 
@@
 
+However, because the number of teams that need to be accurately ordered
+differs based on the tournament organizers, a second weighting function
+was found to portray the importance, or remarkability, of each value in
+$\hat{r}$. We call these "remarkable weights" ($m(r)$) to differentiate
+from the other weighting function. "Remarkable weights" are set to allow
+for a subjective belief of what determines a good tournament. For a
+four-team tournament, if the user decides that only the winner is
+important, then $m(r) = \{1,0,0,0\}$. However, if the organizers want an
+accurate leaderboard with the winner as the most important, then a
+potential $m(r)$ could be $\{3,1,1,0\}$. With the addition of the
+weighting functions, the formula for mutual information is updated:
+
 $$
-I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
+I(X;Y)=\sum_{r}\sum_{\hat{r}} w(r,\hat{r}) * m(r) * p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
 $$
 
-Let $x$ be
+Because this mutual information is unitless, the formula can be
+standardized to be between 0 and 1 using the complement of Rajski's
+distance: $$
+1 - (1 - \frac{I(X,Y)}{\text{max\{}H(X), H(Y)\}})
+$$
 
-x is the true ranking and y is the output
+where $H(X,Y)$ is entropy given by:
 
-Rajski's distance: $$
-1 - \frac{I(X,Y)}{H(X,Y)}
 $$
+max\{H(X), H(Y)\} = \sum_{r}  p(r) \log p(r) = n! \log n!
+$$
+
+The complement is taken so that the more accurate tournaments are closer
+to 1 and the less accurate are closer to 0.
 
@@
 
-$r = \{1,2,3,4\}$ which is the
-
-Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n)$ be the permutation of
-indexes of $\boldsymbol{\theta}$ such that
-$x_1 = {i | \theta_i = max(\boldsymbol{\theta})}$
-
-Let $x$ be
-
-x is the true ranking and y is the output
-
-$$
-I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
-$$
-
-Rajski's distance: $$
-1 - \frac{I(X,Y)}{H(X,Y)}
-$$
+<!-- Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n)$ be the permutation of indexes of $\boldsymbol{\theta}$ such that $x_1 = {i | \theta_i = max(\boldsymbol{\theta})}$ -->
 
@@
 
+This study introduced a weighted mutual information framework to
+evaluate the effectiveness of different tournament formats in preserving
+the true underlying rankings of teams. By simulating tournaments with
+varying seeding structures and applying an information-theoretic
+approach, we quantified how accurately each format conveys ranking
+information. Our findings confirm the intuitive advantage of round robin
+tournaments, which consistently outperformed single and double
+elimination formats in ranking accuracy. Additionally, we observed that
+seeding plays a critical role: poor seeding structures can significantly
+degrade performance, in some cases making outcomes less informative than
+random chance.
+
+These results underscore the trade-off between tournament accuracy and
+practical constraints such as time, cost, and entertainment value. While
+round robin offers the highest fidelity to true rankings, it is often
+impractical for large tournaments, highlighting the need for hybrid or
+adaptive structures that balance accuracy with logistical feasibility.
+
+Future work should extend this analysis to scenarios with more
+tournament structures, more teams, incorporate probabilistic strength
+models reflecting real-world uncertainty, and explore alternative
+weighting schemes for different competitive priorities (e.g., top-4
+accuracy versus full ranking). Applying these methods to actual
+tournament data could further validate their usefulness for organizers
+aiming to design fair and informative competitions.
+
 # Acknowledgements {.unnumbered}
+
+We thank the Department of Mathematics and Statistics at Loyola
+University Chicago for their support and resources in conducting this
+study. No external funding was received for this research.
 
@@
 
+All supplementary material available at
+<https://github.com/gjm112/tournaments>.
+
 # References
+
+Rajski, C. (1961). "A metric space of discrete probability
+distributions". Information and Control. 4 (4): 371–377.
+[doi:10.1016/S0019-9958(61)80055-7](doi:10.1016/S0019-9958(61)80055-7){.uri}.
+
+McShane, R. (2019). *Modeling Stochastically Intransitive Relationships
+in Paired Comparison Data.* (Doctoral dissertation). Southern Methodist
+University. Retried from
+<https://scholar.smu.edu/hum_sci_statisticalscience_etds/13/>
+
+Karel Devriesere, László Csató, Dries Goossens (2024). "Tournament
+design: A review from an operational research perspective". European
+Journal of Operational Research.
+<https://doi.org/10.1016/j.ejor.2024.10.044>.

### 2025-08-29 manuscript/paper.tex
First draft paper

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 \setcounter{secnumdepth}{5}
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
-  \fi
-  % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 \usepackage{setspace} \setstretch{1.15} \usepackage{float} \floatplacement{figure}{t}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \usepackage{bookmark}
@@
 \author{\large Zach Culp \vspace{-1.1mm}\\
-\normalsize Epi \vspace{-1mm}\\
-\normalsize Oxford something at Emory \vspace{-1mm}\\
-\normalsize Georgia? \vspace{-1mm}\\
-\normalsize \href{mailto:ypu@something.edu}{\texttt{email}}
-\vspace{-1mm}\\
+\normalsize  \vspace{-1mm}\\
+\normalsize Loyola University Chicago \vspace{-1mm}\\
+\normalsize Chicago, IL 60660 \vspace{-1mm}\\
 \strut \\
-\large Josie Peterburgs \vspace{-1.1mm}\\
-\normalsize College of Education \vspace{-1mm}\\
-\normalsize University of Texas at Austin \vspace{-1mm}\\
-\normalsize Austin, TX 78712 \vspace{-1mm}\\
-\normalsize \href{mailto:brian.mills@austin.utexas.edu}{\texttt{brian.mills@austin.utexas.edu}}
-\vspace{-1mm}\\
+\large Josie Peterburs \vspace{-1.1mm}\\
+\normalsize \vspace{-1mm}\\
+\normalsize Loyola University Chicago \vspace{-1mm}\\
+\normalsize Chicago, IL 60660 \vspace{-1mm}\\
 \strut \\
@@
 \normalsize Chicago, IL 60660 \vspace{-1mm}\\
-\normalsize \href{mailto:lderango@luc.edu}{\texttt{lderango@luc.edu}}
-\vspace{-1mm}\\
 \strut \\
@@
 \normalsize Chicago, IL 60660 \vspace{-1mm}\\
+\normalsize \href{mailto:ypu@something.edu}{\texttt{email}}
+\vspace{-1mm}\\
+\normalsize \href{mailto:zculp@luc.edu}{\texttt{zculp@luc.edu}}
+\vspace{-1mm}\\
+\normalsize \href{mailto:jpeterburs@luc.edu}{\texttt{jpeterburs@luc.edu}}
+\vspace{-1mm}\\
+\normalsize \href{mailto:rmcshane@uchicago.edu}{\texttt{rmcshane@uchicago.edu}}
+\vspace{-1mm}\\
 \normalsize \href{mailto:gmatthews1@luc.edu}{\texttt{gmatthews1@luc.edu}}
@@
 \begin{abstract}
-Here is the abstract.\\
+We evaluate multiple tournament formats to assess their effectiveness in
+accurately reproducing the true underlying ranking of participating
+teams. Using repeated simulations for each structure, we compare teams'
+true performance ranks to their simulated outcomes. To quantify the
+accuracy of each format, we calculate weighted mutual information
+between true and simulated rankings across a large number of replicates.
+``Remarkable'' weights are applied to look at only the top n
+teams/competitors in a tournament. This information-theoretic approach
+allows us to objectively compare formats and identify which structures
+most effectively preserve rank order. Results are visualized through
+comparative plots, providing clear insights into the trade-offs and
+strengths of each tournament design.\\
 
-\vspace{2mm} \textbar{} Keywords: survival analysis, hockey
+\vspace{2mm} \textbar{} Keywords: tournament structures, mutual
+information
 \end{abstract}
@@
 
-Popkin et al. (\citeproc{ref-Popkin2023}{2023})
-
 There are hundreds of different tournament structures that dictate how a
 set of teams compete against each other to determine an overall ranking.
-Different tournaments have unique strengths and weaknesses, affecting
-how well they reflect the true rankings of the teams. The goal of a
-tournament is to find the best teams, but how can we quantify how well a
-tournament performs? In some instances, the tournament organizers only
-award the overall winner, but other times, the top three or ten teams
-are awarded. Because of this, organizers may prefer one structure over
-another based on their needs. Tournament organizers must also take into
-account factors like cost, timeliness, entertainment value, and fairness
-when choosing a tournament structure.
+Different tournament structures have unique strengths and weaknesses,
+affecting how well they reflect the true rankings of the teams. The goal
+of a tournament is to find the best teams, but how can we quantify how
+well a tournament performs? In some instances, the tournament organizers
+only award the overall winner, but other times, the top three or ten
+teams are awarded. Because of this, organizers may prefer one structure
+over another based on their needs. Tournament organizers must also take
+into account factors like cost, timeliness, entertainment value, and
+fairness when choosing a tournament structure.
 
-\begin{verbatim}
-To evaluate a tournament’s effectiveness, we propose a numerical metric that quantifies how accurately it orders teams based on their true rankings.
-Tournament results can be viewed as a “message” attempting to convey the true team rankings.
-However, various factors introduce noise, leading to information loss.
-We use principles from information theory to measure this information loss across multiple tournament simulations to assess the reliability of different formats.
-\end{verbatim}
+To evaluate a tournament's effectiveness, we propose a numerical metric
+that quantifies how accurately it orders teams based on their true
+rankings. Tournament results can be viewed as a ``message'' attempting
+to convey the true team rankings. However, various factors introduce
+noise, leading to information loss. We use principles from information
+theory to measure this information loss across multiple tournament
+simulations to assess the reliability of different formats.
 
@@
 
+Consider a set of \(n\) teams indexed from \(i = 1, 2, \cdots, n\) each
+with an associated true strength parameter
+\(\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)\) such
+that \((\theta_1 > \theta_2 > \ldots > \theta_n)\). Next, let
+\(r(\boldsymbol{\theta})\) be the indexes of \(\boldsymbol{\theta}\)
+when theta is sorted from largest to smallest so that
+\(r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}\). Next define
+\(T(\phi, s)\) as a tournament with schedule \(\phi\) and seeding
+structure \(s\). \(\phi\) contains all the information about the
+scheduling of teams, which could be fully known prior to the tournament
+(i.e.~round robin) or determined as the tournament progresses
+(i.e.~single elimination tournament). We then let
+\(\hat{r}(T(\phi, s),\boldsymbol{\theta})\) be an \(n\)- dimensional
+vector valued random variable that gives the results of a tournament as
+a vector of the indexes of the vector \(\boldsymbol{\theta}\). The index
+in the first position of the vector
+\(\hat{r}(T(\phi, s),\boldsymbol{\theta})\) indicates the index of the
+team that finished first, the index in the second position of the vector
+\(\hat{r}(T(\phi, s),\boldsymbol{\theta})\) indicates the index of the
+team that finished second, and so on.
+
+For example, in a 4 team tournament,
+\(\hat{r}(T(\phi, s),\boldsymbol{\theta})\) = (3, 2, 1, 4) indicates
+that the team with index 3 (i.e.~the true third best team) won the
+tournament, the true second best team finished second, the true best
+team finished third and the true 4th team finished 4th in that
+particular tournament. If \(\hat{r}(T(\phi, s),\boldsymbol{\theta})\) =
+(1, 2, 3, 4), this means that the random outcome of the tournament was
+the same as the true ordering of the teams.
+
+If one views the true ranking of the teams in a tournament as a message
+to be sent to a receiver and the outcome of the tournament as a message
+that is received, we can measure the ``goodness'' of a tournament in
+terms of its ability to accurately transmit the true ranking. We can
+then use concepts from information theory to assess the ability of a
+tournament to correctly rank teams. Specifically, we start with the
+concept of mutual information (CITE). For two random variables \(X\) and
+\(Y\) mutual information is defined as:
+
+\[
+I(X;Y)=\sum_{y\in Y}\sum_{x\in X}p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
+\]
+
+In our setting, we replace \(X\) and \(Y\) with
+\(r(\boldsymbol{\theta})\) and
+\(\hat{r}(T(\phi, s),\boldsymbol{\theta})\), and we could seek to
+estimate
+\(I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))\).
+For simplicity, we drop the arguments from the functions \(r\) and
+\(\hat{r}\) for ease of exposition.
+
+We note here that \(r\) is not a random variable, however, since the
+indexing of the teams in the vector \(\boldsymbol{\theta}\) is arbitrary
+(we use 1, 2, 3, etc. for convenience, but any indexing is valid), we
+can view this a random variable where all permutation so of the \(n\)
+teams are equally likely. Therefore when calculating the mutual
+information in this setting, we set \(p(r) = \frac{1}{n!}\). By a
+similar argument we set \(p(\hat{r}) = \frac{1}{n!}\).
+
+We define the mutual information of \(r\) and \(\hat{r}\) to be:
+
+\[
+I(r,\hat{r})=\sum_{r}\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})} = n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{(\frac{1}{n!})^2}
+\]
+
+While there are \(n!\) different permutations for the result of \(r\),
+we don't need to sum across these as any specific choice of \(r\) is
+just an arbitrary labeling of the true strength parameters vector
+\(\boldsymbol{\theta}\). So given \(r\), the distribution of \(\hat{r}\)
+is the same up to the labeling. Therefore, we only need to consider a
+single permutation of \(r\), we compute the the quantity inside the
+summation, and then multiple by \(n!\) (effectively summing across
+\(r\)).
+
+In order to compute this quantity, we need to compute \(p(r,\hat{r})\).
+This probability is found by calculating the probability of a given
+permutation of \(\hat{r}\) and \(r\) is assumed to be the permutation
+from \(\{1, 2, \ldots, n\}\). We estimate \(p(r,\hat{r})\) empirically
+through simulation.
+
+However, using mutual information in this form does not suit our needs
+in this setting. The problem is that mutual information in this form
+will yield high values of mutual information when the output from the
+tournament \(\hat{r}\) is highly consistent \emph{even if the ranking
+from the tournament is incorrect}. As an example, if \(n=4\) and the
+true order of \(\boldsymbol{\theta}\) is \(r = \{1, 2, 3, 4\}\) and
+\(p(\hat{r} = \{4, 3, 2, 1\}) = 1\) will be the same mutual information
+as when \(r = \{1, 2, 3, 4\}\) and \(p(\hat{r} = \{1, 2, 3, 4\}) = 1\)
+and in both of these cases the mutual information will be maximized at:
+
+\[
+I(r,\hat{r})= n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{\left(\frac{1}{n!}\right)^2} = n! log((n!)^2)
+\] and for the specific case when \(n=4\) would be: \[
+4! log_2(4!^2) = 24*log_2(24^2) = 220.0782
+\]
+
+In order to alleviate this problem, we instead consider \emph{weighted}
+mutual information. We want to give more weight to permutations from
+\(\hat{r}\) that are ``closer'' to \(r\). There are any number of
+reasonable choices for this weighting function, and we choose to use
+squared error. Specifically,
+
+\[
+w(r, \hat{r}) =
+\begin{cases}
+\frac{1}{\hat{r}'r}, & r \ne \hat{r} \\
+1, & r = \hat{r}
+\end{cases}
+\]
+
+In general, though any loss function \(l(r, \hat{r})\) can be used:
+
+\[
+w(r, \hat{r}) =
+\begin{cases}
+\frac{1}{l(r, \hat{r})}, & r \ne \hat{r} \\
+1, & r = \hat{r}
+\end{cases}
+\]
+
+Note that one is 1/N!
+
+And this does work because consistently getting the order incorrect is
+perfect information.
+
+However, because the number of teams that need to be accurately ordered
+differs based on the tournament organizers, a second weighting function
+was found to portray the importance, or remarkability, of each value in
+\(\hat{r}\). We call these ``remarkable weights'' (\(m(r)\)) to
+differentiate from the other weighting function. ``Remarkable weights''
+are set to allow for a subjective belief of what determines a good
+tournament. For a four-team tournament, if the user decides that only
+the winner is important, then \(m(r) = \{1,0,0,0\}\). However, if the
+organizers want an accurate leaderboard with the winner as the most
+important, then a potential \(m(r)\) could be \(\{3,1,1,0\}\). With the
+addition of the weighting functions, the formula for mutual information
+is updated:
+
+\[
+I(X;Y)=\sum_{r}\sum_{\hat{r}} w(r,\hat{r}) * m(r) * p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
+\]
+
+Because this mutual information is unitless, the formula can be
+standardized to be between 0 and 1 using the complement of Rajski's
+distance: \[
+1 - (1 - \frac{I(X,Y)}{\text{max\{}H(X), H(Y)\}})
+\]
+
+where \(H(X,Y)\) is entropy given by:
+
+\[
+max\{H(X), H(Y)\} = \sum_{r}  p(r) \log p(r) = n! \log n!
+\]
+
+The complement is taken so that the more accurate tournaments are closer
+to 1 and the less accurate are closer to 0.
+
+\section{Results}\label{results}
+
+In order to fairly compare tournament structures, we assume the true
+strengths of the teams follow a normal distribution, where each team is
+assigned an equally spaced quantile and each team's strength is the
+z-score of its respective quantile. Using this assumption, 10,000
+simulations, all with 8 total teams, were ran for a single game round
+robin (each team plays each other once), a single elimination structure
+with the usual seeding structure (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7), a
+single elimination structure with a poor seeding structure (1 vs 2, 3 vs
+4, 5 vs 6, 7 vs 8). For a better understanding of the results, the best
+possible results (every structure has the best in rank 1, the 2nd best
+in rank 2, etc), the worst possible results (every structure has the
+worst in rank 1, the 2nd worst in rank 2, etc), every possible
+permutation of the ranks once, and a single elimination tournament with
+the usual seeding structure where every team is of equal strengths.
+Because the metric found is unitless, we were able to normalize the
+values using the best and worst results to make the range from 0 to 1,
+with 0 as the worst and 1 as the best. After normalizing, we get the
+following graph (note the y-axis is on a log10 scale):
+
+\includegraphics[width=4.25in,height=\textheight,keepaspectratio]{images/Normal_Information_Curve_Graph.png}
+
+Based on the graph, the intuitive best tournament, round robin, does
+show to be the best. Double elimination is slightly better than single
+elimination, with the gap growing wider as the top number of teams
+looked at increases. However, those are under the assumption that the
+true strengths were ordered correctly in the seeding. If there is a poor
+seeding structure, as seen in the graph, it can drastically impact the
+effectiveness of the tournament. In the extreme case seen, the
+tournament can actually become worse than deciding games by random
+chance (like flipping a coin). Another important discovery is that the
+single elimination structure with equal strengths is better than every
+possible permutation. The line for every possible permutation would
+likely be similar to a round robin structure with all equal strengths,
+so perhaps the closer the teams are in true strength, the closer round
+robin and single elimination become.
+
 \section{Conclusion}\label{sec:conclusion}
+
+This study introduced a weighted mutual information framework to
+evaluate the effectiveness of different tournament formats in preserving
+the true underlying rankings of teams. By simulating tournaments with
+varying seeding structures and applying an information-theoretic
+approach, we quantified how accurately each format conveys ranking
+information. Our findings confirm the intuitive advantage of round robin
+tournaments, which consistently outperformed single and double
+elimination formats in ranking accuracy. Additionally, we observed that
+seeding plays a critical role: poor seeding structures can significantly
+degrade performance, in some cases making outcomes less informative than
+random chance.
+
+These results underscore the trade-off between tournament accuracy and
+practical constraints such as time, cost, and entertainment value. While
+round robin offers the highest fidelity to true rankings, it is often
+impractical for large tournaments, highlighting the need for hybrid or
+adaptive structures that balance accuracy with logistical feasibility.
+
+Future work should extend this analysis to scenarios with more
+tournament structures, more teams, incorporate probabilistic strength
+models reflecting real-world uncertainty, and explore alternative
+weighting schemes for different competitive priorities (e.g., top-4
+accuracy versus full ranking). Applying these methods to actual
+tournament data could further validate their usefulness for organizers
+aiming to design fair and informative competitions.
 
@@
 
+We thank the Department of Mathematics and Statistics at Loyola
+University Chicago for their support and resources in conducting this
+study. No external funding was received for this research.
+
 \section*{Supplementary Material}\label{supplementary-material}
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
+All supplementary material available at
+\url{https://github.com/gjm112/tournaments}.
 
-\phantomsection\label{refs}
-\begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-Popkin2023}
-Popkin, Charles A., Cole R. Morrissette, Thomas A. Fortney, Kyle L.
-McCormick, Prakash Gorroochurn, and Michael J. Stuart. 2023. {``Fighting
-and Penalty Minutes Associated with Long-Term Mortality Among National
-Hockey League Players, 1967 to 2022.''} \emph{JAMA Network Open} 6 (5):
-e2311308--8.
+\section{References}\label{references}
 
-\end{CSLReferences}
+Rajski, C. (1961). ``A metric space of discrete probability
+distributions''. Information and Control. 4~(4):~371--377.
+\url{doi:10.1016/S0019-9958(61)80055-7}.
+
+McShane, R. (2019). \emph{Modeling Stochastically Intransitive
+Relationships in Paired Comparison Data.} (Doctoral dissertation).
+Southern Methodist University. Retried from
+\url{https://scholar.smu.edu/hum_sci_statisticalscience_etds/13/}
+
+Karel Devriesere, László Csató, Dries Goossens (2024). ``Tournament
+design: A review from an operational research perspective''. European
+Journal of Operational Research.
+\url{https://doi.org/10.1016/j.ejor.2024.10.044}.
 

### 2025-09-04 manuscript/paper.Rmd
Update paper.Rmd

@@
   | \large Ryan McShane \vspace{-1.1mm}
+  |	\normalsize  \vspace{-1mm}
   | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
@@
   | \large Gregory J. Matthews \vspace{-1.1mm}
+  |	\normalsize  \vspace{-1mm}
   | \normalsize Department of Mathematics and Statistics \vspace{-1mm}

### 2025-09-04 manuscript/paper.Rmd
bib update.

--- 2025-08-29 manuscript/paper.Rmd
@@
 title: |
-  |  \large Tournaments 
+  |  \large Assessing Ranking Fidelity of Tournament Formats via Weighted Mutual Information
 author: | 
   | \large Zach Culp \vspace{-1.1mm}
-  | \normalsize  \vspace{-1mm}
+  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
   | \normalsize Loyola University Chicago \vspace{-1mm}
@@
   | \large Josie Peterburs \vspace{-1.1mm}
-  | \normalsize \vspace{-1mm}
+  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
   | \normalsize Loyola University Chicago \vspace{-1mm}
@@
   | \large Ryan McShane \vspace{-1.1mm}
-  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
-  | \normalsize Loyola University Chicago \vspace{-1mm}
+  | \normalsize University of Chicago \vspace{-1mm}
   | \normalsize Chicago, IL 60660 \vspace{-1mm}
@@
 # Introduction
+
+@Rajski1961
+@Goosens2024
+@McShane2019
 
@@
 # References
-
-Rajski, C. (1961). "A metric space of discrete probability
-distributions". Information and Control. 4 (4): 371–377.
-[doi:10.1016/S0019-9958(61)80055-7](doi:10.1016/S0019-9958(61)80055-7){.uri}.
-
-McShane, R. (2019). *Modeling Stochastically Intransitive Relationships
-in Paired Comparison Data.* (Doctoral dissertation). Southern Methodist
-University. Retried from
-<https://scholar.smu.edu/hum_sci_statisticalscience_etds/13/>
-
-Karel Devriesere, László Csató, Dries Goossens (2024). "Tournament
-design: A review from an operational research perspective". European
-Journal of Operational Research.
-<https://doi.org/10.1016/j.ejor.2024.10.044>.

### 2025-09-05 manuscript/paper.Rmd
Updated paper and bib file.

@@
 title: |
-  |  \large Assessing Ranking Fidelity of Tournament Formats via Weighted Mutual Information
+  | \large Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information
 author: | 
@@
 abstract: | 
- | We evaluate multiple tournament formats to assess their effectiveness in accurately reproducing the true underlying ranking of participating teams.
+ | We evaluate multiple competitions structures to assess their effectiveness in accurately reproducing the true underlying ranking of participating teams.
 Using repeated simulations for each structure, we compare teams’ true performance ranks to their simulated outcomes.
 To quantify the accuracy of each format, we calculate weighted mutual information between true and simulated rankings across a large number of replicates.
-"Remarkable" weights are applied to look at only the top n teams/competitors in a tournament.
+"Remarkable" weights are applied to look at only the top n teams/competitors in a given competition structure.
 This information-theoretic approach allows us to objectively compare formats and identify which structures most effectively preserve rank order.
-Results are visualized through comparative plots, providing clear insights into the trade-offs and strengths of each tournament design.
+Results are visualized through comparative plots, providing clear insights into the trade-offs and strengths of each competition design.
  | 
@@
 
-@Rajski1961
-@Goosens2024
-@McShane2019
+<!-- @Rajski1961 -->
+<!-- @McShane2019 -->
+<!-- Csato 2020: The incentive (in)compatibility of group-based qualification systems. World cup qualifying has problems where teams are better off exerting lower effort -->
 
-There are hundreds of different tournament structures that dictate how a
-set of teams compete against each other to determine an overall ranking.
-Different tournament structures have unique strengths and weaknesses,
-affecting how well they reflect the true rankings of the teams. The goal
-of a tournament is to find the best teams, but how can we quantify how
-well a tournament performs? In some instances, the tournament organizers
-only award the overall winner, but other times, the top three or ten
-teams are awarded. Because of this, organizers may prefer one structure
-over another based on their needs. Tournament organizers must also take
-into account factors like cost, timeliness, entertainment value, and
-fairness when choosing a tournament structure.
+<!-- @Sziklai2022 -->
+
+There are hundreds of different competitions structures that dictate how a set of teams compete against each other to determine an overall ranking.
+Common structures include bracket tournaments where if a team wins their competition they advance to the next round.
+Bracket tournaments can be run as single elimination where competitors advance or are eliminated based on a single match (e.g.
+NCAA College Basketball Tournament, National Football League (NFL) playoffs, major tennis tournaments) or where each round is contested as a best of $k$ series (e.g.
+$k$ = 7: National Basketball Association (NBA) playoffs, National Hockey League (NHL) playoffs, $k$ varies by round: Major League Baseball (MLB), Women's National Basketball Association (WNBA)) and the series winner advance while the series loser is eliminated.
+Additional styles of the bracket tournaments include variations where competitors are not eliminated after a single loss such as double elimination tournaments where competitors continue until they have lost twice.
+Within double elimination tournament, there are several variations.
+For instance, the NCAA College Wrestling tournament is a double elimination tournament, however, if a competitor loses a match they fall into the consolation bracket and the best they can finish after a loss before the finals is 3rd place making this not a "true" double elimination format.
+Olympic judo uses a variant of this system with single elimination until the quarterfinals followed by a consolation bracket consisting of the four quarterfinal losers who still have a chance for a bronze medal (in Olympic judo they award two bronze medals to the two  competitors who each win their side of the consolation bracket).
+
+"True" double elimination - a bracket tournament where a competitor can still finish ranked first as long as a competitor has not lost two games - is used in NCAA baseball and softball for certain rounds in their postseason.
+Additional variants of double elimination include double elimination brackets with a "true third" match which includes an additional match contested between the competitors ranked 2nd (loser of the finals) and 3rd (winner of the consolation bracket) at the end of the tournament *if they have not already competed against each other*.
+There is also a variation of double elimination where only those competitors who have lost to a competitor that makes the finals are placed into the consolation bracket and given the chance to compete for third place.
+This style of competition is used in Olympic wrestling (both freestyle and Greco-Roman).
+As in Olympic judo, Olympic wrestling also award two bronze medals to the winners of each side of the consolation bracket.
+
+<!-- A paragraph about Swiss, Dutch, and Monrad systems -->
+
+In bracket style competitions, only a handful of the match-ups that are possible actually occur.
+When every possible competitive pairing occurs, this is referred to as a round-robin structure where every competitor plays every other competitor the same number of times.
+Single round-robin, where was possible match-up occurs exactly once, is used in, for examples, the group stage of the FIFA World Cup to create a ranking to determine which teams advance to the next round.
+Double round-robin, where each possible pairing occurs twice, is used in association football in English Premier League (EPL) and in chess during the FIDE candidates tournament (a tournament that decides who will challenge the world champion).
+
+For the sake of brevity, we review only a small number of possible competition structures, but the interested reader can find a more complete list of competition structures in @Goosens2024 and @Csato2021.
+
+A natural question of interest after reviewing different types of competition structures is: which tournament structure is best?
+
+Ideally, the final rankings produced by a competition structure should closely match the true rankings of the strength of the competitors.
+
+Different competition structures have unique strengths and weaknesses, affecting how well they reflect the true rankings of the teams.
+The goal of a tournament is to find the best teams, but how can we quantify how well a tournament performs?
+In some instances, the tournament organizers only award the overall winner, but other times, the top three or ten teams are awarded.
+Because of this, organizers may prefer one structure over another based on their needs.
+Tournament organizers must also take into account factors like cost, timeliness, entertainment value, and fairness when choosing a tournament structure.
 

### 2025-09-05 manuscript/paper.Rmd
updated.

@@
 For instance, the NCAA College Wrestling tournament is a double elimination tournament, however, if a competitor loses a match they fall into the consolation bracket and the best they can finish after a loss before the finals is 3rd place making this not a "true" double elimination format.
-Olympic judo uses a variant of this system with single elimination until the quarterfinals followed by a consolation bracket consisting of the four quarterfinal losers who still have a chance for a bronze medal (in Olympic judo they award two bronze medals to the two  competitors who each win their side of the consolation bracket).
+Olympic judo uses a variant of this system with single elimination until the quarterfinals followed by a consolation bracket consisting of the four quarterfinal losers who still have a chance for a bronze medal^[in Olympic judo they award two bronze medals to the two  competitors who each win their side of the consolation bracket].
 
@@
 There is also a variation of double elimination where only those competitors who have lost to a competitor that makes the finals are placed into the consolation bracket and given the chance to compete for third place.
-This style of competition is used in Olympic wrestling (both freestyle and Greco-Roman).
-As in Olympic judo, Olympic wrestling also award two bronze medals to the winners of each side of the consolation bracket.
+This style of competition is used in Olympic wrestling (both freestyle and Greco-Roman)^[As in Olympic judo, Olympic wrestling also award two bronze medals to the winners of each side of the consolation bracket].
 
@@
 Double round-robin, where each possible pairing occurs twice, is used in association football in English Premier League (EPL) and in chess during the FIDE candidates tournament (a tournament that decides who will challenge the world champion).
+For the sake of brevity, we review only a small number of possible competition structures, but the interested reader can find a more complete list of competition structures in @Goosens2024, @Appleton1995-zn and @Csato2021.
 
-For the sake of brevity, we review only a small number of possible competition structures, but the interested reader can find a more complete list of competition structures in @Goosens2024 and @Csato2021.
+A natural question of interest after reviewing different types of competition structures is: which competition structure is best?
+As with most complicated questions, the answer is "it depends".
+One commonsense conceptualization for the best competition structure is that the final rankings produced by a competition structure should closely match the true rankings of the strength of the competitors.
+That is, a good competition structure should rank the true best team first, the true second best team second, and so on, until finally ranking the true worst team last.
+We refer to this property of a competition structure as the efficacy of the structure (@Lasek2018-de, @Sziklai2022-kf).
+If one were only interested in correctly ranking the true best competitor as first, with the ranks of the remaining competitors viewed as inconsequential, we refer to this here are effectivity (@Glenn1960-mn).
+While there are many competition structures where efficacy and effectivity are useful measures^[efficacy is a useful measure for the EPL, for example, as the ranking of the teams matters at both the high end (i.e.
+entry into special tournaments) and the low end (i.e.
+relegation status) and effectivity is a useful measure for the NCAA College Basketball tournament (i.e.
+main goal is to crown a champion)], there are many other competitions where the correct ranking of more than one competitor, but not all the competitors is of import.
+For instance, in Olympic competition, the goal of a competition structure is to correctly rank the top 3 competitors to award gold, silver, and bronze medals to^[Some sports award two bronze medals (e.g.
+wrestling, judo, boxing, tae kwon do)].
+In many American sports, for instance the NFL, the goal of the competition structure of the regular season is to rank the top 7 teams in each conference to choose the playoff teams with the ranking of the teams 8 through 16 in each conference inconsequential in determining a Super Bowl champion^[The ranking of teams outside of the playoffs does matter for draft purposes though].
+We coin the term "salient ranks" to refer to the ranks of a competition structure that matter.
+That is the ranks 1, 2, and 3 in Olympic competition are the salient ranks in that competition structure.
+In this notation, efficacy is a measure with with salient ranks 1 through $n$ where $n$ is the number of competitors, and effectivity is a measure where 1 is the only salient rank.
+<!-- Efficacy: @Lasek2018-de, @Sziklai2022-kf -->
+<!-- Effectivity: @Glenn1960-mn -->
 
-A natural question of interest after reviewing different types of competition structures is: which tournament structure is best?
+<!-- From Goosens 2024: The ability of a tournament to rank contestants according to their true strength is called efficacy (Lasek and Gagolewski, -->
+<!-- 2018; Sziklai et al., 2022). Efficacy with respect to only the best player is also known as effectivity (Glenn, 1960) -->
+<!-- or predictive power (Ryvkin and Ortmann, 2008; Vu and Shoham, 2011). -->
 
-Ideally, the final rankings produced by a competition structure should closely match the true rankings of the strength of the competitors.
+If the only goal of a competition structure was high efficacy of the salient ranks, the best way to achieve this would be to play each pair of possible match ups of competitors as many times as possible.
+Of course, this is impossible due to many factors including time constraints, cost constraints, and wear and tear on competitors in high impact sports (e.g.
+boxing, American football, etc.).
+In addition, competitions where the true best team is guaranteed or nearly guaranteed to win may be less appealing to fans and may drive down demand as discussed in @Johnson2022-wz.
+
+In light of this, many different competition structures have been proposed, several of which are mentioned previously.
+Given these many competition structures, ideally there would be a way comparing the performance of different competition structures to each other.
+
+Here, we propose a novel metric for evaluating different competition structures.
+
+<!-- @Marchand2002-tq only cares about the probability of winning in seeded vs random draw knockout tournemnts.   -->
+
+<!-- @Johnson2022-wz: Match outcome uncertainty and the Rottenberg hypothesis -->
 
@@
 
-![](images/Normal_Information_Curve_Graph.png){width="408"}
+![Inofrmartion Curves \label{fig:fig1}](images/Normal_Information_Curve_Graph.png){width="408"}
+
+Figure 1 \ref{fig:fig1}
 
@@
 We thank the Department of Mathematics and Statistics at Loyola
-University Chicago for their support and resources in conducting this
-study. No external funding was received for this research.
+University Chicago for their support and resources in conducting this study.
+We thank Kailey Marie Lum for the suggestion of the term "salient ranks".
+No external funding was received for this research.
 

### 2025-09-05 manuscript/paper.Rmd
ueahrsfdhdfgj

@@
 To quantify the accuracy of each format, we calculate weighted mutual information between true and simulated rankings across a large number of replicates.
-"Remarkable" weights are applied to look at only the top n teams/competitors in a given competition structure.
+"Salient" weights are applied to look at only the top n teams/competitors in a given competition structure.
 This information-theoretic approach allows us to objectively compare formats and identify which structures most effectively preserve rank order.
@@
 
-Here, we propose a novel metric for evaluating different competition structures.
+Here, we propose a novel metric for evaluating different competition structures using an information theoretic approach.
+As an analogy, we view the results of a tournament as a transmitted message, and we use information theory to assess how accurately the message (i.e.
+the tournament results) matches the true transmission (i.e.
+the true ranking of competitors).
+
+Specifically, we use weighted mutual information with a weights based on squared error loss^[other loss functions can easily be incorporated into our method] between the ranks generated by a competition structure and the true ranks of the competitors.
+Additionally, we include an indicator vector to specify the salient ranks, allowing us to compare different competition structures with different choices for salient ranks.
+
+The remainder of this manuscript presents the mathematical foundation of the proposed measure followed by a simulation study estimating out proposed measure for different types of competition structures with varying choices of salient ranks.
+We close by offering a summary of the work and ideas for future directions.
 
@@
 
-Different competition structures have unique strengths and weaknesses, affecting how well they reflect the true rankings of the teams.
+<!-- Different competition structures have unique strengths and weaknesses, affecting how well they reflect the true rankings of the teams.
 The goal of a tournament is to find the best teams, but how can we quantify how well a tournament performs?
@@
 Tournament organizers must also take into account factors like cost, timeliness, entertainment value, and fairness when choosing a tournament structure.
+-->
 
-To evaluate a tournament’s effectiveness, we propose a numerical metric
-that quantifies how accurately it orders teams based on their true
-rankings. Tournament results can be viewed as a "message" attempting to
-convey the true team rankings. However, various factors introduce noise,
-leading to information loss. We use principles from information theory
-to measure this information loss across multiple tournament simulations
-to assess the reliability of different formats.
+<!-- To evaluate a tournament’s effectiveness, we propose a numerical metric that quantifies how accurately it orders teams based on their true rankings.
+Tournament results can be viewed as a "message" attempting to convey the true team rankings.
+However, various factors introduce noise, leading to information loss.
+We use principles from information theory to measure this information loss across multiple tournament simulations to assess the reliability of different formats.
+-->
 
@@
 
-Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with
-an associated true strength parameter
-$\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that
-$(\theta_1 > \theta_2 > \ldots > \theta_n)$. Next, let
-$r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when
-theta is sorted from largest to smallest so that
+Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 > \theta_2 > \ldots > \theta_n)$.
+Next, let
+$r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when $\theta$ is sorted from largest to smallest so that
 $r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}$. Next define
-$T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure
-$s$. $\phi$ contains all the information about the scheduling of teams,
-which could be fully known prior to the tournament (i.e. round robin) or
-determined as the tournament progresses (i.e. single elimination
-tournament). We then let $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an
-$n$- dimensional vector valued random variable that gives the results of
-a tournament as a vector of the indexes of the vector
-$\boldsymbol{\theta}$. The index in the first position of the vector
-$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
-team that finished first, the index in the second position of the vector
-$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
-team that finished second, and so on.
+$T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure $s$.
+$\phi$ contains all the information about the scheduling of teams, which could be fully known prior to the tournament (*e.g.* round robin) or determined as the tournament progresses (*e.g.* single elimination tournament).
+We then let $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an $n$-dimensional vector-valued random variable that gives the results of a tournament as a vector of the indexes of the vector $\boldsymbol{\theta}$.
+The index in the first position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished first, the index in the second position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished second, and so on.
 
-For example, in a 4 team tournament,
-$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that
-the team with index 3 (i.e. the true third best team) won the
-tournament, the true second best team finished second, the true best
-team finished third and the true 4th team finished 4th in that
-particular tournament. If $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ =
-(1, 2, 3, 4), this means that the random outcome of the tournament was
-the same as the true ordering of the teams.
+For example, in a 4 team tournament, $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that the team with index 3 (i.e. the true third best team) won the
+tournament, the true second best team finished second, the true best team finished third and the true 4th team finished 4th in that particular tournament.
+If $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (1, 2, 3, 4), this means that the random outcome of the tournament was the same as the true ordering of the teams.
 
-If one views the true ranking of the teams in a tournament as a message
-to be sent to a receiver and the outcome of the tournament as a message
-that is received, we can measure the "goodness" of a tournament in terms
-of its ability to accurately transmit the true ranking. We can then use
-concepts from information theory to assess the ability of a tournament
-to correctly rank teams. Specifically, we start with the concept of
-mutual information (CITE). For two random variables $X$ and $Y$ mutual
+If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the "goodness" of a tournament in terms of its ability to accurately transmit the true ranking.
+We can then use concepts from information theory to assess the ability of a tournament to correctly rank teams.
+Specifically, we start with the concept of mutual information (@Guiasu1977-nu).
+For two random variables $X$ and $Y$ mutual
 information is defined as:
@@
 
-In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and
-$\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate
-$I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$. For
-simplicity, we drop the arguments from the functions $r$ and $\hat{r}$
-for ease of exposition.
+In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and $\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate
+$I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$. For simplicity, we drop the arguments from the functions $r$ and $\hat{r}$ for ease of exposition.
 
 We note here that $r$ is not a random variable, however, since the
-indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary
-(we use 1, 2, 3, etc. for convenience, but any indexing is valid), we
-can view this a random variable where all permutation so of the $n$
-teams are equally likely. Therefore when calculating the mutual
-information in this setting, we set $p(r) = \frac{1}{n!}$. By a similar
-argument we set $p(\hat{r}) = \frac{1}{n!}$.
+indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary (we use 1, 2, 3, etc.
+for convenience, but any indexing is valid), we can view this a random variable where all permutation so of the $n$ teams are equally likely.
+Therefore when calculating the mutual information in this setting, we set $p(r) = \frac{1}{n!}$.
+By a similar argument we set $p(\hat{r}) = \frac{1}{n!}$.
 
@@
 
-While there are $n!$ different permutations for the result of $r$, we
-don't need to sum across these as any specific choice of $r$ is just an
-arbitrary labeling of the true strength parameters vector
-$\boldsymbol{\theta}$. So given $r$, the distribution of $\hat{r}$ is
-the same up to the labeling. Therefore, we only need to consider a
-single permutation of $r$, we compute the the quantity inside the
-summation, and then multiple by $n!$ (effectively summing across $r$).
+While there are $n!$ different permutations for the result of $r$, we don't need to sum across these as any specific choice of $r$ is just an arbitrary labeling of the true strength parameters vector
+$\boldsymbol{\theta}$.
+So given $r$, the distribution of $\hat{r}$ is the same up to the labeling.
+Therefore, we only need to consider a single permutation of $r$, we compute the the quantity inside the summation, and then multiple by $n!$ (effectively summing across $r$).
 
 In order to compute this quantity, we need to compute $p(r,\hat{r})$.
-This probability is found by calculating the probability of a given
-permutation of $\hat{r}$ and $r$ is assumed to be the permutation from
-$\{1, 2, \ldots, n\}$. We estimate $p(r,\hat{r})$ empirically through
-simulation.
+This probability is found by calculating the probability of a given permutation of $\hat{r}$ and $r$ is assumed to be the permutation from $\{1, 2, \ldots, n\}$.
+We estimate $p(r,\hat{r})$ empirically through simulation.
 
-However, using mutual information in this form does not suit our needs
-in this setting. The problem is that mutual information in this form
-will yield high values of mutual information when the output from the
-tournament $\hat{r}$ is highly consistent *even if the ranking from the
-tournament is incorrect*. As an example, if $n=4$ and the true order of
-$\boldsymbol{\theta}$ is $r = \{1, 2, 3, 4\}$ and
-$p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as
-when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in
-both of these cases the mutual information will be maximized at:
+However, using mutual information in this form does not suit our needs in this setting.
+The problem is that mutual information in this form will yield high values of mutual information when the output from the tournament $\hat{r}$ is highly consistent *even if the ranking from the tournament is incorrect*.
+As an example, if $n=4$ and the true order of $\boldsymbol{\theta}$ is $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in both of these cases the mutual information will be maximized at:
 
@@
 In order to alleviate this problem, we instead consider *weighted*
-mutual information. We want to give more weight to permutations from
-$\hat{r}$ that are "closer" to $r$. There are any number of reasonable
-choices for this weighting function, and we choose to use squared error.
-Specifically,
+mutual information (@Guiasu1977-nu). We want to give more weight to permutations from $\hat{r}$ that are "closer" to $r$.
 
-$$
-w(r, \hat{r}) =
-\begin{cases}
-\frac{1}{\hat{r}'r}, & r \ne \hat{r} \\
-1, & r = \hat{r}
-\end{cases}
-$$
-
-In general, though any loss function $l(r, \hat{r})$ can be used:
+In general, any loss function $l(r, \hat{r})$ can be used to define the weigting function:
 
@@
 
-Note that one is 1/N!
+Here though, we choose to use squared error loss:
 
-And this does work because consistently getting the order incorrect is
-perfect information.
+$$
+w(r, \hat{r}) =
+\begin{cases}
+\frac{1}{\hat{r}'r}, & r \ne \hat{r} \\
+1, & r = \hat{r}
+\end{cases}
+$$
+<!-- Note that one is 1/N! -->
 
-However, because the number of teams that need to be accurately ordered
-differs based on the tournament organizers, a second weighting function
-was found to portray the importance, or remarkability, of each value in
-$\hat{r}$. We call these "remarkable weights" ($m(r)$) to differentiate
-from the other weighting function. "Remarkable weights" are set to allow
-for a subjective belief of what determines a good tournament. For a
-four-team tournament, if the user decides that only the winner is
-important, then $m(r) = \{1,0,0,0\}$. However, if the organizers want an
-accurate leaderboard with the winner as the most important, then a
-potential $m(r)$ could be $\{3,1,1,0\}$. With the addition of the
-weighting functions, the formula for mutual information is updated:
+<!-- And this does work because consistently getting the order incorrect is perfect information. -->
+
+However, because the number of teams that need to be accurately ordered differs based on the choice of salient ranks, a second weighting function, $m(r)$, is used to indicate which ranks are salient.
+We call these the "salient weights" to differentiate from the other weighting function.
+As an example, for a four-team tournament, if the organizer decides that only the winner is important, then $m(r) = \{1,0,0,0\}$ (i.e.
+essentially measuring effectivity).
+While we only consider indicators (i.e.
+0's or 1's) in the salient weights vector, there is no reason why you couldn't weight the ranks by relative levels of importance.
+For instance, letting $m(r) = \{3,1,0,0\}$ indicates that the salient ranks are 1 and 2, but getting the true best competitor as the winner is more important than ranking the true second best competitor second.
+
+With the addition of the weighting functions, the formula for mutual information becomes:
 
@@
 standardized to be between 0 and 1 using the complement of Rajski's
-distance: $$
+distance (@Rajski1961): $$
 1 - (1 - \frac{I(X,Y)}{\text{max\{}H(X), H(Y)\}})

### 2025-09-05 manuscript/paper.Rmd
updated.

@@
 To quantify the accuracy of each format, we calculate weighted mutual information between true and simulated rankings across a large number of replicates.
-"Salient" weights are applied to look at only the top n teams/competitors in a given competition structure.
+"Salient" weights are applied to look at only the top i out of n total teams/competitors in a given competition structure.
 This information-theoretic approach allows us to objectively compare formats and identify which structures most effectively preserve rank order.
@@
 
-For example, in a 4 team tournament, $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that the team with index 3 (i.e. the true third best team) won the
+For example, in a $n$ = 4 team tournament, $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that the team with index 3 (i.e. the true third best team) won the
 tournament, the true second best team finished second, the true best team finished third and the true 4th team finished 4th in that particular tournament.
@@
 
-We note here that $r$ is not a random variable, however, since the
-indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary (we use 1, 2, 3, etc.
+We note here that $r$ is not a random variable, however, since the indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary (we use 1, 2, 3, etc.
 for convenience, but any indexing is valid), we can view this a random variable where all permutation so of the $n$ teams are equally likely.
@@
 
-While there are $n!$ different permutations for the result of $r$, we don't need to sum across these as any specific choice of $r$ is just an arbitrary labeling of the true strength parameters vector
-$\boldsymbol{\theta}$.
+While there are $n!$ different permutations for the result of $r$, we don't need to sum across these as any specific choice of $r$ is just an arbitrary labeling of the true strength parameters vector $\boldsymbol{\theta}$.
 So given $r$, the distribution of $\hat{r}$ is the same up to the labeling.
@@
 
-In order to alleviate this problem, we instead consider *weighted*
-mutual information (@Guiasu1977-nu). We want to give more weight to permutations from $\hat{r}$ that are "closer" to $r$.
+In order to alleviate this problem, we instead consider *weighted* mutual information (@Guiasu1977-nu). We want to give more weight to permutations from $\hat{r}$ that are "closer" to $r$.
 
@@
 
-However, because the number of teams that need to be accurately ordered differs based on the choice of salient ranks, a second weighting function, $m(r)$, is used to indicate which ranks are salient.
+However, because the number of teams that need to be accurately ordered differs based on the choice of salient ranks, a second set of weights is defined, $m$, is used to indicate which ranks are salient.
 We call these the "salient weights" to differentiate from the other weighting function.
-As an example, for a four-team tournament, if the organizer decides that only the winner is important, then $m(r) = \{1,0,0,0\}$ (i.e.
+As an example, for a four-team tournament, if the organizer decides that only the winner is important, then $m = \{1,0,0,0\}$ (i.e.
 essentially measuring effectivity).
-While we only consider indicators (i.e.
+While here we only consider indicators (i.e.
 0's or 1's) in the salient weights vector, there is no reason why you couldn't weight the ranks by relative levels of importance.
-For instance, letting $m(r) = \{3,1,0,0\}$ indicates that the salient ranks are 1 and 2, but getting the true best competitor as the winner is more important than ranking the true second best competitor second.
+For instance, letting $m = \{3,1,0,0\}$ indicates that the salient ranks are 1 and 2, but getting the true best competitor as the winner is more important than ranking the true second best competitor second.
 
@@
 $$
-I(X;Y)=\sum_{r}\sum_{\hat{r}} w(r,\hat{r}) * m(r) * p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
+wI_m(r, \hat{r})=\sum_{r}\sum_{\hat{r}} w(r,\hat{r}) * m * p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
 $$
 
-Because this mutual information is unitless, the formula can be
-standardized to be between 0 and 1 using the complement of Rajski's
-distance (@Rajski1961): $$
-1 - (1 - \frac{I(X,Y)}{\text{max\{}H(X), H(Y)\}})
+Because this mutual information is unitless, the formula can be standardized to be between 0 and 1 as follows: 
+$$
+\frac{wI_m(r,\hat{r})}{\text{max\{}H(r), H(\hat{r})\}}
 $$
 
-where $H(X,Y)$ is entropy given by:
+where $H(.)$ is the usual Shannon entropy.
+Competition structures with higher levels of efficacy are indicated when our measure is closer to 1.
+Likewise, competition structures that are near 0 by our proposed metric indicate low levels of efficacy.
 
-$$
-max\{H(X), H(Y)\} = \sum_{r}  p(r) \log p(r) = n! \log n!
-$$
-
-The complement is taken so that the more accurate tournaments are closer
-to 1 and the less accurate are closer to 0.
+<!-- $$ -->
+<!-- max\{H(X), H(Y)\} = \sum_{r}  p(r) \log p(r) = n! \log n! -->
+<!-- $$ -->
 
@@
 
-In order to fairly compare tournament structures, we assume the true
-strengths of the teams follow a normal distribution, where each team is
-assigned an equally spaced quantile and each team's strength is the
-z-score of its respective quantile. Using this assumption, 10,000
-simulations, all with 8 total teams, were ran for a single game round
-robin (each team plays each other once), a single elimination structure
-with the usual seeding structure (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7), a
-single elimination structure with a poor seeding structure (1 vs 2, 3 vs
-4, 5 vs 6, 7 vs 8). For a better understanding of the results, the best
-possible results (every structure has the best in rank 1, the 2nd best
-in rank 2, etc), the worst possible results (every structure has the
-worst in rank 1, the 2nd worst in rank 2, etc), every possible
-permutation of the ranks once, and a single elimination tournament with
-the usual seeding structure where every team is of equal strengths.
+What's interesting: 
+1. Single and double are really close at 1 or 2, but they separate past 3.  
+2.  The black curve crosses the yellow and green curve. 
+3.  The yellow and black curves actually increases towards the end.
+
+We examined the following structure: 
+ - Single elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
+ - Double elimination
+
+In order to fairly compare tournament structures, we assume the true strengths of the teams follow a normal distribution, where each team is assigned an equally spaced quantile and each team's strength is the z-score of its respective quantile.
+Using this assumption, 10,000 simulations, all with 8 total teams, were ran for a single game round robin (each team plays each other once), a single elimination structure
+with the usual seeding structure (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7), a single elimination structure with a poor seeding structure (1 vs 2, 3 vs 4, 5 vs 6, 7 vs 8).
+For a better understanding of the results, the best possible results (every structure has the best in rank 1, the 2nd best in rank 2, etc), the worst possible results (every structure has the worst in rank 1, the 2nd worst in rank 2, etc), every possible permutation of the ranks once, and a single elimination tournament with the usual seeding structure where every team is of equal strengths.
 Because the metric found is unitless, we were able to normalize the
@@
 
-![Inofrmartion Curves \label{fig:fig1}](images/Normal_Information_Curve_Graph.png){width="408"}
+![Zach add a caption \label{fig:fig1}](images/Normal_Information_Curve_Graph.png){width="408"}
 
-Figure 1 \ref{fig:fig1}
+Figure \ref{fig:fig1}
 
-Based on the graph, the intuitive best tournament, round robin, does
-show to be the best. Double elimination is slightly better than single
+Based on the graph, the intuitive best tournament, round robin, does show to be the best. Double elimination is slightly better than single
 elimination, with the gap growing wider as the top number of teams

### 2025-09-06 manuscript/paper.Rmd
paper edits

@@
 <!-- @Rajski1961 -->
+
 <!-- @McShane2019 -->
+
 <!-- Csato 2020: The incentive (in)compatibility of group-based qualification systems. World cup qualifying has problems where teams are better off exerting lower effort -->
@@
 
-There are hundreds of different competitions structures that dictate how a set of teams compete against each other to determine an overall ranking.
-Common structures include bracket tournaments where if a team wins their competition they advance to the next round.
-Bracket tournaments can be run as single elimination where competitors advance or are eliminated based on a single match (e.g.
-NCAA College Basketball Tournament, National Football League (NFL) playoffs, major tennis tournaments) or where each round is contested as a best of $k$ series (e.g.
-$k$ = 7: National Basketball Association (NBA) playoffs, National Hockey League (NHL) playoffs, $k$ varies by round: Major League Baseball (MLB), Women's National Basketball Association (WNBA)) and the series winner advance while the series loser is eliminated.
-Additional styles of the bracket tournaments include variations where competitors are not eliminated after a single loss such as double elimination tournaments where competitors continue until they have lost twice.
-Within double elimination tournament, there are several variations.
-For instance, the NCAA College Wrestling tournament is a double elimination tournament, however, if a competitor loses a match they fall into the consolation bracket and the best they can finish after a loss before the finals is 3rd place making this not a "true" double elimination format.
-Olympic judo uses a variant of this system with single elimination until the quarterfinals followed by a consolation bracket consisting of the four quarterfinal losers who still have a chance for a bronze medal^[in Olympic judo they award two bronze medals to the two  competitors who each win their side of the consolation bracket].
+There are hundreds of different competitions structures that dictate how
+a set of teams compete against each other to determine an overall
+ranking. Common structures include bracket tournaments where if a team
+wins their competition they advance to the next round. Bracket
+tournaments can be run as single elimination where competitors advance
+or are eliminated based on a single match (e.g. NCAA College Basketball
+Tournament, National Football League (NFL) playoffs, major tennis
+tournaments) or where each round is contested as a best of $k$ series
+(e.g. $k$ = 7: National Basketball Association (NBA) playoffs, National
+Hockey League (NHL) playoffs, $k$ varies by round: Major League Baseball
+(MLB), Women's National Basketball Association (WNBA)) and the series
+winner advance while the series loser is eliminated. Additional styles
+of the bracket tournaments include variations where competitors are not
+eliminated after a single loss such as double elimination tournaments
+where competitors continue until they have lost twice. Within double
+elimination tournament, there are several variations. For instance, the
+NCAA College Wrestling tournament is a double elimination tournament,
+however, if a competitor loses a match they fall into the consolation
+bracket and the best they can finish after a loss before the finals is
+3rd place making this not a "true" double elimination format. Olympic
+judo uses a variant of this system with single elimination until the
+quarterfinals followed by a consolation bracket consisting of the four
+quarterfinal losers who still have a chance for a bronze medal[^1].
 
-"True" double elimination - a bracket tournament where a competitor can still finish ranked first as long as a competitor has not lost two games - is used in NCAA baseball and softball for certain rounds in their postseason.
-Additional variants of double elimination include double elimination brackets with a "true third" match which includes an additional match contested between the competitors ranked 2nd (loser of the finals) and 3rd (winner of the consolation bracket) at the end of the tournament *if they have not already competed against each other*.
-There is also a variation of double elimination where only those competitors who have lost to a competitor that makes the finals are placed into the consolation bracket and given the chance to compete for third place.
-This style of competition is used in Olympic wrestling (both freestyle and Greco-Roman)^[As in Olympic judo, Olympic wrestling also award two bronze medals to the winners of each side of the consolation bracket].
+[^1]: in Olympic judo they award two bronze medals to the two
+    competitors who each win their side of the consolation bracket
+
+"True" double elimination - a bracket tournament where a competitor can
+still finish ranked first as long as a competitor has not lost two
+games - is used in NCAA baseball and softball for certain rounds in
+their postseason. Additional variants of double elimination include
+double elimination brackets with a "true third" match which includes an
+additional match contested between the competitors ranked 2nd (loser of
+the finals) and 3rd (winner of the consolation bracket) at the end of
+the tournament *if they have not already competed against each other*.
+There is also a variation of double elimination where only those
+competitors who have lost to a competitor that makes the finals are
+placed into the consolation bracket and given the chance to compete for
+third place. This style of competition is used in Olympic wrestling
+(both freestyle and Greco-Roman)[^2].
+
+[^2]: As in Olympic judo, Olympic wrestling also award two bronze medals
+    to the winners of each side of the consolation bracket
 
@@
 
-In bracket style competitions, only a handful of the match-ups that are possible actually occur.
-When every possible competitive pairing occurs, this is referred to as a round-robin structure where every competitor plays every other competitor the same number of times.
-Single round-robin, where was possible match-up occurs exactly once, is used in, for examples, the group stage of the FIFA World Cup to create a ranking to determine which teams advance to the next round.
-Double round-robin, where each possible pairing occurs twice, is used in association football in English Premier League (EPL) and in chess during the FIDE candidates tournament (a tournament that decides who will challenge the world champion).
-For the sake of brevity, we review only a small number of possible competition structures, but the interested reader can find a more complete list of competition structures in @Goosens2024, @Appleton1995-zn and @Csato2021.
+In bracket style competitions, only a handful of the match-ups that are
+possible actually occur. When every possible competitive pairing occurs,
+this is referred to as a round-robin structure where every competitor
+plays every other competitor the same number of times. Single
+round-robin, where was possible match-up occurs exactly once, is used
+in, for examples, the group stage of the FIFA World Cup to create a
+ranking to determine which teams advance to the next round. Double
+round-robin, where each possible pairing occurs twice, is used in
+association football in English Premier League (EPL) and in chess during
+the FIDE candidates tournament (a tournament that decides who will
+challenge the world champion). For the sake of brevity, we review only a
+small number of possible competition structures, but the interested
+reader can find a more complete list of competition structures in
+@Goosens2024, @Appleton1995-zn and @Csato2021.
 
-A natural question of interest after reviewing different types of competition structures is: which competition structure is best?
-As with most complicated questions, the answer is "it depends".
-One commonsense conceptualization for the best competition structure is that the final rankings produced by a competition structure should closely match the true rankings of the strength of the competitors.
-That is, a good competition structure should rank the true best team first, the true second best team second, and so on, until finally ranking the true worst team last.
-We refer to this property of a competition structure as the efficacy of the structure (@Lasek2018-de, @Sziklai2022-kf).
-If one were only interested in correctly ranking the true best competitor as first, with the ranks of the remaining competitors viewed as inconsequential, we refer to this here are effectivity (@Glenn1960-mn).
-While there are many competition structures where efficacy and effectivity are useful measures^[efficacy is a useful measure for the EPL, for example, as the ranking of the teams matters at both the high end (i.e.
-entry into special tournaments) and the low end (i.e.
-relegation status) and effectivity is a useful measure for the NCAA College Basketball tournament (i.e.
-main goal is to crown a champion)], there are many other competitions where the correct ranking of more than one competitor, but not all the competitors is of import.
-For instance, in Olympic competition, the goal of a competition structure is to correctly rank the top 3 competitors to award gold, silver, and bronze medals to^[Some sports award two bronze medals (e.g.
-wrestling, judo, boxing, tae kwon do)].
-In many American sports, for instance the NFL, the goal of the competition structure of the regular season is to rank the top 7 teams in each conference to choose the playoff teams with the ranking of the teams 8 through 16 in each conference inconsequential in determining a Super Bowl champion^[The ranking of teams outside of the playoffs does matter for draft purposes though].
-We coin the term "salient ranks" to refer to the ranks of a competition structure that matter.
-That is the ranks 1, 2, and 3 in Olympic competition are the salient ranks in that competition structure.
-In this notation, efficacy is a measure with with salient ranks 1 through $n$ where $n$ is the number of competitors, and effectivity is a measure where 1 is the only salient rank.
+A natural question of interest after reviewing different types of
+competition structures is: which competition structure is best? As with
+most complicated questions, the answer is "it depends". One commonsense
+conceptualization for the best competition structure is that the final
+rankings produced by a competition structure should closely match the
+true rankings of the strength of the competitors. That is, a good
+competition structure should rank the true best team first, the true
+second best team second, and so on, until finally ranking the true worst
+team last. We refer to this property of a competition structure as the
+efficacy of the structure (@Lasek2018-de, @Sziklai2022-kf). If one were
+only interested in correctly ranking the true best competitor as first,
+with the ranks of the remaining competitors viewed as inconsequential,
+we refer to this here are effectivity (@Glenn1960-mn). While there are
+many competition structures where efficacy and effectivity are useful
+measures[^3], there are many other competitions where the correct
+ranking of more than one competitor, but not all the competitors is of
+import. For instance, in Olympic competition, the goal of a competition
+structure is to correctly rank the top 3 competitors to award gold,
+silver, and bronze medals to[^4]. In many American sports, for instance
+the NFL, the goal of the competition structure of the regular season is
+to rank the top 7 teams in each conference to choose the playoff teams
+with the ranking of the teams 8 through 16 in each conference
+inconsequential in determining a Super Bowl champion[^5]. We coin the
+term "salient ranks" to refer to the ranks of a competition structure
+that matter. That is the ranks 1, 2, and 3 in Olympic competition are
+the salient ranks in that competition structure. In this notation,
+efficacy is a measure with with salient ranks 1 through $n$ where $n$ is
+the number of competitors, and effectivity is a measure where 1 is the
+only salient rank.\
 <!-- Efficacy: @Lasek2018-de, @Sziklai2022-kf -->
@@
 
+[^3]: efficacy is a useful measure for the EPL, for example, as the
+    ranking of the teams matters at both the high end (i.e. entry into
+    special tournaments) and the low end (i.e. relegation status) and
+    effectivity is a useful measure for the NCAA College Basketball
+    tournament (i.e. main goal is to crown a champion)
+
+[^4]: Some sports award two bronze medals (e.g. wrestling, judo, boxing,
+    tae kwon do)
+
+[^5]: The ranking of teams outside of the playoffs does matter for draft
+    purposes though
+
 <!-- From Goosens 2024: The ability of a tournament to rank contestants according to their true strength is called efficacy (Lasek and Gagolewski, -->
+
 <!-- 2018; Sziklai et al., 2022). Efficacy with respect to only the best player is also known as effectivity (Glenn, 1960) -->
+
 <!-- or predictive power (Ryvkin and Ortmann, 2008; Vu and Shoham, 2011). -->
 
-If the only goal of a competition structure was high efficacy of the salient ranks, the best way to achieve this would be to play each pair of possible match ups of competitors as many times as possible.
-Of course, this is impossible due to many factors including time constraints, cost constraints, and wear and tear on competitors in high impact sports (e.g.
-boxing, American football, etc.).
-In addition, competitions where the true best team is guaranteed or nearly guaranteed to win may be less appealing to fans and may drive down demand as discussed in @Johnson2022-wz.
+If the only goal of a competition structure was high efficacy of the
+salient ranks, the best way to achieve this would be to play each pair
+of possible match ups of competitors as many times as possible. Of
+course, this is impossible due to many factors including time
+constraints, cost constraints, and wear and tear on competitors in high
+impact sports (e.g. boxing, American football, etc.). In addition,
+competitions where the true best team is guaranteed or nearly guaranteed
+to win may be less appealing to fans and may drive down demand as
+discussed in @Johnson2022-wz.
 
-In light of this, many different competition structures have been proposed, several of which are mentioned previously.
-Given these many competition structures, ideally there would be a way comparing the performance of different competition structures to each other.
+In light of this, many different competition structures have been
+proposed, several of which are mentioned previously. Given these many
+competition structures, ideally there would be a way comparing the
+performance of different competition structures to each other.
 
-Here, we propose a novel metric for evaluating different competition structures using an information theoretic approach.
-As an analogy, we view the results of a tournament as a transmitted message, and we use information theory to assess how accurately the message (i.e.
-the tournament results) matches the true transmission (i.e.
-the true ranking of competitors).
+Here, we propose a novel metric for evaluating different competition
+structures using an information theoretic approach. As an analogy, we
+view the results of a tournament as a transmitted message, and we use
+information theory to assess how accurately the message (i.e. the
+tournament results) matches the true transmission (i.e. the true ranking
+of competitors).
 
-Specifically, we use weighted mutual information with a weights based on squared error loss^[other loss functions can easily be incorporated into our method] between the ranks generated by a competition structure and the true ranks of the competitors.
-Additionally, we include an indicator vector to specify the salient ranks, allowing us to compare different competition structures with different choices for salient ranks.
+Specifically, we use weighted mutual information with a weights based on
+squared error loss[^6] between the ranks generated by a competition
+structure and the true ranks of the competitors. Additionally, we
+include an indicator vector to specify the salient ranks, allowing us to
+compare different competition structures with different choices for
+salient ranks.
 
-The remainder of this manuscript presents the mathematical foundation of the proposed measure followed by a simulation study estimating out proposed measure for different types of competition structures with varying choices of salient ranks.
-We close by offering a summary of the work and ideas for future directions.
+[^6]: other loss functions can easily be incorporated into our method
+
+The remainder of this manuscript presents the mathematical foundation of
+the proposed measure followed by a simulation study estimating out
+proposed measure for different types of competition structures with
+varying choices of salient ranks. We close by offering a summary of the
+work and ideas for future directions.
 
@@
 
-Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 > \theta_2 > \ldots > \theta_n)$.
-Next, let
-$r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when $\theta$ is sorted from largest to smallest so that
+Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with
+an associated true strength parameter
+$\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that
+$(\theta_1 > \theta_2 > \ldots > \theta_n)$. Next, let
+$r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when
+$\theta$ is sorted from largest to smallest so that
 $r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}$. Next define
-$T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure $s$.
-$\phi$ contains all the information about the scheduling of teams, which could be fully known prior to the tournament (*e.g.* round robin) or determined as the tournament progresses (*e.g.* single elimination tournament).
-We then let $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an $n$-dimensional vector-valued random variable that gives the results of a tournament as a vector of the indexes of the vector $\boldsymbol{\theta}$.
-The index in the first position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished first, the index in the second position of the vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the team that finished second, and so on.
+$T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure
+$s$. $\phi$ contains all the information about the scheduling of teams,
+which could be fully known prior to the tournament (*e.g.* round robin)
+or determined as the tournament progresses (*e.g.* single elimination
+tournament). We then let $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an
+$n$-dimensional vector-valued random variable that gives the results of
+a tournament as a vector of the indexes of the vector
+$\boldsymbol{\theta}$. The index in the first position of the vector
+$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
+team that finished first, the index in the second position of the vector
+$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
+team that finished second, and so on.
 
-For example, in a $n$ = 4 team tournament, $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that the team with index 3 (i.e. the true third best team) won the
-tournament, the true second best team finished second, the true best team finished third and the true 4th team finished 4th in that particular tournament.
-If $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (1, 2, 3, 4), this means that the random outcome of the tournament was the same as the true ordering of the teams.
+For example, in a $n$ = 4 team tournament,
+$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that
+the team with index 3 (i.e. the true third best team) won the
+tournament, the true second best team finished second, the true best
+team finished third and the true 4th team finished 4th in that
+particular tournament. If $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ =
+(1, 2, 3, 4), this means that the random outcome of the tournament was
+the same as the true ordering of the teams.
 
-If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the "goodness" of a tournament in terms of its ability to accurately transmit the true ranking.
-We can then use concepts from information theory to assess the ability of a tournament to correctly rank teams.
-Specifically, we start with the concept of mutual information (@Guiasu1977-nu).
-For two random variables $X$ and $Y$ mutual
-information is defined as:
+If one views the true ranking of the teams in a tournament as a message
+to be sent to a receiver and the outcome of the tournament as a message
+that is received, we can measure the "goodness" of a tournament in terms
+of its ability to accurately transmit the true ranking. We can then use
+concepts from information theory to assess the ability of a tournament
+to correctly rank teams. Specifically, we start with the concept of
+mutual information (@Guiasu1977-nu). For two random variables $X$ and
+$Y$ mutual information is defined as:
 
@@
 
-In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and $\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate
-$I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$. For simplicity, we drop the arguments from the functions $r$ and $\hat{r}$ for ease of exposition.
+In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and
+$\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate
+$I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$. For
+simplicity, we drop the arguments from the functions $r$ and $\hat{r}$
+for ease of exposition.
 
-We note here that $r$ is not a random variable, however, since the indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary (we use 1, 2, 3, etc.
-for convenience, but any indexing is valid), we can view this a random variable where all permutation so of the $n$ teams are equally likely.
-Therefore when calculating the mutual information in this setting, we set $p(r) = \frac{1}{n!}$.
-By a similar argument we set $p(\hat{r}) = \frac{1}{n!}$.
+We note here that $r$ is not a random variable, however, since the
+indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary
+(we use 1, 2, 3, etc. for convenience, but any indexing is valid), we
+can view this a random variable where all permutation so of the $n$
+teams are equally likely. Therefore when calculating the mutual
+information in this setting, we set $p(r) = \frac{1}{n!}$. By a similar
+argument we set $p(\hat{r}) = \frac{1}{n!}$.
 
@@
 
-While there are $n!$ different permutations for the result of $r$, we don't need to sum across these as any specific choice of $r$ is just an arbitrary labeling of the true strength parameters vector $\boldsymbol{\theta}$.
-So given $r$, the distribution of $\hat{r}$ is the same up to the labeling.
-Therefore, we only need to consider a single permutation of $r$, we compute the the quantity inside the summation, and then multiple by $n!$ (effectively summing across $r$).
+While there are $n!$ different permutations for the result of $r$, we
+don't need to sum across these as any specific choice of $r$ is just an
+arbitrary labeling of the true strength parameters vector
+$\boldsymbol{\theta}$. So given $r$, the distribution of $\hat{r}$ is
+the same up to the labeling. Therefore, we only need to consider a
+single permutation of $r$, we compute the the quantity inside the
+summation, and then multiple by $n!$ (effectively summing across $r$).
 
 In order to compute this quantity, we need to compute $p(r,\hat{r})$.
-This probability is found by calculating the probability of a given permutation of $\hat{r}$ and $r$ is assumed to be the permutation from $\{1, 2, \ldots, n\}$.
-We estimate $p(r,\hat{r})$ empirically through simulation.
+This probability is found by calculating the probability of a given
+permutation of $\hat{r}$ and $r$ is assumed to be the permutation from
+$\{1, 2, \ldots, n\}$. We estimate $p(r,\hat{r})$ empirically through
+simulation.
 
-However, using mutual information in this form does not suit our needs in this setting.
-The problem is that mutual information in this form will yield high values of mutual information when the output from the tournament $\hat{r}$ is highly consistent *even if the ranking from the tournament is incorrect*.
-As an example, if $n=4$ and the true order of $\boldsymbol{\theta}$ is $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in both of these cases the mutual information will be maximized at:
+However, using mutual information in this form does not suit our needs
+in this setting. The problem is that mutual information in this form
+will yield high values of mutual information when the output from the
+tournament $\hat{r}$ is highly consistent *even if the ranking from the
+tournament is incorrect*. As an example, if $n=4$ and the true order of
+$\boldsymbol{\theta}$ is $r = \{1, 2, 3, 4\}$ and
+$p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as
+when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in
+both of these cases the mutual information will be maximized at:
 
@@
 
-In order to alleviate this problem, we instead consider *weighted* mutual information (@Guiasu1977-nu). We want to give more weight to permutations from $\hat{r}$ that are "closer" to $r$.
+In order to alleviate this problem, we instead consider *weighted*
+mutual information (@Guiasu1977-nu). We want to give more weight to
+permutations from $\hat{r}$ that are "closer" to $r$.
 
-In general, any loss function $l(r, \hat{r})$ can be used to define the weigting function:
+In general, any loss function $l(r, \hat{r})$ can be used to define the
+weigting function:
 
@@
 \end{cases}
-$$
-<!-- Note that one is 1/N! -->
+$$ <!-- Note that one is 1/N! -->
 
@@
 
-However, because the number of teams that need to be accurately ordered differs based on the choice of salient ranks, a second set of weights is defined, $m$, is used to indicate which ranks are salient.
-We call these the "salient weights" to differentiate from the other weighting function.
-As an example, for a four-team tournament, if the organizer decides that only the winner is important, then $m = \{1,0,0,0\}$ (i.e.
-essentially measuring effectivity).
-While here we only consider indicators (i.e.
-0's or 1's) in the salient weights vector, there is no reason why you couldn't weight the ranks by relative levels of importance.
-For instance, letting $m = \{3,1,0,0\}$ indicates that the salient ranks are 1 and 2, but getting the true best competitor as the winner is more important than ranking the true second best competitor second.
+However, because the number of teams that need to be accurately ordered
+differs based on the choice of salient ranks, a second set of weights is
+defined, $m$, is used to indicate which ranks are salient. We call these
+the "salient weights" to differentiate from the other weighting
+function. As an example, for a four-team tournament, if the organizer
+decides that only the winner is important, then $m = \{1,0,0,0\}$ (i.e.
+essentially measuring effectivity). While here we only consider
+indicators (i.e. 0's or 1's) in the salient weights vector, there is no
+reason why you couldn't weight the ranks by relative levels of
+importance. For instance, letting $m = \{3,1,0,0\}$ indicates that the
+salient ranks are 1 and 2, but getting the true best competitor as the
+winner is more important than ranking the true second best competitor
+second.
 
-With the addition of the weighting functions, the formula for mutual information becomes:
+With the addition of the weighting functions, the formula for mutual
+information becomes:
 
@@
 
-Because this mutual information is unitless, the formula can be standardized to be between 0 and 1 as follows: 
-$$
+Because this mutual information is unitless, the formula can be
+standardized to be between 0 and 1 as follows: $$
 \frac{wI_m(r,\hat{r})}{\text{max\{}H(r), H(\hat{r})\}}
@@
 
-where $H(.)$ is the usual Shannon entropy.
-Competition structures with higher levels of efficacy are indicated when our measure is closer to 1.
-Likewise, competition structures that are near 0 by our proposed metric indicate low levels of efficacy.
+where $H(.)$ is the usual Shannon entropy. Competition structures with
+higher levels of efficacy are indicated when our measure is closer to 1.
+Likewise, competition structures that are near 0 by our proposed metric
+indicate low levels of efficacy.
 
 <!-- $$ -->
+
 <!-- max\{H(X), H(Y)\} = \sum_{r}  p(r) \log p(r) = n! \log n! -->
+
 <!-- $$ -->
@@
 
-What's interesting: 
-1. Single and double are really close at 1 or 2, but they separate past 3.  
-2.  The black curve crosses the yellow and green curve. 
-3.  The yellow and black curves actually increases towards the end.
+```{=html}
+<!--
+
+| Structure/Ranks | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 |
+|:----------------------:|:----:|:----:|:----:|:----:|:----:|:----:|:----:|:----:|
+| Double Elimination | 0.4067 | 0.2517 | 0.1443 | 0.0916 | 0.0532 | 0.0390 | 0.0340 | 0.0290 |
+| Single Elimination | 0.4023 | 0.2439 | 0.1338 | 0.0837 | 0.0478 | 0.0356 | 0.0312 | 0.0270 |
+| Equal Strengths Single Elimination | 0.1297 | 0.0474 | 0.0224 | 0.0129 | 0.0072 | 0.0051 | 0.0050 | 0.0050 |
+| Bad Seeding Single Elimination | 0.2679 | 0.0146 | 0.0102 | 0.0059 | 0.0032 | 0.0022 | 0.0020 | 0.0026 |
+| Every Possible Permutation | 0.1036 | 0.0302 | 0.0113 | 0.0047 | 0.0018 | 0.0008 | 0.0011 | 0.0014 |
+| Round Robin | 0.8851 | 0.8473 | 0.8096 | 0.7708 | 0.7337 | 0.7013 | 0.6755 | 0.6697 |
+| Worst Possible | 0 | 0 | 0 | 0 | 0 | 0 | 0 | 0 |
+| Best Possible | 1 | 1 | 1 | 1 | 1 | 1 | 1 | 1 |
+
+-->
+```
 
 We examined the following structure:
+
 -   Single elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
- - Double elimination
+    and normally distributed strengths
 
-In order to fairly compare tournament structures, we assume the true strengths of the teams follow a normal distribution, where each team is assigned an equally spaced quantile and each team's strength is the z-score of its respective quantile.
-Using this assumption, 10,000 simulations, all with 8 total teams, were ran for a single game round robin (each team plays each other once), a single elimination structure
-with the usual seeding structure (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7), a single elimination structure with a poor seeding structure (1 vs 2, 3 vs 4, 5 vs 6, 7 vs 8).
-For a better understanding of the results, the best possible results (every structure has the best in rank 1, the 2nd best in rank 2, etc), the worst possible results (every structure has the worst in rank 1, the 2nd worst in rank 2, etc), every possible permutation of the ranks once, and a single elimination tournament with the usual seeding structure where every team is of equal strengths.
+-   Single elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
+    and equal strengths (effectively each game is determined by a coin
+    toss)
+
+-   Single elimination, bad seeding (1 vs 2, 3 vs 4, 5 vs 6, 7 vs 8) and
+    normally distributed strengths
+
+-   Every Possible Permutation: One simulation for each $8! = 40320$
+    possible combinations that final rankings may occur
+
+-   Double elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
+
+-   Round Robin, normally distributed strengths
+
+In order to fairly compare tournament structures, we assume the true
+strengths of the teams follow a normal distribution, where each team is
+assigned an equally spaced quantile and each team's strength is the
+z-score of its respective quantile. Using this assumption, 10,000
+simulations, all with 8 total teams, were ran for a single game round
+robin (each team plays each other once), a single elimination structure
+with the usual seeding structure (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7), a
+single elimination structure with a poor seeding structure (1 vs 2, 3 vs
+4, 5 vs 6, 7 vs 8). For a better understanding of the results, the best
+possible results (every structure has the best in rank 1, the 2nd best
+in rank 2, etc), the worst possible results (every structure has the
+worst in rank 1, the 2nd worst in rank 2, etc), every possible
+permutation of the ranks once, and a single elimination tournament with
+the usual seeding structure where every team is of equal strengths.
 Because the metric found is unitless, we were able to normalize the
@@
 
-![Zach add a caption \label{fig:fig1}](images/Normal_Information_Curve_Graph.png){width="408"}
+![This chart compares the normalized mutual information (on log10 scale)
+of various tournament structures against the number of top teams being
+considered, where the team strengths are equally distanced based on a
+normal distribution.
+\label{fig:fig1}](images/Normal_Information_Curve_Graph.png){width="408"}
 
-Figure \ref{fig:fig1}
+Based on the Figure 1 \ref{fig:fig1} , the intuitive best tournament,
+round robin, does show to be the best. It is important to note that, due
+to the weighting structure, the worst possible score approaches 0 as the
+$i$ number of teams increases, but never reaches 0. So, the plot is
+normalized so that the worst possible is 0 and the best possible is 1.
+This explains the increase for a few of the lines at the end of the
+plot. If there is a poor seeding structure, as seen in the graph, it can
+drastically impact the effectiveness of the tournament. In the extreme
+case seen, the tournament can actually become worse than deciding games
+by random chance (like flipping a coin). Another important discovery is
+that the single elimination structure with equal strengths is better
+than every possible permutation. This is likely because, for a single
+elimination structure, teams that play in the first round cannot both
+advance to the next round in each simulation, leading to slightly more
+information. Additionally, the bad seeding structure overlaps with the
+equal strengths single elimination. This shows that, given that the best
+team is the 1-seed, even a poorly structured single elimination
+structure still results in the best team winning more than by random
+chance. However, because the 2-seed plays the 1-seed in the first round,
+the poor structure results in a less effective result because the top 2
+can never be truly accurate according to the known strengths.
 
-Based on the graph, the intuitive best tournament, round robin, does show to be the best. Double elimination is slightly better than single
-elimination, with the gap growing wider as the top number of teams
-looked at increases. However, those are under the assumption that the
-true strengths were ordered correctly in the seeding. If there is a poor
-seeding structure, as seen in the graph, it can drastically impact the
-effectiveness of the tournament. In the extreme case seen, the
-tournament can actually become worse than deciding games by random
-chance (like flipping a coin). Another important discovery is that the
-single elimination structure with equal strengths is better than every
-possible permutation. The line for every possible permutation would
-likely be similar to a round robin structure with all equal strengths,
-so perhaps the closer the teams are in true strength, the closer round
-robin and single elimination become.
+![Plot to further explore the difference between single elimination and
+double elimination competition structures
+\label{fig:fig2}](images/tournament_comparison_plot_se_de.png){width="409"}
+
+Double elimination is slightly better than single elimination, with the
+gap growing wider as the top number of teams looked at increases.
+However, those are under the assumption that the true strengths were
+ordered correctly in the seeding. While double elimination was expected
+to be slightly better than single elimination, the increase in the gap
+is likely due to noise that comes with how ties are handled. Ties led to
+randomly assigning teams to a ranking, but with double elimination,
+there were less ties that had to be dealt with, leading to more accurate
+rankings.
 
@@
 We thank the Department of Mathematics and Statistics at Loyola
-University Chicago for their support and resources in conducting this study.
-We thank Kailey Marie Lum for the suggestion of the term "salient ranks".
-No external funding was received for this research.
+University Chicago for their support and resources in conducting this
+study. We thank Kailey Marie Lum for the suggestion of the term "salient
+ranks". No external funding was received for this research.
 

### 2025-09-17 sources_summary.qmd
summaries of different sources

---
title: "sources_summary"
format: html
editor: visual
---

# Tournament Papers - Summaries

### Csato 2021 - The incentive (in)compatability of group-based qualification systems

- A mechanism is called incentive-compatible if every participant can achieve their own best outcome by reporting their true preferences.

- All sporting contests should provide players with appropriate incentives to perform.

- In sports, incentive INcompatibility can come from the fact that being ranked lower in a group stage leads to facing a preferred competitor in the subsequent stage, giving an advantage.

- Motivating real-world example is the European qualification tournament for the 2018 FIFA World Cup.

### Elo 1978 - The Rating of Chess Players, Past and Present

- Explains the Elo system and includes calculated ratings of many players in the past.

### Aziz and Others 2018 - Fixing Balanced Knockout and Double Elimination Tournaments

- Balanced knockout tournaments are one of the most widely-used formats for sports competitions. They are considered to be fair and allow a large number of matches to be played in parallel.

- Consider the computational problem of finding the optimal draw for a particular player in this kind of tournament.

- Proved that checking whether there exists a draw in which a player wins is NP-complete, which settles an outstanding open problem. The NP-completeness result also extends to double-elimination tournaments.

### Coates, Parshakov 2016 - Team vs Individual Tournaments: Evidence from Prize Structure in eSports

- There is literature saying that rank order tournaments (and elimination tournaments) are optimal labor contracts; and large salary dispersion can lead to greater effort and higher productivity.

- Also literature on examining production in teams: need to get incentives right such that team members do not shrink their responsibilities or sabotage the efforts of the team.

- This paper incorporates team production with the theory of rank order elimination tournaments. Existing literature focuses on individual sports or on team sports which are not elimination tournaments. In this analysis, competitors in an elimination tournament are groups rather than individuals (competition studied is video games).

### Devriesere, Csato, Goossens 2024 - Tournament Design: A Review from an Operational Research Perspective

- Tournament design refers to the rules that determine how a tournament (a series of games between a number of competitors) is organized.

- Provides an overview of the tournament design literature from the perspective of operational research. Discusses three important design criteria: efficacy and effectivity, fairness, and attractiveness.

- Survey classifies the papers discussing these properties according to the main components of tournament design: format, seeding, draw, scheduling, and ranking.

### Glickman, Jensen 2003 - Adaptive Paired Comparative Design

### 2025-10-06 partial_kendall.R
added kendall's tau paper.

@@ -1,0 +1,57 @@
+#vec <- c(1,2,4,3)
+partial_kendall <- function(vec, k){
+
+  n <- length(vec)
+  if (k == n){k <- n-1}
+  ref <- c(1:n)
+  count <- 0
+  for (i in 1:k){
+    for (j in (i+1):n){
+      add <-  (((vec[i] < vec[j]) & (ref[i] < ref[j])) |
+        ((vec[i] > vec[j]) & (ref[i] > ref[j]))) + 0
+      #print(c(i, j, add))
+      count <- count + add
+    }
+  }
+  weight <-  (count / sum(n - (1:k)))
+ return(weight) 
+}
+
+#Squared distance
+#Truncated kendall's
+#Number of discordant pairs. 
+#Spearman's rho
+
+#Spearman rho
+(cor(c(1,2,3), c(3,2,1), method = "spearman") + 1)/2
+(cor(c(1,2,3), c(1,2,3), method = "spearman") + 1)/2
+(cor(c(1,2), c(1,2), method = "spearman") + 1)/2
+
+(cor(c(1,2,3), c(20,40,60), method = "spearman") + 1)/2
+
+
+
+1/(sum((c(1,2,3) - c(3,2,1))^2) + 1)
+partial_kendall(c(3,2,1),k= 3)
+
+partial_kendall(c(1,2,3,4),k= 1)
+partial_kendall(c(1,2,3,4),k= 2)
+partial_kendall(c(1,2,3,4),k= 3)
+partial_kendall(c(1,2,3,4),k= 4)
+
+partial_kendall(c(1,2,3,4),k= 1)
+
+
+partial_kendall(c(4,3,2,1),k= 3)
+partial_kendall(c(1,3,2,4),k= 4)
+partial_kendall(c(4,2,3,1),k= 1)
+partial_kendall(c(3,2,4,1),k= 4)
+
+partial_kendall(c(1,3,2,4:50),k= 3)
+partial_kendall(c(1,3,2),k= 3)
+
+partial_kendall(c(1,3,4:50,2),k= 3)
+partial_kendall(c(1,3,2,4:50),k= 3)
+partial_kendall(c(50,3,2,4:49,1),k= 3)
+partial_kendall(c(2,3,50,4:49,1),k= 3)
+

### 2025-10-06 weighted_kendall.R

@@ -1,0 +1,156 @@
+n <- 4
+k <- 2
+#Greg/Zach method
+w <- matrix(0, ncol = n, nrow = 4)
+w[1:k, ] <- 1
+w[, 1:k] <- 1
+
+#Ryan method
+w <- matrix(0, ncol = n, nrow = 4)
+w[1:k, 1:k] <- 1
+salient_weights <- c(5, 3, 1)
+#salient_weights <- c(1,1,1)
+k <- 3
+
+weighted_kendall <- function(rhat, k, wtype = NULL, salient_weights = NULL) {
+  n <- length(rhat)
+  #salient_weights needs to be a vector of length k
+  
+  if(is.null(salient_weights)){
+    salient_weights <- rep(1,k)
+  }
+  
+  if (is.null(wtype)) {
+    w <- matrix(0, ncol = n, nrow = n)
+    for (q in k:1){
+      w[q, ] <- salient_weights[q]
+      w[, q] <- salient_weights[q]
+    }
+    w <- w/2
+  }
+  
+  if (wtype == "gregzach") {
+    w <- matrix(0, ncol = n, nrow = n)
+    for (q in k:1) {
+      w[q, ] <- salient_weights[q]
+      w[, q] <- salient_weights[q]
+    }
+    #w[upper.tri(w)] <- w[upper.tri(w)]/2
+    #w[lower.tri(w)] <- w[lower.tri(w)]/2
+  }
+  
+  #Ryan method
+  if (wtype == "rydog"){
+  w <- matrix(0, ncol = n, nrow = n)
+#  w[1:k, 1:k] <- 1
+  for (q in k:1) {
+    w[1:q, 1:q] <- salient_weights[q]
+    w[1:q, 1:q] <- salient_weights[q]
+  }
+  #w[upper.tri(w)] <- w[upper.tri(w)]/2
+  #w[lower.tri(w)] <- w[lower.tri(w)]/2
+  }
+  
+  
+  summ <- 0
+  for (i in 1:n) {
+    for (j in 1:n) {
+      if (i != j) {
+        summ <- summ + w[i, j] * sign(i - j) * sign(rhat[i] - rhat[j])
+      }
+      
+    }
+  }
+  
+  out <- summ / (sum(w) - sum(diag(w)))
+  weight <- (out + 1) / 2
+  return(weight)
+  
+}
+
+
+
+
[…]

### 2025-10-06 manuscript/paper.Rmd
man uscript.

@@
   | \large Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information
-author: | 
-  | \large Zach Culp \vspace{-1.1mm}
-  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
-  | \normalsize Loyola University Chicago \vspace{-1mm}
-  | \normalsize Chicago, IL 60660 \vspace{-1mm}
-  |
-  | \large Josie Peterburs \vspace{-1.1mm}
-  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
-  | \normalsize Loyola University Chicago \vspace{-1mm}
-  | \normalsize Chicago, IL 60660 \vspace{-1mm}
-  |
-  | \large Ryan McShane \vspace{-1.1mm}
-  | \normalsize University of Chicago \vspace{-1mm}
-  | \normalsize Chicago, IL 60660 \vspace{-1mm}
-  |
-  | \large Gregory J. Matthews \vspace{-1.1mm}
-  |	\normalsize  \vspace{-1mm}
-  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
-  | \normalsize Center for Data Science and Consulting \vspace{-1mm}
-  | \normalsize Loyola University Chicago \vspace{-1mm}
-  | \normalsize Chicago, IL 60660 \vspace{-1mm}
-  | \normalsize [`email`](mailto:ypu@something.edu) \vspace{-1mm}
-  | \normalsize [`zculp@luc.edu`](mailto:zculp@luc.edu) \vspace{-1mm}
-  | \normalsize [`jpeterburs@luc.edu`](mailto:jpeterburs@luc.edu) \vspace{-1mm}
-  | \normalsize [`rmcshane@uchicago.edu`](mailto:rmcshane@uchicago.edu) \vspace{-1mm}
-  | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm}
 abstract: | 
@@
 
-In order to alleviate this problem, we instead consider *weighted*
-mutual information (@Guiasu1977-nu). We want to give more weight to
-permutations from $\hat{r}$ that are "closer" to $r$.
+In order to alleviate this problem, we instead consider *weighted* mutual information (@Guiasu1977-nu).
+We want to give more weight to permutations from $\hat{r}$ that are "closer" to $r$.
+Essentially, this problem comes down to choosing a metric that measures the distance between the vectors $\hat{r}$ and $r$.
+One then gives more weight to combinations of $\hat{r}$ and $r$ that are closer and less weight to combinations that are far apart.
+A natural choice for distance between vectors the sum of squared deviations between the two vectors.
+Specifically, we define the loss function between $\hat{r}$ and $r$ to be $l(r, \hat{r}) = \sum_{i=1}^n (\hat{r}_i - r_i)^2$.
+Then, since small losses should be associated with large weights, we focus on the recipricol of this loss function.
 
-In general, any loss function $l(r, \hat{r})$ can be used to define the
-weigting function:
+In general, any loss function $l(r, \hat{r})$ can be used to define the weighting function as follows:
 
@@
 
-Here though, we choose to use squared error loss:
+Note that in order to avoid dividing by 0, we define the weight to be 1 when $r$ and  $\hat{r}$ are exactly equal.
+
+Specifically, when squared error loss is chosen, one gets the following weight function:
 
@@
 \begin{cases}
-\frac{1}{\hat{r}'r}, & r \ne \hat{r} \\
+\frac{1}{\sum_{i=1}^n (\hat{r}_i - r_i)^2}, & r \ne \hat{r} \\
 1, & r = \hat{r}
@@
 
-However, because the number of teams that need to be accurately ordered
-differs based on the choice of salient ranks, a second set of weights is
-defined, $m$, is used to indicate which ranks are salient. We call these
-the "salient weights" to differentiate from the other weighting
-function. As an example, for a four-team tournament, if the organizer
-decides that only the winner is important, then $m = \{1,0,0,0\}$ (i.e.
-essentially measuring effectivity). While here we only consider
-indicators (i.e. 0's or 1's) in the salient weights vector, there is no
-reason why you couldn't weight the ranks by relative levels of
-importance. For instance, letting $m = \{3,1,0,0\}$ indicates that the
-salient ranks are 1 and 2, but getting the true best competitor as the
-winner is more important than ranking the true second best competitor
+An alternative to defining weight this way is to consider looking at the correlation between $\hat{r}$ and $r$.
+Since, $\hat{r}$ and $r$ are already ranks, the Pearson and Spearman correlation coefficients will be computationally exactly the same in this setting if that was the measure that someone chose to use.
+
+Spearman correlation can be written as follows:
+
+$$
+\rho_{s} =  1 - \frac{6  \sum_{i=1}^n (\hat{r}_i - r_i)^2}{n(n^2 - 1)}
+$$,
+
+which means that Spearman correlation is simply (one minus) a scaling of squared error loss.
+Since $\rho_{s} \in [-1,1]$, but we desire weights $w(\hat{r},r) \in [0,1]$, we define $w(\hat{r},r) = \frac{(\rho_s + 1)}{2}$ guaranteeing weights in the desired range.
+
+While squared error loss and Spearman/Pearson correlation work well, in this setting we argue that Kendall's $\tau$ correlation is the most appropriate for weighting due to its interpretation and how this related to competitions.
+One way of interpreting Kendall's $\tau$ is that it's a scaling of the number of adjacent swaps required to go from one permutation of ranks to another.
+Recall that Kendall's $\tau$ is defined as: 
+$$
+\rho_\tau = 1 - \frac{2 \times \text{# discordant pairs}}{n \choose 2 }
+$$.
+To convert $\rho_\tau$ back to the range between $[0,1]$, which we want for weighting purposes, we subtract 1 and divide by 2 giving us $\frac{\rho_\tau + 1}{2}$.  This then becomes:
+
+$$
+w(\hat{r},r) = \frac{\rho_\tau + 1}{2} = 1 - \frac{\text{# discordant pairs}}{n \choose 2}
+$$ 
+and the number of discordant pairs is equivalent to the number of adjacent swaps needed to get the order of $\hat{r}$ correct.
+
+As an example, consider $r = (1,2,3,4)$ and $\hat{r} = (1,3,4,2)$.
+This requires two adjacent swaps to get $\hat{r}$ to equal $r$ (i.e swap 4 with 2 and then 3 with 2).
+The Kendall's $\tau$ value between these two vectors is `r cor(c(1,2,3,4),c(1,3,4,2), method = "kendall")`.
+We would then define the weight between the two vectors to be $(\rho_{\tau} + 1)/2 =$ `r (cor(c(1,2,3,4),c(1,3,4,2), method = "kendall") + 1)/2`.
+This is exactly equivalent to $1 - \frac{\text{# adj swaps}}{n \choose 2}$.
+
+<!-- ```{r} -->
+<!-- cor(c(1,2,3,4),c(1,3,4,2), method = "kendall") -->
+<!-- 1 - (2*2)/choose(4, 2) -->
+
+<!-- cor(c(1,2,3,4),c(1,2,4,3), method = "kendall") -->
+<!-- 1 - (2*1)/choose(4, 2) -->
+
+<!-- #Converting this back to 0,1 gives (by adding 1 and dividing by 2):  -->
+<!-- #(Number of swaps)/choose(4, 2).   -->
+<!-- #This is what we use for weights -->
+<!-- ``` -->
+
+@shieh1998weighted
+Now, regardless of which method one chooses for defining weights between $\hat{r}$ and $r$, a second set of weights needs to be defined because the number of teams that need to be accurately ordered differs based on the choice of salient ranks.
+Therefore, we define a second set of weights, $m$, which is used to indicate which ranks are salient.
+We call these the "salient weights" to differentiate from the other weighting function, $w$.
+As an example, for a four-team tournament, if the organizer decided that only the correctly choosing the best team as the winner was important, then $m = \{1,0,0,0\}$ (i.e.
+essentially measuring effectivity).
+One can imagine that in an event like the Olympics, the salient weights would be 1's for the first three elements and 0's elsewhere.
+
+In addition, you could go beyond 0's and 1's.  
+While here we only consider indicators (i.e.
+0's or 1's) in the salient weights vector, there is no reason why you couldn't weight the ranks by relative levels of importance.
+For instance, letting $m = \{3,1,0,0\}$ indicates that the salient ranks are 1 and 2, but getting the true best competitor as the winner is more important than ranking the true second best competitor
 second.
@@
 
+While we have explored many more competition scenarios, for the sake of space we present only results of tournaments with $n$ = 8 with the distribution of true strengths distributed as equally spaced quantiles of a standard normal distribution unless otherwise noted.
+In order to simulate each game we compute the probability that team $j$ will beat team $k$ by taking the inverse logit transformation of the differences in their strength parameters such that
+$$
+p_{jk} = \frac{e^{\theta_j - \theta_k}}{1 + e^{\theta_j - \theta_k}}
+$$
+where $p_{jk}$ is the probability that competitor $j$ deats competitor $k$ and $\theta_j$ and $\theta_k$ are the strength parameters of teams $j$ and $k$, respectively.
+
+We focus on several simple competition structures focused mainly on variations of bracket tournaments to demonstrate the proof of concept of our proposed metric.
+In all cases, where seeding is needed (i.e.
+all the bracket structures), the seeds are considered to be known and correct reflecting each team's true rank.
+Note that in bracket structures, the final rankings of teams not reaching the finals was randomly assigned among possible ranks.
+For example, if a team in a single elimination bracket lost in the semi-finals, their possible ranks are either 3 or 4 and are equally likely; if a team loses in the quarterfinals in that same tournament , their final rank is randomly assigned from the set of ranks 5 through 8.
+Each competition structure was simulated 10,000 times and those results were used to estimate $wI_m(r,\hat{r})$.
+
+Specifically, we examined the following structures:
+
+-   Single elimination bracket, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
+-   Single elimination bracket, bad seeding (1 vs 2, 3 vs 4, 5 vs 6, 7 vs 8)
+-   Double elimination bracket, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
+-   Single elimination bracket, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7) matches determined by a coin toss.  
+-   Round Robin, normally distributed strengths
+-   Every Possible Permutation: All possible $8! = 40320$ permutations of rankings are equally likely.
+
 ```{=html}
@@
 
-We examined the following structure:
+<!-- We examined the following structure: -->
 
--   Single elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
-    and normally distributed strengths
+<!-- -   Single elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7) -->
+<!--     and normally distributed strengths -->
 
--   Single elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
-    and equal strengths (effectively each game is determined by a coin
-    toss)
+<!-- -   Single elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7) -->
+<!--     and equal strengths (effectively each game is determined by a coin -->
+<!--     toss) -->
 
--   Single elimination, bad seeding (1 vs 2, 3 vs 4, 5 vs 6, 7 vs 8) and
-    normally distributed strengths
+<!-- -   Single elimination, bad seeding (1 vs 2, 3 vs 4, 5 vs 6, 7 vs 8) and -->
+<!--     normally distributed strengths -->
 
--   Every Possible Permutation: One simulation for each $8! = 40320$
-    possible combinations that final rankings may occur
+<!-- -   Every Possible Permutation: One simulation for each $8! = 40320$ -->
+<!--     possible combinations that final rankings may occur -->
 
--   Double elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
+<!-- -   Double elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7) -->
 
--   Round Robin, normally distributed strengths
+<!-- -   Round Robin, normally distributed strengths -->
 
-In order to fairly compare tournament structures, we assume the true
-strengths of the teams follow a normal distribution, where each team is
-assigned an equally spaced quantile and each team's strength is the
-z-score of its respective quantile. Using this assumption, 10,000
-simulations, all with 8 total teams, were ran for a single game round
-robin (each team plays each other once), a single elimination structure
-with the usual seeding structure (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7), a
-single elimination structure with a poor seeding structure (1 vs 2, 3 vs
-4, 5 vs 6, 7 vs 8). For a better understanding of the results, the best
-possible results (every structure has the best in rank 1, the 2nd best
-in rank 2, etc), the worst possible results (every structure has the
-worst in rank 1, the 2nd worst in rank 2, etc), every possible
-permutation of the ranks once, and a single elimination tournament with
-the usual seeding structure where every team is of equal strengths.
-Because the metric found is unitless, we were able to normalize the
-values using the best and worst results to make the range from 0 to 1,
-with 0 as the worst and 1 as the best. After normalizing, we get the
-following graph (note the y-axis is on a log10 scale):
+<!-- In order to fairly compare tournament structures, we assume the true -->
+<!-- strengths of the teams follow a normal distribution, where each team is -->
+<!-- assigned an equally spaced quantile and each team's strength is the -->
+<!-- z-score of its respective quantile. Using this assumption, 10,000 -->
+<!-- simulations, all with 8 total teams, were ran for a single game round -->
+<!-- robin (each team plays each other once), a single elimination structure -->
+<!-- with the usual seeding structure (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7), a -->
+<!-- single elimination structure with a poor seeding structure (1 vs 2, 3 vs -->
+<!-- 4, 5 vs 6, 7 vs 8). For a better understanding of the results, the best -->
+<!-- possible results (every structure has the best in rank 1, the 2nd best -->
+<!-- in rank 2, etc), the worst possible results (every structure has the -->
+<!-- worst in rank 1, the 2nd worst in rank 2, etc), every possible -->
+<!-- permutation of the ranks once, and a single elimination tournament with -->
+<!-- the usual seeding structure where every team is of equal strengths. -->
+<!-- Because the metric found is unitless, we were able to normalize the -->
+<!-- values using the best and worst results to make the range from 0 to 1, -->
+<!-- with 0 as the worst and 1 as the best.  -->
 
-![This chart compares the normalized mutual information (on log10 scale)
-of various tournament structures against the number of top teams being
-considered, where the team strengths are equally distanced based on a
-normal distribution.
+![This figure compares the normalized weight mutual information (on log10 scale) of various competition structures against the number of salient ranks. 
 \label{fig:fig1}](images/Normal_Information_Curve_Graph.png){width="408"}
 
-Based on the Figure 1 \ref{fig:fig1} , the intuitive best tournament,
-round robin, does show to be the best. It is important to note that, due
-to the weighting structure, the worst possible score approaches 0 as the
-$i$ number of teams increases, but never reaches 0. So, the plot is
-normalized so that the worst possible is 0 and the best possible is 1.
-This explains the increase for a few of the lines at the end of the
-plot. If there is a poor seeding structure, as seen in the graph, it can
-drastically impact the effectiveness of the tournament. In the extreme
-case seen, the tournament can actually become worse than deciding games
-by random chance (like flipping a coin). Another important discovery is
-that the single elimination structure with equal strengths is better
-than every possible permutation. This is likely because, for a single
-elimination structure, teams that play in the first round cannot both
-advance to the next round in each simulation, leading to slightly more
-information. Additionally, the bad seeding structure overlaps with the
-equal strengths single elimination. This shows that, given that the best
-team is the 1-seed, even a poorly structured single elimination
-structure still results in the best team winning more than by random
-chance. However, because the 2-seed plays the 1-seed in the first round,
-the poor structure results in a less effective result because the top 2
-can never be truly accurate according to the known strengths.
+Figure \ref{fig:fig1} shows the results of the estimated values of $wI_m(r,\hat{r})$ on the y-axis (note the y-axis is on a $log_{10}$ scale) while the x-axis represents the number of salient weights that were considered.
+Intuitively, of the competition structures considered here, the round robin (shown in orange) should be the best as it consists of the most match-ups by far of any structure considered here.
+Note that it dominates all of the bracket structures regardless of the choice of the number of salient ranks.
 
-![Plot to further explore the difference between single elimination and
-double elimination competition structures
+The single and double elimination bracket structures with traditional seeding (red and light blue, respectively) are nearly identical and both drastically better than the single elimination tournament with a bad seeding structure (shown in black).
+The green curve preserves the structure of a single elimination bracket and has the correct seeding, but picks the winner based on a coin flip rather than weighting simulated outcomes based on the strength of the competitors, whereas the yellow curves shows results where every possible permutation of rankings are equally likely.
+The green curve dominates the yellow curve here because there are certain final rank orderings that are not possible in bracket determined by coin flips because of the structure.
+So the information that is generated from the tournament is based solely on the seeding structure of the tournament.
+Whereas the yellow curve is based on ranks generated from a tournament with essentially no structure at all that happens to generate reasonable rankings occasionally.
+
+Probably most notable in this figure is the black curve briefly dipping below the yellow and green curves.
+What is happening here is that the single elimination tournament with bad seeding (in black) is better when the there is only a single salient rank.
+However, when the number of salient ranks is 2, this drops *below* the two curves that are generates based on coin flip outcomes and completely random outcomes.
+This is because in this structure the 1 and 2 seed both play each other in the first round meaning that one of them will be eliminated immediately.
+Thus under this seeding mechanism, the competition structure simply cannot generate a correct result when the number of salient ranks is 2.
+This structure is still below both green and yellow curves when the number of salient ranks is 3, then recovers and climbs above those curves when the number of salient ranks is 4 and continues for the values of salient ranks 5 through 8.
+
+<!-- , the intuitive best tournament, -->
+<!-- round robin, does show to be the best. It is important to note that, due -->
+<!-- to the weighting structure, the worst possible score approaches 0 as the -->
+<!-- $i$ number of teams increases, but never reaches 0. So, the plot is -->
+<!-- normalized so that the worst possible is 0 and the best possible is 1. -->
+<!-- This explains the increase for a few of the lines at the end of the -->
+<!-- plot. If there is a poor seeding structure, as seen in the graph, it can -->
+<!-- drastically impact the effectiveness of the tournament. In the extreme -->
+<!-- case seen, the tournament can actually become worse than deciding games -->
+<!-- by random chance (like flipping a coin). Another important discovery is -->
+<!-- that the single elimination structure with equal strengths is better -->
+<!-- than every possible permutation. This is likely because, for a single -->
+<!-- elimination structure, teams that play in the first round cannot both -->
+<!-- advance to the next round in each simulation, leading to slightly more -->
+<!-- information. Additionally, the bad seeding structure overlaps with the -->
+<!-- equal strengths single elimination. This shows that, given that the best -->
+<!-- team is the 1-seed, even a poorly structured single elimination -->
+<!-- structure still results in the best team winning more than by random -->
+<!-- chance. However, because the 2-seed plays the 1-seed in the first round, -->
+<!-- the poor structure results in a less effective result because the top 2 -->
+<!-- can never be truly accurate according to the known strengths. -->
+
+![This figure focuses on the difference between single elimination and double elimination competition structures
 \label{fig:fig2}](images/tournament_comparison_plot_se_de.png){width="409"}
 
-Double elimination is slightly better than single elimination, with the
-gap growing wider as the top number of teams looked at increases.
-However, those are under the assumption that the true strengths were
-ordered correctly in the seeding. While double elimination was expected
-to be slightly better than single elimination, the increase in the gap
-is likely due to noise that comes with how ties are handled. Ties led to
-randomly assigning teams to a ranking, but with double elimination,
-there were less ties that had to be dealt with, leading to more accurate
-rankings.
+Figure \ref{fig:fig2} shows the same red and light blue curves representing single and double elimination, respectively, in more detail as show in figure \ref{fig:fig1}.
+Note that in terms of declaring a champion (i.e.
+when the number of salient ranks is 1), single and double elimination brackets are nearly identical is their ability to accomplish this.
+However, double elimination is slightly better than single elimination, with the
+gap growing wider as the number of salient ranks increases.
+One clear reason for this is that with double elimination the ranking of teams is "clearer".
+That is, in a single elimination tournament the two teams that lose in the semi-finals can equally make a claim as the third best team and in our schema they are randomly assigned to a rank fo 3 or 4.
+However, in a double elimination tournament, teams are definitively ranked 3 and 4 leading to beter outcomes when the number of salient ranks is in the range of 3 to 4 (or more).
+
+<!-- While double elimination was expected -->
+<!-- to be slightly better than single elimination, the increase in the gap is likely due to noise that comes with how ties are handled. Ties led to -->
+<!-- randomly assigning teams to a ranking, but with double elimination, -->
+<!-- there were less ties that had to be dealt with, leading to more accurate -->
+<!-- rankings. -->
 
@@
 
-This study introduced a weighted mutual information framework to
-evaluate the effectiveness of different tournament formats in preserving
-the true underlying rankings of teams. By simulating tournaments with
-varying seeding structures and applying an information-theoretic
-approach, we quantified how accurately each format conveys ranking
-information. Our findings confirm the intuitive advantage of round robin
-tournaments, which consistently outperformed single and double
-elimination formats in ranking accuracy. Additionally, we observed that
-seeding plays a critical role: poor seeding structures can significantly
-degrade performance, in some cases making outcomes less informative than
-random chance.
+This study introduced a weighted mutual information framework to evaluate the effectiveness of different competition formats in preserving the true underlying rankings of teams.
+We also introduce the concept of salient ranks, thereby extending and generalizing the concepts of efficacy and effectivity.
+By simulating outcomes of competition strucutres with varying seeding methods and applying an information-theoretic approach, we quantified how accurately each format conveys ranking information.
+Our findings confirm the intuitive advantage of round robin tournaments, which consistently outperformed single and double elimination formats in ranking accuracy, but also quantify the difference between these structures.
+Additionally, we demonstrated the degree to which poor seeding structures can significantly degrade performance of a competition structure, in some cases making competition outcomes even worse than simply randomly generating a set of ranks.
 
-These results underscore the trade-off between tournament accuracy and
-practical constraints such as time, cost, and entertainment value. While
-round robin offers the highest fidelity to true rankings, it is often
-impractical for large tournaments, highlighting the need for hybrid or
-adaptive structures that balance accuracy with logistical feasibility.
+These results underscore the trade-off between tournament accuracy and practical constraints such as time, cost, and entertainment value.
+While round robin offers the highest fidelity to true rankings, it is often impractical for large tournaments, highlighting the need for hybrid or adaptive structures that balance accuracy with logistical feasibility.
 
-Future work should extend this analysis to scenarios with more
-tournament structures, more teams, incorporate probabilistic strength
-models reflecting real-world uncertainty, and explore alternative
-weighting schemes for different competitive priorities (e.g., top-4
-accuracy versus full ranking). Applying these methods to actual
-tournament data could further validate their usefulness for organizers
-aiming to design fair and informative competitions.
+Future work should extend this analysis to scenarios with more tournament structures, varying numbers of competitors, incorporate probabilistic strength
+models reflecting real-world uncertainty, and explore alternative weighting schemes for different competitive priorities (e.g., 4 salient ranks versus 8 salient ranking).
+Applying these methods to actual tournament data could further validate their usefulness for organizers aiming to design fair and informative competitions.
 
@@
 
-We thank the Department of Mathematics and Statistics at Loyola
-University Chicago for their support and resources in conducting this
-study. We thank Kailey Marie Lum for the suggestion of the term "salient
-ranks". No external funding was received for this research.
+We thank the Department of Mathematics and Statistics at Loyola University Chicago for their support and resources in conducting this study.
+We thank Kailey Marie Lum for the suggestion of the term "salient ranks".
+No external funding was received for this research.
 
@@
 # References
+
+<!-- author: |  -->
+<!--   | \large Zach Culp \vspace{-1.1mm} -->
+<!--   | \normalsize Department of Mathematics and Statistics \vspace{-1mm} -->
+<!--   | \normalsize Loyola University Chicago \vspace{-1mm} -->
+<!--   | \normalsize Chicago, IL 60660 \vspace{-1mm} -->
+<!--   | -->
+<!--   | \large Josie Peterburs \vspace{-1.1mm} -->
+<!--   | \normalsize Department of Mathematics and Statistics \vspace{-1mm} -->
+<!--   | \normalsize Loyola University Chicago \vspace{-1mm} -->
+<!--   | \normalsize Chicago, IL 60660 \vspace{-1mm} -->
+<!--   | -->
+<!--   | \large Ryan McShane \vspace{-1.1mm} -->
+<!--   | \normalsize University of Chicago \vspace{-1mm} -->
+<!--   | \normalsize Chicago, IL 60660 \vspace{-1mm} -->
+<!--   | -->
+<!--   | \large Gregory J. Matthews \vspace{-1.1mm} -->
+<!--   |	\normalsize  \vspace{-1mm} -->
+<!--   | \normalsize Department of Mathematics and Statistics \vspace{-1mm} -->
+<!--   | \normalsize Center for Data Science and Consulting \vspace{-1mm} -->
+<!--   | \normalsize Loyola University Chicago \vspace{-1mm} -->
+<!--   | \normalsize Chicago, IL 60660 \vspace{-1mm} -->
+<!--   | \normalsize [`email`](mailto:ypu@something.edu) \vspace{-1mm} -->
+<!--   | \normalsize [`zculp@luc.edu`](mailto:zculp@luc.edu) \vspace{-1mm} -->
+<!--   | \normalsize [`jpeterburs@luc.edu`](mailto:jpeterburs@luc.edu) \vspace{-1mm} -->
+<!--   | \normalsize [`rmcshane@uchicago.edu`](mailto:rmcshane@uchicago.edu) \vspace{-1mm} -->
+<!--   | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm} -->

### 2025-10-06 manuscript/paper.Rmd
How the FUCK do we write out the formula for the salient weight function for weighted kendall's tau.  It's pretty easy for spearman/pearson correlation or for squared error loss.  But for Kendall's tau you need a MATRIX of weights and not just a vector.  The matrix is DERIVED from the vector, but how do we write this in general so that both thse cases have the same notation?

@@
 will yield high values of mutual information when the output from the
-tournament $\hat{r}$ is highly consistent *even if the ranking from the
-tournament is incorrect*. As an example, if $n=4$ and the true order of
+tournament $\hat{r}$ is highly consistent *even if the ranking from the tournament is incorrect* (*i.e.* the rankings produced by the tournament are very different from the true rankings).
+As an example, if $n=4$ and the true order of
 $\boldsymbol{\theta}$ is $r = \{1, 2, 3, 4\}$ and
@@
 $$
+# Weighted Mutual Information
 
@@
 $$
-\rho_\tau = 1 - \frac{2 \times \text{# discordant pairs}}{n \choose 2 }
+\rho_\tau = 1 - \frac{2 \times \# \text{ discordant pairs}}{{n \choose 2}}
 $$.
@@
 $$
-w(\hat{r},r) = \frac{\rho_\tau + 1}{2} = 1 - \frac{\text{# discordant pairs}}{n \choose 2}
+w(\hat{r},r) = \frac{\rho_\tau + 1}{2} = 1 - \frac{\# \text{ discordant pairs}}{{n \choose 2}}
 $$ 
@@
 We would then define the weight between the two vectors to be $(\rho_{\tau} + 1)/2 =$ `r (cor(c(1,2,3,4),c(1,3,4,2), method = "kendall") + 1)/2`.
-This is exactly equivalent to $1 - \frac{\text{# adj swaps}}{n \choose 2}$.
+This is exactly equivalent to $1 - \frac{\#\text{ adj swaps}}{{n \choose 2}}$.
 
@@
 
-@shieh1998weighted
-Now, regardless of which method one chooses for defining weights between $\hat{r}$ and $r$, a second set of weights needs to be defined because the number of teams that need to be accurately ordered differs based on the choice of salient ranks.
+# Salient Weights (and ranks)
+
+@shieh1998weighted: Weighted Kendall's Tau
+
+Regardless of which method one chooses for defining weights between $\hat{r}$ and $r$, a second set of weights needs to be defined because the number of teams that need to be accurately ordered differs based on the choice of salient ranks.
 Therefore, we define a second set of weights, $m$, which is used to indicate which ranks are salient.
@@
 As an example, for a four-team tournament, if the organizer decided that only the correctly choosing the best team as the winner was important, then $m = \{1,0,0,0\}$ (i.e.
-essentially measuring effectivity).
-One can imagine that in an event like the Olympics, the salient weights would be 1's for the first three elements and 0's elsewhere.
+essentially measuring effectivity)^[7].
 
-In addition, you could go beyond 0's and 1's.  
-While here we only consider indicators (i.e.
-0's or 1's) in the salient weights vector, there is no reason why you couldn't weight the ranks by relative levels of importance.
-For instance, letting $m = \{3,1,0,0\}$ indicates that the salient ranks are 1 and 2, but getting the true best competitor as the winner is more important than ranking the true second best competitor
-second.
+^[7]: One can imagine that in an event like the Olympics, the salient weights would be 1's for the first three elements and 0's elsewhere.
 
-With the addition of the weighting functions, the formula for mutual
-information becomes:
+In addition, one can also weight the salient ranks.
+For example, in the Olmypics where getting the top three correct is the main objective, one could argue that it is more important to get the first place correct than the third place.
+This could lead to, for instance, a salient weight vector of $m = \{5,3,1,0, \cdots, 0\}$ giving 5 times the weight to getting the first place team correct as getting the third team correct.
+
+With the addition of the salient weighting function, $m$, the formula for weighted mutual information becomes:
 
 $$
-wI_m(r, \hat{r})=\sum_{r}\sum_{\hat{r}} w(r,\hat{r}) * m * p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
+w(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}}) = \sum_{i = 1}^n m_i \times d(\hat{r}_i, r_i)
+
+\\
+
+wI_m(r_{\{{i|m_i >0}\}}, \hat{r_{\{{i|m_i >0}\}}})=\sum_{r_{\{{i|m_i >0}\}}}\sum_{\hat{r}_{\{{i|m_i >0}\}}} w(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}}) \times \\
+p(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}})\times \\ \log \frac {p(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}})}{p(r_{\{{i  \mid m_i >0}\}})p(\hat{r}_{\{{i \mid m_i >0}\}})}
 $$

### 2025-10-06 partial_kendall.R
fucking citation.

@@ -27,7 +27,16 @@ partial_kendall <- function(vec, k){
 (cor(c(1,2,3), c(1,2,3), method = "spearman") + 1)/2
 (cor(c(1,2), c(1,2), method = "spearman") + 1)/2
 
-(cor(c(1,2,3), c(20,40,60), method = "spearman") + 1)/2
+(cor(c(1,2,3,4,5,6), c(1,2,3,4,6,5), method = "spearman") + 1)/2
+(cor(c(1,2,3,4,5,6), c(2,1,3,4,5,6), method = "spearman") + 1)/2
+(cor(c(1,2,3,4,5,6), c(1,2,4,3,5,6), method = "spearman") + 1)/2
+(cor(c(1,2,3,4,5,6), c(1,2,4,5,3,6), method = "spearman") + 1)/2
+
+rhat <- c(1,2,3,4,6,5,7)
+n <- length(c(1,2,4,5,3,6,7))
+1 - (6*sum((c(1,2,3,4,5,6,7) -  rhat)^2))/(n*(n^2-1))
+
+cor(c(1,2,3,4,5,6,7), rhat, method = "spearman")
 
 
 

### 2025-10-07 sources_summary.qmd
fucking citation.
lit review notes

@@
 
-- Balanced knockout tournaments are one of the most widely-used formats for sports competitions. They are considered to be fair and allow a large number of matches to be played in parallel.
+-   Balanced knockout tournaments (knockout tournaments where each player has to play the same number of matches to win the tournament) are one of the most widely-used formats for sports competitions.
+They are considered to be fair and allow a large number of matches to be played in parallel.
 
@@
 ### Glickman, Jensen 2003 - Adaptive Paired Comparative Design
+
+- An important part of paired comparison experiments is deciding how the pairs are formed.
+It is usually difficult to incorporate prior information in typical paired comparison experimental design (prior information is relevant for the design of tournament schedules for players of games and sports).
+
+- The problem of pairing objects can be formalized as a Bayesian optimal design.
+The performance of this method is tested on simulated data and is compared to random pairings and to a system used in tournament chess.
+
+### Groeneveld, Siu 2017 - Comparing different tournament systems in their probability that the best player wins.
+
+- Paper that compares the performance of different tournament systems.
+The performance is measured by looking at the probability that the best player wins and the average rank of the winning player.
+Results are determined by computer simulations.
+
+- Many variables that contribute to the result of a tournament.
+Two major categories are tournament setup and player skill.
+There are also some uncontrollable variables that some would contribute to luck, such as fatigue.
+
+- The variable of tournament setup is measured in the paper. For the player skill, they emulate that through Elo rankings as used for chess.
+
+- Compare single elimination, double elimination, single round-robin, double round-robin, Swiss-system, and the multistage system.
+
+- Each tournament is simulated 100,000 times with the same players and same rankings.
+Each tournament simulation outputs the ranking of the winner, which are counted for each player.
+At the end of the simulation, the result will be the number of wins per player sorted on their Elo rank.
+They divide this by the total number of simulations to get a win probability.
+
+- Results show a convergence to 0 for the lowest ranked players.
+As they focus on the probability that the best player wins, the lower end of the rankings are less significant.
+They put the full results in a logarithmic scale.
+
+- **Graphs are similar in concept to ours**
+
+### Hartiagn 1968 - Inference from a Knockout Tournament
+
+- Develop a parametric probability model for order data on a set of objects, which is appropriate for analysis of data with tree structure (ex. results of knockout tournaments).
+
+- The parameters of the model are the probabilities that the various objects will be ranked highest.
+
+### Ikhwani, et al. 2023 - Single Elimination Tournament Design Using Dynamic Programming Algorithm
+
+- Aims to create an optimal single-elimination tournament design using binary tree modeling with dummy techniques.
+
+- Uses various methods of sub-optimal algorithms and then compares their performance in terms of runtime and optimality as a solution to measure the comparison of sub-algorithms.
+
+### Kendall, Smith 1940 - On the Method of Paired Comparisons
+
+- One method of comparison that is widely used is ranking.
+Ranking method suffers from a serious drawback when the quality considered is not known with certainty to be representable by a linear variable.
+
+- Consider a more general method of investigating preferences.
+With n objects, we suppose that each of the (n choose 2) pairs are presented to an observer and their preference noted.
+With m observers, there are m * (n choose 2) preferences.
+
+### Khalid, et al. 2023 - A Decision Support System for Simulating and Predicting the Impacts of Various Tournament Structures on Tournament Outcomes
+
+- Tournament outcomes can be influenced by several factors, such as attack, defensse, and home advantage strength values, as well as tournament structures. 
+- However, the claim that different structures have their own time restraints and requirements has limited the evaluation of the best structure for a particular type of sports tournament using quantitative approaches.
+
+- This study develops a decision support system (DSS) using Microsoft Virtual Basic to simulate and forecast the impact of the various tournament structures on soccer tournament outcomes.
+
+- The DSS utilized the attack, defense, and home advantage values of the teams involved in the Malaysia Super League 2018 to make better prediction.
+The rankings produced by the DSS were then compared to the actual rankings using Spearman correlation to reveal the simulated accuracy level.
+
+### Krautmann, et al. 2011 - Playoff Uncertainty and Pennant Races
+
+- Outcome uncertainty is a fundamental principle associated with Rottenburg. It has many meanings in sports literature.
+
+- In this article, the authors consider a type of within-season uncertainty termed "playoff uncertainty" (PLU).
+This concept looks at how demanders respond to the probability that their team has a chance of winning over the course of the season.
+
+- More specifically, PLU focuses on whether attendance is affected by a team's chance of making the playoffs.
+Looking at the attendance data on a monthly basis, the authors find that a tighter pennant race does enhance league-wide attendance, but only for the months toward the end of the season when the pennant race is heating up.
+
+### Lee, Fort 2023 - Division Play and Outcome Uncertainty in Sports Leagues
+
+- The analysis of outcome uncertainty (OU) and competitive balance (CB) has been important in sports economics.
+Surprisingly, there is little work on the impact of the structure of play on either OU or CB.
+
+- The impact of the introduction of both unbalanced schedules and division play on OU has not been analyzed.
+
+- In this paper, they assess the impacts on OU, for given choices of CB (OU/CB) of moving to division play.
+This includes the impact of schedule imbalance, division strength, the number of teams, and the number of divisions.
+Also obtain estimates of their marginal impacts on OU/CB via numerical analysis and regression.
+
+### Manurangsi, Suksompong 2022 - Fixing Knockout Tournaments with Seeds
+
+- Prior results have shown that it is often possible to manipulate a knockout tournament by fixing the bracket, these results ignore the prevalent aspect of player seeds, which can significantly constrain the chosen bracket.
+
+- They show that certain structural conditions that guarantee that a player can win a knockout tournament without seeds are no longer sufficient in light of seed constraints.
+
+- Also prove that when the pairwise match outcomes are generated randomly, all players are still likely to be knockout winners under the same probability threshold with seeds as without seeds.
+
+- In addition, investigate the complexity of deciding whether a manipulation is possible when seeds are present.
+
+### Monsuur, Storcken 1997 - Measuring intransitivity
+
+- transitivity definition: In graph theory, a tournament is a directed graph where every pair of vertices has a single directed edge between them, representing a competition or comparison.
+Transitivity in a tournament means that if vertex A beats B, and B beats C, then A must beat C.
+
+- Present a framework for measures of non-transitivity in finite tournaments. Axiomatic characterizations for five well-known intransitivity measures are derived.
+
+- An intransitivity measure expresses the 'amount' of intransitivity in a preference relation in a non-negative number. There are various ways to measure intransitivity.
+
+### Mosteller 1952 - The World Series Competition
+
+- Suppose a number of pairs of teams are compared on the basis of n binomial trials.
+Although we cannot know from the outcomes which teams were actually better, we wish to estimate the average probability that the better team wins a given trial, and thus to measure the discrimination provided by the test.
+
+- Estimate the overall effectiveness of the decision technique. Answer the question: Will the Series be very effective in identifying the better team?
+
+### Daniels 1969 - Round-robin tournament scores
+
+- In a paired comparisons experiment, preferences are expressed between all possible pairs of a set of objects.
+The problem of analyzing such an experiment is basically to assign optimum scores to the objects, often with the more limited aim of ranking them in order of preference.
+
+- Round-robin tournament is another example of this situation. The paper proposes some methods of scoring players, based on a concept of fair allocation of rewards to each player.
+
+### Scarf, Yusof 2010 - A numerical study of the tournament structure and seeding policy for the soccer World Cup Finals
+
+- Tournament outcome uncertainty depends on the design of the tournament and the relative strengths of the competitors--the competitive balance.
+
+- A tournament design comprises the arrangement of the individual matches, which we call the tournament structure, the seeding policy, and the progression rules.
+
+- In this paper, they investigate the effect of seeding policies for various tournament structures, while taking account of competitive balance.
+The methodology uses tournament outcome uncertainty to consider the effect of seeding policy and other design changes.
+
+### Shader 1992 - On Tournament Matrices
+
+- Let T be a tournament of order n with adjacency matrix M.
+We find several conditions that are equivalent to M being singular.
+A correlation between the number of 3-cycles in T and the rank of M is established.
+
+- It is shown that asymptotically at least half of the tournament matrices are nonsingular. Also derive bounds on the spectral radius of tournament matrices with a given row-sum vector.
+
+### Lambers, Pendavingh, Spieksma 2022 - How to Design a Stable Serial Knockout Competition
+
+- Investigate a new tournament format that consists of a series of individual knockout tournaments; call this new format a Serial Knockout Competition (SKC).
+This format was recently adopted by the Professional Darts Corporation.
+
+- Depending on the seedings of the players used for each of the knockout tournaments, players can meet in various rounds of the tournaments.
+
+- Following a fairness principle of treating all players equal, they identify an attractive property of an SKC: each pair of players should potentially meet equally often in each of the rounds of the SKC.
+
+- If the seedings are such that this property is present, they call the resulting SKC stable.
+Then address the question: do there exist seedings for each of the knockout tournaments such that the resulting SKC is stable?
+
+### Sukompong 2015 - Scheduling asynchronous round-robin tournaments
+
+- Study the problem of scheduling asynchronous round-robin tournaments, i.e. round-robin tournaments in which no two games take place at the same time.
+
+- Consider three measures of a schedule that concern the quality and fairness of a tournament: the guaranteed rest time, the games-played difference index, and the rest difference index.
+
+- Show that the well-known 'circle-design' performs well with respect to all three measures when the number of teams is even, but not when the number of teams is odd.
+They propose a difference schedule that performs optimally when the number of teams is odd.
+
+### Sziklai, Biro, Csato 2022 - The efficacy of tournament designs
+
+- Efficacy can be defined as the capability of a tournament to reproduce the ranking of the players according to their strength.
+
+- Investigates a fundamental issue of economics in tournament design: what is the best usage of limited resources--that is, how should the alternatives be compared pairwise to best approximate their true but latent ranking.
+
+- Consider formats including knockout tournaments, multistage championships consisting of round-robin groups followed by single elimination, and the Swiss-system.
+They are evaluated via Monte-Carlo simulation under six different assumptions on winning probabilities.
+
+- Comparing the same pair of alternatives multiple times turns out to be an inefficacious policy.
+
+- While seeding can increase the efficacy of the knockout and group-based designs, its influence remains marginal unless one has an unrealistically good estimation on the true ranking of the players.
+
+- The Swiss-system was found to be the most accurate among these formats, especially in its ability to rank all participants.
+
+### Vu, Shoham 2011 - Fair Seeding in Knockout Tournaments
+
+- Seeding can significantly influence the results of a tournament.
+Most papers focus on how to find a seeding that maximizes the winning probability of the strongest player (so-called predictive power).
+This usually means giving the strongest player the easiest schedule.
+
+- Popular 1-8, 4-5, 2-7, 3-6 seeding seems fair because confrontations between stronger competitors are delayed to later rounds, which increases the chance that one of these strong players will win the tournament.
+
+- Raises several questions: How do we capture this intuition about fairness? Is there any other seeding that is also fair? Can we always find such a fair seeding?
+
+- To answer these questions, the authors consider two alternative fairness criteria adapted from the literature: envy-freeness and order preservation.
+They analyze these criterion separately across different tournament settings, providing both possibility and impossibility results.
+
+### Wright 2013 - OR analysis of sporting rules - A survey
+
+- This paper surveys the academic OR (Operational Research)/analytics literature describing research into the laws and rules of sports and sporting competitions.
+
+- Outline a large number of studies covering 21 sports in many parts of the world.
+
+- The analytical approaches most commonly used are found to be various forms of regression analysis and simulation.
+Issues highlighted by this survey include the different views of what constitutes fairness and the frequency with which changes produce unintended consequences.

### 2025-10-10 CMSAC_presentation.Rmd
CMSAC presentation

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory Matthews, Dr. Ryan McShane"
date: "`r Sys.Date()`"
bibliography: references.bib
fontsize: 12pt
output:
  beamer_presentation:
    theme: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tinytex)
library(tidyverse)
load("curves_data.RData")
#load("tournament_comparison_plot.png")
#load("tournament_comparison_plot_close.png")
```

## Introduction
<!-- Overall comments 
I think the presentation is too long for 12 minutes (+ questions), but
I feel like I didn't cover everything we want. What is plausable to cover
in a 12-minute period? 
Also, we need to figure out the notation for the final metric because I think
the weight should have the subscript m, not I. And I probably was not consistent
with the usage of theta or calling it true rank/true strength. -->

-   There exists numerous competition formats across all professional sports
-   Which competition structure is best?
-   The "best" competition structure depends on how well it reflects the competitors' true strengths
-   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors' true strengths

<!-- Should we talk about Efficacy vs Effectivity here -->

## Game Simulation

-   Given the assigned "known strength", $\theta$, we simulate a game using the Bradley-Terry Model

$$\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$$

## Tournament Simulations
- Each competition structure is simulated given the number of teams, $n$,  and $\theta$
- Below is an example of the output from one simulation of a single elimination tournament structure

```{r}
renamed_df <- all_results_norm %>% 
  select(-c(simulation, distribution, seed, teams)) %>%
  rename(theta = true_strength, r = true_rank, r_hat = rank_hat) %>%
  mutate(theta = round(theta,3))

head(renamed_df,8)
```

## Mutual Information

- If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the “goodness” of a tournament in terms of its ability to accurately transmit the true ranking.

- The notation for mutual information is: 
$$ I(X;Y) = \sum_{y \in Y}\sum_{x \in X} p(x,y) \log{\frac{p(x,y)}{p(x)p(y)}} $$
- In our case, it can be simplified to:
$$ I(r, \hat{r}) = n! \sum_{\hat{r}} p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{(\frac{1}{n!})^2}} = n! \log{((n!)^2)} $$

## Weighted Mutual Information
- A limitation of mutual information is that it does not distinguish between the direction of information
- In other words, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
- Because of this, we proposed a weighting function $w(r, \hat{r})$
- So, the new formula is:
$$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$

## Weighting Function
- In our original paper submission, we used squared error loss as the weighting function:
$$ w(r, \hat{r}) =
\begin{cases}
\frac{1}{\hat{r}'r}, & r \neq \hat{r} \\
1, & r = \hat{r}
\end{cases} $$
- However, after comments and reviews from people at CSAS and CMSAC, we have decided to use a weighted Kendall's Tau

## Weighted Kendall
$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \times \text{sgn}(i-j)\times \text{sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$

- $W$ is a $n\times n$ lower-triangular matrix, where

<!-- Should this be capital W??? -->

$$W_{ij} = w_{\min(i,j)}, \text{ for } i,j = m,2m, \dots, n\times m$$

- $m$: vector of length $n$ representing the "salient weights", or the importance of correctly getting each ranking

- If $\tau_w = 1$, then the ranking is identical to the true ranking; if $\tau_w = 0$, then the order of teams is perfectly inversely proportional to the true rankings

## Weighted Kendall Example
- If $\hat{r} = (2,1,3,4)$ and m = $(3,1,1,0)$, then:
$$ W = \begin{pmatrix}
3 & 3 & 3 & 3 \\
3 & 1 & 1 & 1 \\
3 & 1 & 1 & 1 \\
3 & 1 & 1 & 0
\end{pmatrix} $$

$$\sum_{i,j} W_{ij} = 29; \sum_{i} W_{ii} = 5; \sum_{i\neq j} W_{ij} \times \text{sgn}(i-j)\times \text{sgn}(R_i - R_j) = 18 $$

- So, $\tau_w$ for this example would be $\frac{18}{29-5} = 0.75$

## Normalizing Mutual Information
- Mutual Information can be normalized by dividing by the joint entropy

- Entropy: $-p(x)\log{p(x)}$

- Assuming $r$ and $\hat{r}$ are independent, joint entropy can be written as $H(r, \hat{r}) = H(r) + H(\hat{r})$

- $H(r)$ and $H(\hat{r})$ are theoretically equivalent, leaving $H(r,\hat{r}) = -2\sum_{r} p(r)\log(p(r)) = -2(n!) \times (1/n!) \times\log_2(1/n!) = \log_2((n!)^2)$
<!-- Is this true?? I can't remember if they are equivalent or their maximums were equivalent -->

- So, the final formula for the metric is:
$$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$

## Results

<!-- Should these plots be on separate slides? -->
```{r out.width="49%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot.png")
knitr::include_graphics("tournament_comparison_plot_close.png")
```

## Limitations
- The "known" strengths are assumptions that can be estimated but never truly known
- Ties are not addressed in any of the games
- Ties in the $\hat{r}$ are randomly assigned

## Future Work
- Evaluate structures using various distributions and "known" weights
- Estimate the $\theta$s in a real-world example and compare recent changes to the old format

## References
<!-- I don't know if the references.bib would work here or not -->
- Appleton, David R. 1995. “May the Best Man Win?” Statistician 44 (4): 529.
- Csato, L. 2021. Tournament Design: How Operations Research Can Improve Sports Rules.
1st Ed. Cham, Switzerland: Palgrave Macmillan.
- Devriesere, Csató, K., and D. Goossens. 2025. “Tournament Design: A Review from an
Operational Research Perspective.” European Journal of Operational Research 324 (1):
1–21.
- Glenn, W A. 1960. “A Comparison of the Effectiveness of Tournaments.” Biometrika 47
(3-4): 253–62.

## References
- Guiasu, Silviu. 1977. Information Theory with Applications. New York, NY: McGraw-Hill.
- Johnson, Sidney, and Rodney Fort. 2022. “Match Outcome Uncertainty and Sports Fan
Demand: An Agnostic Review and the Standard Economic Theory of Ports Leagues.”
Int. J. Emp. Econ. 01 (02).
- Lasek, Jan, and Marek Gagolewski. 2018. “The Efficacy of League Formats in Ranking
Teams.” Stat. Modelling 18 (5-6): 411–35.
- Sziklai, Balázs R, Péter Biró, and László Csató. 2022. “The Efficacy of Tournament Designs.”
Comput. Oper. Res. 144 (105821): 105821.

### 2025-10-10 group_stage.qmd
CMSAC presentation

@@
 ```{r}
-simulate_group_play <- function(strengths, rounds=1){
-  teams <- strengths$true_rank
-  n <- length(teams)
-  win_counts <- numeric(n)
-  loss_counts <- numeric(n)
-  for (i in 1:rounds){
-    for (i in seq_along(teams)){
-      for (j in seq_along(teams)){
-        if (i != j){
-          team1 <- teams[i]
-          team2 <- teams[j]
-          strength1 <- strengths$true_strength[strengths$true_rank == team1]
-          strength2 <- strengths$true_strength[strengths$true_rank == team2]
-          match_winner <- simulate_match(teams[i],teams[j],strength1, strength2)
-          if (match_winner == teams[i]) {
-            win_counts[i] <- win_counts[i] + 1
-            loss_counts[j] <- loss_counts[j] + 1
+simulate_group_stage_tournament <- function(num_teams, distribution, group_length = 4, ties = TRUE, rounds = 1, seeded=F) {
+
+  if (num_teams <= 3) {
+    stop("Number of Teams must be greater than 3.")
+  }
+
+  if (num_teams %% group_length != 0) {
+    stop("Number of teams must be divisible by group length.")
+  }
+
+  teams <- paste(num_teams:1)
+
+  # Create team strengths based on distribution
+  if (distribution == "Normal"){
+    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
+    team_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      theta = strengths,
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- arrange(team_strengths, true_rank)
+  }
+  else if (distribution == "Same"){
+    strengths <- sapply(num_teams, function(n) {rep(0, n)})
+    team_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      theta = strengths,
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- arrange(team_strengths, true_rank)
+  }
+  else if (distribution == "Uniform"){
+    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1), 0, sqrt(12)) })
+    team_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      theta = strengths,
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams)),
+      rank_hat = rep(NA, length(teams))
+    )
+    df <- arrange(team_strengths, true_rank)
+  }
+  else if (distribution == "Manual") {
+    cat("Please enter the strengths for each team in descending rank order:\n")
+    strengths <- numeric(num_teams)
+    for (i in num_teams:1) {
+      repeat {
+        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
+        if (!is.na(input)) {
+          strengths[i] <- input
+          break
         } else {
-            win_counts[j] <- win_counts[j] + 1
-            loss_counts[i] <- loss_counts[i] + 1
+          cat("Invalid input. Please enter a numeric value.\n")
+        }
+      }
+    }
+    team_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      theta = strengths,
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- arrange(team_strengths, true_rank)
+  }
+  else {
+    stop("Distribution not found: Enter Manual to input your own strengths")
+  }
+
+  num_groups <- num_teams / group_length
+
+  # Create balanced groups using snake draft seeding
+  # Sort teams by strength (best to worst)
+  df <- df %>% arrange(desc(theta))
+
+  # Assign groups using snake draft method
+  df$groups <- NA
+  for (i in 1:nrow(df)) {
+    group_round <- ceiling(i / num_groups)
+    if (group_round %% 2 == 1) {
+      # Forward direction for odd rounds
+      group_id <- ((i - 1) %% num_groups) + 1
+    } else {
+      # Reverse direction for even rounds
+      group_id <- num_groups - ((i - 1) %% num_groups)
+    }
+    df$groups[i] <- group_id
+  }
+
+  # Reset game stats for group play
+  df$game_wins <- 0
+  df$game_losses <- 0
+  df$rank_hat <- NA
+
+  all_group_results <- list()
+
+  # Simulate each group
+  for (g in 1:num_groups) {
+    group_teams <- df[df$groups == g, ]
+    n_group_teams <- nrow(group_teams)
+
+    # Play round robin within the group
+    for (r in 1:rounds) {
+      for (i in 1:(n_group_teams - 1)) {
+        for (j in (i + 1):n_group_teams) {
+          team1 <- group_teams$true_rank[i]
+          team2 <- group_teams$true_rank[j]
+
+          strength1 <- group_teams$theta[i]
+          strength2 <- group_teams$theta[j]
+
+          match_winner <- simulate_match(team1, team2, strength1, strength2)
+
+          # Update wins and losses
+          if (match_winner == team1) {
+            df$game_wins[df$true_rank == team1] <- df$game_wins[df$true_rank == team1] + 1
+            df$game_losses[df$true_rank == team2] <- df$game_losses[df$true_rank == team2] + 1
+          } else {
+            df$game_wins[df$true_rank == team2] <- df$game_wins[df$true_rank == team2] + 1
+            df$game_losses[df$true_rank == team1] <- df$game_losses[df$true_rank == team1] + 1
           }
@@
     }
-  }
-  results <- data.frame(
-    true_rank = teams,
-    game_wins = win_counts,
-    game_loses = loss_counts,
-    true_strength = strengths$true_strength,
-    groups = strengths$groups
-  )
 
-  # Ranks teams based on number of wins (includes ties)
-  results$rank_hat <- rank(-results$game_wins, ties.method = "random")
-  results <- results[order(results$rank_hat),]
-
-  return(results)
+    # Rank teams within each group
+    group_df <- df[df$groups == g, ]
+    if (ties) {
+      group_df$group_rank <- rank(-group_df$game_wins, ties.method = "average")
+    } else {
+      group_df$group_rank <- rank(-group_df$game_wins, ties.method = "random")
     }
 
-simulate_groups <- function(group_length, num_teams, distribution, num_advance){
-  teams <- paste(num_teams:1)
-  if (distribution == "Normal"){
-    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
-    normal_strengths <- data.frame(
-      true_rank = as.numeric(teams),
-      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
-      rank_hat = rep(NA,length(teams)),
-      game_wins = rep(0,length(teams)),
-      game_losses = rep(0,length(teams))
-    )
-    df <- arrange(normal_strengths, true_rank)
-  }
-  else if (distribution == "Same"){
-    strengths <- sapply(num_teams, function(n) {rep(0,n)})
-    normal_strengths <- data.frame(
-      true_rank = as.numeric(teams),
-      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
-      rank_hat = rep(NA,length(teams)),
-      game_wins = rep(0,length(teams)),
-      game_losses = rep(0,length(teams))
-    )
-    df <- arrange(normal_strengths, true_rank)
-  }
-  else if (distribution == "Uniform"){
-    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
-    unif_strength <- data.frame(
-      true_rank = as.numeric(teams),
-      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
-      game_wins = rep(0,length(teams)),
-      game_losses = rep(0,length(teams)),
-      rank_hat = rep(NA,length(teams))
-    )
-    df <- arrange(unif_strength, true_rank)
+    all_group_results[[g]] <- group_df
   }
 
-  num_groups <- num_teams / group_length
-  # Create tiers based on sorted true_strength
-  df <- df %>% arrange(desc(true_strength))
-  tiers <- split(df, ceiling(seq_along(df$true_rank) / num_groups))
+  # Combine all group results
+  final_df <- do.call(rbind, all_group_results)
 
-  # Initialize group vector
-  df$groups <- NA
+  # Calculate overall tournament ranking
+  # First by group rank, then by total wins, then by true rank as tiebreaker
+  final_df <- final_df %>%
+    arrange(group_rank, desc(game_wins), true_rank)
 
-  # Assign one team from each tier to each group randomly
-  for (tier in tiers) {
-    group_ids <- sample(1:num_groups)  # random group order for this tier
-    df$groups[match(tier$true_rank, df$true_rank)] <- group_ids
+  if (ties) {
+    final_df$rank_hat <- rank(interaction(final_df$group_rank, -final_df$game_wins), ties.method = "average")
+  } else {
+    final_df$rank_hat <- rank(interaction(final_df$group_rank, -final_df$game_wins, final_df$true_rank), ties.method = "random")
   }
 
-  all_results <- list()
-  for (g in unique(df$groups)){
-    group_df <- df[df$groups == g, ]
-    group_results <- simulate_group_play(strengths = group_df)
-    group_results$groups <- g
-    all_results[[g]] <- group_results
+  final_df <- final_df %>%
+    mutate(distribution = distribution) %>%
+    arrange(true_rank) %>%
+    mutate(rank_hat = ifelse(group_rank <= 2, group_rank, num_groups + group_rank)) %>%
+    select(true_rank, theta, game_wins, game_losses, groups, group_rank, rank_hat, distribution)
+
+  group_winners <- final_df %>%
+    filter(group_rank == 1) %>%
+    arrange(groups)
+
+  group_seconds <- final_df %>%
+    filter(group_rank == 2) %>%
+    arrange(groups)
+
+  n_groups <- length(unique(final_df$groups))
+  matchup_vector <- c()
+
+  # Build the alternating order
+  for (i in seq(1, n_groups, by = 2)) {
+    g1 <- group_winners$groups[i]
+    g2 <- group_winners$groups[i + 1]
+
+    winner_g1 <- group_winners$true_rank[group_winners$groups == g1]
+    second_g2 <- group_seconds$true_rank[group_seconds$groups == g2]
+
+    winner_g2 <- group_winners$true_rank[group_winners$groups == g2]
+    second_g1 <- group_seconds$true_rank[group_seconds$groups == g1]
+
+    # append in alternating order
+    matchup_vector <- c(matchup_vector, winner_g1, second_g2, winner_g2, second_g1)
   }
-  print(all_results)
-  final_df <- do.call(rbind, all_results)
-  teams_in_knockout <- final_df %>%
-    arrange(rank_hat, game_wins) %>%
-    head(num_advance * num_groups)
-  return(teams_in_knockout)
-}
-create_knockout_matches <- function(df) {
-  group_list <- sort(unique(df$groups))
-  matches <- list()
+  matchup_vector
 
-  for (i in seq(1, length(group_list), by = 2)) {
-    g1 <- group_list[i]
-    g2 <- group_list[i + 1]
+  knockout_results <- simulate_knockout(df = final_df, matchup_vector = matchup_vector, series = 1, third_place = TRUE)
 
-    # Get the winner from group g1 and runner-up from group g2
-    team1 <- df[df$groups == g1 & df$rank_hat == 1, ]
-    team2 <- df[df$groups == g2 & df$rank_hat == 2, ]
+  knockout_results <- knockout_results %>%
+    filter(!is.na(rank_hat))
+  final_df <- final_df %>%
+    filter(rank_hat > num_groups)
 
-    # Get the winner from group g2 and runner-up from group g1
-    team3 <- df[df$groups == g2 & df$rank_hat == 1, ]
-    team4 <- df[df$groups == g1 & df$rank_hat == 2, ]
+  combined_df <- bind_rows(
+  final_df %>% filter(rank_hat >= num_groups),   # group-stage only teams
+  knockout_results %>% filter(!is.na(rank_hat))  # knockout-stage teams
+) %>%
+  arrange(true_rank) %>%
+    mutate(rank_hat = rank(rank_hat, ties.method = "random"))
 
-    matches[[length(matches) + 1]] <- data.frame(
-      Match = paste("Match", length(matches) + 1),
-      TeamA = team1$true_rank,
-      TeamB = team2$true_rank
-    )
-
-    matches[[length(matches) + 1]] <- data.frame(
-      Match = paste("Match", length(matches) + 1),
-      TeamA = team3$true_rank,
-      TeamB = team4$true_rank
-    )
+  return(combined_df)
 }
 
-  do.call(rbind, matches)
-}
-simulate_knockout <- function(df, series=1, third_place = T){
+# Example usage:
+# Single simulation
+result <- simulate_group_stage_tournament(num_teams = 8, distribution = "Normal", group_length = 4, ties=F)
 
-  seeding_structure <- create_knockout_matches(df)
-
-  for (i in nrow(df)){
-    for (g in unique(df$groups)){
-
+group_stage_results <- list()
+for (teams in c(8)) {
+  group_stage_results[[as.character(teams)]] <- list()
+  for (replicate in 1:10000) {
+    # Store each simulation result in the list with unique indexing
+    group_stage_results[[as.character(teams)]][[replicate]] <- simulate_group_stage_tournament(
+      num_teams = teams,
+      distribution = "Normal",
+      group_length = 4,
+      ties = F
+    )
   }
@@
 
-  results <- list()  # Initialize the results list
+group_stage_results <- do.call(rbind, lapply(names(group_stage_results), function(teams) {
+  team_results <- do.call(rbind, group_stage_results[[teams]])
+  team_results$teams <- as.numeric(teams)
+  return(team_results)
+}))
 
-  for (i in 1:nrow(seeding_structure)) {
-    # Create a new data frame based on the current permutation in 'test'
-    permutation <- seeding_structure[i, ]
+group_stage_results %>% select(true_rank, game_wins, game_losses, groups, group_rank, rank_hat)
+
+# Proportion of true #1 teams that finish group_rank > 2
+sum(group_stage_results$true_rank == 1 & group_stage_results$group_rank > 2) /
+sum(group_stage_results$true_rank == 1)
+
+```
+
+## Simulate Knockout Round
+
+```{r}
+simulate_knockout <- function(df, matchup_vector, series = 1, third_place = TRUE) {
+
+  teams <- matchup_vector
   permuted_df <- df
-    permuted_df$true_rank <- df$true_rank[permutation]  # Update team order 
-    teams <- permuted_df$true_rank
-    permuted_df$true_strength <- df$true_strength[permutation]
-    permuted_df$rank_hat <- NA  # Initialize ranks column
-    permuted_df$game_wins <- 0  # Initialize wins column
-    permuted_df$game_losses <- 0
+  permuted_df$rank_hat <- NA
 
-    strengths <- permuted_df$true_strength
   round_number <- ceiling(log2(length(teams)))
-
-    # Track the losers of the semifinals for the third-place match
   losers_semis <- c()
@@
   while (length(teams) > 1) {
+
     #cat("\n--- New Round ---\n")
-      next_round <- c() # Makes sure only winners are included in next round
+    next_round <- c()
 
-      for (j in seq(1, length(teams), by = 2)) { # Gets 2 teams for each match
+    for (j in seq(1, length(teams), by = 2)) {
       team1 <- teams[j]
@@
 
-        strength1 <- strengths[which(permuted_df$true_rank == team1)]
-        strength2 <- strengths[which(permuted_df$true_rank == team2)]
+      #cat(sprintf("Matchup: Team1 = %s | Team2 = %s\n", team1, team2))
+
+      strength1 <- permuted_df$theta[which(permuted_df$true_rank == team1)]
+      strength2 <- permuted_df$theta[which(permuted_df$true_rank == team2)]
+
+      #cat(sprintf("Strengths: %.2f vs %.2f\n", strength1, strength2))
 
@@
       losses_team2 <- 0
-        for (game in 1:series){
+
+      for (game in 1:series) {
         match_winner <- simulate_match(team1, team2, strength1, strength2)
-          #cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
-          if(!is.na(team1)) {
+        #cat(sprintf("Game %d Winner: %s\n", game, match_winner))
+
         if (match_winner == team1) {
@@
         }
-          }
-          else {wins_team2 <- series / 2 + 1}
-
-          # Determine the overall winner if a majority is reached
-          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
-            break
-          }
+        if (wins_team1 > series / 2 || wins_team2 > series / 2) break
       }
 
-        # adds wins and losses of the series to the teams unless they have a bye
-        if (!any(is.na(c(team1, team2)))) {
+      # update df
       permuted_df <- permuted_df %>%
@@
         mutate(game_losses = ifelse(true_rank == team2, game_losses + losses_team2, game_losses))
+
+      # winner moves to next round
+      series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
+      #cat(sprintf("Series Winner: %s\n", series_winner))
+      next_round <- c(next_round, series_winner)
+
+      # assign ranks
+      if (round_number > 2) {  # early rounds
+        loser <- ifelse(series_winner == team1, team2, team1)
+        permuted_df$rank_hat[permuted_df$true_rank == loser] <- 2^round_number
+      } else if (round_number == 2) {  # semifinals
+        loser <- ifelse(series_winner == team1, team2, team1)
+        permuted_df$rank_hat[permuted_df$true_rank == loser] <- 4
+        losers_semis <- c(losers_semis, loser)
+      } else if (round_number == 1) {  # finals
+        winner <- series_winner
+        loser <- ifelse(series_winner == team1, team2, team1)
+        permuted_df$rank_hat[permuted_df$true_rank == winner] <- 1
+        permuted_df$rank_hat[permuted_df$true_rank == loser] <- 2
       }
 
-        # Determine the match winner based on series results
-        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
-        #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
-        next_round <- c(next_round, series_winner)
-
-        if (round_number > 2){
-          # Assign ranks based on the current round
-          n <- 2^(round_number)
-          #print(series_winner)
-            if (!is.na(team1) & series_winner == team1) {
-                permuted_df$rank_hat[permuted_df$true_rank == team2] <- n
-            } else {
-                permuted_df$rank_hat[permuted_df$true_rank == team1] <- n
-            }
-        } else if (round_number == 2) {  # Semifinals
-          n <- 2^(round_number)
-          if (match_winner == team1) {
-              permuted_df$rank_hat[permuted_df$true_rank == team2] <- n
-            losers_semis <- c(losers_semis, team2)
-          } else {
-              permuted_df$rank_hat[permuted_df$true_rank == team1] <- n
-            losers_semis <- c(losers_semis, team1)
-          }
-        }else if (round_number == 1) { # Finals
-          permuted_df <- permuted_df %>%
-            filter(!is.na(true_rank))
-          n <- 2^(round_number)
-          if (series_winner == team1) {
-            permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
-            permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
-          } else {
-            permuted_df$rank_hat[permuted_df$true_rank == team1] <- 2
-            permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
-          }
-        }
     }
 
-      teams <- next_round  # Winners move to the next round
+    teams <- next_round
     round_number <- round_number - 1
@@
 
-    # Third place match between the losers of the semifinals
-    if (third_place == TRUE){
+  # third place match
+  if (third_place & length(losers_semis) == 2) {
     team1 <- losers_semis[1]
     team2 <- losers_semis[2]
-      strength1 <- permuted_df$true_strength[permuted_df$true_rank == team1]
-      strength2 <- permuted_df$true_strength[permuted_df$true_rank == team2]
+    strength1 <- permuted_df$theta[permuted_df$true_rank == team1]
+    strength2 <- permuted_df$theta[permuted_df$true_rank == team2]
+
+    #cat(sprintf("\nThird Place Match: %s vs %s | Strengths: %.2f vs %.2f\n", team1, team2, strength1, strength2))
+
     wins_team1 <- 0
     wins_team2 <- 0
+
     for (game in 1:series) {
       match_winner <- simulate_match(team1, team2, strength1, strength2)
-        if (match_winner == team1) {
-          wins_team1 <- wins_team1 + 1
-        } else {
-          wins_team2 <- wins_team2 + 1
+      #cat(sprintf("Game %d Winner: %s\n", game, match_winner))
+
+      if (match_winner == team1) wins_team1 <- wins_team1 + 1 else wins_team2 <- wins_team2 + 1
+      if (wins_team1 > series / 2 || wins_team2 > series / 2) break
     }
-        # Determine the overall winner if a majority is reached
-        if (wins_team1 > series / 2 || wins_team2 > series / 2) {
-          break
-        }
-      }
-      # Assign third place and fourth place ranks
+
     if (wins_team1 > wins_team2) {
-        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team1))
       permuted_df$rank_hat[permuted_df$true_rank == team1] <- 3
@@
     } else {
-        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team2))
       permuted_df$rank_hat[permuted_df$true_rank == team2] <- 3
@@
 
-    permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
-    results[[i]] <- permuted_df # Store the results of this permutation
+  permuted_df <- permuted_df %>% arrange(true_rank)
+  return(permuted_df)
 }
 
-  final_results <- do.call(rbind, results)
-
-  # Remove ties if ties = FALSE
-  if (!ties) {
-    final_results <- final_results %>%
-      filter(!is.na(true_rank)) %>%
-      arrange(true_rank) %>%
-      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
-  } else{
-    final_results <- final_results %>%
-      filter(!is.na(true_rank)) %>%
-      arrange(rank_hat) %>%
-      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
-  }
-
-  final_results <- final_results %>%
-    mutate(distribution = distribution) %>%
-    mutate(true_rank = as.numeric(true_rank)) %>%
-    arrange(true_rank)
-  return(final_results)
-}
-
-knockout_teams <- simulate_groups(4,8,"Normal",2)
-create_knockout_matches(knockout_teams)
 ```

### 2025-10-10 mutual_information_curves.qmd
CMSAC presentation

--- 2025-02-27 mutual_information_curves.qmd
@@
 
-From Wikipedia, weight mutual information formula:
+From Wikipedia, weighted mutual information formula:
 
 $I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}$
+
+Rajski's distance: $1 - \frac{I(X,Y)}{H(X,Y)}$
+
+Interpretation: the fraction of uncertainty not explained by mutual information
+
+<https://en.wikipedia.org/wiki/Inverse_distance_weighting>
+
+Every permutation is not possible to elimination tournaments, so should we divide by every possible permutation for 8 teams, or just every possible permutation in the tournament structure?
+
+## Weight Structure (Partial Kendall or Weighted Kendall)
+
+```{r}
+library(gtools)
+library(tidyverse)
+partial_kendall <- function(vec, k, weights = NULL){
+  n <- length(vec) 
+  if (k == n){k <- n-1} 
+  ref <- c(1:n)
+
+  if(is.null(weights)){
+    weights <- rep(1,n)
+  }
+
+  count <- 0 
+  #denom <- sum(n - (1:k))
+
+  for (i in 1:k){ 
+    for (j in (i+1):n){
+      #pair_weight <- weights[i] + weights[j]
+
+        add <- (((vec[i] < vec[j]) & (ref[i] < ref[j])) | 
+                  ((vec[i] > vec[j]) & (ref[i] > ref[j]))) + 0
+        #print(c(i, j, add, pair_weight)) 
+        #if (add == 1){
+          count <- count + add #+ pair_weight
+        #} 
+        denom <- denom + pair_weight
+    } 
+  } 
+  weight <- count / denom
+
+  return(weight) 
+  }
+
+weighted_kendall <- function(rhat, k, wtype = "symmetric_salience", salient_weights = NULL) {
+  n <- length(rhat)
+  #salient_weights needs to be a vector of length k
+
+  if(is.null(salient_weights)){
+    salient_weights <- rep(1,k)
+  }
+
+  if (is.null(wtype)) {
+    w <- matrix(0, ncol = n, nrow = n)
+    for (q in k:1){
+      w[q, ] <- salient_weights[q]
+      w[, q] <- salient_weights[q]
+    }
+    w <- w/2
+  }
+
+  if (wtype == "symmetric_salience") {
+    w <- matrix(0, ncol = n, nrow = n)
+    for (q in k:1) {
+      w[q, ] <- salient_weights[q]
+      w[, q] <- salient_weights[q]
+    }
+    #w[upper.tri(w)] <- w[upper.tri(w)]/2
+    #w[lower.tri(w)] <- w[lower.tri(w)]/2
+  }
+
+  #Ryan method
+  if (wtype == "rydog"){
+  w <- matrix(0, ncol = n, nrow = n)
+#  w[1:k, 1:k] <- 1
+  for (q in k:1) {
+    w[1:q, 1:q] <- salient_weights[q]
+    w[1:q, 1:q] <- salient_weights[q]
+  }
+  #w[upper.tri(w)] <- w[upper.tri(w)]/2
+  #w[lower.tri(w)] <- w[lower.tri(w)]/2
+  }
+
+  summ <- 0
+  for (i in 1:n) {
+    for (j in 1:n) {
+      if (i != j) {
+        summ <- summ + w[i, j] * sign(i - j) * sign(rhat[i] - rhat[j])
+        print(summ)
+      }
+
+    }
+  }
+
+  out <- summ / (sum(w) - sum(diag(w)))
+  weight <- (out + 1) / 2
+  print(sum(w))
+  print(sum(diag(w)))
+  return(weight)
+
+}
+weighted_kendall(c(2,1,3,4), k=4, salient_weights = c(3,1,1,0))
+weighted_kendall(c(2,1,3,4), k=2)
+
+perms_list <- permn(1:5)
+
+# Apply weighted_kendall to each permutation
+results <- sapply(perms_list, function(p) {
+  weighted_kendall(rhat = p, k = 5, wtype = "symmetric_salience")
+})
+
+# Put into a dataframe
+kendall_df <- data.frame(
+  permutation = sapply(perms_list, paste, collapse = ""),
+  weighted_kendall = results
+)
+
+kendall_df
+
+sum(kendall_df$weighted_kendall)
+mean(kendall_df$weighted_kendall)
+```
 
@@
 library(tidyverse)
+library(viridis)
 
-mut_info <- function(sims, num_teams, rank_weights = NULL){
+mut_info <- function(sims, num_teams, rank_weights = NULL, wtype = "symmetric_salience"){
   # Handle missing weights (default to equal weights if not provided)
@@
   for (i in 1:num_teams) {
-    x_i <- paste0("x", i)
-    y_i <- paste0("y", i)
+    #x_i <- paste0("x", i)
+    #y_i <- paste0("y", i)
 
+    #max_footrule <- sum(rank_weights * abs((1:num_teams) - (num_teams:1)))
     # Loop through each row of prob_df
     for (j in 1:nrow(prob_df)) {
+      x_vals <- as.numeric(prob_df[j, paste0("x", 1:num_teams)])
+      y_vals <- as.numeric(prob_df[j, paste0("y", 1:num_teams)])
+
       # Calculate the squared difference between x_i and y_i for each row j
-      diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
-    w[j] <- w[j] + rank_weights[i]* diff^2  # Sum squared differences for each row
+      #diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
+      #rank_diff <- abs(rank(x_vals) - rank(y_vals))
+      #w[j] <- w[j] + rank_weights[i] * diff^2  # Sum squared differences for each row
+
+      # Weighted Kendall's Tau correlation
+      #x_rep <- rep(x_vals, rank_weights)
+      #y_rep <- rep(y_vals, rank_weights)
+      #tau <- cor(x_rep, y_rep, method="kendall")
+      #if(!is.na(tau)){
+        #w[j] <- (1 - tau) / 2 * choose(length(x_rep),2)
+      #} else{
+      #  diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
+      #  w[j] <- w[j] <- w[j] + rank_weights[i] * diff^2
+      #}
+
+      # Truncated Normalized Kendall's Tau Distance
+      # Function to compute normalized "swap similarity" for top-K
+      w[j] <- weighted_kendall(rhat = y_vals, k=num_teams, salient_weights = rank_weights, wtype = wtype)
+
+      # Initialize weighted concordance / discordance sum
+      #weighted_inversions <- 0
+
+      #for (l in 1:i) {
+      #  for (k in (l+1):num_teams) {
+      #    concordant <- (y_vals[l] - y_vals[k]) > 0
+      #    discordant <- (y_vals[l] - y_vals[k]) < 0
+
+          # Weight each pair by sum of their rank_weights
+      #    pair_weight <- rank_weights[l] + rank_weights[k]
+
+      #    if (discordant) weighted_inversions <- weighted_inversions + pair_weight
+      #  }
+      #}
+    #w[j] <- weighted_inversions
+
+        # Calculate NDCG (Normalized discounted cummulative gain)
+      # x_vals tells us the true ranking positions
+      # y_vals tells us the predicted ranking positions
+      #relevance <- rank_weights * (num_teams - x_vals + 1)
+      # Calculate DCG based on predicted ranking (y_vals)
+      #dcg <- 0
+      # for (pos in 1:num_teams) {
+      #   # Find which team is at position 'pos' in the predicted ranking
+      #   team_idx <- which(y_vals == pos)
+      #   if (length(team_idx) > 0) {
+      #     # Use that team's relevance weight
+      #     dcg <- dcg + relevance[team_idx] / log2(pos + 1)
+      #   }
+      # }
+      # # Calculate ideal DCG (teams sorted by their rank_weights in descending order)
+      # ideal_order <- order(relevance, decreasing = TRUE)
+      # idcg <- 0
+      # for (pos in 1:num_teams) {
+      #   idcg <- idcg + relevance[ideal_order[pos]] / log2(pos + 1)
+      # }
+      # # NDCG = DCG / IDCG
+      # if (idcg > 0) {
+      #   ndcg <- dcg / idcg
+      #   # Convert NDCG (0 to 1, where 1 is perfect) to a distance measure
+      #   # Use (1 - NDCG) so that perfect match = 0 distance
+      #   w[j] <- 1 - ndcg
+      # } else {
+      #   w[j] <- 1  # Maximum distance if IDCG is 0
+      # }
+
+      # Calculate weighted Spearman's footrule distance
+      # Footrule = sum of |true_rank - predicted_rank| for each item
+      # Weighted by rank_weights for each team
+
+      #footrule_distance <- 0
+      #for (i in 1:num_teams) {
+        # Team i has true rank x_vals[i] and predicted rank y_vals[i]
+        #footrule_distance <- footrule_distance + rank_weights[i] * abs(x_vals[i] - y_vals[i])
+      #}
+
+      # Normalize by maximum possible distance
+      #if (max_footrule > 0) {
+      #  w[j] <- footrule_distance / max_footrule
+      #} else {
+      #  w[j] <- 0  # All weights are 0, so distance is 0
+      #}
+
     }
   }
-  w <- ifelse(w == 0.0000, 1, 1 / w)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
-  w <- w/sum(w)
+  #w <- 1 / (w+1)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
+  #w <- w/sum(w)
   # Compute weights dynamically for each row
@@
 
-  mutual_info <- factorial(num_teams) * sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/nrow(weighted_df) * weighted_df$probability)))
+  numerator <- sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/factorial(num_teams) * (1/factorial(num_teams)))))
+  denominator <- log2(factorial(num_teams)^2)
 
+  Raj_dist <- 1-numerator/denominator
   # Return the weighted data frame with mutual information weights
-  return(mutual_info)
+  return(Raj_dist)
 }
-mut_info(all_results_norm,8, c(10,5,1,0,0,0,0,0))
-mut_info(de_norm_results_eight,8, c(10,5,1,0,0,0,0,0))
-mut_info(test_rr_noTies,8, c(10,5,1,0,0,0,0,0))
-mut_info(seven_games_se_results,8, c(10,5,1,0,0,0,0,0))
 
+# Single Elimination with reseeding
+reseeded_results_eight <- reseeded_results %>% filter(teams == 8)
+reseeded_results_eight$simulation <- rep(1:10000, each = nrow(reseeded_results_eight) / 10000)
+
+# Group Stage (like World Cup)
+group_stage_results_eight <- group_stage_results %>% filter(teams == 8)
+group_stage_results_eight$simulation <- rep(1:10000, each = nrow(group_stage_results_eight) / 10000)
+
+# Double Elimination
 de_norm_results_eight <- de_norm_results %>% filter(teams == 8)
 de_norm_results_eight$simulation <- rep(1:10000, each = nrow(de_norm_results_eight) / 10000)
+
+# Single Elimination with Normal Distribution
 all_results_norm$simulation <- rep(1:10000, each = nrow(all_results_norm) / 10000)
+
+# Single Elimination with 7 game series
 seven_games_se_results$simulation <- rep(1:10000, each = nrow(seven_games_se_results) / 10000)
-# with randomly assigning ties
+
+# ROund Robin with randomly assigning ties
 test_rr_noTies <- test_rr %>%
@@
 
-# Initialize variables
-num_vars <- 8  # Total length of the vector
-indices <- 1:num_vars  # Store the number of 1's in each step
+# All possible orderings of teams one time
+perms_list <- permn(8)
+all_data <- list()
+# Loop through each permutation and create the data frame
+for (i in 1:length(perms_list)) {
+# Create a temporary data frame for the current permutation
+temp_df <- data.frame(
+true_rank = rep(1:8, each = 1),  # Each row gets a true_rank from 1 to 8
+rank_hat = perms_list[[i]],      # The current permutation's ranks
+simulation = rep(i, 8)           # Assign the simulation number to each row
+)
+# Add the temporary data frame to the list
+all_data[[i]] <- temp_df
+}
+# Combine all the data frames into one
+perms_df <- do.call(rbind, all_data)
+
+# Best Possible (every team in correct order)
+best <- data.frame(
+  true_rank = rep(1:8, times = 10000),  # Repeat 1-8 for 10,000 simulations
+  rank_hat = rep(1:8, times = 10000),   # Ensure rank_hat matches true_rank
+  simulation = rep(1:10000, each = 8)   # Assign each set to a simulation
+)
+# Worst Possible (worst team always wins; best last)
+worst <- data.frame(
+  true_rank = rep(1:8, times = 10000),  # Repeat 1-8 for 10,000 simulations
+  rank_hat = rep(8:1, times = 10000),   # Ensure rank_hat matches true_rank
+  simulation = rep(1:10000, each = 8)   # Assign each set to a simulation
+)
 
@@
 # Compute mutual information for each dataset
+mi_se_reseeded <- compute_mut_plots(reseeded_results_eight)
+mi_group_stage <- compute_mut_plots(group_stage_results_eight)
 mi_de_norm <- compute_mut_plots(de_norm_results_eight)
@@
 mi_seven_se <- compute_mut_plots(seven_games_se_results)
+mi_random <- compute_mut_plots(equal_strengths)
+mi_se_unif <- compute_mut_plots(test_unif)
+mi_bad <- compute_mut_plots(bad_seeding_structure)
+mi_perms <- compute_mut_plots(perms_df)
+mi_best <- compute_mut_plots(best)
+mi_worst <- compute_mut_plots(worst)
 
-# Plot results
-plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm, mi_test_rr, mi_seven_se)),
-     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Mutual Information vs. Number of 1's")
+mi_plots_df <- data.frame(
+indices = 1:8,
+"Single Elimination (Reseeded)" = 1-mi_se_reseeded,
+"Group Stage" = 1-mi_group_stage,
+"Double Elimination" = 1-mi_de_norm,
+"Single Elimination (Set Structure)" = 1-mi_se_norm,
+"Coin Flip Single Elimination" = 1-mi_random,
+#"Unif_SE" = mi_se_unif,
+"Bad Seeding Structure Single Elimination" = 1-mi_bad,
+"Coin Flip Round Robin" = 1-mi_perms,
+"Round Robin" = 1-mi_test_rr,
+"Worst Possible" = 1-mi_worst,
+"Best Possible" = 1-mi_best,
+check.names = FALSE
+)
 
-# Add other datasets
-lines(indices, mi_se_norm, type="b", pch=19, col="red")
-lines(indices, mi_test_rr, type="b", pch=19, col="green")
-lines(indices, mi_seven_se, type = "b", pch=19, col="black")
+# Reshape data into long format for ggplot
+df_long <- pivot_longer(mi_plots_df, cols = -indices, names_to = "Method", values_to = "MI")
 
-# Add legend
-legend("topleft", legend=c("Double Elimination", "Single Elimination", "Round Robin", "7 Game Single Elimination"), 
-       col=c("blue", "red", "green", "black"), pch=19, lty=1)
+okabe_ito <- c(
+  "#E69F00", # orange
+  "#56B4E9", # sky blue
+  "#009E73", # bluish green
+  "#F0E442", # yellow
+  "#0072B2", # blue
+  "#D55E00", # vermillion
+  "#CC79A7", # reddish purple
+  "#999999", # gray
+  "#8A2BE2", # violet (extra)
+  "#20B2AA"  # teal (extra)
+)
+# Define consistent color mapping
+methods <- unique(df_long$Method)
+method_colors <- setNames(okabe_ito, methods)
 
-### Single Elimination vs Double Elimination ###
+windowsFonts(Arial = windowsFont("Arial"))
+ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
+geom_line(linewidth = 2.5) +           # Add lines
+geom_point(size=3) +          # Add points
+labs(
+title = "Comparison of Tournament Structures with Normal Strengths",
+x = "Top i Number of Teams",
+y = "Normalized Mutual Information \\ Entropy (log10 scale)"
+) +
+theme_minimal() +       # Use a clean theme
+theme(text = element_text(family = "Arial", size = 32),
+  legend.position = "bottom",
+  legend.box = "vertical",
+  plot.title = element_text(hjust = 0.5, size = 36),
+  axis.title = element_text(size = 30),
+  axis.text = element_text(size=30)) +
+  guides(color = guide_legend(nrow = 5)) +
+  scale_y_log10(
+    labels = scales::label_number() 
+  ) +
+  scale_color_manual(values = method_colors)
+ggsave(
+  filename = "tournament_comparison_plot.png",  # Change file type if needed
+  plot = last_plot(),  # Saves the last ggplot created
+  dpi = 600,        
+  bg = "white",
+  units = "in",
+  width = 15.28,
+  height = 13.43
+)
 
-# Plot results
-plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm)),
-     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Single Elimination vs Double Elimination")
-
-# Add other datasets
-lines(indices, mi_se_norm, type="b", pch=19, col="red")
-
-# Add legend
-legend("bottomright", legend=c("Double Elimination", "Single Elimination"), 
-       col=c("blue", "red"), pch=19, lty=1)
+# Apply normalization to all columns except "indices"
+df_close <- df_long %>% filter(Method %in% c("Single Elimination (Reseeded)", "Double Elimination", "Group Stage", "Single Elimination (Set Structure)" ))
+windowsFonts(Arial = windowsFont("Arial"))
+ggplot(df_close, aes(x = indices, y = MI, color = Method, group = Method)) +
+geom_line(linewidth = 2.5) +           # Add lines
+geom_point(size=3) +          # Add points
+labs(
+title = "Comparison of Tournament Structures with Normal Strengths",
+x = "Top i Number of Teams",
+y = "Normalized Mutual Information \\ Entropy (log10 scale)"
+) +
+theme_minimal() +       # Use a clean theme
+theme(text = element_text(family = "Arial", size = 32),
+  legend.position = "bottom",
+  legend.box = "vertical",
+  plot.title = element_text(hjust = 0.5, size = 36),
+  axis.title = element_text(size = 30),
+  axis.text = element_text(size=30)) +
+  guides(color = guide_legend(nrow = 5)) +
+  scale_y_log10(
+    labels = scales::label_number() 
+  ) +
+  scale_color_manual(values = method_colors)
+ggsave(
+  filename = "tournament_comparison_plot_close.png",  # Change file type if needed
+  plot = last_plot(),  # Saves the last ggplot created
+  dpi = 600,        
+  bg = "white",
+  units = "in",
+  width = 15.28,
+  height = 13.43
+)
 
 ```
+
+```{r}
+compute_mut_plots_rydog <- function(dataset) {
+  num_teams <- length(unique(dataset$true_rank))
+  results <- numeric(num_teams)
+  for (i in 1:num_teams) {
+    binary_vector <- c(rep(1, i), rep(0, num_teams - i))  # Construct vector
+    results[i] <- mut_info(dataset, num_teams, binary_vector, wtype = "rydog")  # Compute MI
+  }
+  return(results)
+}
+
+# Compute mutual information for each dataset
+mi_se_reseeded_rydog <- compute_mut_plots_rydog(reseeded_results_eight)
+mi_group_stage_rydog <- compute_mut_plots_rydog(group_stage_results_eight)
+mi_de_norm_rydog <- compute_mut_plots_rydog(de_norm_results_eight)
+mi_se_norm_rydog <- compute_mut_plots_rydog(all_results_norm)
+mi_test_rr_rydog <- compute_mut_plots_rydog(test_rr_noTies)
+mi_seven_se_rydog <- compute_mut_plots_rydog(seven_games_se_results)
+mi_random_rydog <- compute_mut_plots_rydog(equal_strengths)
+mi_se_unif_rydog <- compute_mut_plots_rydog(test_unif)
+mi_bad_rydog <- compute_mut_plots_rydog(bad_seeding_structure)
+mi_perms_rydog <- compute_mut_plots_rydog(perms_df)
+mi_best_rydog <- compute_mut_plots_rydog(best)
+mi_worst_rydog <- compute_mut_plots_rydog(worst)
+
+mi_plots_df_rydog <- data.frame(
+indices = 1:8,
+"Single Elimination (Reseeded)" = 1-mi_se_reseeded_rydog,
+"Group Stage" = 1-mi_group_stage_rydog,
+"Double Elimination" = 1-mi_de_norm_rydog,
+"Single Elimination (Set Structure)" = 1-mi_se_norm_rydog,
+"Coin Flip Single Elimination" = 1-mi_random_rydog,
+#"Unif_SE" = mi_se_unif,
+"Bad Seeding Structure Single Elimination" = 1-mi_bad_rydog,
+"Coin Flip Round Robin" = 1-mi_perms_rydog,
+"Round Robin" = 1-mi_test_rr_rydog,
+"Worst Possible" = 1-mi_worst_rydog,
+"Best Possible" = 1-mi_best_rydog
+)
+
+# Reshape data into long format for ggplot
+df_long_rydog <- pivot_longer(mi_plots_df_rydog, cols = -indices, names_to = "Method", values_to = "MI")
+windowsFonts(Arial = windowsFont("Arial"))
+ggplot(df_long_rydog, aes(x = indices, y = MI, color = Method, group = Method)) +
+geom_line(linewidth = 2.5) +           # Add lines
+geom_point(size=3) +          # Add points
+labs(
+title = "Comparison of Tournament Structures with Normal Strengths (Rydog)",
+x = "Top i Number of Teams",
+y = "Normalized Mutual Information \\ Entropy (log10 scale)"
+) +
+theme_minimal() +       # Use a clean theme
+theme(text = element_text(family = "Arial", size = 32),
+  legend.position = "bottom",
+  legend.box = "vertical",
+  plot.title = element_text(hjust = 0.5, size = 36),
+  axis.title = element_text(size = 30),
+  axis.text = element_text(size=30)) +
+  guides(color = guide_legend(nrow = 5)) +
+  scale_y_log10(
+    labels = scales::label_number() 
+  ) +
+  scale_color_viridis_d(option = "H")
+ggsave(
+  filename = "tournament_comparison_plot_rydog.png",  # Change file type if needed
+  plot = last_plot(),  # Saves the last ggplot created
+  dpi = 600,        
+  bg = "white",
+  units = "in",
+  width = 20,
+  height = 13.43
+)
+
+# Apply normalization to all columns except "indices"
+df_close_rydog <- df_long_rydog %>% filter(Method %in% c("Single.Elimination..Reseeded.", "Double.Elimination", "Group.Stage", "Single.Elimination..Set.Structure." ))
+windowsFonts(Arial = windowsFont("Arial"))
+ggplot(df_close_rydog, aes(x = indices, y = MI, color = Method, group = Method)) +
+geom_line(linewidth = 2.5) +           # Add lines
+geom_point(size=3) +          # Add points
+labs(
+title = "Comparison of Tournament Structures with Normal Strengths (Rydog)",
+x = "Top i Number of Teams",
+y = "Normalized Mutual Information \\ Entropy (log10 scale)"
+) +
+theme_minimal() +       # Use a clean theme
+theme(text = element_text(family = "Arial", size = 32),
+  legend.position = "bottom",
+  legend.box = "vertical",
+  plot.title = element_text(hjust = 0.5, size = 36),
+  axis.title = element_text(size = 30),
+  axis.text = element_text(size=30)) +
+  guides(color = guide_legend(nrow = 5)) +
+  scale_y_log10(
+    labels = scales::label_number() 
+  ) +
+  scale_color_viridis_d(option = "H")
+ggsave(
+  filename = "tournament_comparison_plot_close_rydog.png",  # Change file type if needed
+  plot = last_plot(),  # Saves the last ggplot created
+  dpi = 600,        
+  bg = "white",
+  units = "in",
+  width = 20,
+  height = 13.43
+)
+```
+
+```{r}
+make_perms_df <- function(num_teams){
+  perms_list <- permn(1:num_teams)
+  all_data <- vector("list", length(perms_list))
+
+  for (i in seq_along(perms_list)) {
+    all_data[[i]] <- data.frame(
+      true_rank  = 1:num_teams,     # True rank positions
+      rank_hat   = perms_list[[i]], # The current permutation
+      simulation = i
+    )
+  }
+  do.call(rbind, all_data)
+}
+
+# Generate datasets for 4, 6, and 10 teams
+res <- data.frame(
+  num_teams = 2:8,
+  mutual_info = sapply(2:8, function(i){
+    df <- make_perms_df(i)
+    mut_info(df, num_teams = i)
+  })
+)
+
+res
+
+```
+
+Why is it always 0.25?
+
+$$
+I(R;\hat{R}) = \sum_{R\in R} \sum_{\hat{R} \in \hat{R}} w(R,\hat{R})\times p(R,\hat{R})\times \log_2(\frac{p(R,\hat{R})}{p(R)\times p(\hat{R})})
+$$
+
+since every outcome is equally likely: $p(R,\hat{R}) = \frac{1}{n!}$
+
+because $R$ and $\hat{R}$ are independent and have equal probabilities,
+
+$H(R,\hat{R}) = H(R) + H(\hat{R}) = -2\sum_{x \in X} p(x)\log(p(x)) = -2(n!) \times (1/n!) \times\log_2(1/n!) = \log_2((n!)^2)$
+
+$$
+\frac{I(R,\hat{R})}{H(R,\hat{R})} = \frac{\sum \sum w(R,\hat{R}) \times \frac{1}{n!} \times \log_2 (\frac{1/n!}{(1/n!)^2})} {\log_2((n!)^2)}
+$$
+
+$\sum \sum w(R,\hat{R}) = 0.5\times n!$
+
+$$
+= \frac{0.5 \times n!\times \frac{1}{n!} \times \log_2(n!)}{2\times \log_2(n!)} =  \frac{0.5}{2} = 0.25
+$$
+
+include number of games (might be a range)
+
+Single Elimination - third place and without third place; series = 3,5,7
+
+Double Elimination
+
+-   Repechage
+
+-   True Double Elimination
+
+-   Consolation Bracket (loser's bracket best place is third)
+
+Triple Elimination
+
+Deterministic Outcomes
+
+Round Robin - with k rounds (1,2,3,5,7)
+
+Stepladder - 8 vs 7, winner plays 6, etc.
+
+Group Stage - winners of group advance, top 2 advance
+
+Staged Round Robin/Jacobian Ladder - top 2 in each group form new group, bottom 2 new group (each new group is a new round robin)
+
+High School Tennis Ladder - 1 vs 2, 3 vs 4, etc., Re-rank 2 vs 3, 4 vs 5, 6 vs 7, number of teams - 1 iterations

### 2025-10-10 single_elimination_shiny.qmd
CMSAC presentation

@@
 # Simulate the single-elimination tournament
-simulate_single_elimination_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1, third_place = T) {
+simulate_single_elimination_tournament <- function(num_teams, distribution, seeding_structure = NULL, ties = T, series = 1, third_place = T, reseeded = F) {
 
@@
   }
+
   if (distribution == "Normal"){
@@
       true_rank = as.numeric(teams),
-      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      theta = c(strengths, rep(NA, length(teams) - num_teams)),
       rank_hat = rep(NA,length(teams)),
@@
       true_rank = as.numeric(teams),
-      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      theta = c(strengths, rep(NA, length(teams) - num_teams)),
       game_wins = rep(0,length(teams)),
@@
       true_rank = as.numeric(teams),
-      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      theta = c(strengths, rep(NA, length(teams) - num_teams)),
       rank_hat = rep(NA, length(teams)),
@@
     teams <- permuted_df$true_rank
-    permuted_df$true_strength <- df$true_strength[permutation]
+    permuted_df$theta <- df$theta[permutation]
     permuted_df$rank_hat <- NA  # Initialize ranks column
@@
 
-    strengths <- permuted_df$true_strength
+    strengths <- permuted_df$theta
     round_number <- ceiling(log2(length(teams)))
@@
       #cat("\n--- New Round ---\n")
+
+      # RESEEDING LOGIC: If reseeded is TRUE, sort teams by their true rank before each round
+      if (reseeded && length(teams) > 1) {
+        # Remove NA teams for sorting
+        valid_teams <- teams[!is.na(teams)]
+        na_teams <- teams[is.na(teams)]
+
+        # Sort valid teams by their true rank (best to worst)
+        sorted_teams <- valid_teams[order(as.numeric(valid_teams))]
+
+        # Create reseeded matchups: 1 vs worst, 2nd vs 2nd worst, etc.
+        n_valid <- length(sorted_teams)
+        if (n_valid > 1) {
+          reseeded_teams <- c()
+          for (j in 1:(n_valid/2)) {
+            reseeded_teams <- c(reseeded_teams, sorted_teams[j], sorted_teams[n_valid - j + 1])
+          }
+          # Add back any NA teams (byes)
+          teams <- c(reseeded_teams, na_teams)
+        }
+      }
+
       next_round <- c() # Makes sure only winners are included in next round
@@
           }
-        }else if (round_number == 1) { # Finals
+        } else if (round_number == 1) { # Finals
           permuted_df <- permuted_df %>%
@@
       team2 <- losers_semis[2]
-      strength1 <- permuted_df$true_strength[permuted_df$true_rank == team1]
-      strength2 <- permuted_df$true_strength[permuted_df$true_rank == team2]
+      strength1 <- permuted_df$theta[permuted_df$true_rank == team1]
+      strength2 <- permuted_df$theta[permuted_df$true_rank == team2]
       wins_team1 <- 0
@@
 }
-}
-test <- simulate_single_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, third_place=T, seeding_structure = c(1,2,3,4,5,6,7,8))
+
+test <- simulate_single_elimination_tournament(num_teams = 8, distribution = "Same", ties=F, series=1, third_place=T, seeding_structure = c(1,2,3,4,5,6,7,8), reseeded = TRUE)
 test
@@
 }))
+
+reseeded_results <- list()
+for (teams in c(8)) {
+  reseeded_results[[as.character(teams)]] <- list()
+  for (replicate in 1:10000) {
+    # Store each simulation result in the list with unique indexing
+    reseeded_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
+      num_teams = teams,
+      distribution = "Normal",
+      ties = FALSE,
+      series = 1,
+      third_place = TRUE,
+      reseeded = TRUE
+    )
+  }
+}
+
+reseeded_results <- do.call(rbind, lapply(names(reseeded_results), function(teams) {
+  team_results <- do.call(rbind, reseeded_results[[teams]])
+  team_results$teams <- as.numeric(teams)
+  return(team_results)
+}))
+
+reseeded_results
 

### 2025-10-10 CMSAC_presentation.Rmd
new intro slides draft

@@
 -   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors' true strengths
+
+## New Slide - Extreme Tournament
+
+- 4 team single elimination tournament:
+  - Dream Team vs Work Group vs Middle School Girls Team vs Babies
+- If we only care about the top team winning, seeding structure does not matter
+- If we care about who gets second and third, seeding becomes important
+  - Would not pair up Work Group and Dream Team in the first round
+
+## New
+
+- Comparing different competition structures, which competition structure is best?
+
+- Efficacy: Final rankings should match the true rankings of the strength of the competitors
+
+- Effectivity: Interested in correctly ranking the best competitor as first
+
+CITEE
+
+## New
+
+- Other competitions where the correct ranking of some, but not all, of the competitors is of import
+  - Olympic competition ranks the top 3 competitors
+
+- "Salient ranks": the ranks of a competition structure that matter
+  - Efficacy: 1 through $n$ matter
+  - Effectivity: 1 is only salient rank
+
+- We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths
 
@@
 knitr::include_graphics("tournament_comparison_plot.png")
-knitr::include_graphics("tournament_comparison_plot_close.png")
+#knitr::include_graphics("tournament_comparison_plot_close.png")
 ```

### 2025-10-10 CMSAC_reveal_js.qmd
Reveal_js Presentation

--- 2025-10-10 CMSAC_presentation.Rmd
@@
 ---
-title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
-author: "Zach Culp, Josie Peterburs, Dr. Gregory Matthews, Dr. Ryan McShane"
-date: "`r Sys.Date()`"
-bibliography: references.bib
-fontsize: 12pt
-output:
-  beamer_presentation:
-    theme: default
+title: "CMSAC_reveal_js"
+author: "Zach Culp, Josie Peterburs, Dr. Gregory J Matthews, Dr. Ryan P.A. McShane"
+format: revealjs
+editor: visual
 ---
@@
 
+##
+
 ## Introduction
+
+```{=html}
 <!-- Overall comments 
@@
 with the usage of theta or calling it true rank/true strength. -->
+```
 
@@
 ## Tournament Simulations
+
 -   Each competition structure is simulated given the number of teams, $n$, and $\theta$
@@
 
-- The notation for mutual information is: 
-$$ I(X;Y) = \sum_{y \in Y}\sum_{x \in X} p(x,y) \log{\frac{p(x,y)}{p(x)p(y)}} $$
+-   The notation for mutual information is: $$ I(X;Y) = \sum_{y \in Y}\sum_{x \in X} p(x,y) \log{\frac{p(x,y)}{p(x)p(y)}} $$
+
+<!-- -->
+
 -   In our case, it can be simplified to:
+
     $$ I(r, \hat{r}) = n! \sum_{\hat{r}} p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{(\frac{1}{n!})^2}} = n! \log{((n!)^2)} $$
@@
 ## Weighted Mutual Information
+
 -   A limitation of mutual information is that it does not distinguish between the direction of information
@@
 -   Because of this, we proposed a weighting function $w(r, \hat{r})$
-- So, the new formula is:
-$$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$
+-   So, the new formula is: $$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$
 
 ## Weighting Function
-- In our original paper submission, we used squared error loss as the weighting function:
-$$ w(r, \hat{r}) =
+
+-   In our original paper submission, we used squared error loss as the weighting function: $$ w(r, \hat{r}) =
     \begin{cases}
-\frac{1}{\hat{r}'r}, & r \neq \hat{r} \\
+    \frac{1}{(r-\hat{r})^2}, & r \neq \hat{r} \\
     1, & r = \hat{r}
     \end{cases} $$
+
 -   However, after comments and reviews from people at CSAS and CMSAC, we have decided to use a weighted Kendall's Tau
@@
 $$W_{ij} = w_{\min(i,j)}, \text{ for } i,j = m,2m, \dots, n\times m$$
-
 - $m$: vector of length $n$ representing the "salient weights", or the importance of correctly getting each ranking
@@
 $$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$
-
 ## Results

### 2025-10-10 CMSAC_reveal_js.qmd
new intro slides

@@
 ```
-
-##
 
@@
 <!-- Should we talk about Efficacy vs Effectivity here -->
+
+## Extreme Tournament
+
+- 4 team single elimination tournament:
+  - Dream Team vs Work Group vs Middle School Girls Team vs Babies
+- If we only care about the top team winning, seeding structure does not matter
+- If we care about who gets second and third, seeding becomes important
+  - Would not pair up Work Group and Dream Team in the first round
+
+## Comparing Structures
+
+- Comparing different structures like round robin and single eliminations, which is best?
+
+- Efficacy: Final rankings should match the true rankings of the strength of the competitors
+
+- Effectivity: Interested in correctly ranking the best competitor as first
+
+CITEE
+
+## Efficacy and Effectivity
+
+- Other competitions where the correct ranking of some, but not all, of the competitors is of import
+  - Olympic competition ranks the top 3 competitors
+
+- "Salient ranks": the ranks of a competition structure that matter
+  - Efficacy: 1 through $n$ matter
+  - Effectivity: 1 is only salient rank
+
+- We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths
 
@@
 ## Weighted Kendall
+
 $$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \times \text{sgn}(i-j)\times \text{sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$
@@
 
-$$W_{ij} = w_{\min(i,j)}, \text{ for } i,j = m,2m, \dots, n\times m$$
-- $m$: vector of length $n$ representing the "salient weights", or the importance of correctly getting each ranking
+$$W_{ij} = w_{\min(i,j)}, \text{ for } i,j = m,2m, \dots, n\times m$$ - $m$: vector of length $n$ representing the "salient weights", or the importance of correctly getting each ranking
 
@@
 ## Weighted Kendall Example
-- If $\hat{r} = (2,1,3,4)$ and m = $(3,1,1,0)$, then:
-$$ W = \begin{pmatrix}
+
+-   If $\hat{r} = (2,1,3,4)$ and m = $(3,1,1,0)$, then: $$ W = \begin{pmatrix}
     3 & 3 & 3 & 3 \\
@@
 ## Normalizing Mutual Information
+
 -   Mutual Information can be normalized by dividing by the joint entropy
@@
 
-- $H(r)$ and $H(\hat{r})$ are theoretically equivalent, leaving $H(r,\hat{r}) = -2\sum_{r} p(r)\log(p(r)) = -2(n!) \times (1/n!) \times\log_2(1/n!) = \log_2((n!)^2)$
-<!-- Is this true?? I can't remember if they are equivalent or their maximums were equivalent -->
+-   $H(r)$ and $H(\hat{r})$ are theoretically equivalent, leaving $H(r,\hat{r}) = -2\sum_{r} p(r)\log(p(r)) = -2(n!) \times (1/n!) \times\log_2(1/n!) = \log_2((n!)^2)$ <!-- Is this true??
+I can't remember if they are equivalent or their maximums were equivalent -->
 
-- So, the final formula for the metric is:
-$$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$
-## Results
+-   So, the final formula for the metric is: $$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$ \## Results
 
 <!-- Should these plots be on separate slides? -->
+
 ```{r out.width="49%", fig.show='hold', fig.align='center'}
@@
 ## Limitations
+
 -   The "known" strengths are assumptions that can be estimated but never truly known
@@
 ## Future Work
+
 -   Evaluate structures using various distributions and "known" weights
@@
 ## References
+
 <!-- I don't know if the references.bib would work here or not -->
+
 -   Appleton, David R. 1995. “May the Best Man Win?” Statistician 44 (4): 529.
-- Csato, L. 2021. Tournament Design: How Operations Research Can Improve Sports Rules.
-1st Ed. Cham, Switzerland: Palgrave Macmillan.
-- Devriesere, Csató, K., and D. Goossens. 2025. “Tournament Design: A Review from an
-Operational Research Perspective.” European Journal of Operational Research 324 (1):
-1–21.
-- Glenn, W A. 1960. “A Comparison of the Effectiveness of Tournaments.” Biometrika 47
-(3-4): 253–62.
+-   Csato, L. 2021. Tournament Design: How Operations Research Can Improve Sports Rules. 1st Ed. Cham, Switzerland: Palgrave Macmillan.
+-   Devriesere, Csató, K., and D. Goossens. 2025. “Tournament Design: A Review from an Operational Research Perspective.” European Journal of Operational Research 324 (1): 1–21.
+-   Glenn, W A. 1960. “A Comparison of the Effectiveness of Tournaments.” Biometrika 47 (3-4): 253–62.
 
 ## References
+
 -   Guiasu, Silviu. 1977. Information Theory with Applications. New York, NY: McGraw-Hill.
-- Johnson, Sidney, and Rodney Fort. 2022. “Match Outcome Uncertainty and Sports Fan
-Demand: An Agnostic Review and the Standard Economic Theory of Ports Leagues.”
-Int. J. Emp. Econ. 01 (02).
-- Lasek, Jan, and Marek Gagolewski. 2018. “The Efficacy of League Formats in Ranking
-Teams.” Stat. Modelling 18 (5-6): 411–35.
-- Sziklai, Balázs R, Péter Biró, and László Csató. 2022. “The Efficacy of Tournament Designs.”
-Comput. Oper. Res. 144 (105821): 105821.
+-   Johnson, Sidney, and Rodney Fort. 2022. “Match Outcome Uncertainty and Sports Fan Demand: An Agnostic Review and the Standard Economic Theory of Ports Leagues.” Int. J. Emp. Econ. 01 (02).
+-   Lasek, Jan, and Marek Gagolewski. 2018. “The Efficacy of League Formats in Ranking Teams.” Stat. Modelling 18 (5-6): 411–35.
+-   Sziklai, Balázs R, Péter Biró, and László Csató. 2022. “The Efficacy of Tournament Designs.” Comput. Oper. Res. 144 (105821): 105821.

### 2025-10-10 CMSAC_reveal_js.qmd
intro slides

@@
 title: "CMSAC_reveal_js"
-author: "Zach Culp, Josie Peterburs, Dr. Gregory J Matthews, Dr. Ryan P.A. McShane"
+author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
 format: revealjs
@@
 
--   So, the final formula for the metric is: $$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$ \## Results
+-   So, the final formula for the metric is: $$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$
+
+## Results
 

### 2025-10-10 josie slides.qmd
update intro slides

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
format: revealjs
editor: visual
date: "`r Sys.Date()`"
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tinytex)
library(tidyverse)
library(combinat)
load("curves_data.RData")
#load("tournament_comparison_plot.png")
#load("tournament_comparison_plot_close.png")
```

##

```{r}
knitr::include_graphics("4 bracket good.png")
```

##

:::: {.columns}

::: {.column width="50%"}
Good 
:::

::: {.column width="50%"}
Bad
:::

::::

```{r}
knitr::include_graphics("4 seeded tourn w arrow.png")
```

::: {.notes}
- 4 team single elimination tournament:
  - Dream Team vs Work Group vs Middle School Girls Team vs Babies
- If we only care about the top team winning, seeding structure does not matter
- If we care about who gets second and third, seeding becomes important
  - Would not pair up Work Group and Dream Team in the first round
:::

## Comparing Structures

- Which competition structure is best at reflecting the true strengths of competitors?

- **Efficacy**^[Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)]: Final rankings should match the true rankings of the strength of the competitors

- **Effectivity**^[Glenn (1960)]: Interested in correctly ranking the best competitor as first

## Efficacy and Effectivity

- Decide how many rankings are important
  - Olympic competition ranks the top 3 competitors
  - Premier League: top 4 (Champions League) and bottom 3 (Relegation)

- **Salient ranks**: the ranks that matter
  - Efficacy: 1 through $n$ matter
  - Effectivity: 1 is only salient rank

## Goal

- Want to quantify how well a structure can accurately order the competitors
- We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths

## Distribution of $\hat{r}$

```{r}
ranks <- seq(1:3)
permn(ranks) %>% unlist() %>% matrix(ncol = 3, byrow = TRUE)

```

### 2025-10-17 CMSAC_presentation/josie_slides.qmd
josie slides but having issues

--- 2025-10-10 josie slides.qmd
@@
 date: "`r Sys.Date()`"
-bibliography: references.bib
+bibliography: ../references.bib
 ---
@@
 library(combinat)
-load("curves_data.RData")
+# load("curves_data.RData")
 #load("tournament_comparison_plot.png")
@@
 
+::::: columns
+::: {.column width="50%"}
 ```{r}
@@
 ```
+:::
+
+::::: columns
+::: {.column width="50%"}
+```{r}
+knitr::include_graphics("Mens-Basketball-Olympics-Getty-H-2024.jpeg")
+knitr::include_graphics("five_rec_league_men_playing_basketball.png")
+knitr::include_graphics("jv_girls_basketball.png")
+knitr::include_graphics("five_babies_playing_basketball.png")
+
+```
+:::
+:::::
 
@@
 
-:::: {.columns}
-
+::::: columns
 ::: {.column width="50%"}
@@
 :::
-
-::::
+:::::
 
@@
 
-- **Efficacy**^[Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)]: Final rankings should match the true rankings of the strength of the competitors
+-   **Efficacy**[^1]: Final rankings should match the true rankings of the strength of the competitors
 
-- **Effectivity**^[Glenn (1960)]: Interested in correctly ranking the best competitor as first
+-   **Effectivity**[^2]: Interested in correctly ranking the best competitor as first
+
+[^1]: Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)
+
+[^2]: Glenn (1960)
 
@@
     -   Premier League: top 4 (Champions League) and bottom 3 (Relegation)
-
 -   **Salient ranks**: the ranks that matter

### 2025-10-17 CMSAC_presentation/josie_slides.qmd
updated with columns

@@
 
-::::: columns
 ::: {.column width="50%"}
@@
 
-::: {.notes}
+::: notes
 -   4 team single elimination tournament:
@@
 -   **Effectivity**[^2]: Interested in correctly ranking the best competitor as first
-
-[^1]: Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)
-
-[^2]: Glenn (1960)
 
@@
 ```
+::::::::::
+
+[^1]: Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)
+
+[^2]: Glenn (1960)

### 2025-10-17 CMSAC_presentation/CMSAC_reveal_js.qmd
moved into the correct folder.

--- 2025-10-10 CMSAC_reveal_js.qmd
@@
 
-```{=html}
 <!-- Overall comments 
@@
 with the usage of theta or calling it true rank/true strength. -->
-```
 
@@
     -   Olympic competition ranks the top 3 competitors
-
 -   "Salient ranks": the ranks of a competition structure that matter
@@
     -   Effectivity: 1 is only salient rank
-
 -   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths

### 2025-10-17 CMSAC_presentation/CMSAC_reveal_js.qmd
Updated with fragments.

@@
 library(tidyverse)
+library(combinat)
 load("curves_data.RData")
@@
 #load("tournament_comparison_plot_close.png")
+```
+
+##
+
+::::: columns
+::: {.column width="50%"}
+```{r}
+knitr::include_graphics("4 bracket good.png")
+```
+:::
+
+::: {.column width="50%"}
+::: {.fragment}
+```{r}
+knitr::include_graphics("Mens-Basketball-Olympics-Getty-H-2024.jpeg")
+```
+
+:::
+
+::: {.fragment}
+
+```{r}
+knitr::include_graphics("five_rec_league_men_playing_basketball.png")
+```
+
+:::
+
+::: {.fragment}
+
+```{r}
+knitr::include_graphics("jv_girls_basketball.png")
+```
+
+:::
+
+::: {.fragment}
+
+```{r}
+knitr::include_graphics("five_babies_playing_basketball.png")
+
+```
+:::
+
+:::
+:::::
+
+##
+
+::::: columns
+::: {.column width="50%"}
+Good
+:::
+
+::: {.column width="50%"}
+Bad
+:::
+:::::
+
+```{r}
+knitr::include_graphics("4 seeded tourn w arrow.png")
 ```

### 2025-10-17 CMSAC_presentation/CMSAC_reveal_js.qmd
updated.

@@
 
--   Comparing different structures like round robin and single eliminations, which is best?
+-   Which competition structure is best at reflecting the true strengths of competitors?
 
--   Efficacy: Final rankings should match the true rankings of the strength of the competitors
+-   **Efficacy**[^1]: Final rankings should match the true rankings of the strength of the competitors
 
--   Effectivity: Interested in correctly ranking the best competitor as first
+-   **Effectivity**[^2]: Interested in correctly ranking the best competitor as first
 
-CITEE
+[^1]: Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)
+
+[^2]: Glenn (1960)
 

### 2025-10-17 mutual_information_curves.qmd
Zach's slides

@@
 library(tidyverse)
+library(combinat)
+
+squared_error <- function(vec, weights = NULL) {
+  n <- length(vec)
+  ref <- 1:n
+
+  # Default weights = 1 for each element
+  if (is.null(weights)) {
+    weights <- rep(1, n)
+  }
+
+  # Initialize sum
+  summ <- 0
+
+  # Accumulate weighted squared error
+  for (i in 1:n) {
+    summ <- summ + (weights[i] * (vec[i] - ref[i]))^2
+  }
+
+  # Handle near-zero sum safely
+  if (abs(summ) < 1e-10) {
+    out <- 1
+  } else {
+    out <- 1 / summ
+  }
+
+  return(out)
+}
+
 partial_kendall <- function(vec, k, weights = NULL){
@@
 
-perms_list <- permn(1:5)
+make_weights <- function(k) {
+  w <- rep(0, 4)
+  w[1:k] <- 1
+  return(w)
+}
 
-# Apply weighted_kendall to each permutation
-results <- sapply(perms_list, function(p) {
-  weighted_kendall(rhat = p, k = 5, wtype = "symmetric_salience")
+# Store all results
+all_results <- list()
+
+# Loop over k = 1 to 4
+for (k in 1:4) {
+  weights <- make_weights(k)
+
+  # Weighted Kendall (using provided function)
+  kendall_results <- sapply(perms_list, function(p) {
+    weighted_kendall(rhat = p, k = 4, wtype = "symmetric_salience")
   })
 
-# Put into a dataframe
-kendall_df <- data.frame(
+  # Squared error using current weights
+  sq_results <- sapply(perms_list, function(p) {
+    squared_error(vec = p, weights = weights)
+  })
+
+  # Combine into one dataframe for this k
+  rhat_df <- data.frame(
+    k = k,
     permutation = sapply(perms_list, paste, collapse = ""),
-  weighted_kendall = results
+    weighted_kendall = kendall_results,
+    squared_error = sq_results
   )
 
-kendall_df
+  all_results[[k]] <- rhat_df
+}
 
-sum(kendall_df$weighted_kendall)
-mean(kendall_df$weighted_kendall)
+# Combine across all k
+rhat_df <- bind_rows(all_results)
 ```

### 2025-10-17 CMSAC_presentation/CMSAC_reveal_js.qmd
updated josie slides

@@
 ---
-title: "CMSAC_reveal_js"
+title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
 author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
@@
 
-::::: columns
-::: {.column width="50%"}
-```{r}
-knitr::include_graphics("4 bracket good.png")
-```
-:::
+::::::::: columns
 
-::: {.column width="50%"}
+::::::: {.column width="40%"}
 ::: {.fragment}
@@
 ```
-
 :::
@@
 ::: {.fragment}
-
 ```{r}
@@
 :::
+:::::::
 
+::::::: {.column width="60%"}
+::: {.fragment}
+```{r, out.width="80%"}
+knitr::include_graphics("4 bracket good.png")
+```
 :::
-:::::
+:::::::
+:::::::::
 
@@
 
-## Introduction
-
-<!-- Overall comments 
-I think the presentation is too long for 12 minutes (+ questions), but
-I feel like I didn't cover everything we want. What is plausable to cover
-in a 12-minute period? 
-Also, we need to figure out the notation for the final metric because I think
-the weight should have the subscript m, not I. And I probably was not consistent
-with the usage of theta or calling it true rank/true strength. -->
-
--   There exists numerous competition formats across all professional sports
--   Which competition structure is best?
--   The "best" competition structure depends on how well it reflects the competitors' true strengths
--   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors' true strengths
-
-<!-- Should we talk about Efficacy vs Effectivity here -->
-
-## Extreme Tournament
-
+::: notes
 -   4 team single elimination tournament:
@@
     -   Would not pair up Work Group and Dream Team in the first round
+:::
 
@@
 
--   Other competitions where the correct ranking of some, but not all, of the competitors is of import
+-   Decide how many rankings are important
     -   Olympic competition ranks the top 3 competitors
--   "Salient ranks": the ranks of a competition structure that matter
+    -   Premier League: top 4 (Champions League) and bottom 3 (Relegation)
+-   **Salient ranks**: the ranks that matter
     -   Efficacy: 1 through $n$ matter
     -   Effectivity: 1 is only salient rank
+
+## Goal
+
+-   Want to quantify how well a structure can accurately order the competitors
 -   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths
+
+## Distribution of $\hat{r}$
+
+```{r}
+knitr::include_graphics("r_dist.png")
+```
 

### 2025-10-17 CMSAC_presentation/Zach_slides.qmd

--- 2025-10-10 CMSAC_reveal_js.qmd
@@
 
-## Introduction
-
-```{=html}
-<!-- Overall comments 
-I think the presentation is too long for 12 minutes (+ questions), but
-I feel like I didn't cover everything we want. What is plausable to cover
-in a 12-minute period? 
-Also, we need to figure out the notation for the final metric because I think
-the weight should have the subscript m, not I. And I probably was not consistent
-with the usage of theta or calling it true rank/true strength. -->
-```
-
--   There exists numerous competition formats across all professional sports
--   Which competition structure is best?
--   The "best" competition structure depends on how well it reflects the competitors' true strengths
--   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors' true strengths
-
-<!-- Should we talk about Efficacy vs Effectivity here -->
-
-## Extreme Tournament
-
-- 4 team single elimination tournament:
-  - Dream Team vs Work Group vs Middle School Girls Team vs Babies
-- If we only care about the top team winning, seeding structure does not matter
-- If we care about who gets second and third, seeding becomes important
-  - Would not pair up Work Group and Dream Team in the first round
-
-## Comparing Structures
-
-- Comparing different structures like round robin and single eliminations, which is best?
-
-- Efficacy: Final rankings should match the true rankings of the strength of the competitors
-
-- Effectivity: Interested in correctly ranking the best competitor as first
-
-CITEE
-
-## Efficacy and Effectivity
-
-- Other competitions where the correct ranking of some, but not all, of the competitors is of import
-  - Olympic competition ranks the top 3 competitors
-
-- "Salient ranks": the ranks of a competition structure that matter
-  - Efficacy: 1 through $n$ matter
-  - Effectivity: 1 is only salient rank
-
-- We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths
-
-## Game Simulation
-
--   Given the assigned "known strength", $\theta$, we simulate a game using the Bradley-Terry Model
-
-$$\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$$
-
-## Tournament Simulations
-
--   Each competition structure is simulated given the number of teams, $n$, and $\theta$
--   Below is an example of the output from one simulation of a single elimination tournament structure
-
-```{r}
-renamed_df <- all_results_norm %>% 
-  select(-c(simulation, distribution, seed, teams)) %>%
-  rename(theta = true_strength, r = true_rank, r_hat = rank_hat) %>%
-  mutate(theta = round(theta,3))
-
-head(renamed_df,8)
-```
-
 ## Mutual Information
 
--   If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the “goodness” of a tournament in terms of its ability to accurately transmit the true ranking.
+::: notes
+If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the "goodness" of a tournament in terms of its ability to accurately transmit the true ranking.
+:::
 
 -   The notation for mutual information is: $$ I(X;Y) = \sum_{y \in Y}\sum_{x \in X} p(x,y) \log{\frac{p(x,y)}{p(x)p(y)}} $$
-
-<!-- -->
 
@@
 -   A limitation of mutual information is that it does not distinguish between the direction of information
--   In other words, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
--   Because of this, we proposed a weighting function $w(r, \hat{r})$
--   So, the new formula is: $$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$
+-   In other words, given equal probabilities, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
+-   Because of this, we proposed a new weighting function $w(r, \hat{r})$\
+    $$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$
 
@@
 
--   In our original paper submission, we used squared error loss as the weighting function: $$ w(r, \hat{r}) =
-    \begin{cases}
-    \frac{1}{(r-\hat{r})^2}, & r \neq \hat{r} \\
-    1, & r = \hat{r}
-    \end{cases} $$
-
+-   In our original paper submission, we used squared error loss as the weighting function: $w(r, \hat{r}) = \begin{cases} \frac{1}{(r-\hat{r})^2}, & r \neq \hat{r} \\ 1, & r = \hat{r} \end{cases}$
 -   However, after comments and reviews from people at CSAS and CMSAC, we have decided to use a weighted Kendall's Tau
@@
 
-$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \times \text{sgn}(i-j)\times \text{sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$
+$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$
 
--   $W$ is a $n\times n$ lower-triangular matrix, where
+-   If $\hat{r} = (2,1,3,4)$, then:
 
-<!-- Should this be capital W??? -->
+$$\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j) = 10$$
 
-$$W_{ij} = w_{\min(i,j)}, \text{ for } i,j = m,2m, \dots, n\times m$$ - $m$: vector of length $n$ representing the "salient weights", or the importance of correctly getting each ranking
+-   So, $\tau_w$ for this example would be $\frac{10}{15-3} = 0.833$
 
--   If $\tau_w = 1$, then the ranking is identical to the true ranking; if $\tau_w = 0$, then the order of teams is perfectly inversely proportional to the true rankings
+## Top K-teams Weighted Kendall
 
-## Weighted Kendall Example
+```{r}
+rhat_df %>% 
+  select(-squared_error) %>%   
+  filter(permutation %in% c("1234", "4321", "1324", "2143")) %>%   
+  arrange(permutation,k) %>%   
+  rename("Weighted Kendall" = weighted_kendall)
+```
 
--   If $\hat{r} = (2,1,3,4)$ and m = $(3,1,1,0)$, then: $$ W = \begin{pmatrix}
-    3 & 3 & 3 & 3 \\
-    3 & 1 & 1 & 1 \\
-    3 & 1 & 1 & 1 \\
-    3 & 1 & 1 & 0
-    \end{pmatrix} $$
+$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$
 
-$$\sum_{i,j} W_{ij} = 29; \sum_{i} W_{ii} = 5; \sum_{i\neq j} W_{ij} \times \text{sgn}(i-j)\times \text{sgn}(R_i - R_j) = 18 $$
+## Comparing Weighted Kendall's Tau to Squared Error
 
--   So, $\tau_w$ for this example would be $\frac{18}{29-5} = 0.75$
+```{r}
+rhat_df %>%    
+  filter (k==4) %>%   
+  select(-k) %>%   
+  arrange(desc(weighted_kendall)) %>%    
+  rename("Weighted Kendall"= weighted_kendall, "Squared Error" = squared_error)
+```
 
@@
 
--   Mutual Information can be normalized by dividing by the joint entropy
+-   To make comparisons across different values of $n$, we normalize by the joint entropy
 
--   Entropy: $-p(x)\log{p(x)}$
+-   So, the final formula for the metric is:
+
+$$\frac{wI_m(r,\hat{r})}{H(r,\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$
 
@@
 
--   $H(r)$ and $H(\hat{r})$ are theoretically equivalent, leaving $H(r,\hat{r}) = -2\sum_{r} p(r)\log(p(r)) = -2(n!) \times (1/n!) \times\log_2(1/n!) = \log_2((n!)^2)$ <!-- Is this true??
-I can't remember if they are equivalent or their maximums were equivalent -->
+## Tournament Simulations
 
--   So, the final formula for the metric is: $$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$
+-   Given the assigned "known strength", $\theta$, we simulate a game using the Bradley-Terry Model: $\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$
+-   Below is output from one simulation of a single elimination tournament structure
+
+```{r}
+renamed_df <- all_results_norm %>%    
+  select(-c(simulation, distribution, seed, teams)) %>%   
+  rename(theta = true_strength, r = true_rank, r_hat = rank_hat) %>%   
+  mutate(theta = round(theta,3))  
+head(renamed_df,8)
+```
 
@@
 
-<!-- Should these plots be on separate slides? -->
+```{r out.width="49%", fig.show='hold', fig.align='center'}
+knitr::include_graphics("tournament_comparison_plot.png")
+```
+
+## Close-Up Results
 
 ```{r out.width="49%", fig.show='hold', fig.align='center'}
-knitr::include_graphics("tournament_comparison_plot.png")
 knitr::include_graphics("tournament_comparison_plot_close.png")

### 2025-10-17 CMSAC_presentation/CMSAC_reveal_js.qmd
Updates

@@
 
-::::::::: columns
-
+:::::::::: columns
 ::::::: {.column width="40%"}
-::: {.fragment}
+::: fragment
 ```{r}
@@
 
-::: {.fragment}
+::: fragment
 ```{r}
@@
 
-::::::: {.column width="60%"}
-::: {.fragment}
+:::: {.column width="60%"}
+::: fragment
 ```{r, out.width="80%"}
@@
 :::
-:::::::
-:::::::::
+::::
+::::::::::
 
@@
 
-## Game Simulation
-
--   Given the assigned "known strength", $\theta$, we simulate a game using the Bradley-Terry Model
-
-$$\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$$
-
-## Tournament Simulations
-
--   Each competition structure is simulated given the number of teams, $n$, and $\theta$
--   Below is an example of the output from one simulation of a single elimination tournament structure
-
-```{r}
-renamed_df <- all_results_norm %>% 
-  select(-c(simulation, distribution, seed, teams)) %>%
-  rename(theta = true_strength, r = true_rank, r_hat = rank_hat) %>%
-  mutate(theta = round(theta,3))
-
-head(renamed_df,8)
-```
-
 ## Mutual Information
 
--   If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the “goodness” of a tournament in terms of its ability to accurately transmit the true ranking.
+::: notes
+If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the "goodness" of a tournament in terms of its ability to accurately transmit the true ranking.
+:::
 
 -   The notation for mutual information is: $$ I(X;Y) = \sum_{y \in Y}\sum_{x \in X} p(x,y) \log{\frac{p(x,y)}{p(x)p(y)}} $$
-
-<!-- -->
 
@@
 
-<!-- I don't know if the references.bib would work here or not -->
-
 -   Appleton, David R. 1995. “May the Best Man Win?” Statistician 44 (4): 529.

### 2025-10-17 CMSAC_presentation/CMSAC_reveal_js.qmd
reorder pics

@@
 ```{r}
-knitr::include_graphics("five_rec_league_men_playing_basketball.png")
+knitr::include_graphics("five_babies_playing_basketball.png")
 ```
@@
 ```{r}
-knitr::include_graphics("five_babies_playing_basketball.png")
-
+knitr::include_graphics("five_rec_league_men_playing_basketball.png")
 ```

### 2025-10-17 CMSAC_presentation/CMSAC_reveal_js.qmd
Updated dataframe format

@@
 author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
-format: revealjs
+format: 
+  revealjs: 
+    df-print: paged
 editor: visual
+
 ---
@@
 library(combinat)
+library(knitr)
+library(kableExtra)
 load("curves_data.RData")
@@
 ```{r}
-rhat_df %>% 
+# Get the filtered data
+table_data <- rhat_df %>% 
   select(-squared_error) %>%   
   filter(permutation %in% c("1234", "4321", "1324", "2143")) %>%   
-  arrange(permutation,k) %>%   
+  arrange(permutation, k) %>%   
   rename("Weighted Kendall" = weighted_kendall)
+
+# Split into two halves
+n <- nrow(table_data)
+left_table <- table_data %>% head(ceiling(n/2))
+right_table <- table_data %>% tail(floor(n/2))
+
+# Create side-by-side tables
+left_html <- left_table %>% 
+  kable(format = "html", row.names = FALSE) %>% 
+  kableExtra::kable_styling(font_size = 24)
+
+right_html <- right_table %>% 
+  kable(format = "html", row.names = FALSE) %>% 
+  kableExtra::kable_styling(font_size = 24)
 ```
+::: {.columns}
+::: {.column width="48%"}
+```{r}
+#| echo: false
+left_html
+```
+:::
+::: {.column width="%48%"}
+```{r}
+#| echo: false
+right_html
+```
+:::
+:::
 
@@
 ```{r}
-rhat_df %>%    
-  filter (k==4) %>%   
+# Get the filtered data
+table_data <- rhat_df %>%    
+  filter(k == 4) %>%   
   select(-k) %>%   
   arrange(desc(weighted_kendall)) %>%    
-  rename("Weighted Kendall"= weighted_kendall, "Squared Error" = squared_error)
+  rename("Weighted Kendall" = weighted_kendall, "Squared Error" = squared_error)
+
+# Split into two halves
+n <- nrow(table_data)
+left_table <- table_data %>% head(ceiling(n/2))
+right_table <- table_data %>% tail(floor(n/2))
+
+# Create side-by-side tables
+left_html <- left_table %>% 
+  kable(format = "html", row.names = FALSE) %>% 
+  kableExtra::kable_styling(font_size = 20)
+right_html <- right_table %>% 
+  kable(format = "html", row.names = FALSE) %>% 
+  kableExtra::kable_styling(font_size = 20)
+
 ```
+
+::: {.columns}
+::: {.column width="48%"}
+```{r}
+#| echo: false
+left_html
+```
+:::
+::: {.column width="%48%"}
+```{r}
+#| echo: false
+right_html
+```
+:::
+:::
 
@@
 -   Given the assigned "known strength", $\theta$, we simulate a game using the Bradley-Terry Model: $\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$
--   Below is output from one simulation of a single elimination tournament structure
+- Below is an example of the output of a single tournament simulation:
 
@@
   select(-c(simulation, distribution, seed, teams)) %>%   
-  rename(theta = true_strength, r = true_rank, r_hat = rank_hat) %>%   
-  mutate(theta = round(theta,3))  
-head(renamed_df,8)
+  rename(theta = true_strength, r = true_rank, r_hat = rank_hat, "Wins" = game_wins, "Losses" = game_losses) %>%   
+  mutate(theta = round(theta,3)) %>%
+  head(8)
+
+# Split into two halves
+n <- nrow(renamed_df)
+left_table <- renamed_df %>% head(ceiling(n/2))
+right_table <- renamed_df %>% tail(floor(n/2))
+
+# Create side-by-side tables
+left_html <- left_table %>% 
+  kable(format = "html", row.names = FALSE) %>% 
+  kableExtra::kable_styling(font_size = 24)
+right_html <- right_table %>% 
+  kable(format = "html", row.names = FALSE) %>% 
+  kableExtra::kable_styling(font_size = 24)
+
 ```
+
+::: {.columns}
+::: {.column width="48%"}
+```{r}
+#| echo: false
+left_html
+```
+:::
+::: {.column width="%48%"}
+```{r}
+#| echo: false
+right_html
+```
+:::
+:::
 
@@
 
-```{r out.width="49%", fig.show='hold', fig.align='center'}
+::: {style="position: relative;"}
+::: {.fade-out fragment-index="1"}
+```{r out.width="60%", fig.show='hold', fig.align='center'}
+knitr::include_graphics("boxed_tournament_comparison_plot.png")
+```
+:::
+
+::: {.fragment .fade-in fragment-index="1" .fade-out fragment-index="2" style="position: absolute; top: 50%; left: 0; right: 0; transform: translateY(-50%);"}
+```{r out.width="60%", fig.show='hold', fig.align='center'}
 knitr::include_graphics("tournament_comparison_plot.png")
 ```
+:::
 
-## Close-Up Results
-
-```{r out.width="49%", fig.show='hold', fig.align='center'}
+::: {.fragment .fade-in fragment-index="2" style="position: absolute; top: 50%; left: 0; right: 0; transform: translateY(-50%);"}
+```{r out.width="60%", fig.show='hold', fig.align='center'}
 knitr::include_graphics("tournament_comparison_plot_close.png")
 ```
+:::
+:::
 
-## Limitations
+## Conclusion
 
--   The "known" strengths are assumptions that can be estimated but never truly known
--   Ties are not addressed in any of the games
--   Ties in the $\hat{r}$ are randomly assigned
-
-## Future Work
-
--   Evaluate structures using various distributions and "known" weights
--   Estimate the $\theta$s in a real-world example and compare recent changes to the old format
+-   These results underscore the trade-off between tournament accuracy and practical constraints such as time, cost, and entertainment value.
+-   Limitations: The "known" strengths are assumptions that can be estimated but never truly known
+-   Future Work: Estimate the $\theta$s in a real-world example and compare recent changes to the old format
 

### 2025-10-17 mutual_information_curves.qmd
more changes

@@
         summ <- summ + w[i, j] * sign(i - j) * sign(rhat[i] - rhat[j])
-        print(summ)
+        #print(summ)
       }
@@
   weight <- (out + 1) / 2
-  print(sum(w))
-  print(sum(diag(w)))
+  #print(sum(w))
+  #print(sum(diag(w)))
   return(weight)
@@
 
+perms_list <- permn(1:4)
+
 # Loop over k = 1 to 4
@@
   kendall_results <- sapply(perms_list, function(p) {
-    weighted_kendall(rhat = p, k = 4, wtype = "symmetric_salience")
+    weighted_kendall(rhat = p, k = k, wtype = "symmetric_salience")
   })

### 2025-10-20 CMSAC_presentation/CMSAC_reveal_js.qmd
presentation updates

@@
 editor: visual
-
 ---
@@
 
+## Distribution of $\hat{r}$
+
+```{r}
+knitr::include_graphics("rhat_dist.png")
+```
+
 ## Comparing Structures
+::: {.incremental}
+-   Which competition structure is best?
 
--   Which competition structure is best at reflecting the true strengths of competitors?
+    -   What is considered "best"?
 
@@
 -   **Effectivity**[^2]: Interested in correctly ranking the best competitor as first
+:::
+
+::: {.fragment}
+```{r}
+knitr::include_graphics("salient bar.png")
+```
 
@@
 
+:::
+
 ## Efficacy and Effectivity
 
--   Decide how many rankings are important
-    -   Olympic competition ranks the top 3 competitors
-    -   Premier League: top 4 (Champions League) and bottom 3 (Relegation)
--   **Salient ranks**: the ranks that matter
+```{r}
+knitr::include_graphics("salient examples.png")
+```
+
+-   **Salient ranks**: output ranks of the tournament you want to get correct
     -   Efficacy: 1 through $n$ matter
@@
 -   Want to quantify how well a structure can accurately order the competitors
--   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths
-
-## Distribution of $\hat{r}$
-
-```{r}
-knitr::include_graphics("r_dist.png")
-```
+-   We propose a metric that quantifies how effectively different tournament structures recover the true ranking
 
@@
 ```
-::: {.columns}
+
+::::: columns
 ::: {.column width="48%"}
@@
 :::
+
 ::: {.column width="%48%"}
@@
 :::
-:::
+:::::
 
@@
 
-::: {.columns}
+::::: columns
 ::: {.column width="48%"}
@@
 
-::: {style="position: relative;"}
+:::::: {style="position: relative;"}
 ::: {.fade-out fragment-index="1"}
@@
 
-::: {.fragment .fade-in fragment-index="1" .fade-out fragment-index="2" style="position: absolute; top: 50%; left: 0; right: 0; transform: translateY(-50%);"}
+::: {.fragment .fade-in .fade-out fragment-index="1" fragment-index="2" style="position: absolute; top: 50%; left: 0; right: 0; transform: translateY(-50%);"}
 ```{r out.width="60%", fig.show='hold', fig.align='center'}
@@
 :::
-:::
+::::::
 

### 2025-10-21 CMSAC_presentation/CMSAC_reveal_js.qmd
updated method to mathjax.
josie slides updates

@@
 ## Comparing Structures
+
+-   Which competition structure is best?
+
     ::: {.incremental}
+
+    -   What is considered "best"?
+
+    - A good tournament: top *k* produced by the tournament matches closely to the true top *k*
+
+    :::
+
+::: {.incremental}
+
+-   **Efficacy**: Final rankings should match the true rankings of the strength of the competitors
+
+-   **Effectivity**: Interested in correctly ranking the best competitor as first
+:::
+
+## Comparing Structures
+
 -   Which competition structure is best?
@@
 
+    - A good tournament: top *k* produced by the tournament matches closely to the true top *k*
+
 -   **Efficacy**[^1]: Final rankings should match the true rankings of the strength of the competitors
@@
 -   **Effectivity**[^2]: Interested in correctly ranking the best competitor as first
-:::
-
-::: {.fragment}
-```{r}
-knitr::include_graphics("salient bar.png")
-```
 
@@
 
-:::
-
-## Efficacy and Effectivity
+## Effectivity and Efficacy
 
 ```{r}
+knitr::include_graphics("salient bar.png")
+```
+
+## Effectivity and Efficacy
+
+```{r, out.width="100%"}
 knitr::include_graphics("salient examples.png")
@@
 
+::: {.fragment}
 -   **Salient ranks**: output ranks of the tournament you want to get correct
@@
     -   Effectivity: 1 is only salient rank
+:::
 

### 2025-10-21 CMSAC_presentation/CMSAC_reveal_js.qmd
Trimmed Slides

@@
   revealjs: 
+    html-math-method: 
+      method: mathjax
+      url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
     df-print: paged
 editor: visual
+
 ---
@@
 -   In other words, given equal probabilities, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
--   Because of this, we proposed a new weighting function $w(r, \hat{r})$\
+-   Because of this, we proposed a new weighting function\
     $$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$
+-   The final metric we used normalizes the results: $\frac{wI_m(r,\hat{r})}{H(r,\hat{r})}$
 
@@
 -   However, after comments and reviews from people at CSAS and CMSAC, we have decided to use a weighted Kendall's Tau
-
-## Weighted Kendall
-
-$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$
-
--   If $\hat{r} = (2,1,3,4)$, then:
-
-$$\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j) = 10$$
-
--   So, $\tau_w$ for this example would be $\frac{10}{15-3} = 0.833$
 
@@
 $$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$
-
-## Comparing Weighted Kendall's Tau to Squared Error
-
-```{r}
-# Get the filtered data
-table_data <- rhat_df %>%    
-  filter(k == 4) %>%   
-  select(-k) %>%   
-  arrange(desc(weighted_kendall)) %>%    
-  rename("Weighted Kendall" = weighted_kendall, "Squared Error" = squared_error)
-
-# Split into two halves
-n <- nrow(table_data)
-left_table <- table_data %>% head(ceiling(n/2))
-right_table <- table_data %>% tail(floor(n/2))
-
-# Create side-by-side tables
-left_html <- left_table %>% 
-  kable(format = "html", row.names = FALSE) %>% 
-  kableExtra::kable_styling(font_size = 20)
-right_html <- right_table %>% 
-  kable(format = "html", row.names = FALSE) %>% 
-  kableExtra::kable_styling(font_size = 20)
-
-```
-
-::::: columns
-::: {.column width="48%"}
-```{r}
-#| echo: false
-left_html
-```
-:::
-
-::: {.column width="%48%"}
-```{r}
-#| echo: false
-right_html
-```
-:::
-:::::
-
-## Normalizing Mutual Information
-
--   To make comparisons across different values of $n$, we normalize by the joint entropy
-
--   So, the final formula for the metric is:
-
-$$\frac{wI_m(r,\hat{r})}{H(r,\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$
-
--   Assuming $r$ and $\hat{r}$ are independent, joint entropy can be written as $H(r, \hat{r}) = H(r) + H(\hat{r})$
 

### 2025-10-21 CMSAC_presentation/CMSAC_reveal_js.qmd
slides

@@
 -   **Salient ranks**: output ranks of the tournament you want to get correct
-    -   Efficacy: 1 through $n$ matter
     -   Effectivity: 1 is only salient rank
+    -   Efficacy: 1 through $n$ are salient ranks
 :::

### 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd
slides

@@
 
-::: {.incremental}
-
--   **Efficacy**: Final rankings should match the true rankings of the strength of the competitors
-
--   **Effectivity**: Interested in correctly ranking the best competitor as first
-:::
-
 ## Comparing Structures
@@
 
--   **Efficacy**[^1]: Final rankings should match the true rankings of the strength of the competitors
+-   **Efficacy**[^1]: Final rankings should match the true rankings of all the competitors
 

### 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd
slides
sldies

@@
 
-    - A good tournament: top *k* produced by the tournament matches closely to the true top *k*
+    - A good tournament: top *k* teams produced by the tournament matches closely to the true top *k*
 
@@
 
-    - A good tournament: top *k* produced by the tournament matches closely to the true top *k*
+    - A good tournament: top *k* teams produced by the tournament matches closely to the true top *k*
 
--   **Efficacy**[^1]: Final rankings should match the true rankings of all the competitors
+-   **Effectivity**[^1]: Interested in correctly ranking the best competitor as first
 
--   **Effectivity**[^2]: Interested in correctly ranking the best competitor as first
+-   **Efficacy**[^2]: Final rankings should match the true rankings of all the competitors
 
-[^1]: Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)
+[^1]: Glenn (1960)
 
-[^2]: Glenn (1960)
+[^2]: Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)
 
@@
 -   **Salient ranks**: output ranks of the tournament you want to get correct
-    -   Effectivity: 1 is only salient rank
-    -   Efficacy: 1 through $n$ are salient ranks
+    -   Effectivity: 1 is only salient rank (*k*=1)
+    -   Efficacy: 1 through $n$ are salient ranks (*k*=n)
 :::
@@
 
--   Want to quantify how well a structure can accurately order the competitors
--   We propose a metric that quantifies how effectively different tournament structures recover the true ranking
+-   Quantify how well a structure can accurately order the competitors
+-   We propose a metric that quantifies how effectively different tournament structures recover the true ranking of the teams
 

### 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd
sldies

@@
 
--   Quantify how well a structure can accurately order the competitors
+-   Quantify how well a structure can accurately order the competitors using weighted mutual information
 -   We propose a metric that quantifies how effectively different tournament structures recover the true ranking of the teams

### 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd
Updates

--- 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd
@@
 
+::: {.incremental}
+
+-   **Efficacy**: Final rankings should match the true rankings of the strength of the competitors
+
+-   **Effectivity**: Interested in correctly ranking the best competitor as first
+:::
+
 ## Comparing Structures
@@
 
--   **Efficacy**[^1]: Final rankings should match the true rankings of all the competitors
+-   **Efficacy**[^1]: Final rankings should match the true rankings of the strength of the competitors
 
@@
 ::: {.fragment}
--   **Salient ranks**: output ranks of the tournament you want to get correct
+-   **Salient ranks**: output ranks of the competition you want to get correct
     -   Effectivity: 1 is only salient rank
@@
 -   Want to quantify how well a structure can accurately order the competitors
--   We propose a metric that quantifies how effectively different tournament structures recover the true ranking
+-   We propose a metric that quantifies how effectively different competition structures recover the true ranking
 
@@
 ::: notes
-If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the "goodness" of a tournament in terms of its ability to accurately transmit the true ranking.
+If one views the true ranking of the teams in a competition as a message to be sent to a receiver and the outcome of the competition as a message that is received, we can measure the "goodness" of a competition in terms of its ability to accurately transmit the true ranking.
 :::
@@
 
-    $$ I(r, \hat{r}) = n! \sum_{\hat{r}} p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{(\frac{1}{n!})^2}} = n! \log{((n!)^2)} $$
+    $$ I(r, \hat{r}) = n! \sum_{\hat{r}} p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{(\frac{1}{n!})^2}} $$
 
@@
 -   A limitation of mutual information is that it does not distinguish between the direction of information
--   In other words, given equal probabilities, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
+-   Given equal probabilities, the information found from a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
 -   Because of this, we proposed a new weighting function\
-    $$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$
--   The final metric we used normalizes the results: $\frac{wI_m(r,\hat{r})}{H(r,\hat{r})}$
-
-## Weighting Function
-
--   In our original paper submission, we used squared error loss as the weighting function: $w(r, \hat{r}) = \begin{cases} \frac{1}{(r-\hat{r})^2}, & r \neq \hat{r} \\ 1, & r = \hat{r} \end{cases}$
--   However, after comments and reviews from people at CSAS and CMSAC, we have decided to use a weighted Kendall's Tau
+    $$ I_w(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$
+-   The final metric we used normalizes the results: $\frac{I_w(r,\hat{r})}{H(r,\hat{r})}$
 
@@
 # Get the filtered data
-table_data <- rhat_df %>% 
-  select(-squared_error) %>%   
-  filter(permutation %in% c("1234", "4321", "1324", "2143")) %>%   
-  arrange(permutation, k) %>%   
-  rename("Weighted Kendall" = weighted_kendall)
+#table_data <- rhat_df %>% 
+#  select(-squared_error) %>%   
+#  filter(permutation %in% c("1234", "4321", "1324", "2143")) %>%   
+#  arrange(permutation, k) %>%   
+#  rename("Weighted Kendall" = weighted_kendall)
 
 # Split into two halves
-n <- nrow(table_data)
-left_table <- table_data %>% head(ceiling(n/2))
-right_table <- table_data %>% tail(floor(n/2))
+#n <- nrow(table_data)
+#left_table <- table_data %>% head(ceiling(n/2))
+#right_table <- table_data %>% tail(floor(n/2))
 
 # Create side-by-side tables
-left_html <- left_table %>% 
-  kable(format = "html", row.names = FALSE) %>% 
-  kableExtra::kable_styling(font_size = 24)
+#left_html <- left_table %>% 
+#  kable(format = "html", row.names = FALSE) %>% 
+#  kableExtra::kable_styling(font_size = 24)
 
-right_html <- right_table %>% 
-  kable(format = "html", row.names = FALSE) %>% 
-  kableExtra::kable_styling(font_size = 24)
+#right_html <- right_table %>% 
+#  kable(format = "html", row.names = FALSE) %>% 
+#  kableExtra::kable_styling(font_size = 24)
+knitr::include_graphics("Top_k_Kendall.png")
+
 ```
-
-::::: columns
-::: {.column width="48%"}
-```{r}
-#| echo: false
-left_html
-```
-:::
-
-::: {.column width="%48%"}
-```{r}
-#| echo: false
-right_html
-```
-:::
-:::::
-
-$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$
 
@@
 
--   Given the assigned "known strength", $\theta$, we simulate a game using the Bradley-Terry Model: $\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$
--   Below is an example of the output of a single tournament simulation:
+-   Given the assigned "known strength", $\theta$, we simulate a game using: $\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$
 
 ```{r}
-renamed_df <- all_results_norm %>%    
-  select(-c(simulation, distribution, seed, teams)) %>%   
+renamed_df <- single_run_example %>%    
+  select(-c(distribution, seed)) %>%   
   rename(theta = true_strength, r = true_rank, r_hat = rank_hat, "Wins" = game_wins, "Losses" = game_losses) %>%   
-  mutate(theta = round(theta,3)) %>%
-  head(8)
-
-# Split into two halves
-n <- nrow(renamed_df)
-left_table <- renamed_df %>% head(ceiling(n/2))
-right_table <- renamed_df %>% tail(floor(n/2))
+  mutate(theta = round(theta,3))
 
 # Create side-by-side tables
-left_html <- left_table %>% 
-  kable(format = "html", row.names = FALSE) %>% 
-  kableExtra::kable_styling(font_size = 24)
-right_html <- right_table %>% 
+html_df <- renamed_df %>% 
   kable(format = "html", row.names = FALSE) %>% 
@@
 #| echo: false
-left_html
+html_df
 ```
@@
 #| echo: false
-right_html
+#| out.width: "100%"
+knitr::include_graphics("simulation_example.png")
 ```
@@
 
-## Conclusion
+## Discussion
 
--   These results underscore the trade-off between tournament accuracy and practical constraints such as time, cost, and entertainment value.
--   Limitations: The "known" strengths are assumptions that can be estimated but never truly known
--   Future Work: Estimate the $\theta$s in a real-world example and compare recent changes to the old format
+-   We used mutual information to analyze competition structures in a range from efficacy to effectivity.
+-   Using the method, we found that Round Robin is the best structure, followed by Reseeded Single Elimination, Double Elimination, and Group Stage.
+-   Limitations: The "known" ranks are assigned using $\theta$ and, for some structure, the order of the ranks matter in the formatting of the structure
+-   Future Work: Examine how the distribution of $\theta$s influence the score of each competition structure
 

### 2025-10-22 mutual_information_curves.qmd
Updates

@@
 df_long <- pivot_longer(mi_plots_df, cols = -indices, names_to = "Method", values_to = "MI")
+df_long <- df_long %>% mutate(Method = recode( Method, "Single.Elimination..Reseeded." = "Single Elimination (Reseeded)", "Group.Stage" = "Group Stage", "Double.Elimination" = "Double Elimination", "Single.Elimination..Set.Structure." = "Single Elimination (Set Structure)", "Coin.Flip.Single.Elimination" = "Coin Flip Single Elimination", "Bad.Seeding.Structure.Single.Elimination" = "Bad Seeding Structure Single Elimination", "Coin.Flip.Round.Robin" = "Coin Flip Round Robin", "Round.Robin" = "Round Robin", "Worst.Possible" = "Worst Possible", "Best.Possible" = "Best Possible" ))
 
@@
 )
-# Define consistent color mapping
-methods <- unique(df_long$Method)
-method_colors <- setNames(okabe_ito, methods)
+# Compute the ordering of methods by mean MI (descending) 
+method_order <- df_long %>% 
+  group_by(Method) %>% 
+  summarize(mean_MI = mean(MI, na.rm = TRUE)) %>% 
+  arrange(desc(mean_MI)) %>% 
+  pull(Method)
+
+# Reorder Method factor levels by MI 
+df_long <- df_long %>% 
+  mutate(Method = factor(Method, levels = method_order))
+
+# Define color mapping in the new order 
+methods <- levels(df_long$Method) 
+method_colors <- setNames(okabe_ito[seq_along(methods)], methods)
 
 windowsFonts(Arial = windowsFont("Arial"))
-ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
+comparison_plot <- ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
 geom_line(linewidth = 2.5) +           # Add lines
@@
 x = "Top i Number of Teams",
-y = "Normalized Mutual Information \\ Entropy (log10 scale)"
+y = "Normalized Mutual Information \\ Entropy"
 ) +
@@
   guides(color = guide_legend(nrow = 5)) +
-  scale_y_log10(
-    labels = scales::label_number() 
-  ) +
   scale_color_manual(values = method_colors)
@@
   filename = "tournament_comparison_plot.png",  # Change file type if needed
-  plot = last_plot(),  # Saves the last ggplot created
+  plot = comparison_plot,  # Saves the last ggplot created
   dpi = 600,        

### 2025-10-22 single_elimination_shiny.qmd
Updates

@@
       true_rank = as.numeric(teams),
-      theta = c(strengths, rep(NA, length(teams) - num_teams)),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
       rank_hat = rep(NA,length(teams)),
@@
       true_rank = as.numeric(teams),
-      theta = c(strengths, rep(NA, length(teams) - num_teams)),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
       game_wins = rep(0,length(teams)),
@@
       true_rank = as.numeric(teams),
-      theta = c(strengths, rep(NA, length(teams) - num_teams)),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
       rank_hat = rep(NA, length(teams)),
@@
     teams <- permuted_df$true_rank
-    permuted_df$theta <- df$theta[permutation]
+    permuted_df$true_strength <- df$true_strength[permutation]
     permuted_df$rank_hat <- NA  # Initialize ranks column
@@
 
-    strengths <- permuted_df$theta
+    strengths <- permuted_df$true_strength
     round_number <- ceiling(log2(length(teams)))
@@
       team2 <- losers_semis[2]
-      strength1 <- permuted_df$theta[permuted_df$true_rank == team1]
-      strength2 <- permuted_df$theta[permuted_df$true_rank == team2]
+      strength1 <- permuted_df$true_strength[permuted_df$true_rank == team1]
+      strength2 <- permuted_df$true_strength[permuted_df$true_rank == team2]
       wins_team1 <- 0
       wins_team2 <- 0
+      losses_team1 <- 0
+      losses_team2 <- 0
       for (game in 1:series) {
@@
           wins_team1 <- wins_team1 + 1
+          losses_team2 <- losses_team2 + 1
+
         } else {
           wins_team2 <- wins_team2 + 1
+          losses_team1 <- losses_team1 + 1
         }
@@
       }
+
+      # Update game_wins / game_losses in permuted_df for both teams
+        permuted_df <- permuted_df %>%
+          mutate(game_wins = ifelse(true_rank == team1, game_wins + wins_team1, game_wins)) %>%
+          mutate(game_wins = ifelse(true_rank == team2, game_wins + wins_team2, game_wins)) %>%
+          mutate(game_losses = ifelse(true_rank == team1, game_losses + losses_team1, game_losses)) %>%
+          mutate(game_losses = ifelse(true_rank == team2, game_losses + losses_team2, game_losses))
+
       # Assign third place and fourth place ranks
@@
 
-test <- simulate_single_elimination_tournament(num_teams = 8, distribution = "Same", ties=F, series=1, third_place=T, seeding_structure = c(1,2,3,4,5,6,7,8), reseeded = TRUE)
-test
+single_run_example <- simulate_single_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, third_place=T, reseeded=F)
 
@@
       series = 1,
-      third_place = FALSE
+      third_place = TRUE
     )

### 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd

@@
 
-    - A good tournament: top *k* produced by the tournament matches closely to the true top *k*
+    - A good tournament: top *k* teams produced by the tournament matches closely to the true top *k*
 
-    :::
-
-::: {.incremental}
-
--   **Efficacy**: Final rankings should match the true rankings of the strength of the competitors
-
--   **Effectivity**: Interested in correctly ranking the best competitor as first
     :::
@@
 
-    - A good tournament: top *k* produced by the tournament matches closely to the true top *k*
+    - A good tournament: top *k* teams produced by the tournament matches closely to the true top *k*
 
--   **Efficacy**[^1]: Final rankings should match the true rankings of the strength of the competitors
+-   **Effectivity**[^1]: Interested in correctly ranking the best competitor as first
 
--   **Effectivity**[^2]: Interested in correctly ranking the best competitor as first
+-   **Efficacy**[^2]: Final rankings should match the true rankings of all the competitors
 
-[^1]: Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)
+[^1]: Glenn (1960)
 
-[^2]: Glenn (1960)
+[^2]: Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)
 
@@
     -   Efficacy: 1 through $n$ are salient ranks
+=======
+-   **Salient ranks**: output ranks of the tournament you want to get correct
+    -   Effectivity: 1 is only salient rank (*k*=1)
+    -   Efficacy: 1 through $n$ are salient ranks (*k*=n)
 :::
@@
 -   We propose a metric that quantifies how effectively different competition structures recover the true ranking
+=======
+-   Quantify how well a structure can accurately order the competitors using weighted mutual information
+-   We propose a metric that quantifies how effectively different tournament structures recover the true ranking of the teams
 

### 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd
Fixed

@@
 -   We propose a metric that quantifies how effectively different competition structures recover the true ranking
-=======
+
 -   Quantify how well a structure can accurately order the competitors using weighted mutual information

### 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd
slides

@@
 ::: {.fragment}
--   **Salient ranks**: output ranks of the competition you want to get correct
-    -   Effectivity: 1 is only salient rank
-    -   Efficacy: 1 through $n$ are salient ranks
-=======
 -   **Salient ranks**: output ranks of the tournament you want to get correct
@@
 ## Goal
-
--   Want to quantify how well a structure can accurately order the competitors
--   We propose a metric that quantifies how effectively different competition structures recover the true ranking
 

### 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd

@@
   revealjs: 
+    embed-resources: true
+    theme: [default, styles.scss]
+    # theme: [serif, styles.scss]
     html-math-method: 
@@
 
-    ::: {.incremental}
-
+    ::: incremental
     -   What is considered "best"?
@@
     -   A good tournament: top *k* teams produced by the tournament matches closely to the true top *k*
-
     :::
@@
 
-::: {.fragment}
+::: fragment
 -   **Salient ranks**: output ranks of the tournament you want to get correct
@@
 
+<!-- ## References {.smaller} -->
+
 -   Appleton, David R. 1995. “May the Best Man Win?” Statistician 44 (4): 529.

### 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd
Image Size fix

@@
 ::::: columns
-::: {.column width="48%"}
+::: {.column width="37%"}
 ```{r}
@@
 
-::: {.column width="%48%"}
+::: {.column width="63%"}
 ```{r}
 #| echo: false
-#| out.width: "100%"
+#| out-width: "125%"
+#| out-length: "150%"
+#| fig-width: 8
+#| fig-height: 16
 knitr::include_graphics("simulation_example.png")

### 2025-10-22 CMSAC_presentation/CMSAC_reveal_js.qmd
Updates

@@
 -   Because of this, we proposed a new weighting function\
-    $$ I_w(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$
+```{r out.width="150%", fig.show='hold', fig.align='center'}
+knitr::include_graphics("Weighted_Mutual_Information_Formula.png")
+```
 -   The final metric we used normalizes the results: $\frac{I_w(r,\hat{r})}{H(r,\hat{r})}$
@@
 #  kableExtra::kable_styling(font_size = 24)
-knitr::include_graphics("Top_k_Kendall.png")
+#knitr::include_graphics("Top_k_Kendall.png")
 
 ```
+
+$$\tau_w = \frac{\sum_{i\neq j} W_{ij} \times \text{sgn}(i-j)\times \text{sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$
+
+:::::: {style="position: relative; display: flex; height: 100%;"}
+::: {.fragment .fade-out fragment-index="1" style="position: absolute; width: 100%;"}
+```{r out.width="100%", fig.show='hold', fig.align='center'}
+knitr::include_graphics("Top_2_Kendall.png")
+```
+:::
+::: {.fragment .fade-in fragment-index="1" style="position: absolute; width: 100%;"}
+```{r out.width="100%", fig.show='hold', fig.align='center'}
+knitr::include_graphics("Top_3_Kendall.png")
+```
+:::
+::::::
 
@@
 
-:::::: {style="position: relative;"}
-::: {.fade-out fragment-index="1"}
-```{r out.width="60%", fig.show='hold', fig.align='center'}
-knitr::include_graphics("boxed_tournament_comparison_plot.png")
-```
-:::
-
-::: {.fragment .fade-in .fade-out fragment-index="1" fragment-index="2" style="position: absolute; top: 50%; left: 0; right: 0; transform: translateY(-50%);"}
-```{r out.width="60%", fig.show='hold', fig.align='center'}
+:::::: {style="position: relative; display: flex; align-items: center; justify-content: center; height: 100%;"}
+::: {.fragment .fade-out fragment-index="1" style="position: absolute; width: 100%;"}
+```{r out.width="125%", fig.show='hold', fig.align='center'}
 knitr::include_graphics("tournament_comparison_plot.png")
@@
 :::
-
-::: {.fragment .fade-in fragment-index="2" style="position: absolute; top: 50%; left: 0; right: 0; transform: translateY(-50%);"}
-```{r out.width="60%", fig.show='hold', fig.align='center'}
+::: {.fragment .fade-in fragment-index="1" style="position: absolute; width: 100%;"}
+```{r out.width="150%", fig.show='hold', fig.align='center'}
 knitr::include_graphics("tournament_comparison_plot_close.png")

### 2025-10-22 mutual_information_curves.qmd
Updates

@@
 windowsFonts(Arial = windowsFont("Arial"))
-comparison_plot <- ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
+ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
 geom_line(linewidth = 2.5) +           # Add lines
@@
   filename = "tournament_comparison_plot.png",  # Change file type if needed
-  plot = comparison_plot,  # Saves the last ggplot created
+  plot = last_plot(),  # Saves the last ggplot created
   dpi = 600,        

### 2025-10-31 manuscript/paperCMSACversion.Rmd
cmsac version of paper.

--- 2025-10-06 manuscript/paper.Rmd
@@
 
-In order to alleviate this problem, we instead consider *weighted* mutual information (@Guiasu1977-nu).
-We want to give more weight to permutations from $\hat{r}$ that are "closer" to $r$.
-Essentially, this problem comes down to choosing a metric that measures the distance between the vectors $\hat{r}$ and $r$.
-One then gives more weight to combinations of $\hat{r}$ and $r$ that are closer and less weight to combinations that are far apart.
-A natural choice for distance between vectors the sum of squared deviations between the two vectors.
-Specifically, we define the loss function between $\hat{r}$ and $r$ to be $l(r, \hat{r}) = \sum_{i=1}^n (\hat{r}_i - r_i)^2$.
-Then, since small losses should be associated with large weights, we focus on the recipricol of this loss function.
+In order to alleviate this problem, we instead consider *weighted*
+mutual information (@Guiasu1977-nu). We want to give more weight to
+permutations from $\hat{r}$ that are "closer" to $r$.
 
-In general, any loss function $l(r, \hat{r})$ can be used to define the weighting function as follows:
+In general, any loss function $l(r, \hat{r})$ can be used to define the
+weigting function:
 
@@
 
-Note that in order to avoid dividing by 0, we define the weight to be 1 when $r$ and  $\hat{r}$ are exactly equal.
-
-Specifically, when squared error loss is chosen, one gets the following weight function:
+Here though, we choose to use squared error loss:
 
@@
 \begin{cases}
-\frac{1}{\sum_{i=1}^n (\hat{r}_i - r_i)^2}, & r \ne \hat{r} \\
+\frac{1}{\hat{r}'r}, & r \ne \hat{r} \\
 1, & r = \hat{r}
@@
 
-An alternative to defining weight this way is to consider looking at the correlation between $\hat{r}$ and $r$.
-Since, $\hat{r}$ and $r$ are already ranks, the Pearson and Spearman correlation coefficients will be computationally exactly the same in this setting if that was the measure that someone chose to use.
-
-Spearman correlation can be written as follows:
-
-$$
-\rho_{s} =  1 - \frac{6  \sum_{i=1}^n (\hat{r}_i - r_i)^2}{n(n^2 - 1)}
-$$,
-
-which means that Spearman correlation is simply (one minus) a scaling of squared error loss.
-Since $\rho_{s} \in [-1,1]$, but we desire weights $w(\hat{r},r) \in [0,1]$, we define $w(\hat{r},r) = \frac{(\rho_s + 1)}{2}$ guaranteeing weights in the desired range.
-
-While squared error loss and Spearman/Pearson correlation work well, in this setting we argue that Kendall's $\tau$ correlation is the most appropriate for weighting due to its interpretation and how this related to competitions.
-One way of interpreting Kendall's $\tau$ is that it's a scaling of the number of adjacent swaps required to go from one permutation of ranks to another.
-Recall that Kendall's $\tau$ is defined as: 
-$$
-\rho_\tau = 1 - \frac{2 \times \text{# discordant pairs}}{n \choose 2 }
-$$.
-To convert $\rho_\tau$ back to the range between $[0,1]$, which we want for weighting purposes, we subtract 1 and divide by 2 giving us $\frac{\rho_\tau + 1}{2}$.  This then becomes:
-
-$$
-w(\hat{r},r) = \frac{\rho_\tau + 1}{2} = 1 - \frac{\text{# discordant pairs}}{n \choose 2}
-$$ 
-and the number of discordant pairs is equivalent to the number of adjacent swaps needed to get the order of $\hat{r}$ correct.
-
-As an example, consider $r = (1,2,3,4)$ and $\hat{r} = (1,3,4,2)$.
-This requires two adjacent swaps to get $\hat{r}$ to equal $r$ (i.e swap 4 with 2 and then 3 with 2).
-The Kendall's $\tau$ value between these two vectors is `r cor(c(1,2,3,4),c(1,3,4,2), method = "kendall")`.
-We would then define the weight between the two vectors to be $(\rho_{\tau} + 1)/2 =$ `r (cor(c(1,2,3,4),c(1,3,4,2), method = "kendall") + 1)/2`.
-This is exactly equivalent to $1 - \frac{\text{# adj swaps}}{n \choose 2}$.
-
-<!-- ```{r} -->
-<!-- cor(c(1,2,3,4),c(1,3,4,2), method = "kendall") -->
-<!-- 1 - (2*2)/choose(4, 2) -->
-
-<!-- cor(c(1,2,3,4),c(1,2,4,3), method = "kendall") -->
-<!-- 1 - (2*1)/choose(4, 2) -->
-
-<!-- #Converting this back to 0,1 gives (by adding 1 and dividing by 2):  -->
-<!-- #(Number of swaps)/choose(4, 2).   -->
-<!-- #This is what we use for weights -->
-<!-- ``` -->
-
-@shieh1998weighted
-Now, regardless of which method one chooses for defining weights between $\hat{r}$ and $r$, a second set of weights needs to be defined because the number of teams that need to be accurately ordered differs based on the choice of salient ranks.
-Therefore, we define a second set of weights, $m$, which is used to indicate which ranks are salient.
-We call these the "salient weights" to differentiate from the other weighting function, $w$.
-As an example, for a four-team tournament, if the organizer decided that only the correctly choosing the best team as the winner was important, then $m = \{1,0,0,0\}$ (i.e.
-essentially measuring effectivity).
-One can imagine that in an event like the Olympics, the salient weights would be 1's for the first three elements and 0's elsewhere.
-
-In addition, you could go beyond 0's and 1's.  
-While here we only consider indicators (i.e.
-0's or 1's) in the salient weights vector, there is no reason why you couldn't weight the ranks by relative levels of importance.
-For instance, letting $m = \{3,1,0,0\}$ indicates that the salient ranks are 1 and 2, but getting the true best competitor as the winner is more important than ranking the true second best competitor
+However, because the number of teams that need to be accurately ordered
+differs based on the choice of salient ranks, a second set of weights is
+defined, $m$, is used to indicate which ranks are salient. We call these
+the "salient weights" to differentiate from the other weighting
+function. As an example, for a four-team tournament, if the organizer
+decides that only the winner is important, then $m = \{1,0,0,0\}$ (i.e.
+essentially measuring effectivity). While here we only consider
+indicators (i.e. 0's or 1's) in the salient weights vector, there is no
+reason why you couldn't weight the ranks by relative levels of
+importance. For instance, letting $m = \{3,1,0,0\}$ indicates that the
+salient ranks are 1 and 2, but getting the true best competitor as the
+winner is more important than ranking the true second best competitor
 second.

### 2025-11-01 mutual_information_curves.qmd
Fixed combinations issue in weighted Kendall's

@@
 
-weighted_kendall <- function(rhat, k, wtype = "symmetric_salience", salient_weights = NULL) {
+weighted_kendall <- function(rhat, k, wtype = "symmetric_salience", salient_weights = NULL, r = NULL) {
   n <- length(rhat)
-  #salient_weights needs to be a vector of length k
+  if (is.null(r)) r <- 1:n
+  if (is.null(salient_weights)) salient_weights <- rep(1, k)
 
-  if(is.null(salient_weights)){
-    salient_weights <- rep(1,k)
-  }
-
-  if (is.null(wtype)) {
+  # Weight matrix
   w <- matrix(0, ncol = n, nrow = n)
-    for (q in k:1){
-      w[q, ] <- salient_weights[q]
-      w[, q] <- salient_weights[q]
-    }
-    w <- w/2
-  }
-
-  if (wtype == "symmetric_salience") {
-    w <- matrix(0, ncol = n, nrow = n)
+  if (is.null(wtype) || wtype == "symmetric_salience") {
     for (q in k:1) {
@@
     }
-    #w[upper.tri(w)] <- w[upper.tri(w)]/2
-    #w[lower.tri(w)] <- w[lower.tri(w)]/2
   }
-
-  #Ryan method
-  if (wtype == "rydog"){
-  w <- matrix(0, ncol = n, nrow = n)
-#  w[1:k, 1:k] <- 1
+  if (wtype == "rydog") {
     for (q in k:1) {
       w[1:q, 1:q] <- salient_weights[q]
-    w[1:q, 1:q] <- salient_weights[q]
-  }
-  #w[upper.tri(w)] <- w[upper.tri(w)]/2
-  #w[lower.tri(w)] <- w[lower.tri(w)]/2
-  }
-
-  summ <- 0
-  for (i in 1:n) {
-    for (j in 1:n) {
-      if (i != j) {
-        summ <- summ + w[i, j] * sign(i - j) * sign(rhat[i] - rhat[j])
-        #print(summ)
-      }
-
     }
@@
 
-  out <- summ / (sum(w) - sum(diag(w)))
+  # Compute weighted Kendall numerator and denominator consistently
+  summ <- 0
+  denom <- 0
+  for (i in 1:n) {
+    for (j in 1:n) {
+      if (i != j && w[i, j] != 0) {
+        # Skip uninformative pairs (ties in either vector)
+        if (r[i] == r[j] || rhat[i] == rhat[j]) next
+
+        denom <- denom + abs(w[i, j])
+        summ <- summ + w[i, j] * sign(r[i] - r[j]) * sign(rhat[i] - rhat[j])
+      }
+    }
+  }
+
+  # Handle case where denom = 0 (e.g., all ties)
+  if (denom == 0) return(1)
+
+  out <- summ / denom
   weight <- (out + 1) / 2
-  #print(sum(w))
-  #print(sum(diag(w)))
   return(weight)
+}
 
-}
 weighted_kendall(c(2,1,3,4), k=4, salient_weights = c(3,1,1,0))
 weighted_kendall(c(2,1,3,4), k=2)
+
+weighted_kendall(c(1,1,1,1,5,6), k = 6, r = c(1,1,1,1,5,6))
+# Only works for k = # of top teams tied
 
@@
 High School Tennis Ladder - 1 vs 2, 3 vs 4, etc., Re-rank 2 vs 3, 4 vs 5, 6 vs 7, number of teams - 1 iterations
+
+## Combinations
+
+-   Manually setting the top-K teams with r=1 to look at if the top K teams were in the top K spots
+
+```{r}
+compute_mut_plots_combinations <- function(dataset, num_teams) {
+
+  results <- numeric(num_teams)
+
+  for (k in 1:num_teams) {
+    df <- dataset %>%
+      mutate(
+        true_rank = ifelse(true_rank <= k, 1, true_rank),
+        rank_hat = ifelse(rank_hat <= k, 1, rank_hat)
+      )
+    binary_vector <- c(rep(1, k), rep(0, num_teams - k))
+    results[k] <- mut_info(df, num_teams, binary_vector)
+  }
+  return(results)
+}
+
+combinations_se <- compute_mut_plots_combinations(all_results_norm, num_teams=8)
+combinations_rr <- compute_mut_plots_combinations(test_rr_noTies, num_teams=8)
+
+# Define x values
+x <- 1:8
+
+# Combine into one data frame for plotting
+plot_data <- data.frame(
+  x = rep(x, 2),
+  y = c(1-mi_se_norm, 1-combinations_se, 1-mi_test_rr, 1-combinations_rr),
+  type = rep(c("mi_se_norm", "combinations_se", "mi_test_rr", "combinatinos_rr"), each = 8)
+)
+
+# Create plot
+ggplot(plot_data, aes(x = x, y = y, color = type, group = type)) +
+  geom_line(size = 1.2) +
+  geom_point(size = 3) +
+  scale_x_continuous(breaks = 1:8) +
+  labs(
+    x = "Salient Ranks",
+    y = "Value",
+    color = "Series",
+    title = "Comparison of MI vs Top K-teams"
+  ) +
+  theme_minimal(base_size = 14) +
+  theme(
+    legend.position = "top",
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```

### 2025-11-06 mutual_information_curves.qmd
updated weighted kendall

@@
 ## Weight Structure (Partial Kendall or Weighted Kendall)
+
+##
+
+```{r}
+r <- c(1,2,3,4,5)
+rhat <- c(4,1,1,1,5)
+n <- length(r)
+k <- sum(rhat == 1)
+count <- 0
+for (i in 1:(length(r)-1)){
+  for (j in ((i+1):(length(r)))){
+    print(sign(r[i] - r[j])*sign(rhat[i] - rhat[j]))
+    count <- count + sign(r[i] - r[j])*sign(rhat[i] - rhat[j])
+  }
+}
+
+count
+#Correlation
+count / (choose(n,2) - choose(k,2))
+#Weight
+((count / (choose(n,2) - choose(k,2))) + 1)/2
+
+```
 
@@
 
-weighted_kendall <- function(rhat, k, wtype = "symmetric_salience", salient_weights = NULL, r = NULL) {
+weighted_kendall <- function(rhat, k, wtype = "symmetric_salience", salient_weights = NULL) {
   n <- length(rhat)
-  if (is.null(r)) r <- 1:n
+  #r <- 1:n
   if (is.null(salient_weights)) salient_weights <- rep(1, k)
@@
   denom <- 0
-  for (i in 1:n) {
-    for (j in 1:n) {
-      if (i != j && w[i, j] != 0) {
-        # Skip uninformative pairs (ties in either vector)
-        if (r[i] == r[j] || rhat[i] == rhat[j]) next
-
-        denom <- denom + abs(w[i, j])
-        summ <- summ + w[i, j] * sign(r[i] - r[j]) * sign(rhat[i] - rhat[j])
+  for (i in 1:(n-1)) {
+    for (j in (i+1):n) {
+      if (i != j) {
+        sign_ij <- sign(i - j)
+        sign_rhat <- sign(rhat[i] - rhat[j])
+        sign_rhat <- ifelse(sign_rhat == 0, -1, sign_rhat)
+        summ <- summ + w[i, j] * sign_ij * sign_rhat
+        denom <- denom + w[i, j]
+        #print(c(sign_r, sign_rhat, summ, denom))
       }
@@
   }
-
-  # Handle case where denom = 0 (e.g., all ties)
-  if (denom == 0) return(1)
 
@@
 weighted_kendall(c(2,1,3,4), k=4, salient_weights = c(3,1,1,0))
-weighted_kendall(c(2,1,3,4), k=2)
+weighted_kendall(c(1.5,3,1.5), k=2)
+weighted_kendall(c(1,3,4,2), k=3)
+weighted_kendall(c(1,1,4,1), k=3)
 
-weighted_kendall(c(1,1,1,1,5,6), k = 6, r = c(1,1,1,1,5,6))
+# Define the function calls
+results <- data.frame(
+  rhat = rep(c("(1,1,3)", "(1,3,1)", "(3,1,1)"), each = 3),
+  k = rep(1:3, times = 3),
+  weighted_kendall = c(
+    weighted_kendall(c(1,1,3), k=1),
+    weighted_kendall(c(1,1,3), k=2),
+    weighted_kendall(c(1,1,3), k=3),
+    weighted_kendall(c(1,3,1), k=1),
+    weighted_kendall(c(1,3,1), k=2),
+    weighted_kendall(c(1,3,1), k=3),
+    weighted_kendall(c(3,1,1), k=1),
+    weighted_kendall(c(3,1,1), k=2),
+    weighted_kendall(c(3,1,1), k=3)
+  )
+)
+
+results
+
+weighted_kendall(c(1,1,1,1,5,6), k = 3, r = c(1,1,1,1,5,6))
 # Only works for k = # of top teams tied
@@
   )
-
 ```

### 2025-11-06 NFl2010theta.R
thetas

@@ -1,0 +1,22 @@
+#https://www.pro-football-reference.com/years/2010/games.htm
+library(tidyverse)
+nfl <- read.csv("./NFL2010.csv")
+nfl <- nfl %>% mutate(Week = as.numeric(Week)) %>% filter(Week <= 17) %>% mutate(home = ifelse(X == "@",Loser.tie, Winner.tie),
+                                                                                 away = ifelse(X== "@", Winner.tie, Loser.tie))
+teams <- sort(unique(nfl$Winner.tie))
+x <- matrix(0, nrow = nrow(nfl), ncol = length(teams))
+for (i in 1:nrow(nfl)){
+  for (j in 1:length(teams)){
+    x[i,which(nfl$home[i] == teams)] <- 1
+    x[i,which(nfl$away[i] == teams)] <- -1
+  }
+}
+
+a <- glm((nfl$home == nfl$Winner.tie) ~ x, family = "binomial")
+theta <- data.frame(
+  team = c("int",teams),
+  theta = a$coefficients)
+theta$theta[length(theta$theta)] <- 0
+theta %>% arrange(-theta)
+
+

### 2025-11-07 manuscript/paper.Rmd
fucking citation.
paper update.

@@
 $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that
-$(\theta_1 > \theta_2 > \ldots > \theta_n)$. Next, let
+$(\theta_1 \ge \theta_2 \ge \ldots \ge \theta_n)$. Next, let
 $r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when
 $\theta$ is sorted from largest to smallest so that
-$r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}$. Next define
+$r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}$ (Note: While we allow for two values of $\theta$ to be equal to each other, we choose to assign unique ranks to each $\theta_i$).  Next define
 $T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure
@@
 $n$-dimensional vector-valued random variable that gives the results of
-a tournament as a vector of the indexes of the vector
-$\boldsymbol{\theta}$. The index in the first position of the vector
+a tournament as a vector of the indexes of
+$\boldsymbol{\theta}$. The number in the first position of the vector
 $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
-team that finished first, the index in the second position of the vector
+team that finished first, the number in the second position of the vector
 $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
-team that finished second, and so on.
+team that finished second, and so on.  (Note: We do allow for ties in $\hat{r}$.)
 
-For example, in a $n$ = 4 team tournament,
-$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that
+For example, in a $n$ = 4 team tournament, $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that
 the team with index 3 (i.e. the true third best team) won the
@@
 
+As noted above we allow for ties in $\hat{r}$.
+Continuing the above example, if $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 3) this indicates that the teams that are truly ranked 1st and 4th tied for 3rd.
+
+We need a transition: 
 If one views the true ranking of the teams in a tournament as a message
@@
 to correctly rank teams. Specifically, we start with the concept of
-mutual information (@Guiasu1977-nu). For two random variables $X$ and
-$Y$ mutual information is defined as:
+mutual information (@Guiasu1977-nu).
+
+For two random variables $X$ and $Y$ mutual information is defined as:
 
@@
 
-We note here that $r$ is not a random variable, however, since the
-indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary
-(we use 1, 2, 3, etc. for convenience, but any indexing is valid), we
-can view this a random variable where all permutation so of the $n$
-teams are equally likely. Therefore when calculating the mutual
-information in this setting, we set $p(r) = \frac{1}{n!}$. By a similar
-argument we set $p(\hat{r}) = \frac{1}{n!}$.
+In order to compute mutual information, marginal distributions of $r$ and $\hat{r}$ are required.
+Without knowing $r$, all possible permutations of $\hat{r}$ are equally likely so we set $p(\hat{r}) = \frac{1}{n!}$.
+By a similar argument we set $p(r) = \frac{1}{n!}$.
+
+we set . By a similar argument we set $p(\hat{r}) = \frac{1}{n!}$.
 
@@
 I(r,\hat{r})= n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{\left(\frac{1}{n!}\right)^2} = n! log((n!)^2)
-$$ and for the specific case when $n=4$ would be: $$
+$$ and for the specific case when $n=4$ would be: 
+$$
 4! log_2(4!^2) = 24*log_2(24^2) = 220.0782
 $$
+
 # Weighted Mutual Information
@@
 We want to give more weight to permutations from $\hat{r}$ that are "closer" to $r$.
-Essentially, this problem comes down to choosing a metric that measures the distance between the vectors $\hat{r}$ and $r$.
+Essentially, this problem comes down to choosing a metric that measures the closeness between the vectors $\hat{r}$ and $r$.
 One then gives more weight to combinations of $\hat{r}$ and $r$ that are closer and less weight to combinations that are far apart.
@@
 This requires two adjacent swaps to get $\hat{r}$ to equal $r$ (i.e swap 4 with 2 and then 3 with 2).
-The Kendall's $\tau$ value between these two vectors is `r cor(c(1,2,3,4),c(1,3,4,2), method = "kendall")`.
-We would then define the weight between the two vectors to be $(\rho_{\tau} + 1)/2 =$ `r (cor(c(1,2,3,4),c(1,3,4,2), method = "kendall") + 1)/2`.
+The Kendall's $\tau$ value between these two vectors is $\frac{1}{3}$.
+We would then define the weight between the two vectors to be $(\rho_{\tau} + 1)/2 = \frac{2}{3}$ .
 This is exactly equivalent to $1 - \frac{\#\text{ adj swaps}}{{n \choose 2}}$.
+
+<!-- `r cor(c(1,2,3,4),c(1,3,4,2), method = "kendall")` -->
+<!-- `r (cor(c(1,2,3,4),c(1,3,4,2), method = "kendall") + 1)/2` -->
 
@@
 
+One and appropriate weight function was chosen, we get the following formula for weighted information:
+
+$$
+I_w(r,\hat{r})= \sum_{r}\sum_{\hat{r}}w(r, \hat{r})p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
+$$
+
 # Salient Weights (and ranks)
 
-@shieh1998weighted: Weighted Kendall's Tau
+@Shieh1998: Weighted Kendall's Tau
 
@@
 $$
-w(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}}) = \sum_{i = 1}^n m_i \times d(\hat{r}_i, r_i)
-
-\\
-
-wI_m(r_{\{{i|m_i >0}\}}, \hat{r_{\{{i|m_i >0}\}}})=\sum_{r_{\{{i|m_i >0}\}}}\sum_{\hat{r}_{\{{i|m_i >0}\}}} w(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}}) \times \\
-p(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}})\times \\ \log \frac {p(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}})}{p(r_{\{{i  \mid m_i >0}\}})p(\hat{r}_{\{{i \mid m_i >0}\}})}
+wI_m(r_{\{{i|m_i >0}\}}, \hat{r}_{\{{i|m_i >0}\}})=\sum_{r_{\{{i|m_i >0}\}}}\sum_{\hat{r}_{\{{i|m_i >0}\}}} w_m(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}}) \times
 $$
 
-Because this mutual information is unitless, the formula can be
-standardized to be between 0 and 1 as follows: $$
-\frac{wI_m(r,\hat{r})}{\text{max\{}H(r), H(\hat{r})\}}
+<!-- $$ -->
+<!-- p(\hat{r}_{\{{i|m_i > 0}\}}, r_{\{{i|m_i >0}\}})\times \\  -->
+
+<!-- \log \frac{p(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}})}{p(r_{\{{i  \mid m_i >0}\}})p(\hat{r}_{\{{i \mid m_i >0}\}})} -->
+<!-- $$ -->
+
+Because this mutual information is unitless, the formula can be standardized to be between 0 and 1 as follows: 
+$$
+\frac{wI_m(r,\hat{r})}{\text{max}\{H(r), H(\hat{r})\}}
 $$

### 2025-11-07 manuscript/paper.Rmd
updated.

@@
 
-One and appropriate weight function was chosen, we get the following formula for weighted information:
+Once an appropriate weight function $w(r, \hat{r})$ is chosen, we get the following formula for weighted information:
 
@@
 Regardless of which method one chooses for defining weights between $\hat{r}$ and $r$, a second set of weights needs to be defined because the number of teams that need to be accurately ordered differs based on the choice of salient ranks.
-Therefore, we define a second set of weights, $m$, which is used to indicate which ranks are salient.
-We call these the "salient weights" to differentiate from the other weighting function, $w$.
+Therefore, we define a second vector of weights, $m$, which is used to indicate which ranks are salient.
+We call these the "salient weights" to differentiate from the other weighting function, $w$, becomes a function of $m$ (and also, as before, $r$ and $hat{r}$).
 As an example, for a four-team tournament, if the organizer decided that only the correctly choosing the best team as the winner was important, then $m = \{1,0,0,0\}$ (i.e.
@@
 
-^[7]: One can imagine that in an event like the Olympics, the salient weights would be 1's for the first three elements and 0's elsewhere.
+^[7]: One can imagine that in an event like the Olympics, the salient weights would be 1's for the first three elements and 0's elsewhere (i.e. $m = \{1,1,1,0\}$).
 
@@
 $$
-wI_m(r_{\{{i|m_i >0}\}}, \hat{r}_{\{{i|m_i >0}\}})=\sum_{r_{\{{i|m_i >0}\}}}\sum_{\hat{r}_{\{{i|m_i >0}\}}} w_m(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}}) \times
+I^m_w(r,\hat{r})= \sum_{r}\sum_{\hat{r}}w(r, \hat{r}, m)p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
 $$
+.
+
+If we choose to use the weighted Kendall's $\tau$ for our weights, $w(r, \hat{r}, m)$ would be:
+
+$$
+w(r, \hat{r}, m) = \frac{\sum_{i \ne j} m^{\star}_{ij}\text{sign}_{(0+)}(r_i - r_j)sign_{(0+)}(\hat{r}_i - \hat{r}_j)}{\sum_{i,j}m^{\star}_{i,j}- \sum_i m^{\star}_{i,i}}
+$$
+where $m^{\star}_{i,j} = m_{min(i,j)}$ and $sign_{(0+)}(.)$ is the sign of the input with the special case that a 0 maps to 1.
+
+<!-- $$ -->
+<!-- wI_m(r_{\{{i|m_i >0}\}}, \hat{r}_{\{{i|m_i >0}\}})=\sum_{r_{\{{i|m_i >0}\}}}\sum_{\hat{r}_{\{{i|m_i >0}\}}} w_m(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}}) \times -->
+<!-- $$ -->
 
@@
 $$
-\frac{wI_m(r,\hat{r})}{\text{max}\{H(r), H(\hat{r})\}}
+\frac{I^m_w(r,\hat{r})}{\text{max}\{H(r), H(\hat{r})\}}
 $$

### 2025-11-07 brackets_pictures.R

@@ -1,0 +1,51 @@
+plot(0,0,xlim = c(0,3), ylim = c(0,10), frame.plot= FALSE, xlab= "", ylab = "",xaxt= 'n', yaxt = 'n', col = "white", asp = 1/2)
+
+points(c(0,1), c(0,1), type = "l")
+points(c(0,1), c(2,1), type = "l")
+points(c(0,1), c(4,5), type = "l")
+points(c(0,1), c(6,5), type = "l")
+
+points(c(1,2), c(1,3), type = "l")
+points(c(1,2), c(5,3), type = "l")
+
+
+
+plot(0,0,xlim = c(0,3), ylim = c(0,30), frame.plot= FALSE, xlab= "", ylab = "",xaxt= 'n', yaxt = 'n', col = "white", asp = 1/3)
+
+points(c(0,1), c(0,2), type = "l")
+points(c(0,1), c(4,2), type = "l")
+points(c(0,1), c(8,10), type = "l")
+points(c(0,1), c(12,10), type = "l")
+points(c(0,1), c(16,18), type = "l")
+points(c(0,1), c(20,18), type = "l")
+points(c(0,1), c(24,26), type = "l")
+points(c(0,1), c(28,26), type = "l")
+
+points(c(1,2), c(2,6), type = "l")
+points(c(1,2), c(10,6), type = "l")
+points(c(1,2), c(26,22), type = "l")
+points(c(1,2), c(18,22), type = "l")
+
+points(c(2,3), c(22,14), type = "l")
+points(c(2,3), c(6,14), type = "l")
+
+#step ladder
+plot(0,0,xlim = c(0,3), ylim = c(0,10), frame.plot= FALSE, xlab= "", ylab = "",xaxt= 'n', yaxt = 'n', col = "white", asp = 1/2)
+
+points(c(0,-1), c(0,0), type = "l")
+points(c(0,-1), c(4,4), type = "l")
+points(c(1,0), c(6,6), type = "l")
+points(c(2,1), c(8,8), type = "l")
+
+points(c(0,1), c(0,2), type = "l")
+points(c(0,1), c(4,2), type = "l")
+
+points(c(1,2), c(2,4), type = "l")
+points(c(1,2), c(6,4), type = "l")
+
+points(c(2,3), c(8,6), type = "l")
+points(c(2,3), c(4,6), type = "l")
+
+
+
+

### 2025-11-07 group_stage.qmd
Updates

@@
 ```{r}
-simulate_group_stage_tournament <- function(num_teams, distribution, group_length = 4, ties = TRUE, rounds = 1, seeded=F) {
+simulate_group_stage_tournament <- function(num_teams, distribution, group_length = 4, ties = TRUE, rounds = 1, third_place = T, theta_hat = NULL) {
 
@@
   else if (distribution == "Manual") {
-    cat("Please enter the strengths for each team in descending rank order:\n")
-    strengths <- numeric(num_teams)
-    for (i in num_teams:1) {
-      repeat {
-        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
-        if (!is.na(input)) {
-          strengths[i] <- input
-          break
-        } else {
-          cat("Invalid input. Please enter a numeric value.\n")
-        }
-      }
-    }
-    team_strengths <- data.frame(
-      true_rank = as.numeric(teams),
-      theta = strengths,
+    strengths <- theta_hat
+    manual_strengths <- data.frame(
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
       rank_hat = rep(NA, length(teams)),
@@
     )
-    df <- arrange(team_strengths, true_rank)
+    df <- manual_strengths %>% arrange(true_strength) %>%
+      mutate(true_rank = as.numeric(teams)) %>%
+      arrange(true_strength, true_rank)
   }
@@
 
+  # Get top finishers from each group (group_rank <= 1.5 covers both rank 1 and tied 1.5)
   group_winners <- final_df %>%
-    filter(group_rank == 1) %>%
+    group_by(groups) %>%
+    filter(group_rank == min(group_rank)) %>%  # Get best finisher(s)
+    slice(1) %>%  # If still tied, take first one
+    ungroup() %>%
     arrange(groups)
 
+  # Get second place finishers (next best after winners)
   group_seconds <- final_df %>%
-    filter(group_rank == 2) %>%
+    group_by(groups) %>%
+    arrange(group_rank, true_rank) %>%  # Sort by rank, then true_rank as tiebreaker
+    slice(2) %>%  # Take second team
+    ungroup() %>%
     arrange(groups)
@@
 
-  knockout_results <- simulate_knockout(df = final_df, matchup_vector = matchup_vector, series = 1, third_place = TRUE)
+  knockout_results <- simulate_knockout(df = final_df, matchup_vector = matchup_vector, series = 1, third_place = third_place)
 
@@
 ) %>%
-  arrange(true_rank) %>%
+  arrange(true_rank)
+  if (ties == F){
+    final_df <- final_df %>%
       mutate(rank_hat = rank(rank_hat, ties.method = "random"))
+  } else{
+    final_df <- final_df %>%
+      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
+  }
 
@@
 # Single simulation
-result <- simulate_group_stage_tournament(num_teams = 8, distribution = "Normal", group_length = 4, ties=F)
+result <- simulate_group_stage_tournament(num_teams = 16, distribution = "Same", group_length = 4, ties=T)
 

### 2025-11-07 mutual_information_curves.qmd
Updates

@@
 
-weighted_kendall(c(2,1,3,4), k=4, salient_weights = c(3,1,1,0))
+weighted_kendall(rhat=c(2,1,3,4), k=4)
 weighted_kendall(c(1.5,3,1.5), k=2)
@@
 
-Triple Elimination
+~~Triple Elimination~~
 
@@
 
-Round Robin - with k rounds (1,2,3,5,7)
+Round Robin - with m rounds (1,2,4)
 
-Stepladder - 8 vs 7, winner plays 6, etc.
+Stepladder- 8 vs 7, winner plays 6, etc.
+
+Waterfall - stepladder round robin
 
@@
 
-Staged Round Robin/Jacobian Ladder - top 2 in each group form new group, bottom 2 new group (each new group is a new round robin)
+Staged Round Robin - top 2 in each group form new group, bottom 2 new group (each new group is a new round robin)
 

### 2025-11-07 single_elimination_shiny.qmd
Updates

@@
 # Simulate the single-elimination tournament
-simulate_single_elimination_tournament <- function(num_teams, distribution, seeding_structure = NULL, ties = T, series = 1, third_place = T, reseeded = F) {
+simulate_single_elimination_tournament <- function(num_teams, distribution, seeding_structure = NULL, ties = T, series = 1, third_place = T, reseeded = F, theta_hat = NULL) {
 
@@
   }
+
   teams <- paste(num_teams:1)
+
   if (log2(num_teams) %% 1 != 0) {
@@
   else if (distribution == "Manual") {
-    cat("Please enter the strengths for each team in descending rank order:\n")
-    strengths <- numeric(num_teams)
-    for (i in num_teams:1) {
-      repeat {
-        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
-        if (!is.na(input)) {
-          strengths[i] <- input
-          break
-        } else {
-          cat("Invalid input. Please enter a numeric value.\n")
-        }
-      }
-    }
+    strengths <- theta_hat
     manual_strengths <- data.frame(
-      true_rank = as.numeric(teams),
       true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
@@
     )
-    df <- arrange(manual_strengths, true_rank)
+    df <- manual_strengths %>% arrange(true_strength) %>%
+      mutate(true_rank = as.numeric(teams)) %>%
+      arrange(true_strength, true_rank)
   }
@@
         valid_teams <- teams[!is.na(teams)]
-        na_teams <- teams[is.na(teams)]
+        num_byes <- sum(is.na(teams))
 
@@
 
-        # Create reseeded matchups: 1 vs worst, 2nd vs 2nd worst, etc.
+        # Create reseeded matchups with byes going to top seeds
         n_valid <- length(sorted_teams)
+
         if (n_valid > 1) {
           reseeded_teams <- c()
+
+          # If there are byes, give them to the top seeds
+          if (num_byes > 0) {
+            # Top seeds get byes
+            for (j in 1:num_byes) {
+              reseeded_teams <- c(reseeded_teams, sorted_teams[j], NA)
+            }
+            # Remaining teams play each other
+            remaining_teams <- sorted_teams[(num_byes + 1):n_valid]
+            n_remaining <- length(remaining_teams)
+            if (n_remaining > 1) {
+              for (j in 1:(n_remaining/2)) {
+                reseeded_teams <- c(reseeded_teams, remaining_teams[j], remaining_teams[n_remaining - j + 1])
+              }
+            }
+          } else {
+            # No byes: standard reseeding (best vs worst)
             for (j in 1:(n_valid/2)) {
@@
             }
-          # Add back any NA teams (byes)
-          teams <- c(reseeded_teams, na_teams)
+          }
+
+          teams <- reseeded_teams
         }

### 2025-11-07 NFl2010theta.R

@@ -19,4 +19,39 @@ theta <- data.frame(
 theta$theta[length(theta$theta)] <- 0
 theta %>% arrange(-theta)
 
+nfc_theta <- theta %>% filter(team %in%
+  c("Atlanta Falcons", "Chicago Bears", "Philadelphia Eagles", "Seattle Seahawks", "New Orleans Saints", "Green Bay Packers")) %>%
+  pull(theta)
+
+afc_theta <- theta %>% filter(team %in%
+  c("New England Patriots", "Pittsburgh Steelers", "Indianapolis Colts", "Kansas City Chiefs", "Baltimore Ravens", "New York Jets")) %>%
+  pull(theta)
+
+desired_nfc_order <- c(
+  "Atlanta Falcons",
+  "Chicago Bears",
+  "Philadelphia Eagles",
+  "Seattle Seahawks",
+  "New Orleans Saints",
+  "Green Bay Packers"
+)
+
+nfc_teams <- theta %>%
+  filter(team %in% desired_nfc_order) %>%
+  slice(match(desired_nfc_order, team)) %>%
+  pull(team)
+
+desired_afc_order <- c(
+  "New England Patriots",
+  "Pittsburgh Steelers",
+  "Indianapolis Colts",
+  "Kansas City Chiefs",
+  "Baltimore Ravens",
+  "New York Jets"
+)
+
+afc_teams <- theta %>%
+  filter(team %in% desired_afc_order) %>%
+  slice(match(desired_afc_order, team)) %>%
+  pull(team)
 

### 2025-11-07 sources_summary.qmd
add lit review draft

@@
 
-# Tournament Papers - Summaries
+# Tournament Paper - Lit Review
+
+# Lit Review
+
+There are hundreds of different competitions structures that dictate how a set of teams compete against each other to determine an overall ranking.
+Common structures include bracket tournaments where if a team wins their competition, they advance to the next round.
+Bracket tournaments can be run as single elimination where competitors advance or are eliminated based on a single match (e.g.
+NCAA College Basketball Tournament, National Football League (NFL) playoffs, major tennis tournaments) or where each round is contested as a best of $k$ series (e.g.
+$k$ = 7: National Basketball Association (NBA) playoffs, National Hockey League (NHL) playoffs, $k$ varies by round: Major League Baseball (MLB), Women's National Basketball Association (WNBA)) and the series winner advance while the series loser is eliminated.
+Additional styles of the bracket tournaments include variations where competitors are not eliminated after a single loss, such as in double elimination tournaments where competitors continue until they have lost twice.
+Within double elimination tournaments, there are several variations.
+For instance, the NCAA College Wrestling tournament is a double elimination tournament, however, if a competitor loses a match, they fall into the consolation bracket and the best they can finish after a loss before the finals is 3rd place, making this not a "true" double elimination format.
+Olympic judo uses a variant of this system with single elimination until the quarterfinals followed by a consolation bracket consisting of the four quarterfinal losers who still have a chance for a bronze medal[^1].
+
+[^1]: in Olympic judo they award two bronze medals to the two competitors who each win their side of the consolation bracket
+
+"True" double elimination - a bracket tournament where a competitor can still finish ranked first as long as a competitor has not lost two games - is used in NCAA baseball and softball for certain rounds in their postseason.
+Additional variants of double elimination include double elimination brackets with a "true third" match, which includes an additional match contested between the competitors ranked 2nd (loser of the finals) and 3rd (winner of the consolation bracket) at the end of the tournament *if they have not already competed against each other*.
+There is also a variation of double elimination where only those competitors who have lost to a competitor that makes the finals are placed into the consolation bracket and given the chance to compete for third place.
+This style of competition is used in Olympic wrestling (both freestyle and Greco-Roman)[^2].
+
+[^2]: As in Olympic judo, Olympic wrestling also award two bronze medals to the winners of each side of the consolation bracket
+
+<!-- A paragraph about Swiss, Dutch, and Monrad systems -->
+
+In bracket style competitions, only a handful of the match-ups that are possible actually occur.
+When every possible competitive pairing occurs, this is referred to as a round-robin structure where every competitor plays every other competitor the same number of times.
+Single round-robin, where each possible match-up occurs exactly once, is used in, for examples, the group stage of the FIFA World Cup to create a ranking to determine which teams advance to the next round.
+Double round-robin, where each possible pairing occurs twice, is used in association football in English Premier League (EPL) and in chess during the FIDE candidates tournament (a tournament that decides who will challenge the world champion).
+For the sake of brevity, we review only a small number of possible competition structures, but the interested reader can find a more complete list of competition structures in @Goosens2024, @Appleton1995-zn and @Csato2021.
+
+A natural question of interest after reviewing different types of competition structures is: which competition structure is best?
+As with most complicated questions, the answer is "it depends".
+One commonsense conceptualization for the best competition structure is that the final rankings produced by a competition structure should closely match the true rankings of the strength of the competitors.
+That is, a good competition structure should rank the true best team first, the true second best team second, and so on, until finally ranking the true worst team last.
+We refer to this property of a competition structure as the efficacy of the structure (@Lasek2018-de, @Sziklai2022-kf).
+If one were only interested in correctly ranking the true best competitor as first, with the ranks of the remaining competitors viewed as inconsequential, we refer to this here are effectivity (@Glenn1960-mn).
+While there are many competition structures where efficacy and effectivity are useful measures[^3], there are many other competitions where the correct ranking of more than one competitor, but not all the competitors is of import.
+For instance, in Olympic competition, the goal of a competition structure is to correctly rank the top 3 competitors to award gold, silver, and bronze medals to[^4].
+In many American sports, for instance the NFL, the goal of the competition structure of the regular season is to rank the top 7 teams in each conference to choose the playoff teams with the ranking of the teams 8 through 16 in each conference inconsequential in determining a Super Bowl champion[^5].
+We coin the term "salient ranks" to refer to the ranks of a competition structure that matter.
+For example, the ranks 1, 2, and 3 in Olympic competition are the salient ranks in that competition structure.
+In this notation, efficacy is a measure with with salient ranks 1 through $n$ where $n$ is the number of competitors, and effectivity is a measure where 1 is the only salient rank.\
+<!-- Efficacy: @Lasek2018-de, @Sziklai2022-kf --> <!-- Effectivity: @Glenn1960-mn -->
+
+[^3]: efficacy is a useful measure for the EPL, for example, as the ranking of the teams matters at both the high end (i.e.
+entry into special tournaments) and the low end (i.e.
+relegation status) and effectivity is a useful measure for the NCAA College Basketball tournament (i.e.
+main goal is to crown a champion)
+
+[^4]: Some sports award two bronze medals (e.g. wrestling, judo, boxing, tae kwon do)
+
+[^5]: The ranking of teams outside of the playoffs does matter for draft purposes though
+
+<!-- From Goosens 2024: The ability of a tournament to rank contestants according to their true strength is called efficacy (Lasek and Gagolewski, -->
+
+<!-- 2018; Sziklai et al., 2022). Efficacy with respect to only the best player is also known as effectivity (Glenn, 1960) -->
+
+<!-- or predictive power (Ryvkin and Ortmann, 2008; Vu and Shoham, 2011). -->
+
+If the only goal of a competition structure was high efficacy of the salient ranks, the best way to achieve this would be to play each pair of possible match ups of competitors as many times as possible.
+Of course, this is impossible due to many factors including time constraints, cost constraints, and wear and tear on competitors in high impact sports (e.g.
+boxing, American football, etc.).
+In addition, competitions where the true best team is guaranteed or nearly guaranteed to win may be less appealing to fans and may drive down demand as discussed in @Johnson2022-wz.
+
+In light of this, many different competition structures have been proposed, several of which are mentioned previously.
+Given these many competition structures, ideally there would be a way comparing the performance of different competition structures to each other.
+
+In this paper, we propose a novel metric for evaluating different competition structures using an information theoretic approach.
+As an analogy, we view the results of a tournament as a transmitted message, and we use information theory to assess how accurately the message (i.e.
+the tournament results) matches the true transmission (i.e.
+the true ranking of competitors).
+
+# Notes from Other Papers
 
@@
 With n objects, we suppose that each of the (n choose 2) pairs are presented to an observer and their preference noted.
-With m observers, there are m * (n choose 2) preferences.
+With m observers, there are m \* (n choose 2) preferences.
 
@@
 -   Tournament outcomes can be influenced by several factors, such as attack, defensse, and home advantage strength values, as well as tournament structures.
+
 -   However, the claim that different structures have their own time restraints and requirements has limited the evaluation of the best structure for a particular type of sports tournament using quantitative approaches.

### 2025-11-07 double_elimination.qmd

@@
 ```{r}
-simulate_double_elimination_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1) {
-
+simulate_double_elimination_tournament <- function(num_teams, distribution, seeding_structure = NULL, ties = T, series = 1, final = T, theta_hat = NULL) {
   if (num_teams <= 3) {
@@
   }
-  if (distribution == "Normal"){
-    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
+
+  if (distribution == "Normal") {
+    strengths <- sapply(num_teams, function(n) {
+      qnorm(1:n/(n+1))
+    })
+    normal_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- arrange(normal_strengths, true_rank)
+  } else if (distribution == "Uniform") {
+    strengths <- sapply(num_teams, function(n) {
+      qunif(1:n/(n+1), 0, sqrt(12))
+    })
+    unif_strength <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams)),
+      rank_hat = rep(NA, length(teams))
+    )
+    df <- arrange(unif_strength, true_rank)
+  } else if (distribution == "Same"){
+    strengths <- sapply(num_teams, function(n) {rep(0,n)})
     normal_strengths <- data.frame(
@@
     df <- arrange(normal_strengths, true_rank)
-  }
-  else if (distribution == "Uniform"){
-    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
-    unif_strength <- data.frame(
-      true_rank = as.numeric(teams),
-      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
-      game_wins = rep(0,length(teams)),
-      game_losses = rep(0,length(teams)),
-      rank_hat = rep(NA,length(teams))
-    )
-    df <- arrange(unif_strength, true_rank)
-  }
-     else if (distribution == "Manual") {
-    cat("Please enter the strengths for each team in descending rank order:\n")
-    strengths <- numeric(num_teams)
-    for (i in num_teams:1) {
-      repeat {
-        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
-        if (!is.na(input)) {
-          strengths[i] <- input
-          break
-        } else {
-          cat("Invalid input. Please enter a numeric value.\n")
-        }
-      }
-    }
+  } else if (distribution == "Manual") {
+    strengths <- theta_hat
     manual_strengths <- data.frame(
-      true_rank = as.numeric(teams),
       true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
@@
     )
-    df <- arrange(manual_strengths, true_rank)
-  }
-  else {
+    df <- manual_strengths %>%
+      arrange(true_strength) %>%
+      mutate(true_rank = as.numeric(teams)) %>%
+      arrange(true_strength, true_rank)
+  } else {
     stop("Distribution not found: Enter Manual to input your own strengths")
@@
   # Sets seed to that of common best vs worst structure
-  if (is.null(seeding_structure)){
+  if (is.null(seeding_structure)) {
     seeding_structure <- matrix(generate_elimination_tournament_structure(num_teams), nrow = 1)
-  } else{
+  } else {
     seeding_structure <- matrix(seeding_structure, nrow = 1)
   }
-  if(series %% 2 == 0) {
+
+  if (series %% 2 == 0) {
     stop("Series must be an odd number to avoid potential ties.")
@@
     strengths <- permuted_df$true_strength
-
     losers_bracket <- c()
@@
       eliminated <- c()
+
       if (length(winners_bracket) > 1) {
@@
 
-      for (j in seq(1, length(winners_bracket), by = 2)) { # Gets 2 teams for each match
+        for (j in seq(1, length(winners_bracket), by = 2)) {
+          # Gets 2 teams for each match
           if (j + 1 > length(winners_bracket)) break
+
           team1 <- winners_bracket[j]
           team2 <- winners_bracket[j+1]
-
           strength1 <- strengths[which(permuted_df$true_rank == team1)]
@@
           losses_team2 <- 0
-        for (game in 1:series){
+
+          for (game in 1:series) {
             match_winner <- simulate_match(team1, team2, strength1, strength2)
-          cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
-          if(!is.na(team1)) {
+            #cat(sprintf("Matchup: %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
+
+            if (!is.na(team1)) {
               if (match_winner == team1) {
@@
               }
-          }
-          else {wins_team2 <- (series+1) / 2}
+            } else {
+              wins_team2 <- (series+1) / 2
+            }
 
@@
           }
+
           # Update the game_wins and game_losses outside of the loop
           if (!any(is.na(c(team1, team2)))) {
-          cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
+            #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
             permuted_df <- permuted_df %>%
@@
           }
+
           # Determine the match winner based on series results
@@
           series_loser <- ifelse(series_winner == team1, team2, team1)
-        cat(sprintf("Matchup:  %s vs %s | Winner: %s\n: | Loser: %s\n", team1, team2, series_winner, series_loser))
+          #cat(sprintf("Matchup: %s vs %s | Winner: %s\n: | Loser: %s\n", team1, team2, series_winner, series_loser))
+
           next_winners <- c(next_winners, series_winner)
@@
         }
+
         winners_bracket <- next_winners # Winners move to the next round
         losers_bracket <- c(losers_bracket, round_losers)
-     if(length(losers_bracket) %% 2 != 0){losers_bracket <- c(losers_bracket, NA)}
-
+        if (length(losers_bracket) %% 2 != 0) {
+          losers_bracket <- c(losers_bracket, NA)
+        }
       }
@@
       if (length(losers_bracket) > 0) {
-      if(length(losers_bracket) > 1){
-      cat("\n--- Loser's Bracket --- \n")
+        if (length(losers_bracket) > 1) {
+          #cat("\n--- Loser's Bracket --- \n")
         }
@@
         # alternates games for losers bracket
-      if(length(losers_bracket) > 2){
-        if(length(losers_bracket) == 4 && is.na(losers_bracket[4])){
+        if (length(losers_bracket) > 2) {
+          if (length(losers_bracket) == 4 && is.na(losers_bracket[4])) {
             losers_bracket <- losers_bracket
-        } else{
-          top_bracket <- losers_bracket[seq(1,length(losers_bracket), by = 2)]
-          bottom_bracket <- losers_bracket[seq(2,length(losers_bracket), by = 2)]
-          losers_bracket <- c(top_bracket,bottom_bracket)
+          } else {
+            top_bracket <- losers_bracket[seq(1, length(losers_bracket), by = 2)]
+            bottom_bracket <- losers_bracket[seq(2, length(losers_bracket), by = 2)]
+            losers_bracket <- c(top_bracket, bottom_bracket)
           }
@@
           if (j + 1 > length(losers_bracket)) break
+
           team1 <- losers_bracket[j]
@@
             match_winner <- simulate_match(team1, team2, strength1, strength2)
-          if(is.na(match_winner)){
+            if (is.na(match_winner)) {
               wins_team1 == wins_team1 + 1
             }
-          if(!is.na(team1)) {
+
+            if (!is.na(team1)) {
               if (match_winner == team1) {
@@
           }
+
           if (!any(is.na(c(team1, team2)))) {
-          cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
+            #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
             permuted_df <- permuted_df %>%
@@
 
-        if(is.na(team1) && is.na(team2)){
+          if (is.na(team1) && is.na(team2)) {
             series_winner <- NA
             next_losers <- c(next_losers, NA)
-        } else{
+          } else {
             series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
@@
             eliminated <- c(eliminated, series_loser)
-        cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
-        }
-      }
+            #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, series_winner))
+          }
+        }
+
         team_rank <- team_rank - 1
@@
         losers_bracket <- next_losers
-
       }
@@
       if (length(winners_bracket) == 1 && length(losers_bracket) == 1) {
-
+        if (final == F){
+          # No final game played between winner of winner bracket and winner of loser bracket
+          team1 <- winners_bracket[1]
+          team2 <- losers_bracket[1]
+          permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
+          permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
+          break
+        }
         # Final matchup between winners and losers
-    cat("\n--- Final Match ---\n")
+        else if (final == T) {
+          #cat("\n--- Final Match ---\n")
           final_match <- c(winners_bracket, losers_bracket)
-    if(length(final_match) == 2){
+
+          if (length(final_match) == 2) {
             team1 <- final_match[1]
@@
               }
-
               if (wins_team1 > series / 2 || wins_team2 > series / 2) {
@@
             }
+
             if (!any(is.na(c(team1, team2)))) {
-      cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
+              #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
               permuted_df <- permuted_df %>%
@@
             if (wins_team1 > wins_team2) {
-      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
+              #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team1))
               permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
@@
             } else {
-      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
+              #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team2))
               # second game if the loser's bracket team wins
+              # RESET the counters for the second series
+              wins_team1 <- 0
+              wins_team2 <- 0
+              losses_team1 <- 0
+              losses_team2 <- 0
+
               for (game in 1:series) {
@@
                 }
-
                 if (wins_team1 > series / 2 || wins_team2 > series / 2) {
@@
               }
+
               if (!any(is.na(c(team1, team2)))) {
-        cat(sprintf("Wins - %s: %d | %s: %d\n", team1, wins_team1, team2, wins_team2))
+                #cat(sprintf("Wins - %s: %d | %s: %d\n", team1, wins_team1, team2, wins_team2))
                 permuted_df <- permuted_df %>%
@@
                     true_rank == team1 ~ game_wins + wins_team1,
-            true_rank == team2 ~ game_wins + wins_team2 - 1, # wins_team2 = 2 in loop but only need to add one game
+                    true_rank == team2 ~ game_wins + wins_team2,
                     TRUE ~ game_wins
@@
                   mutate(game_losses = case_when(
-            true_rank == team1 ~ game_losses + losses_team1 - 1,
+                    true_rank == team1 ~ game_losses + losses_team1,
                     true_rank == team2 ~ game_losses + losses_team2,
@@
               }
+
               if (wins_team1 > wins_team2) {
-      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
+                #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team1))
                 permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
@@
               } else {
-      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
+                #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team2))
                 permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
@@
       }
+    }
+
     permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
@@
   }
+
   final_results <- do.call(rbind, results)
@@
       mutate(rank_hat = rank(rank_hat, ties.method = "random"))
-  }
-  else{
+  } else {
     final_results <- final_results %>%
@@
     arrange(true_rank)
+
   return(final_results)
@@
 
-test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1)
+test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, final = F)
 #test

### 2025-11-07 nfl_playoffs.qmd

---
title: "NFL Playoffs"
format: html
editor: visual
---

## Standard NFL Playoff Format

```{r}
# Initialize storage for results
nfl_results <- vector("list", 10000)

for (sim in 1:10000){
  nfc <- simulate_single_elimination_tournament(num_teams = 6, distribution = "Manual", 
                                                 ties=T, series=1, 
                                                 seeding_structure = c(1,NA,5,4,6,3,2,NA), 
                                                 third_place=F, reseeded=T, theta_hat = nfc_theta)

  afc <- simulate_single_elimination_tournament(num_teams = 6, distribution = "Manual", 
                                                 ties=T, series=1, 
                                                 seeding_structure = c(1,NA,2,6,4,5,3,NA), 
                                                 third_place=F, reseeded=T, theta_hat = afc_theta)

  nfc_theta_df <- nfc %>%
    left_join(theta, by = c("true_strength" = "theta")) %>%
    select(c(true_strength, true_rank, rank_hat, team))

  afc_theta_df <- afc %>%
    left_join(theta, by = c("true_strength" = "theta")) %>%
    select(c(true_strength, true_rank, rank_hat, team))

  winner_nfc <- nfc_theta_df %>% filter(rank_hat == 1)
  winner_afc <- afc_theta_df %>% filter(rank_hat == 1)
  sb_winner <- simulate_match(winner_nfc$team, winner_afc$team, 
                              winner_nfc$true_strength, winner_afc$true_strength)

  # Combine and adjust rankings
  nfl_playoff_results <- rbind(nfc_theta_df, afc_theta_df) %>%
    mutate(
      # Super Bowl winner gets rank 1, both conference winners get rank 2, others shift down
      rank_hat = case_when(
        team == sb_winner ~ 1,
        rank_hat == 1 ~ 2,  # losing conference champion
        TRUE ~ rank_hat + 1
      )
    ) %>%
    arrange(-true_strength) %>%
    mutate(
      true_rank = 1:12,
      sim_id = sim
    )

  all_results[[sim]] <- nfl_playoff_results
}

# Combine all simulations
nfl_results <- bind_rows(all_results)
```

### 2025-11-07 round_robin.qmd

@@
 ```{r}
-simulate_roundrobin_r <- function(teams, strengths, r){
+simulate_roundrobin_r <- function(num_teams, distribution, ties = TRUE, r = 1, theta_hat = NULL) {
+  teams <- paste(1:num_teams)
   n <- length(teams)
-  win_counts <- numeric(n)
-  for (i in 1:r){
-    for (i in seq_along(teams)){
-      for (j in seq_along(teams)){
-        if (i != j){
-          match_winner <- simulate_match(teams[i],teams[j],strengths)
-          if (match_winner == teams[i]) {
-            win_counts[i] <- win_counts[i] + 1
+
+  # Initialize strength vector
+  if (distribution == "Normal") {
+    strengths <- qnorm(1:num_teams / (num_teams + 1))
+  } else if (distribution == "Same") {
+    strengths <- rep(0, num_teams)
+  } else if (distribution == "Uniform") {
+    strengths <- qunif(1:num_teams / (num_teams + 1), 0, sqrt(12))
+  } else if (distribution == "Manual") {
+    if (is.null(theta_hat)) {
+      stop("theta_hat must be provided for Manual distribution")
+    }
+    strengths <- theta_hat
   } else {
-            win_counts[j] <- win_counts[j] + 1
+    stop("Distribution not found: Enter Manual to input your own strengths")
+  }
+
+  # Initialize win/loss counters (indexed by original team position)
+  game_wins <- rep(0, num_teams)
+  game_losses <- rep(0, num_teams)
+
+  # Simulate round robin tournament r times
+  for (round in 1:r) {
+    for (i in 1:(num_teams - 1)) {
+      for (j in (i + 1):num_teams) {
+        team1 <- teams[i]
+        team2 <- teams[j]
+        strength1 <- strengths[i]
+        strength2 <- strengths[j]
+
+        # Simulate match between team i and team j using their strengths
+        match_winner <- simulate_match(team1, team2, strength1, strength2)
+
+        # Check which team won based on the returned team name
+        if (match_winner == team1) {
+          game_wins[i] <- game_wins[i] + 1
+          game_losses[j] <- game_losses[j] + 1
+        } else {
+          game_wins[j] <- game_wins[j] + 1
+          game_losses[i] <- game_losses[i] + 1
         }
@@
   }
-  }
-  results <- data.frame(
-    Team = teams,
-    Game_Wins = win_counts
+
+  # Create final dataframe with results
+  df <- data.frame(
+    team = teams,
+    true_strength = strengths,
+    game_wins = game_wins,
+    game_losses = game_losses,
+    distribution = distribution
   )
 
-  # Ranks teams based on number of wins (includes ties)
-  results$Rank_Hat <- rank(-results$Game_Wins, ties.method = "average")
-  results <- results[order(results$Rank_Hat),]
+  # Sort by true_strength to assign true rankings
+  df <- df %>% 
+    arrange(desc(true_strength)) %>% 
+    mutate(true_rank = row_number())
 
-  results <- results %>% 
-    inner_join(strengths, by = 'Team') %>% 
-    select(Team, True_Rank, True_Strength, Rank_Hat, Game_Wins)
+  # Rank teams based on number of wins
+  if (ties == TRUE) {
+    df$rank_hat <- rank(-df$game_wins, ties.method = "average")
+  } else {
+    df$rank_hat <- rank(-df$game_wins, ties.method = "random")
+  }
 
-  return(results)
+  # Select and order columns as specified
+  df <- df %>% 
+    select(true_rank, true_strength, rank_hat, game_wins, game_losses, distribution) %>%
+    arrange(true_rank)
+
+  return(df)
 }
+
+simulate_roundrobin_r(16,"Normal", r = 4)
 ```

### 2025-11-07 simulations.qmd

---
title: "Simulations"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(combinat)
library(ggplot2)
library(future)
library(future.apply)
library(progressr)

# Use all cores except one
cl <- parallel::makeCluster(parallel::detectCores() - 2)
plan(cluster, workers = cl)

teams <- 16
```

Note: The naming of each data frame has the following structure: tournament_distribution_ties_r_info

-   tournament is the tournament structure
-   distribution is the name of the distribution for $\theta$ (Normal, Uniform, Same, Manual)
    -   Same is equal strengths or coin flip

    -   Manual is an user created $\theta$ vector passed by adding theta_hat = *user input* at the end of the function
-   ties: ties or noTies
-   r is the number of games in a series or the number of iterations of round robin or group stage
-   info is any additional information like reseeded, third place game, or bad seeding structure

# Single Elimination

```{r}
single_elimination_param_grid <- expand.grid( 
  distribution = c("Normal","Uniform", "Same"), 
  ties = c(FALSE, TRUE), 
  series = c(1, 3, 7), 
  third_place = c(FALSE, TRUE), 
  reseeded = c(FALSE, TRUE), 
  bad_seeding_structure = c(FALSE, TRUE) 
  )

single_elimination_run_sim_block <- function(
  teams, distribution, ties, series, third_place, reseeded,
  bad_seeding_structure, p_block) {

  # Apply seeding structure rule
  seeding_structure <- if (bad_seeding_structure) 1:16 else NULL

  # Build the key
  parts <- c(
  "singleElimination",
  tolower(distribution),
  ifelse(ties, "ties", "noTies"),
  paste0(series),
  if (third_place) "thirdplace",
  if (reseeded) "reseeded",
  if (bad_seeding_structure) "badseed"
  )

  key <- paste(parts, collapse = "_")

  # Run the 10,000 tournament sims in parallel
  sim_list <- future_lapply(
    X = 1:10000,
    FUN = function(rep) {
      simulate_single_elimination_tournament(
        num_teams = teams,
        distribution = distribution,
        ties = ties,
        series = series,
        third_place = third_place,
        reseeded = reseeded,
        seeding_structure = seeding_structure
      )
    },
    future.seed = TRUE   # parallel-safe RNG
  )

  # Bind the simulation output
  final_df <- do.call(rbind, sim_list)
  final_df$teams <- teams

  assign(key, final_df, envir = .GlobalEnv)

  # Update outer progress bar
  p_block()

  message("Finished: ", key)
}

with_progress({
  p_block <- progressor(steps = nrow(single_elimination_param_grid))
  for (i in seq_len(nrow(single_elimination_param_grid))) {
    row <- single_elimination_param_grid[i, ]

    single_elimination_run_sim_block(
      teams        = teams,
      distribution = row$distribution,
      ties         = row$ties,
      series       = row$series,
      third_place  = row$third_place,
      reseeded     = row$reseeded,
      bad_seeding_structure = row$bad_seeding_structure,
      p_block      = p_block
    )
  }
})

```

# Double Elimination

```{r}
double_elimination_param_grid <- expand.grid( 
  distribution = c("Normal","Uniform", "Same"), 
  ties = c(FALSE, TRUE), 
  series = c(1, 3, 7), 
  final = c(FALSE, TRUE), 
  bad_seeding_structure = c(FALSE, TRUE) 
  )

double_elimination_run_sim_block <- function(
  teams, distribution, ties, series, final, bad_seeding_structure, p_block) {

  # Apply bad seeding structure override if required
  seeding_structure <- if (bad_seeding_structure) 1:16 else NULL

  # Build key
    # Build the key
  parts <- c(
    "doubleElimination",
    tolower(distribution),
    ifelse(ties, "ties", "noTies"),
    paste0(series),
    if (!final) "consolation",
    if (bad_seeding_structure) "badseed"
  )

  key <- paste(parts, collapse = "_")

  # Run 10,000 simulations in parallel
  sim_list <- future_lapply(
    X = 1:10000,
    FUN = function(rep) {
      simulate_double_elimination_tournament(
        num_teams = teams,
        distribution = distribution,
        ties = ties,
        series = series,
        final = final,
        seeding_structure = seeding_structure
      )
    },
    future.seed = TRUE    # ensures parallel RNG is valid
  )

  final_df <- do.call(rbind, sim_list)
  final_df$teams <- teams

  assign(key, final_df, envir = .GlobalEnv)

  # UPDATE BLOCK-LEVEL PROGRESS BAR
  p_block()

  message("Finished: ", key)
}

with_progress({
  p_block <- progressor(steps = nrow(double_elimination_param_grid))

  for (i in seq_len(nrow(double_elimination_param_grid))) {
    row <- double_elimination_param_grid[i, ]

    double_elimination_run_sim_block(
      teams        = teams,
      distribution = row$distribution,
      ties         = row$ties,
      series       = row$series,
      final        = row$final,
      bad_seeding_structure = row$bad_seeding_structure,
      p_block = p_block
    )
  }
})

```

# Round Robin

```{r}
round_robin_param_grid <- expand.grid( 
  distribution = c("Normal","Uniform", "Same"), 
  ties = c(FALSE, TRUE), 
  r = c(1,2,4)
  )

round_robin_run_sim_block <- function(
  teams, distribution, ties, r, p_block) {

  # Build key
    # Build the key
  parts <- c(
    "roundRobin",
    tolower(distribution),
    ifelse(ties, "ties", "noTies"),
    paste0(r)
  )

  key <- paste(parts, collapse = "_")

  # Run 10,000 simulations in parallel
  sim_list <- future_lapply(
    X = 1:10,
    FUN = function(rep) {
      simulate_roundrobin_r(
        num_teams = teams,
        distribution = distribution,
        ties = ties,
        r = r
      )
    },
    future.seed = TRUE    # ensures parallel RNG is valid
  )

  final_df <- do.call(rbind, sim_list)
  final_df$teams <- teams

  assign(key, final_df, envir = .GlobalEnv)

  # UPDATE BLOCK-LEVEL PROGRESS BAR
  p_block()

  message("Finished: ", key)
}

with_progress({
  p_block <- progressor(steps = nrow(round_robin_param_grid))

  for (i in seq_len(nrow(round_robin_param_grid))) {
    row <- round_robin_param_grid[i, ]

    round_robin_run_sim_block(
      teams        = teams,
      distribution = row$distribution,
      ties         = row$ties,
      r            = row$r,
      p_block = p_block
    )
  }
})
```

# Group Stage

```{r}
# Group stage is currently set so only the top 2 in each group advance
group_stage_param_grid <- expand.grid( 
  distribution = c("Normal","Uniform", "Same"), 
  ties = c(FALSE, TRUE), 
  rounds = c(1,2,4),
  group_length = c(4,8),
  third_place = c(FALSE, TRUE)
  )

group_stage_run_sim_block <- function(
  teams, distribution, ties, rounds, group_length, p_block) {

  # Build key
    # Build the key
  parts <- c(
    "groupStage",
    tolower(distribution),
    ifelse(ties, "ties", "noTies"),
    paste0(rounds),
    paste0(group_length, "groups"),
    if(third_place) "thirdplace"
  )

  key <- paste(parts, collapse = "_")

  # Run 10,000 simulations in parallel
  sim_list <- future_lapply(
    X = 1:10000,
    FUN = function(rep) {
      simulate_group_stage_tournament(
        num_teams = teams,
        distribution = distribution,
        ties = ties,
        rounds = rounds,
        group_length = group_length
      )
    },
    future.seed = TRUE    # ensures parallel RNG is valid
  )

  final_df <- do.call(rbind, sim_list)
  final_df$teams <- teams

  assign(key, final_df, envir = .GlobalEnv)

  # UPDATE BLOCK-LEVEL PROGRESS BAR
  p_block()

  message("Finished: ", key)
}

with_progress({
  p_block <- progressor(steps = nrow(group_stage_param_grid))

  for (i in seq_len(nrow(group_stage_param_grid))) {
    row <- group_stage_param_grid[i, ]

    group_stage_run_sim_block(
      teams        = teams,
      distribution = row$distribution,
      ties         = row$ties,
      rounds       = row$rounds,
      group_length = row$group_length,
      p_block = p_block
    )
  }
})

```

# Waterfall

# Step Ladder

# Staged Round Robin

# High School Tennis Ladder

### 2025-11-13 mutual_information_curves.qmd
Recommit changes without large RData file

@@
   # Handle missing weights (default to equal weights if not provided)
-  if (is.null(rank_weights)){
-    rank_weights <- rep(1,num_teams)
+  if (is.null(rank_weights)) {
+    rank_weights <- rep(1, num_teams)
   }
-  # Extract y values from sims, assuming `rank_hat` contains a list of rankings per simulation
+
+  # Extract rank_hat lists, widen into columns y1..yN
   y_df <- sims %>%
     group_by(simulation) %>%
-    summarise(rank_hat = list(rank_hat)) %>%  # Ensure rank_hat is a list column
-    mutate(rank_hat = map(rank_hat, unlist)) %>%  # Unlist rank_hat properly
-    unnest_wider(rank_hat, names_sep = "_")  # Expand into separate columns
+    summarise(rank_hat = list(rank_hat), .groups = "drop") %>%
+    mutate(rank_hat = map(rank_hat, unlist)) %>%
+    unnest_wider(rank_hat, names_sep = "_")
 
-  # Rename y columns
   colnames(y_df)[-1] <- paste0("y", 1:num_teams)
 
-  # Create x matrix, representing positions 1:num_teams
-  x_matrix <- matrix(rep(1:num_teams, times = nrow(y_df)), ncol = num_teams, byrow = TRUE)
-  x_df <- as.data.frame(x_matrix)
-  colnames(x_df) <- paste0("x", 1:num_teams)
-
-  # Combine x and y data frames
-  perms_df <- cbind(x_df, y_df %>% select(-simulation))
-  perms_df <- perms_df %>%
-    mutate(across(starts_with("x"), as.numeric)) %>%
-    mutate(across(starts_with("y"), as.numeric))
-
-  # Generate all possible permutations of rankings
-  all_perms <- as.data.frame(do.call(rbind, permn(1:num_teams)))
-  colnames(all_perms) <- paste0("y", 1:num_teams)
-
-  # Count occurrences of each observed ranking
-  prob_df <- perms_df %>%
-    group_by(across(starts_with("x")),across(starts_with("y")))  %>%
-    summarise(count = n(), .groups = "drop")
-
-  # Merge with full permutation set to ensure all permutations exist
-  prob_df <- right_join(all_perms, prob_df, by = colnames(all_perms)) %>%
-    mutate(count = replace_na(count, 0)) %>%
-    mutate(probability = count / sum(count))  # Normalize probabilities
+  # Probability of each observed permutation
+  prob_df <- y_df %>%
+    select(starts_with("y")) %>%
+    group_by(across(everything())) %>%
+    summarise(count = n(), .groups = "drop") %>%
+    mutate(probability = count / sum(count))
 
@@
 
-  for (i in 1:num_teams) {
+  #for (i in 1:num_teams) {
     #x_i <- paste0("x", i)
@@
     for (j in 1:nrow(prob_df)) {
-      x_vals <- as.numeric(prob_df[j, paste0("x", 1:num_teams)])
+      #x_vals <- as.numeric(prob_df[j, paste0("x", 1:num_teams)])
       y_vals <- as.numeric(prob_df[j, paste0("y", 1:num_teams)])
@@
     }
-  }
+  #}
   #w <- 1 / (w+1)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
@@
 
-  Raj_dist <- 1-numerator/denominator
+  Raj_dist <- numerator/denominator
   # Return the weighted data frame with mutual information weights
@@
 all_results_norm$simulation <- rep(1:10000, each = nrow(all_results_norm) / 10000)
-
-# Single Elimination with 7 game series
-seven_games_se_results$simulation <- rep(1:10000, each = nrow(seven_games_se_results) / 10000)
 
@@
 
+# Single Elimination with 7 game series
+seven_games_se_results$simulation <- rep(1:10000, each = nrow(seven_games_se_results) / 10000)
+
 # Function to compute mutual information for a given dataset
@@
 indices = 1:8,
-"Single Elimination (Reseeded)" = 1-mi_se_reseeded,
-"Group Stage" = 1-mi_group_stage,
-"Double Elimination" = 1-mi_de_norm,
-"Single Elimination (Set Structure)" = 1-mi_se_norm,
-"Coin Flip Single Elimination" = 1-mi_random,
+"Single Elimination (Reseeded)" = mi_se_reseeded,
+"Group Stage" = mi_group_stage,
+"Double Elimination" = mi_de_norm,
+"Single Elimination (Set Structure)" = mi_se_norm,
+"Coin Flip Single Elimination" = mi_random,
 #"Unif_SE" = mi_se_unif,
-"Bad Seeding Structure Single Elimination" = 1-mi_bad,
-"Coin Flip Round Robin" = 1-mi_perms,
-"Round Robin" = 1-mi_test_rr,
-"Worst Possible" = 1-mi_worst,
-"Best Possible" = 1-mi_best,
+"Bad Seeding Structure Single Elimination" = mi_bad,
+"Coin Flip Round Robin" = mi_perms,
+"Round Robin" = mi_test_rr,
+"Worst Possible" = mi_worst,
+"Best Possible" = mi_best,
 check.names = FALSE
@@
 
+## MI Calculation and Plotting
+
+These functions can take a list of dataframes and create mutual information that can then be plotted.
+
 ```{r}
-compute_mut_plots_rydog <- function(dataset) {
-  num_teams <- length(unique(dataset$true_rank))
-  results <- numeric(num_teams)
-  for (i in 1:num_teams) {
-    binary_vector <- c(rep(1, i), rep(0, num_teams - i))  # Construct vector
-    results[i] <- mut_info(dataset, num_teams, binary_vector, wtype = "rydog")  # Compute MI
-  }
-  return(results)
+compute_mi_for_datasets <- function(dataset_list) {
+  # Validate input
+  if (!is.list(dataset_list)) {
+    stop("dataset_list must be a list of data frames")
   }
 
-# Compute mutual information for each dataset
-mi_se_reseeded_rydog <- compute_mut_plots_rydog(reseeded_results_eight)
-mi_group_stage_rydog <- compute_mut_plots_rydog(group_stage_results_eight)
-mi_de_norm_rydog <- compute_mut_plots_rydog(de_norm_results_eight)
-mi_se_norm_rydog <- compute_mut_plots_rydog(all_results_norm)
-mi_test_rr_rydog <- compute_mut_plots_rydog(test_rr_noTies)
-mi_seven_se_rydog <- compute_mut_plots_rydog(seven_games_se_results)
-mi_random_rydog <- compute_mut_plots_rydog(equal_strengths)
-mi_se_unif_rydog <- compute_mut_plots_rydog(test_unif)
-mi_bad_rydog <- compute_mut_plots_rydog(bad_seeding_structure)
-mi_perms_rydog <- compute_mut_plots_rydog(perms_df)
-mi_best_rydog <- compute_mut_plots_rydog(best)
-mi_worst_rydog <- compute_mut_plots_rydog(worst)
+  # Get dataset names from the list
+  if (!is.null(names(dataset_list)) && all(names(dataset_list) != "")) {
+    dataset_names <- names(dataset_list)
+  } else {
+    dataset_names <- paste0("Dataset_", seq_along(dataset_list))
+  }
 
-mi_plots_df_rydog <- data.frame(
-indices = 1:8,
-"Single Elimination (Reseeded)" = 1-mi_se_reseeded_rydog,
-"Group Stage" = 1-mi_group_stage_rydog,
-"Double Elimination" = 1-mi_de_norm_rydog,
-"Single Elimination (Set Structure)" = 1-mi_se_norm_rydog,
-"Coin Flip Single Elimination" = 1-mi_random_rydog,
-#"Unif_SE" = mi_se_unif,
-"Bad Seeding Structure Single Elimination" = 1-mi_bad_rydog,
-"Coin Flip Round Robin" = 1-mi_perms_rydog,
-"Round Robin" = 1-mi_test_rr_rydog,
-"Worst Possible" = 1-mi_worst_rydog,
-"Best Possible" = 1-mi_best_rydog
-)
+  # Helper function to add simulation number if needed
+  add_simulation_number <- function(df) {
+    if (!"simulation" %in% colnames(df)) {
+      n_rows <- nrow(df)
+      n_sims <- 10000
+      rows_per_sim <- round(n_rows / n_sims)
 
-# Reshape data into long format for ggplot
-df_long_rydog <- pivot_longer(mi_plots_df_rydog, cols = -indices, names_to = "Method", values_to = "MI")
-windowsFonts(Arial = windowsFont("Arial"))
-ggplot(df_long_rydog, aes(x = indices, y = MI, color = Method, group = Method)) +
-geom_line(linewidth = 2.5) +           # Add lines
-geom_point(size=3) +          # Add points
+      # Memory efficient: create simulation numbers
+      df$simulation <- rep(1:n_sims, each = rows_per_sim)
+    }
+    return(df)
+  }
+
+  # Compute MI for all datasets and store as list column
+  mi_list <- lapply(seq_along(dataset_list), function(i) {
+    df <- dataset_list[[i]]
+    df <- add_simulation_number(df)
+
+    # Compute MI - returns a vector or dataframe
+    result <- compute_mut_plots(df)
+
+    # Convert to vector if it's a dataframe
+    if (is.data.frame(result)) {
+      result <- unlist(result)
+    }
+
+    return(result)
+  })
+
+  # Return dataframe with list column
+  data.frame(
+    dataset = dataset_names,
+    stringsAsFactors = FALSE
+  ) %>%
+    mutate(mi = mi_list)
+}
+
+# Example usage:
+mi_ex <- compute_mi_for_datasets(list(
+   singleElimination_normal_noTies_1 = singleElimination_normal_noTies_1,
+   singleElimination_normal_noTies_1_badseed = singleElimination_normal_noTies_1_badseed
+ ))
+
+plot_mi <- function(mi_results, dataset_names = NULL) {
+  # If no dataset names specified, plot all datasets
+  if (is.null(dataset_names)) {
+    dataset_names <- mi_results$dataset
+  }
+
+  # Filter to only requested datasets
+  filtered_results <- mi_results %>%
+    filter(dataset %in% dataset_names)
+
+  if (nrow(filtered_results) == 0) {
+    stop("No matching datasets found in results.")
+  }
+
+  # Convert list column to long format for plotting
+  plot_df <- filtered_results %>%
+    rowwise() %>%
+    mutate(
+      index = list(1:length(mi)),
+      mi_values = list(mi)
+    ) %>%
+    unnest(cols = c(index, mi_values)) %>%
+    ungroup()
+
+  # Create plot
+  ggplot(plot_df, aes(x = index, y = mi_values, color = dataset, group = dataset)) +
+    geom_line(linewidth = 1) +
+    geom_point(size = 2) +
     labs(
-title = "Comparison of Tournament Structures with Normal Strengths (Rydog)",
-x = "Top i Number of Teams",
-y = "Normalized Mutual Information \\ Entropy (log10 scale)"
+      title = "Mutual Information Comparison",
+      x = "Top K Teams",
+      y = "Mutual Information (MI)",
+      color = "Dataset"
     ) +
-theme_minimal() +       # Use a clean theme
-theme(text = element_text(family = "Arial", size = 32),
+    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+    scale_x_continuous(breaks = 1:max(plot_df$index)) +
+    theme_minimal() +
+    theme(
+      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+      axis.title = element_text(size = 12),
+      axis.text = element_text(size = 10),
       legend.position = "bottom",
-  legend.box = "vertical",
-  plot.title = element_text(hjust = 0.5, size = 36),
-  axis.title = element_text(size = 30),
-  axis.text = element_text(size=30)) +
-  guides(color = guide_legend(nrow = 5)) +
-  scale_y_log10(
-    labels = scales::label_number() 
+      legend.title = element_text(size = 11, face = "bold"),
+      legend.text = element_text(size = 9)
     ) +
-  scale_color_viridis_d(option = "H")
-ggsave(
-  filename = "tournament_comparison_plot_rydog.png",  # Change file type if needed
-  plot = last_plot(),  # Saves the last ggplot created
-  dpi = 600,        
-  bg = "white",
-  units = "in",
-  width = 20,
-  height = 13.43
-)
+    scale_color_brewer(palette = "Set1")
+}
 
-# Apply normalization to all columns except "indices"
-df_close_rydog <- df_long_rydog %>% filter(Method %in% c("Single.Elimination..Reseeded.", "Double.Elimination", "Group.Stage", "Single.Elimination..Set.Structure." ))
-windowsFonts(Arial = windowsFont("Arial"))
-ggplot(df_close_rydog, aes(x = indices, y = MI, color = Method, group = Method)) +
-geom_line(linewidth = 2.5) +           # Add lines
-geom_point(size=3) +          # Add points
-labs(
-title = "Comparison of Tournament Structures with Normal Strengths (Rydog)",
-x = "Top i Number of Teams",
-y = "Normalized Mutual Information \\ Entropy (log10 scale)"
-) +
-theme_minimal() +       # Use a clean theme
-theme(text = element_text(family = "Arial", size = 32),
-  legend.position = "bottom",
-  legend.box = "vertical",
-  plot.title = element_text(hjust = 0.5, size = 36),
-  axis.title = element_text(size = 30),
-  axis.text = element_text(size=30)) +
-  guides(color = guide_legend(nrow = 5)) +
-  scale_y_log10(
-    labels = scales::label_number() 
-  ) +
-  scale_color_viridis_d(option = "H")
-ggsave(
-  filename = "tournament_comparison_plot_close_rydog.png",  # Change file type if needed
-  plot = last_plot(),  # Saves the last ggplot created
-  dpi = 600,        
-  bg = "white",
-  units = "in",
-  width = 20,
-  height = 13.43
-)
+plot_mi(mi_ex)
 ```

### 2025-11-13 simulations.qmd
Recommit changes without large RData file

@@
 
-# Single Elimination
+# Getting Simulations
+
+## Single Elimination
 
@@
 
-# Double Elimination
+## Double Elimination
 
@@
 
-# Round Robin
+## Round Robin
 
@@
   sim_list <- future_lapply(
-    X = 1:10,
+    X = 1:10000,
     FUN = function(rep) {
@@
 
-# Group Stage
+## Group Stage
 
@@
   rounds = c(1,2,4),
-  group_length = c(4,8),
-  third_place = c(FALSE, TRUE)
+  group_length = c(4,8)
   )
@@
     paste0(rounds),
-    paste0(group_length, "groups"),
-    if(third_place) "thirdplace"
+    paste0(group_length, "groups")
   )
@@
 
-# Waterfall
+## Waterfall
 
-# Step Ladder
+## Step Ladder
 
-# Staged Round Robin
+```{r}
+stepLadder_param_grid <- expand.grid( 
+  distribution = c("Normal","Uniform", "Same"), 
+  series = c(1, 3, 7), 
+  bad_seeding_structure = c(FALSE, TRUE) 
+  )
 
-# High School Tennis Ladder
+stepLadder_run_sim_block <- function(
+  teams, distribution, bad_seeding_structure, series, p_block) {
+
+  # Apply bad seeding structure override if required
+  seeding_structure <- if (bad_seeding_structure) 1:16 else NULL
+
+  # Build key
+    # Build the key
+  parts <- c(
+    "stepLadder",
+    tolower(distribution),
+    paste0(series),
+    if (bad_seeding_structure) "badseed"
+
+  )
+
+  key <- paste(parts, collapse = "_")
+
+  # Run 10,000 simulations in parallel
+  sim_list <- future_lapply(
+    X = 1:10000,
+    FUN = function(rep) {
+      simulate_step_ladder(
+        num_teams = teams,
+        distribution = distribution,
+        series = series,
+        seeding_structure = seeding_structure
+      )
+    },
+    future.seed = TRUE    # ensures parallel RNG is valid
+  )
+
+  final_df <- do.call(rbind, sim_list)
+  final_df$teams <- teams
+
+  assign(key, final_df, envir = .GlobalEnv)
+
+  # UPDATE BLOCK-LEVEL PROGRESS BAR
+  p_block()
+
+  message("Finished: ", key)
+}
+
+with_progress({
+  p_block <- progressor(steps = nrow(stepLadder_param_grid))
+
+  for (i in seq_len(nrow(stepLadder_param_grid))) {
+    row <- stepLadder_param_grid[i, ]
+
+    stepLadder_run_sim_block(
+      teams        = teams,
+      distribution = row$distribution,
+      series       = row$series,
+      bad_seeding_structure = row$bad_seeding_structure,
+      p_block = p_block
+    )
+  }
+})
+```
+
+## Staged Round Robin
+
+## High School Tennis Ladder
+
+# Getting Mutual Information
+
+```{r}
+for (df in single_elimination_df){
+  MI_singleElimination <- compute_mi_for_datasets(df)
+}
+
+```

### 2025-11-13 staged_rr.R

@@ -1,0 +1,263 @@
+simulate_staged_rr <-  function(num_teams, distribution, group_length = 4, top_k = 2, stages = 2, ties = TRUE, rounds = 1, true_strength_hat = NULL) {
+
+  if (num_teams <= 3) {
+    stop("Number of Teams must be greater than 3.")
+  }
+
+  if (num_teams %% group_length != 0) {
+    stop("Number of teams must be divisible by group length.")
+  }
+
+  teams <- paste(num_teams:1)
+
+  # Create team strengths based on distribution
+  if (distribution == "Normal"){
+    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
+    team_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = strengths,
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- arrange(team_strengths, true_rank)
+  }
+  else if (distribution == "Same"){
+    strengths <- sapply(num_teams, function(n) {rep(0, n)})
+    team_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = strengths,
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- arrange(team_strengths, true_rank)
+  }
+  else if (distribution == "Uniform"){
+    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1), 0, sqrt(12)) })
+    team_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = strengths,
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams)),
+      rank_hat = rep(NA, length(teams))
+    )
+    df <- arrange(team_strengths, true_rank)
+  }
+  else if (distribution == "Manual") {
+    strengths <- true_strength_hat
+    manual_strengths <- data.frame(
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- manual_strengths %>% arrange(true_strength) %>%
+      mutate(true_rank = as.numeric(teams)) %>%
+      arrange(true_strength, true_rank)
+  }
+  else {
[…]

### 2025-11-13 stepladder.R

@@ -1,0 +1,151 @@
+simulate_step_ladder <- function(num_teams, distribution, seeding_structure = NULL,series = 1, theta_hat = NULL) {
+
+  teams <- paste(num_teams:1)
+
+  if (distribution == "Normal") {
+    strengths <- sapply(num_teams, function(n) {
+      qnorm(1:n/(n+1))
+    })
+    normal_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- arrange(normal_strengths, true_rank)
+  } else if (distribution == "Uniform") {
+    strengths <- sapply(num_teams, function(n) {
+      qunif(1:n/(n+1), 0, sqrt(12))
+    })
+    unif_strength <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams)),
+      rank_hat = rep(NA, length(teams))
+    )
+    df <- arrange(unif_strength, true_rank)
+  } else if (distribution == "Same"){
+    strengths <- sapply(num_teams, function(n) {rep(0,n)})
+    normal_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      rank_hat = rep(NA,length(teams)),
+      game_wins = rep(0,length(teams)),
+      game_losses = rep(0,length(teams))
+    )
+    df <- arrange(normal_strengths, true_rank)
+  } else if (distribution == "Manual") {
+    strengths <- theta_hat
+    manual_strengths <- data.frame(
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- manual_strengths %>%
+      arrange(true_strength) %>%
+      mutate(true_rank = as.numeric(teams)) %>%
+      arrange(true_strength, true_rank)
+  } else {
+    stop("Distribution not found: Enter Manual to input your own strengths")
+  }
+
[…]

### 2025-11-14 group_stage.qmd
.

@@
     manual_strengths <- data.frame(
-      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
-      rank_hat = rep(NA, length(teams)),
-      game_wins = rep(0, length(teams)),
-      game_losses = rep(0, length(teams))
+      theta = strengths,  # Also renamed to match other distributions
+      rank_hat = rep(NA, length(strengths)),
+      game_wins = rep(0, length(strengths)),
+      game_losses = rep(0, length(strengths))
     )
-    df <- manual_strengths %>% arrange(true_strength) %>%
-      mutate(true_rank = as.numeric(teams)) %>%
-      arrange(true_strength, true_rank)
+    df <- manual_strengths %>% 
+      arrange(desc(theta)) %>%
+      mutate(true_rank = row_number()) %>%
+      arrange(true_rank)
   }
@@
     group_df <- df[df$groups == g, ]
-    if (ties) {
-      group_df$group_rank <- rank(-group_df$game_wins, ties.method = "average")
-    } else {
+
     group_df$group_rank <- rank(-group_df$game_wins, ties.method = "random")
-    }
 
@@
 
-  if (ties) {
-    final_df$rank_hat <- rank(interaction(final_df$group_rank, -final_df$game_wins), ties.method = "average")
-  } else {
-    final_df$rank_hat <- rank(interaction(final_df$group_rank, -final_df$game_wins, final_df$true_rank), ties.method = "random")
-  }
-
   final_df <- final_df %>%
@@
     arrange(true_rank) %>%
-    mutate(rank_hat = ifelse(group_rank <= 2, group_rank, num_groups + group_rank)) %>%
+    mutate(rank_hat = ifelse(group_rank <= 2, group_rank, num_groups*group_rank)) %>%
     select(true_rank, theta, game_wins, game_losses, groups, group_rank, rank_hat, distribution)
+
+  #print(final_df)
 
@@
   knockout_results %>% filter(!is.na(rank_hat))  # knockout-stage teams
-) %>%
-  arrange(true_rank)
+) %>% arrange(true_rank)
   if (ties == F){

### 2025-11-14 mutual_information_curves.qmd
.

@@
 
-## MI Calculation and Plotting
-
-These functions can take a list of dataframes and create mutual information that can then be plotted.
-
-```{r}
-compute_mi_for_datasets <- function(dataset_list) {
-  # Validate input
-  if (!is.list(dataset_list)) {
-    stop("dataset_list must be a list of data frames")
-  }
-
-  # Get dataset names from the list
-  if (!is.null(names(dataset_list)) && all(names(dataset_list) != "")) {
-    dataset_names <- names(dataset_list)
-  } else {
-    dataset_names <- paste0("Dataset_", seq_along(dataset_list))
-  }
-
-  # Helper function to add simulation number if needed
-  add_simulation_number <- function(df) {
-    if (!"simulation" %in% colnames(df)) {
-      n_rows <- nrow(df)
-      n_sims <- 10000
-      rows_per_sim <- round(n_rows / n_sims)
-
-      # Memory efficient: create simulation numbers
-      df$simulation <- rep(1:n_sims, each = rows_per_sim)
-    }
-    return(df)
-  }
-
-  # Compute MI for all datasets and store as list column
-  mi_list <- lapply(seq_along(dataset_list), function(i) {
-    df <- dataset_list[[i]]
-    df <- add_simulation_number(df)
-
-    # Compute MI - returns a vector or dataframe
-    result <- compute_mut_plots(df)
-
-    # Convert to vector if it's a dataframe
-    if (is.data.frame(result)) {
-      result <- unlist(result)
-    }
-
-    return(result)
-  })
-
-  # Return dataframe with list column
-  data.frame(
-    dataset = dataset_names,
-    stringsAsFactors = FALSE
-  ) %>%
-    mutate(mi = mi_list)
-}
-
-# Example usage:
-mi_ex <- compute_mi_for_datasets(list(
-   singleElimination_normal_noTies_1 = singleElimination_normal_noTies_1,
-   singleElimination_normal_noTies_1_badseed = singleElimination_normal_noTies_1_badseed
- ))
-
-plot_mi <- function(mi_results, dataset_names = NULL) {
-  # If no dataset names specified, plot all datasets
-  if (is.null(dataset_names)) {
-    dataset_names <- mi_results$dataset
-  }
-
-  # Filter to only requested datasets
-  filtered_results <- mi_results %>%
-    filter(dataset %in% dataset_names)
-
-  if (nrow(filtered_results) == 0) {
-    stop("No matching datasets found in results.")
-  }
-
-  # Convert list column to long format for plotting
-  plot_df <- filtered_results %>%
-    rowwise() %>%
-    mutate(
-      index = list(1:length(mi)),
-      mi_values = list(mi)
-    ) %>%
-    unnest(cols = c(index, mi_values)) %>%
-    ungroup()
-
-  # Create plot
-  ggplot(plot_df, aes(x = index, y = mi_values, color = dataset, group = dataset)) +
-    geom_line(linewidth = 1) +
-    geom_point(size = 2) +
-    labs(
-      title = "Mutual Information Comparison",
-      x = "Top K Teams",
-      y = "Mutual Information (MI)",
-      color = "Dataset"
-    ) +
-    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
-    scale_x_continuous(breaks = 1:max(plot_df$index)) +
-    theme_minimal() +
-    theme(
-      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-      axis.title = element_text(size = 12),
-      axis.text = element_text(size = 10),
-      legend.position = "bottom",
-      legend.title = element_text(size = 11, face = "bold"),
-      legend.text = element_text(size = 9)
-    ) +
-    scale_color_brewer(palette = "Set1")
-}
-
-plot_mi(mi_ex)
-```
-
 ```{r}

### 2025-11-14 simulations.qmd
.

@@
 
+### Functions to get MI and put it in a Data Frame for Plots
+
 ```{r}
-for (df in single_elimination_df){
-  MI_singleElimination <- compute_mi_for_datasets(df)
+library(ggsci)
+compute_mi_for_datasets <- function(dataset_list) {
+  # Validate input
+  if (!is.list(dataset_list)) {
+    stop("dataset_list must be a list of data frames")
   }
 
+  # Get dataset names from the list
+  if (!is.null(names(dataset_list)) && all(names(dataset_list) != "")) {
+    dataset_names <- names(dataset_list)
+  } else {
+    dataset_names <- paste0("Dataset_", seq_along(dataset_list))
+  }
+
+  # Helper function to add simulation number if needed
+  add_simulation_number <- function(df) {
+    if (!"simulation" %in% colnames(df)) {
+      n_rows <- nrow(df)
+      n_sims <- 10000
+      rows_per_sim <- round(n_rows / n_sims)
+
+      # Memory efficient: create simulation numbers
+      df$simulation <- rep(1:n_sims, each = rows_per_sim)
+    }
+    return(df)
+  }
+
+  # Compute MI for all datasets and store as list column
+  mi_list <- lapply(seq_along(dataset_list), function(i) {
+    df <- dataset_list[[i]]
+    df <- add_simulation_number(df)
+
+    # Compute MI - returns a vector or dataframe
+    result <- compute_mut_plots(df)
+
+    # Convert to vector if it's a dataframe
+    if (is.data.frame(result)) {
+      result <- unlist(result)
+    }
+
+    return(result)
+  })
+
+  # Return dataframe with list column
+  data.frame(
+    dataset = dataset_names,
+    stringsAsFactors = FALSE
+  ) %>%
+    mutate(mi = mi_list)
+}
+
+# Example usage:
+mi_ex <- compute_mi_for_datasets(list(
+   singleElimination_normal_noTies_1,
+   singleElimination_normal_noTies_1_badseed
+ ))
+
+plot_mi <- function(mi_results, dataset_names = NULL) {
+  # If no dataset names specified, plot all datasets
+  if (is.null(dataset_names)) {
+    dataset_names <- mi_results$dataset
+  }
+
+  # Filter to only requested datasets
+  filtered_results <- mi_results %>%
+    filter(dataset %in% dataset_names)
+
+  if (nrow(filtered_results) == 0) {
+    stop("No matching datasets found in results.")
+  }
+
+  # Convert list column to long format for plotting
+  plot_df <- filtered_results %>%
+    rowwise() %>%
+    mutate(
+      index = list(1:length(mi)),
+      mi_values = list(mi)
+    ) %>%
+    unnest(cols = c(index, mi_values)) %>%
+    ungroup()
+
+  return(plot_df)
+}
 ```
+
+### Getting MI for All Structures
+
+```{r}
+compute_and_store_mi_by_tournament <- function(tournament_structure) {
+
+  # Validate input
+  valid_structures <- c("singleElimination", "doubleElimination", "roundRobin", "groupStage", "stepLadder")
+  if (!tournament_structure %in% valid_structures) {
+    stop(paste("Invalid tournament_structure. Must be one of:", 
+               paste(valid_structures, collapse = ", ")))
+  }
+
+  # Helper function to collect datasets for a specific tournament type
+  collect_tournament_datasets <- function(param_grid, tournament_type, key_builder) {
+    dataset_list <- list()
+
+    for (i in seq_len(nrow(param_grid))) {
+      row <- param_grid[i, ]
+      parts <- key_builder(row, tournament_type)
+      key <- paste(parts, collapse = "_")
+
+      if (exists(key, envir = .GlobalEnv)) {
+        dataset_list[[key]] <- get(key, envir = .GlobalEnv)
+      } else {
+        warning(paste("Dataset not found:", key))
+      }
+    }
+
+    return(dataset_list)
+  }
+
+  # Key builders for each tournament type
+  single_elim_key <- function(row, type) {
+    c(
+      type,
+      tolower(row$distribution),
+      ifelse(row$ties, "ties", "noTies"),
+      paste0(row$series),
+      if (row$third_place) "thirdplace",
+      if (row$reseeded) "reseeded",
+      if (row$bad_seeding_structure) "badseed"
+    )
+  }
+
+  double_elim_key <- function(row, type) {
+    c(
+      type,
+      tolower(row$distribution),
+      ifelse(row$ties, "ties", "noTies"),
+      paste0(row$series),
+      if (!row$final) "consolation",
+      if (row$bad_seeding_structure) "badseed"
+    )
+  }
+
+  round_robin_key <- function(row, type) {
+    c(
+      type,
+      tolower(row$distribution),
+      ifelse(row$ties, "ties", "noTies"),
+      paste0(row$r)
+    )
+  }
+
+  group_stage_key <- function(row, type) {
+    c(
+      type,
+      tolower(row$distribution),
+      ifelse(row$ties, "ties", "noTies"),
+      paste0(row$rounds),
+      paste0(row$group_length, "groups")
+    )
+  }
+
+   stepLadder_key <- function(row, type) {
+    c(
+    type,
+    tolower(row$distribution),
+    paste0(row$series),
+    if (row$bad_seeding_structure) "badseed"
+    )
+   }
+
+  # Function to process a single tournament type
+  process_tournament_type <- function(type_name, param_grid, key_builder) {
+    message(paste("Collecting", type_name, "datasets..."))
+    datasets <- collect_tournament_datasets(param_grid, type_name, key_builder)
+
+    message(paste("Computing MI for", type_name, "..."))
+    mi_results <- compute_mi_for_datasets(datasets)
+
+    # Store in global environment
+    mi_var_name <- paste0(type_name, "_mi")
+    assign(mi_var_name, mi_results, envir = .GlobalEnv)
+
+    # Create summary
+    summary_df <- data.frame(
+      dataset = names(mi_results),
+      mi = mi_results,
+      stringsAsFactors = FALSE
+    )
+
+    message(paste("- Created", mi_var_name, "with", length(mi_results), "datasets"))
+
+    return(list(mi = mi_results, summary = summary_df))
+  }
+
+  # Process based on tournament_structure input
+  results_list <- list()
+  summary_list <- list()
+
+  if (tournament_structure %in% c("singleElimination")) {
+    result <- process_tournament_type("singleElimination", 
+                                      single_elimination_param_grid, 
+                                      single_elim_key)
+    results_list$singleElimination <- result$mi
+    summary_list$singleElimination <- result$summary
+  }
+
+  if (tournament_structure %in% c("doubleElimination")) {
+    result <- process_tournament_type("doubleElimination", 
+                                      double_elimination_param_grid, 
+                                      double_elim_key)
+    results_list$doubleElimination <- result$mi
+    summary_list$doubleElimination <- result$summary
+  }
+
+  if (tournament_structure %in% c("roundRobin")) {
+    result <- process_tournament_type("roundRobin", 
+                                      round_robin_param_grid, 
+                                      round_robin_key)
+    results_list$roundRobin <- result$mi
+    summary_list$roundRobin <- result$summary
+  }
+
+  if (tournament_structure %in% c("groupStage")) {
+    result <- process_tournament_type("groupStage", 
+                                      group_stage_param_grid, 
+                                      group_stage_key)
+    results_list$groupStage <- result$mi
+    summary_list$groupStage <- result$summary
+  }
+   if (tournament_structure %in% c("stepLadder")) {
+    result <- process_tournament_type("stepLadder", 
+                                      stepLadder_param_grid, 
+                                      stepLadder_key)
+    results_list$stepLadder <- result$mi
+    summary_list$stepLadder <- result$summary
+  }
+
+  message("\n✓ MI computation complete!")
+
+  # Return a list of all results (optional)
+  invisible(results_list)
+}
+
+compute_and_store_mi_by_tournament("groupStage")
+
+all_mi <- bind_rows(
+  singleElimination_mi,
+  doubleElimination_mi,
+  groupStage_mi,
+  roundRobin_mi,
+  stepLadder_mi
+)
+```
+
+## MI Plots
+
+```{r}
+# Example Plot for Round Robin
+roundRobin_plot_df <- plot_mi(roundRobin_mi)
+
+ggplot(roundRobin_plot_df, aes(x = index, y = mi_values, color = dataset, group = dataset)) +
+    geom_line(linewidth = 1) +
+    geom_point(size = 2) +
+    labs(
+      title = "Mutual Information Round Robin Comparison",
+      x = "Top K Teams",
+      y = "Mutual Information (MI)",
+      color = "Dataset"
+    ) +
+    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+    scale_x_continuous(breaks = 1:max(roundRobin_plot_df$index)) +
+    theme_minimal() +
+    theme(
+      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+      axis.title = element_text(size = 12),
+      axis.text = element_text(size = 10),
+      legend.position = "bottom",
+      legend.title = element_text(size = 11, face = "bold"),
+      legend.text = element_text(size = 9)
+    ) +
+    scale_color_d3("category20")
+
+normal_plot_df <- plot_mi(all_mi, c("singleElimination_normal_noTies_1", "doubleElimination_normal_noTies_1", "roundRobin_normal_noTies_1", "groupStage_normal_noTies_1_4groups", "stepLadder_normal_1"))
+
+ggplot(normal_plot_df, aes(x = index, y = mi_values, color = dataset, group = dataset)) +
+    geom_line(linewidth = 1) +
+    geom_point(size = 2) +
+    labs(
+      title = "Mutual Information Normal Comparison",
+      x = "Top K Teams",
+      y = "Mutual Information (MI)",
+      color = "Dataset"
+    ) +
+    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+    scale_x_continuous(breaks = 1:max(normal_plot_df$index)) +
+    theme_minimal() +
+    theme(
+      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+      axis.title = element_text(size = 12),
+      axis.text = element_text(size = 10),
+      legend.position = "right",
+      legend.title = element_text(size = 11, face = "bold"),
+      legend.text = element_text(size = 9)
+    ) +
+    scale_color_d3("category20")
+
+unif_plot_df <- plot_mi(all_mi, c("singleElimination_uniform_noTies_1", "doubleElimination_uniform_noTies_1", "roundRobin_uniform_noTies_1", "groupStage_uniform_noTies_1_4groups", "stepLadder_uniform_1"))
+
+ggplot(unif_plot_df, aes(x = index, y = mi_values, color = dataset, group = dataset)) +
+    geom_line(linewidth = 1) +
+    geom_point(size = 2) +
+    labs(
+      title = "Mutual Information Uniform Comparison",
+      x = "Top K Teams",
+      y = "Mutual Information (MI)",
+      color = "Dataset"
+    ) +
+    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+    scale_x_continuous(breaks = 1:max(unif_plot_df$index)) +
+    theme_minimal() +
+    theme(
+      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+      axis.title = element_text(size = 12),
+      axis.text = element_text(size = 10),
+      legend.position = "right",
+      legend.title = element_text(size = 11, face = "bold"),
+      legend.text = element_text(size = 9)
+    ) +
+    scale_color_d3("category20")
+
+```

### 2025-11-21 double_elimination.qmd
updates

@@
     df <- arrange(normal_strengths, true_rank)
+  } else if (distribution == "Exponential"){
+      strengths <- sapply(num_teams, function(n) { qexp(1:n/(n+1)) })
+      unif_strength <- data.frame(
+        true_rank = as.numeric(teams),
+        true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+        game_wins = rep(0,length(teams)),
+        game_losses = rep(0,length(teams)),
+        rank_hat = rep(NA,length(teams))
+      )
+      df <- arrange(unif_strength, true_rank)
   } else if (distribution == "Manual") {

### 2025-11-21 group_stage.qmd
updates

@@
     df <- arrange(team_strengths, true_rank)
+  } 
+  else if (distribution == "Exponential"){
+      strengths <- sapply(num_teams, function(n) { qexp(1:n/(n+1)) })
+      unif_strength <- data.frame(
+        true_rank = as.numeric(teams),
+        true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+        game_wins = rep(0,length(teams)),
+        game_losses = rep(0,length(teams)),
+        rank_hat = rep(NA,length(teams))
+      )
+      df <- arrange(unif_strength, true_rank)
   }    

### 2025-11-21 round_robin.qmd
Saved Single Elimination and Double Elimination Simulations
.
updates

@@
     strengths <- qunif(1:num_teams / (num_teams + 1), 0, sqrt(12))
+  } else if (distribution == "Exponential") {
+    strengths <- qexp(1:num_teams / (num_teams + 1), 1)
   } else if (distribution == "Manual") {
@@
 
-simulate_roundrobin_r(16,"Normal", r = 4)
+simulate_roundrobin_r(16,"Normal", r = 1)
 ```

### 2025-11-21 simulations.qmd
updates

@@
 single_elimination_param_grid <- expand.grid( 
-  distribution = c("Normal","Uniform", "Same"), 
+  distribution = c("Normal","Uniform", "Same", "Exponential"), 
   ties = c(FALSE, TRUE), 
@@
 
-  # Run the 10,000 tournament sims in parallel
+  # Run the 100,000 tournament sims in parallel
   sim_list <- future_lapply(
-    X = 1:10000,
+    X = 1:100000,
     FUN = function(rep) {
@@
 double_elimination_param_grid <- expand.grid( 
-  distribution = c("Normal","Uniform", "Same"), 
+  distribution = c("Normal","Uniform", "Same", "Exponential"), 
   ties = c(FALSE, TRUE), 
@@
 
-  # Run 10,000 simulations in parallel
+  # Run 100,000 simulations in parallel
   sim_list <- future_lapply(
-    X = 1:10000,
+    X = 1:100000,
     FUN = function(rep) {
@@
 round_robin_param_grid <- expand.grid( 
-  distribution = c("Normal","Uniform", "Same"), 
+  distribution = c("Normal","Uniform", "Same", "Exponential"), 
   ties = c(FALSE, TRUE), 
@@
 
-  # Run 10,000 simulations in parallel
+  # Run 100,000 simulations in parallel
   sim_list <- future_lapply(
-    X = 1:10000,
+    X = 1:1000000,
     FUN = function(rep) {
@@
 group_stage_param_grid <- expand.grid( 
-  distribution = c("Normal","Uniform", "Same"), 
+  distribution = c("Normal","Uniform", "Same", "Exponential"), 
   ties = c(FALSE, TRUE), 
@@
 stepLadder_param_grid <- expand.grid( 
-  distribution = c("Normal","Uniform", "Same"), 
+  distribution = c("Normal","Uniform", "Same", "Exponential"), 
   series = c(1, 3, 7), 
@@
 # Example usage:
-mi_ex <- compute_mi_for_datasets(list(
-   singleElimination_normal_noTies_1,
-   singleElimination_normal_noTies_1_badseed
- ))
+#mi_ex <- compute_mi_for_datasets(list(
+#   singleElimination_normal_noTies_1,
+#   singleElimination_normal_noTies_1_badseed
+# ))
+
+compute_mi_combinations_for_datasets <- function(dataset_list) {
+  # Validate input
+  if (!is.list(dataset_list)) {
+    stop("dataset_list must be a list of data frames")
+  }
+
+  # Get dataset names from the list
+  if (!is.null(names(dataset_list)) && all(names(dataset_list) != "")) {
+    dataset_names <- names(dataset_list)
+  } else {
+    dataset_names <- paste0("Dataset_", seq_along(dataset_list))
+  }
+
+  # Helper function to add simulation number if needed
+  add_simulation_number <- function(df) {
+    if (!"simulation" %in% colnames(df)) {
+      n_rows <- nrow(df)
+      n_sims <- 10000
+      rows_per_sim <- round(n_rows / n_sims)
+
+      # Memory efficient: create simulation numbers
+      df$simulation <- rep(1:n_sims, each = rows_per_sim)
+    }
+    return(df)
+  }
+
+  # Compute MI for all datasets and store as list column
+  mi_list <- lapply(seq_along(dataset_list), function(i) {
+    df <- dataset_list[[i]]
+    df <- add_simulation_number(df)
+
+    # Compute MI - returns a vector or dataframe
+    result <- compute_mut_plots_combinations(df)
+
+    # Convert to vector if it's a dataframe
+    if (is.data.frame(result)) {
+      result <- unlist(result)
+    }
+
+    return(result)
+  })
+
+  # Return dataframe with list column
+  data.frame(
+    dataset = dataset_names,
+    stringsAsFactors = FALSE
+  ) %>%
+    mutate(mi = mi_list)
+}
 
@@
 
+    message(paste("Computing MI combinations for", type_name, "..."))
+    mi_combinations_results <- compute_mi_combinations_for_datasets(datasets)
+
     # Store in global environment
@@
     assign(mi_var_name, mi_results, envir = .GlobalEnv)
+
+    mi_combinations_var_name <- paste0(type_name, "_combinations_mi")
+    assign(mi_combinations_var_name, mi_combinations_results, envir = .GlobalEnv)
 
@@
       mi = mi_results,
+      mi_combinations = mi_combinations_results,
       stringsAsFactors = FALSE
@@
     message(paste("- Created", mi_var_name, "with", length(mi_results), "datasets"))
-
-    return(list(mi = mi_results, summary = summary_df))
+    message(paste("- Created", mi_combinations_var_name, "with", length(mi_combinations_results), "datasets"))
+
+    return(list(mi = mi_results, mi_combinations = mi_combinations_results, summary = summary_df))
   }
@@
                                       single_elim_key)
-    results_list$singleElimination <- result$mi
-    summary_list$singleElimination <- result$summary
+    results_list$singleElimination <- result
   }
@@
                                       double_elim_key)
-    results_list$doubleElimination <- result$mi
-    summary_list$doubleElimination <- result$summary
+    results_list$doubleElimination <- result
   }
@@
                                       round_robin_key)
-    results_list$roundRobin <- result$mi
-    summary_list$roundRobin <- result$summary
+    results_list$roundRobin <- result
   }
@@
                                       group_stage_key)
-    results_list$groupStage <- result$mi
-    summary_list$groupStage <- result$summary
+    results_list$groupStage <- result
   }
@@
                                       stepLadder_key)
-    results_list$stepLadder <- result$mi
-    summary_list$stepLadder <- result$summary
+    results_list$stepLadder <- result
   }
@@
 
+compute_and_store_mi_by_tournament("singleElimination")
+compute_and_store_mi_by_tournament("doubleElimination")
+compute_and_store_mi_by_tournament("roundRobin")
 compute_and_store_mi_by_tournament("groupStage")
+compute_and_store_mi_by_tournament("stepLadder")
 
@@
 
-## MI Plots
+# MI Plots
 
 ```{r}
-# Example Plot for Round Robin
-roundRobin_plot_df <- plot_mi(roundRobin_mi)
-
-ggplot(roundRobin_plot_df, aes(x = index, y = mi_values, color = dataset, group = dataset)) +
+library(ggsci)
+```
+
+## Comparison Plot
+
+```{r}
+combined_plot_df <- plot_mi(all_mi, c(
+  "singleElimination_normal_noTies_1",
+  "singleElimination_uniform_noTies_1", 
+  "singleElimination_same_noTies_1", 
+  #"singleElimination_exponential_noTies_1",
+  "doubleElimination_normal_noTies_1", 
+  "doubleElimination_uniform_noTies_1",
+  "doubleElimination_same_noTies_1", 
+  #"doubleElimination_exponential_noTies_1", 
+  "roundRobin_normal_noTies_1", 
+  "roundRobin_uniform_noTies_1", 
+  "roundRobin_same_noTies_1", 
+  #"roundRobin_exponential_noTies_1", 
+  "groupStage_normal_noTies_1_4groups",
+  "groupStage_uniform_noTies_1_4groups",
+  "groupStage_same_noTies_1_4groups"#,
+  #"groupStage_exponential_noTies_1_4groups"
+  )
+)
+
+parse_dataset_name_combined <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract tournament type (first part)
+  tournament_type <- parts[1]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  # default group length
+  group_length <- NA
+
+  if(tournament_type == "groupStage"){
+    group_length <- as.numeric(str_extract(parts[5], "\\d+"))
+    }
+
+  return(data.frame(
+    tournament_type = tournament_type,
+    distribution = distribution,
+    group_length = group_length
+  ))
+}
+
+combined_plot_df <- combined_plot_df %>%
+  select(-any_of(c("tournament_type", "distribution", "group_length"))) %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_dataset_name_combined(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    # Create a label for the additional parameters
+    param_label = paste(tournament_type, sep = "\n")
+  )
+
+combined_plot_df <- combined_plot_df %>%
+  mutate(
+    color_group = case_when(
+      tournament_type == "groupStage" ~ "groupStage (4 Groups)",
+      TRUE ~ tournament_type
+    )
+  )
+
+combined_plot <- ggplot(combined_plot_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
   geom_line(linewidth = 1) +
-    geom_point(size = 2) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    ~ distribution,
+    labeller = labeller(
+      distribution = function(x) x
+    )
+  ) +
   labs(
-      title = "Mutual Information Round Robin Comparison",
+    title = "Mutual Information (No Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 7),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 2))
+
+print(combined_plot)
+```
+
+## Single Elimination
+
+```{r}
+single_elimination_plot_df <- bind_rows(
+  plot_mi(singleElimination_mi),
+  plot_mi(stepLadder_mi)
+)
+
+# Parse dataset names to extract parameters
+parse_dataset_name_se <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract tournament type (first part)
+  tournament_type <- parts[1]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  if (tournament_type == "singleElimination") {
+    # Extract ties (third part)
+    ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+    # Extract series (fourth part - just the number)
+    series <- as.numeric(parts[4])
+
+    # Check for third_place (could be in position 5+)
+    third_place <- ifelse(any(parts == "thirdplace"), "Third Place", "No Third Place")
+
+    # Check for reseeded
+    reseeded <- ifelse(any(parts == "reseeded"), "Reseeded", "Not Reseeded")
+
+    # Check for badseed
+    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+  } else if (tournament_type == "stepLadder") {
+    # Step ladder doesn't have ties parameter
+    ties <- "No Ties"  # Default for step ladder
+
+    # Extract series (third part for step ladder)
+    series <- as.numeric(parts[3])
+
+    # Step ladder specific defaults
+    third_place <- "No Third Place"
+    reseeded <- "Not Reseeded"
+
+    # Check for badseed
+    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+  }
+
+  return(data.frame(
+    tournament_type = tournament_type,
+    distribution = distribution,
+    ties = ties,
+    series = series,
+    third_place = third_place,
+    reseeded = reseeded,
+    bad_seeding = bad_seeding
+  ))
+}
+
+# Add parameter columns to the plot data
+single_elimination_plot_df <- single_elimination_plot_df %>%
+  select(-any_of(c("tournament_type", "distribution", "ties", "series", "third_place", "reseeded", "bad_seeding", "param_label", "color_group"))) %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_dataset_name_se(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+   filter(third_place == "Third Place" | tournament_type == "stepLadder") %>% 
+  mutate(
+    series = factor(series, levels = c(1, 3, 7)),
+    # Create a label for the additional parameters
+    param_label = paste(tournament_type, reseeded, bad_seeding, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+single_elimination_plot_df <- single_elimination_plot_df %>%
+  mutate(
+    color_group = paste(tournament_type, reseeded, bad_seeding, sep = " | ")
+  )
+
+# Filter data for No Ties
+single_elim_noties_df <- single_elimination_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+single_elim_ties_df <- single_elimination_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties
+single_elimination_noties_plot <- ggplot(single_elim_noties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Series:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Single Elimination: Mutual Information (No Ties)",
     x = "Top K Teams",
     y = "Mutual Information (MI)",
-      color = "Dataset"
+    color = "Parameters"
   ) +
   scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
-    scale_x_continuous(breaks = 1:max(roundRobin_plot_df$index)) +
+  scale_x_continuous(breaks = 1:16) + 
   theme_minimal() +
   theme(
-      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-      axis.title = element_text(size = 12),
-      axis.text = element_text(size = 10),
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
     legend.position = "bottom", 
-      legend.title = element_text(size = 11, face = "bold"),
-      legend.text = element_text(size = 9)
-    ) +
-    scale_color_d3("category20")
-
-normal_plot_df <- plot_mi(all_mi, c("singleElimination_normal_noTies_1", "doubleElimination_normal_noTies_1", "roundRobin_normal_noTies_1", "groupStage_normal_noTies_1_4groups", "stepLadder_normal_1"))
-
-ggplot(normal_plot_df, aes(x = index, y = mi_values, color = dataset, group = dataset)) +
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 7),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 2))
+
+# Plot 2: Ties
+single_elimination_ties_plot <- ggplot(single_elim_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
   geom_line(linewidth = 1) +
-    geom_point(size = 2) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Series:", x),
+      distribution = function(x) x
+    )
+  ) +
   labs(
-      title = "Mutual Information Normal Comparison",
+    title = "Single Elimination: Mutual Information (Ties)",
     x = "Top K Teams",
     y = "Mutual Information (MI)",
-      color = "Dataset"
+    color = "Parameters"
   ) +
   scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
-    scale_x_continuous(breaks = 1:max(normal_plot_df$index)) +
+  scale_x_continuous(breaks = 1:16) + 
   theme_minimal() +
   theme(
-      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-      axis.title = element_text(size = 12),
-      axis.text = element_text(size = 10),
-      legend.position = "right",
-      legend.title = element_text(size = 11, face = "bold"),
-      legend.text = element_text(size = 9)
-    ) +
-    scale_color_d3("category20")
-
-unif_plot_df <- plot_mi(all_mi, c("singleElimination_uniform_noTies_1", "doubleElimination_uniform_noTies_1", "roundRobin_uniform_noTies_1", "groupStage_uniform_noTies_1_4groups", "stepLadder_uniform_1"))
-
-ggplot(unif_plot_df, aes(x = index, y = mi_values, color = dataset, group = dataset)) +
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 7),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 2))
+
+# Display plots
+print(single_elimination_noties_plot)
+print(single_elimination_ties_plot)
+```
+
+## Double Elimination
+
+```{r}
+double_elimination_plot_df <- plot_mi(doubleElimination_mi)
+
+# Parse dataset names to extract parameters
+parse_dataset_name <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  # Extract ties (third part)
+  ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+  # Extract series (fourth part - just the number)
+  series <- as.numeric(parts[4])
+
+  # Check for consolation (fifth part might be "consolation")
+  #final <- ifelse(length(parts) >= 5 && parts[5] == "consolation", "Consolation", "Final")
+
+  # Check for badseed (could be fifth or sixth part)
+  bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+
+  return(data.frame(
+    distribution = distribution,
+    ties = ties,
+    series = series,
+    #final = final,
+    bad_seeding = bad_seeding
+  ))
+}
+
+# Add parameter columns to the plot data
+double_elimination_plot_df <- double_elimination_plot_df %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_dataset_name(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    series = factor(series, levels = c(1, 3, 7)),
+    # Create a label for the additional parameters
+    param_label = paste(bad_seeding, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+double_elimination_plot_df <- double_elimination_plot_df %>%
+  mutate(
+    color_group = paste(bad_seeding, sep = " | ")
+  )
+
+# Filter data for No Ties
+double_elimination_plot_noTies_df <- double_elimination_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+double_elimination_plot_ties_df <- double_elimination_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties
+double_elimination_noties_plot <- ggplot(double_elimination_plot_noTies_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
   geom_line(linewidth = 1) +
-    geom_point(size = 2) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
   labs(
-      title = "Mutual Information Uniform Comparison",
+    title = "Double Elimination: Mutual Information (No Ties)",
     x = "Top K Teams",
     y = "Mutual Information (MI)",
-      color = "Dataset"
-    ) +
-    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
-    scale_x_continuous(breaks = 1:max(unif_plot_df$index)) +
+    color = "Seeding"
+  ) +
+  scale_y_continuous(limits = c(0.25, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
   theme_minimal() +
   theme(
-      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
-      axis.title = element_text(size = 12),
-      axis.text = element_text(size = 10),
-      legend.position = "right",
-      legend.title = element_text(size = 11, face = "bold"),
-      legend.text = element_text(size = 9)
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Plot 2: Ties
+double_elimination_ties_plot <- ggplot(double_elimination_plot_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Double Elimination: Mutual Information (Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Seeding"
+  ) +
+  scale_y_continuous(limits = c(0.25, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Display plots
+print(double_elimination_noties_plot)
+print(double_elimination_ties_plot)
+
+```
+
+## Round Robin
+
+```{r}
+roundRobin_plot_df <- plot_mi(roundRobin_mi)
+
+# Parse dataset names to extract parameters
+parse_dataset_name <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  # Extract ties (third part)
+  ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+  # Extract series (fourth part - just the number)
+  rounds <- as.numeric(parts[4])
+
+  return(data.frame(
+    distribution = distribution,
+    ties = ties,
+    rounds = rounds
+  ))
+}
+
+# Add parameter columns to the plot data
+roundRobin_plot_df <- roundRobin_plot_df %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_dataset_name(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    series = factor(rounds, levels = c(1, 2, 4)),
+  )
+
+# Filter data for No Ties
+roundRobin_plot_noties_df <- roundRobin_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+roundRobin_plot_ties_df <- roundRobin_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties
+roundRobin_noties_plot <- ggplot(roundRobin_plot_noties_df, aes(x = index, y = mi_values, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Round Robin: Mutual Information (No Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
   ) +
@@
 
+# Plot 2: Ties
+roundRobin_ties_plot <- ggplot(roundRobin_plot_ties_df, aes(x = index, y = mi_values, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Round Robin: Mutual Information (Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20")
+
+# Display plots
+print(roundRobin_noties_plot)
+print(roundRobin_ties_plot)
+
 ```
+
+## Group Stage
+
+```{r}
+group_stage_plot_df <- plot_mi(groupStage_mi)
+
+# Parse dataset names to extract parameters
+parse_dataset_name <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  # Extract ties (third part)
+  ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+  # Extract series (fourth part - just the number)
+  rounds <- as.numeric(parts[4])
+
+  # Check for group length (fifth part)
+  group_length <- as.numeric(str_extract(parts[5], "\\d+"))
+
+  return(data.frame(
+    distribution = distribution,
+    ties = ties,
+    rounds = rounds,
+    group_length = group_length
+  ))
+}
+
+# Add parameter columns to the plot data
+group_stage_plot_df <- group_stage_plot_df %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_dataset_name(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    series = factor(rounds, levels = c(1, 2, 4)),
+    # Create a label for the additional parameters
+    param_label = paste(group_length, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+group_stage_plot_df <- group_stage_plot_df %>%
+  mutate(
+    color_group = paste(group_length, sep = " | ")
+  )
+
+# Filter data for No Ties
+group_stage_noties_df <- group_stage_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+group_stage_ties_df <- group_stage_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties
+group_stage_noties_plot <- ggplot(group_stage_noties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Group Stage: Mutual Information (No Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Group Length:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Plot 2: Ties
+group_stage_ties_plot <- ggplot(group_stage_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Group Stage: Mutual Information (Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Group Length:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Display plots
+print(group_stage_noties_plot)
+print(group_stage_ties_plot)
+
+```

### 2025-11-21 single_elimination_shiny.qmd
updates

@@
     strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
+    unif_strength <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      game_wins = rep(0,length(teams)),
+      game_losses = rep(0,length(teams)),
+      rank_hat = rep(NA,length(teams))
+    )
+    df <- arrange(unif_strength, true_rank)
+  }
+  else if (distribution == "Exponential"){
+    strengths <- sapply(num_teams, function(n) { qexp(1:n/(n+1)) })
     unif_strength <- data.frame(

### 2025-11-21 staged_rr.R

@@ -44,6 +44,17 @@ simulate_staged_rr <-  function(num_teams, distribution, group_length = 4, top_k
     )
     df <- arrange(team_strengths, true_rank)
   }
+  else if (distribution == "Exponential"){
+    strengths <- sapply(num_teams, function(n) { qexp(1:n/(n+1)) })
+    unif_strength <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      game_wins = rep(0,length(teams)),
+      game_losses = rep(0,length(teams)),
+      rank_hat = rep(NA,length(teams))
+    )
+    df <- arrange(unif_strength, true_rank)
+  }
   else if (distribution == "Manual") {
     strengths <- true_strength_hat
     manual_strengths <- data.frame(
@@ -247,17 +258,18 @@ simulate_staged_rr <-  function(num_teams, distribution, group_length = 4, top_k
 
   combined_df <- rbind(top_results, bottom_results)
 
-
   if (ties == F){
-    final_df <- final_df %>%
-      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
+    final_df <- combined_df %>%
+      mutate(rank_hat = rank(rank_hat, ties.method = "random")) %>%
+      arrange(true_rank, rank_hat)
   } else{
-    final_df <- final_df %>%
-      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
+    final_df <- combined_df %>%
+      mutate(rank_hat = rank(rank_hat, ties.method = "average")) %>%
+      arrange(true_rank)
   }
 
 
-  return(combined_df)
+  return(final_df)
 }
 
-simulate_staged_rr(16, "Normal", rounds=10)
+simulate_staged_rr(16, "Normal", rounds=1)

### 2025-11-21 stepladder.R

@@ -36,6 +36,16 @@ simulate_step_ladder <- function(num_teams, distribution, seeding_structure = NU
       game_losses = rep(0,length(teams))
     )
     df <- arrange(normal_strengths, true_rank)
+    } else if (distribution == "Exponential"){
+      strengths <- sapply(num_teams, function(n) { qexp(1:n/(n+1)) })
+      unif_strength <- data.frame(
+        true_rank = as.numeric(teams),
+        true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+        game_wins = rep(0,length(teams)),
+        game_losses = rep(0,length(teams)),
+        rank_hat = rep(NA,length(teams))
+      )
+      df <- arrange(unif_strength, true_rank)
   } else if (distribution == "Manual") {
     strengths <- theta_hat
     manual_strengths <- data.frame(
@@ -148,4 +158,4 @@ simulate_step_ladder <- function(num_teams, distribution, seeding_structure = NU
 }
 
 
-simulate_step_ladder(4, "Normal", series = 3, seeding_structure = c(1,2,3,4))
+simulate_step_ladder(4, "Exponential", series = 3, seeding_structure = c(1,2,3,4))

### 2025-12-05 double_elimination.qmd
updates

@@
 
-test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, final = F)
-#test
-#manual_test <- simulate_double_elimination_tournament(num_teams = 4, distribution = "Manual", ties=F, series=1)
-#manual_test
-
-# Gets all results of a normal distribution num_teams between 4 and 64 single-game elimination
-de_norm_results <- list()
-for (teams in c(4,8,16,32,64)) {
-  de_norm_results[[as.character(teams)]] <- list()
-  for (replicate in 1:10000) {
-    # Store each simulation result in the list with unique indexing
-    de_norm_results[[as.character(teams)]][[replicate]] <- simulate_double_elimination_tournament(
-      num_teams = teams,
-      distribution = "Normal",
-      ties = FALSE,
-      series = 1
-    )
-  }
-}
-
-de_norm_results <- do.call(rbind, lapply(names(all_normal_results), function(teams) {
-  team_results <- do.call(rbind, all_normal_results[[teams]])
-  team_results$teams <- as.numeric(teams)
-  return(team_results)
-}))
-de_norm_results
-
 ```
-
-## Shiny User Interface Tournament Simulation
-
-```{r}
-ui <- fluidPage(
-  titlePanel("Double-Elimination Tournament Simulator"),
-
-  sidebarLayout(
-    sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", value = 8, min = 4, step = 1),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
-      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
-      checkboxInput("ties", "Allow Ties?", value = TRUE),
-      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
-      actionButton("run_sim", "Run Simulation")
-    ),
-
-    mainPanel(
-      h4("Tournament Results"),
-      tableOutput("ranked_results")
-    )
-  )
-)
-
-# Define the server logic
-server <- function(input, output) {
-  custom_seeding <- reactive({
-    if (input$seeding_input != "") {
-      # Convert the comma-separated string into a numeric vector
-      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
-      seeding_vector[seeding_vector == "NA"] <- NA
-      seeding_vector <- as.numeric(seeding_vector)
-      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
-      # Validate that it matches the number of teams
-      if (length(seeding_vector) == num_teams_next_power_of_two) {
-        return(seeding_vector)
-      } else {
-        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
-        return(NULL)
-      }
-    } else {
-      return(generate_elimination_tournament_structure(input$num_teams))
-    }
-  })
-
-  tournament_results <- eventReactive(input$run_sim, {
-    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
-
-    seed_structure <- custom_seeding()
-
-    simulate_double_elimination_tournament(
-      num_teams = input$num_teams,
-      distribution = input$distribution,
-      ties = input$ties,
-      series = input$series,
-      seeding_structure = seed_structure
-    )
-  })
-
-  output$ranked_results <- renderTable({
-    results <- tournament_results()
-    results[order(results$true_rank), ]  # Order by Strength descending
-  })
-}
-
-shinyApp(ui = ui, server = server)
-```
-
-## Shiny Number of Wins
-
-```{r}
-ui <- fluidPage(
-  titlePanel("Tournament Simulation with Total Wins Visualization"),
-
-  sidebarLayout(
-    sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
-      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
-      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
-      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
-      checkboxInput("ties", "Allow Ties?", value = TRUE),
-      actionButton("run_sim", "Run Simulation")
-    ),
-
-    mainPanel(
-      h4("Total Wins Summary"),
-      plotOutput("wins_plot")
-    )
-  )
-)
-
-# Shiny server logic
-server <- function(input, output) {
-  custom_seeding <- reactive({
-    if (input$seeding_input != "") {
-      # Convert the comma-separated string into a numeric vector
-      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
-      seeding_vector[seeding_vector == "NA"] <- NA
-      seeding_vector <- as.numeric(seeding_vector)
-      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
-      # Validate that it matches the number of teams
-      if (length(seeding_vector) == num_teams_next_power_of_two) {
-        return(seeding_vector)
-      } else {
-        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
-        return(NULL)
-      }
-    } else {
-      return(generate_elimination_tournament_structure(input$num_teams))
-    }
-  })
-
-  simulation_results <- eventReactive(input$run_sim, {
-    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
-    seed_structure <- custom_seeding()
-    # Run simulations
-    replicate_results <- replicate(
-      input$num_simulations,
-      simulate_double_elimination_tournament(
-        num_teams = input$num_teams,
-        distribution = input$distribution,
-        ties = input$ties,
-        series = input$series,
-        seeding_structure = seed_structure
-      ),
-      simplify = FALSE
-    )
-
-    # Combine all results and summarize total wins
-    all_results <- do.call(rbind, replicate_results)
-    total_wins_summary <- all_results %>%
-      group_by(true_rank) %>%
-      summarise(
-        total_wins = sum(rank_hat==1, na.rm = TRUE),
-        .groups = 'drop'
-      ) %>%
-      arrange(desc(total_wins))
-    return(total_wins_summary)
-  })
-
-  output$wins_plot <- renderPlot({
-    req(simulation_results())
-    results <- simulation_results()
-
-    ggplot(results, aes(x = reorder(true_rank, -total_wins), y = total_wins, fill = total_wins)) +
-      geom_text(aes(label = total_wins), vjust = -0.5, size = 4, color = "black") +
-      geom_bar(stat = "identity") +
-      labs(
-        title = "Total Wins Across Simulations",
-        x = "True Rank",
-        y = "Total Wins"
-      ) +
-      theme_minimal() +
-      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-      scale_fill_gradient(low = "skyblue", high = "blue")
-  })
-}
-
-# Run the Shiny app
-shinyApp(ui = ui, server = server)
-```
-
-# Multiple Simulations for 1 Seeding Structure
-
-```{r}
-library(shiny)
-
-ui <- fluidPage(
-  titlePanel("Tournament Simulation"),
-
-  sidebarLayout(
-    sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", 8, min = 4, step = 1),
-      selectInput("distribution", "Distribution Type:", choices = c("Normal", "Uniform")),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
-      numericInput("replicates", "Number of Replicates:", 100, min = 1, max = 1000, step = 1),
-      numericInput("series", "Number of Games in a Series (odd number):", 1, min = 1, max = 11, step = 2),
-      checkboxInput("ties", "Allow Ties?", value = TRUE),
-      actionButton("run_sim", "Run Tournament Simulation")
-    ),
-
-    mainPanel(
-      tableOutput("summary_results"),
-      verbatimTextOutput("kendall_correlation"),
-      verbatimTextOutput("spearman_correlation")
-    )
-  )
-)
-
-server <- function(input, output, session) {
- custom_seeding <- reactive({
-    if (input$seeding_input != "") {
-      # Convert the comma-separated string into a numeric vector
-      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
-      seeding_vector[seeding_vector == "NA"] <- NA
-      seeding_vector <- as.numeric(seeding_vector)
-      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
-      # Validate that it matches the number of teams
-      if (length(seeding_vector) == num_teams_next_power_of_two) {
-        return(seeding_vector)
-      } else {
-        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
-        return(NULL)
-      }
-    } else {
-      return(generate_elimination_tournament_structure(input$num_teams))
-    }
-  })
-
-  # Function to run tournament simulations
-  observeEvent(input$run_sim, {
-    req(input$num_teams, input$distribution, input$replicates, input$series)
-
-    # Initialize lists to store correlation results
-    kendall_correlations <- numeric(input$replicates)
-    spearman_correlations <- numeric(input$replicates)
-
-    seed_structure <- custom_seeding()
-
-    # Run tournament simulations for the specified number of replicates
-    simulation_results <- list()
-    for (replicate in 1:input$replicates) {
-      simulation_results[[replicate]] <- simulate_double_elimination_tournament(
-        num_teams = input$num_teams,
-        distribution = input$distribution,
-        ties = input$ties,
-        series = input$series,
-        seeding_structure = seed_structure
-      )
-
-      # Combine all results into one data frame
-      all_results <- do.call(rbind, simulation_results)
-
-      # Compute average ranks across all replicates
-      average_ranks <- all_results %>%
-        group_by(true_rank) %>%
-        summarise(avg_rank = mean(rank_hat, na.rm = TRUE), .groups = 'drop') %>%
-        arrange(true_rank)
-
-      # Calculate Kendall Tau correlation for the current replicate
-      kendall_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "kendall", use = "complete.obs")
-
-      # Calculate Spearman correlation for the current replicate
-      spearman_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "spearman", use = "complete.obs")
-    }
-
-    # Compute average Kendall Tau and Spearman correlations across all replicates
-    avg_kendall_cor <- mean(kendall_correlations, na.rm = TRUE)
-    avg_spearman_cor <- mean(spearman_correlations, na.rm = TRUE)
-
-    # Show summary results in table
-    output$summary_results <- renderTable({
-      average_ranks
-    })
-
-    # Display average Kendall Tau correlation
-    output$kendall_correlation <- renderPrint({
-      cat("Average Kendall Tau Correlation:", avg_kendall_cor, "\n")
-    })
-
-    # Display average Spearman correlation
-    output$spearman_correlation <- renderPrint({
-      cat("Average Spearman Correlation:", avg_spearman_cor, "\n")
-    })
-  })
-}
-
-shinyApp(ui, server)
-```
-
-## Shiny Probability of Top 3 Correct
-
-```{r}
-ui <- fluidPage(
-  titlePanel("Probability of Top 3 Teams in the Correct Order"),
-
-  sidebarLayout(
-    sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
-      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
-      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
-      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
-      checkboxInput("ties", "Allow Ties?", value = TRUE),
-      actionButton("run_sim", "Run Simulation")
-    ),
-
-    mainPanel(
-      h4("Top 3 Correct"),
-      verbatimTextOutput("top3_output")
-    )
-  )
-)
-
-# Shiny server logic
-server <- function(input, output) {
-  custom_seeding <- reactive({
-    if (input$seeding_input != "") {
-      # Convert the comma-separated string into a numeric vector
-      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
-      seeding_vector[seeding_vector == "NA"] <- NA
-      seeding_vector <- as.numeric(seeding_vector)
-      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
-      # Validate that it matches the number of teams
-      if (length(seeding_vector) == num_teams_next_power_of_two) {
-        return(seeding_vector)
-      } else {
-        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
-        return(NULL)
-      }
-    } else {
-      return(generate_elimination_tournament_structure(input$num_teams))
-    }
-  })
-
-  simulation_results <- eventReactive(input$run_sim, {
-    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
-    seed_structure <- custom_seeding()
-
-    # Run simulations
-    replicate_results <- replicate(
-      input$num_simulations,
-      simulate_double_elimination_tournament(
-        num_teams = input$num_teams,
-        distribution = input$distribution,
-        ties = input$ties,
-        series = input$series,
-        seeding_structure = seed_structure
-      ),
-      simplify = FALSE
-    )
-
-    # Combine all results into one dataframe and add a Simulation column
-    all_results <- do.call(rbind, replicate_results)
-    all_results$simulation <- rep(1:input$num_simulations, each = nrow(all_results) / input$num_simulations)
-
-    return(all_results)
-  })
-
-  # Identify top-3 ranks for each simulation and calculate the probability
-  top_three_probability <- reactive({
-    all_results <- simulation_results()
-
-    # Group by simulation and filter for top 3 true_rank and rank_hat
-    top_three_results <- all_results %>%
-      group_by(simulation) %>%
-      filter(true_rank %in% 1:3 & rank_hat %in% 1:3) %>%
-      arrange(simulation, rank_hat) %>%
-      summarize(top_3_true_rank = list(true_rank), top_3_rank = list(rank_hat), .groups = 'drop') %>%
-      mutate(match = mapply(function(true, rank) setequal(true, rank), top_3_true_rank, top_3_rank))
-
-    # Calculate probability of matching top 3 true_rank with rank_hat
-    match_count <- sum(top_three_results$match)
-    total_simulations <- n_distinct(all_results$simulation)
-
-    probability <- match_count / total_simulations
-    return(probability)
-  })
-
-  # Output the probability to the UI
-  output$top3_output <- renderText({
-    probability <- top_three_probability()
-
-    paste("Probability of Top 3 Seeds with Ranks 1-3: ", round(probability,4))
-  })
-}
-
-shinyApp(ui=ui,server=server)
-```
-
-## Shiny Conditional Probability Curve
-
-```{r}
-ui <- fluidPage(
-  titlePanel("Conditional Probability of Team Ranks in Tournament"),
-
-  sidebarLayout(
-    sidebarPanel(
-      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
-      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", 
-                placeholder = "e.g. 1 8 4 5 3 6 2 7"),
-      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
-      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
-      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
-      actionButton("run_sim", "Run Simulation")
-    ),
-
-    mainPanel(
-      plotOutput("probability_plot")
-    )
-  )
-)
-
-# Server
-server <- function(input, output) {
-
-  # Helper Function: Parse Custom Seeding Structure
-  parse_seeding <- function(seeding_input, num_teams) {
-    if (seeding_input != "") {
-      seeding_vector <- unlist(strsplit(seeding_input, " "))
-      seeding_vector[seeding_vector == "NA"] <- NA
-      seeding_vector <- as.numeric(seeding_vector)
-
-      num_teams_next_power_of_two <- 2^ceiling(log2(num_teams))
-      if (length(seeding_vector) == num_teams_next_power_of_two) {
-        return(seeding_vector)
-      } else {
-        showNotification(
-          paste("Seeding structure length must match", num_teams_next_power_of_two), 
-          type = "error"
-        )
-        return(NULL)
-      }
-    } else {
-      return(NULL)
-    }
-  }
-
-  # Curve Function to Calculate Conditional Probabilities
-  curve <- function(sim, num_teams) {
-    probability <- numeric(num_teams)
-    for (i in 1:num_teams) {
-      if (i == 1) {
-        probability[i] <- nrow(sim[sim$rank_hat == 1 & sim$true_rank == 1, ]) / length(unique(sim$simulation))
-      } else {
-        top_ranks <- sim[sim$rank_hat <= (i - 1) & sim$true_rank <= (i - 1), ]
-        better_teams <- 1:(i - 1)
-        numerator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
-          sim_subset <- sim[sim$simulation == sim_id, ]
-          if (all(better_teams %in% sim_subset$true_rank[sim_subset$rank_hat <= (i - 1)])) {
-            return(sum(sim_subset$rank_hat == i & sim_subset$true_rank == i))
-          } else {
-            return(0)
-          }
-        }))
-        denominator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
-          sim_subset <- top_ranks[top_ranks$simulation == sim_id, ]
-          all(better_teams %in% sim_subset$true_rank)
-        }))
-        probability[i] <- numerator / denominator
-      }
-    }
-    return(probability)
-  }
-
-  # Reactive Event: Run Simulations
-  simulation_data <- eventReactive(input$run_sim, {
-    num_teams <- input$num_teams
-    distribution <- input$distribution
-    series <- input$series
-    num_simulations <- input$num_simulations
-
-    # Parse custom seeding structure
-    seeding_structure <- parse_seeding(input$seeding_input, num_teams)
-
-    # Run simulations
-    replicate_results <- replicate(
-      num_simulations,
-      simulate_double_elimination_tournament(
-        num_teams = num_teams,
-        distribution = distribution,
-        series = series,
-        ties = FALSE,
-        seeding_structure = seeding_structure
-      ),
-      simplify = FALSE
-    )
-
-    # Combine all results into one dataframe
-    #all_results <- do.call(rbind, replicate_results)
-    #all_results$simulation <- rep(1:num_simulations, each = nrow(all_results) / num_simulations)
-    all_results <- list_rbind(x=replicate_results, names_to = "simulation")
-
-    # Calculate conditional probabilities
-    sim_probs <- curve(all_results, num_teams)
-    probabilities <- data.frame(x = 1:num_teams, y = sim_probs)
-    probabilities$line <- "Simulated"
-
-    # Baseline probabilities
-    baseline_prob <- 1 / (num_teams:1)
-    baseline <- data.frame(x = 1:num_teams, y = baseline_prob)
-    baseline$line <- "Baseline"
-
-    # Combine for plotting
-    rbind(baseline, probabilities)
-  })
-
-  # Plot Output
-  output$probability_plot <- renderPlot({
-    req(simulation_data())
-    ggplot(data = simulation_data(), aes(x = x, y = y, color = line, group = line)) +
-      geom_point() +
-      geom_line() +
-      ylim(0, 1) +
-      labs(
-        x = "Rank",
-        y = "Conditional Probability",
-        title = "Simulated vs Baseline Conditional Probabilities"
-      ) +
-      theme_minimal()
-  })
-}
-
-# Run the App
-shinyApp(ui = ui, server = server)
-```

### 2025-12-05 simulations.qmd
updates

@@
 
-  # Run the 100,000 tournament sims in parallel
+  # Run the 10,000 tournament sims in parallel
   sim_list <- future_lapply(
-    X = 1:100000,
+    X = 1:10000,
     FUN = function(rep) {
@@
 
-  # Run 100,000 simulations in parallel
+  # Run 10,000 simulations in parallel
   sim_list <- future_lapply(
-    X = 1:100000,
+    X = 1:10000,
     FUN = function(rep) {
@@
 
+### Repechege
+
+```{r}
+repechege_param_grid <- expand.grid( 
+  distribution = c("Normal","Uniform", "Same", "Exponential"), 
+  ties = c(FALSE, TRUE), 
+  series = c(1, 3, 7), 
+  bad_seeding_structure = c(FALSE, TRUE) 
+  )
+
+repechege_run_sim_block <- function(
+  teams, distribution, ties, series, bad_seeding_structure, p_block) {
+
+  # Apply bad seeding structure override if required
+  seeding_structure <- if (bad_seeding_structure) 1:16 else NULL
+
+  # Build key
+    # Build the key
+  parts <- c(
+    "Repechage",
+    tolower(distribution),
+    ifelse(ties, "ties", "noTies"),
+    paste0(series),
+    if (bad_seeding_structure) "badseed"
+  )
+
+  key <- paste(parts, collapse = "_")
+
+  # Run 10,000 simulations in parallel
+  sim_list <- future_lapply(
+    X = 1:10000,
+    FUN = function(rep) {
+      simulate_repechege_tournament(
+        num_teams = teams,
+        distribution = distribution,
+        ties = ties,
+        series = series,
+        seeding_structure = seeding_structure
+      )
+    },
+    future.seed = TRUE    # ensures parallel RNG is valid
+  )
+
+  final_df <- do.call(rbind, sim_list)
+  final_df$teams <- teams
+
+  assign(key, final_df, envir = .GlobalEnv)
+
+  # UPDATE BLOCK-LEVEL PROGRESS BAR
+  p_block()
+
+  message("Finished: ", key)
+}
+
+with_progress({
+  p_block <- progressor(steps = nrow(repechege_param_grid))
+
+  for (i in seq_len(nrow(repechege_param_grid))) {
+    row <- repechege_param_grid[i, ]
+
+    repechege_run_sim_block(
+      teams        = teams,
+      distribution = row$distribution,
+      ties         = row$ties,
+      series       = row$series,
+      bad_seeding_structure = row$bad_seeding_structure,
+      p_block = p_block
+    )
+  }
+})
+```
+
 ## Round Robin
@@
 
-  # Run 100,000 simulations in parallel
+  # Run 10,000 simulations in parallel
   sim_list <- future_lapply(
-    X = 1:1000000,
+    X = 1:5000,
     FUN = function(rep) {
@@
 }
+
+compute_mut_plots_combinations <- function(dataset) {
+
+  num_teams <- teams
+
+  results <- numeric(num_teams)
+
+  for (k in 1:num_teams) {
+    df <- dataset %>%
+      mutate(
+        true_rank = ifelse(true_rank <= k, 1, true_rank),
+        rank_hat = ifelse(rank_hat <= k, 1, rank_hat)
+      )
+    binary_vector <- c(rep(1, k), rep(0, num_teams - k))
+    results[k] <- mut_info(df, num_teams, binary_vector)
+  }
+  return(results)
+}
+
 ```
@@
   # Validate input
-  valid_structures <- c("singleElimination", "doubleElimination", "roundRobin", "groupStage", "stepLadder")
+  valid_structures <- c("singleElimination", "doubleElimination", "roundRobin", "groupStage", "stepLadder", "Repechage")
   if (!tournament_structure %in% valid_structures) {
@@
 
+  repechege_key <- function(row, type) {
+    c(
+      type,
+      tolower(row$distribution),
+      ifelse(row$ties, "ties", "noTies"),
+      paste0(row$series),
+      if (row$bad_seeding_structure) "badseed"
+    )
+  }
+
   round_robin_key <- function(row, type) {
@@
 
+   if (tournament_structure %in% c("Repechage")) {
+    result <- process_tournament_type("Repechage", 
+                                      repechege_param_grid, 
+                                      repechege_key)
+    results_list$Repechege <- result
+  }
+
   if (tournament_structure %in% c("roundRobin")) {
@@
 compute_and_store_mi_by_tournament("stepLadder")
+compute_and_store_mi_by_tournament("Repechage")
 
@@
   roundRobin_mi,
-  stepLadder_mi
+  stepLadder_mi,
+  Repechage_mi
+)
+
+all_mi_combinations <- bind_rows(
+  singleElimination_combinations_mi,
+  doubleElimination_combinations_mi,
+  groupStage_combinations_mi,
+  roundRobin_combinations_mi,
+  stepLadder_combinations_mi,
+  Repechage_combinations_mi
 )
@@
   "singleElimination_same_noTies_1", 
-  #"singleElimination_exponential_noTies_1",
+  "singleElimination_exponential_noTies_1",
   "doubleElimination_normal_noTies_1", 
@@
   "doubleElimination_same_noTies_1", 
-  #"doubleElimination_exponential_noTies_1", 
+  "doubleElimination_exponential_noTies_1", 
   "roundRobin_normal_noTies_1", 
@@
   "roundRobin_same_noTies_1", 
-  #"roundRobin_exponential_noTies_1", 
+  "roundRobin_exponential_noTies_1", 
   "groupStage_normal_noTies_1_4groups",
   "groupStage_uniform_noTies_1_4groups",
-  "groupStage_same_noTies_1_4groups"#,
-  #"groupStage_exponential_noTies_1_4groups"
+  "groupStage_same_noTies_1_4groups",
+  "groupStage_exponential_noTies_1_4groups"
   )
@@
 
-print(combined_plot)
-```
-
-## Single Elimination
-
-```{r}
-single_elimination_plot_df <- bind_rows(
-  plot_mi(singleElimination_mi),
-  plot_mi(stepLadder_mi)
-)
-
-# Parse dataset names to extract parameters
-parse_dataset_name_se <- function(dataset_name) {
-  parts <- str_split(dataset_name, "_")[[1]]
-
-  # Extract tournament type (first part)
-  tournament_type <- parts[1]
-
-  # Extract distribution (second part)
-  distribution <- parts[2]
-  distribution <- str_to_title(distribution)
-
-  if (tournament_type == "singleElimination") {
-    # Extract ties (third part)
-    ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
-
-    # Extract series (fourth part - just the number)
-    series <- as.numeric(parts[4])
-
-    # Check for third_place (could be in position 5+)
-    third_place <- ifelse(any(parts == "thirdplace"), "Third Place", "No Third Place")
-
-    # Check for reseeded
-    reseeded <- ifelse(any(parts == "reseeded"), "Reseeded", "Not Reseeded")
-
-    # Check for badseed
-    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
-  } else if (tournament_type == "stepLadder") {
-    # Step ladder doesn't have ties parameter
-    ties <- "No Ties"  # Default for step ladder
-
-    # Extract series (third part for step ladder)
-    series <- as.numeric(parts[3])
-
-    # Step ladder specific defaults
-    third_place <- "No Third Place"
-    reseeded <- "Not Reseeded"
-
-    # Check for badseed
-    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
-  }
-
-  return(data.frame(
-    tournament_type = tournament_type,
-    distribution = distribution,
-    ties = ties,
-    series = series,
-    third_place = third_place,
-    reseeded = reseeded,
-    bad_seeding = bad_seeding
-  ))
-}
-
-# Add parameter columns to the plot data
-single_elimination_plot_df <- single_elimination_plot_df %>%
-  select(-any_of(c("tournament_type", "distribution", "ties", "series", "third_place", "reseeded", "bad_seeding", "param_label", "color_group"))) %>%
+combined_combinations_df <- plot_mi(all_mi_combinations, c(
+  "singleElimination_normal_noTies_1",
+  "singleElimination_uniform_noTies_1", 
+  "singleElimination_same_noTies_1", 
+  "singleElimination_exponential_noTies_1",
+  "doubleElimination_normal_noTies_1", 
+  "doubleElimination_uniform_noTies_1",
+  "doubleElimination_same_noTies_1", 
+  "doubleElimination_exponential_noTies_1", 
+  "roundRobin_normal_noTies_1", 
+  "roundRobin_uniform_noTies_1", 
+  "roundRobin_same_noTies_1", 
+  "roundRobin_exponential_noTies_1", 
+  "groupStage_normal_noTies_1_4groups",
+  "groupStage_uniform_noTies_1_4groups",
+  "groupStage_same_noTies_1_4groups",
+  "groupStage_exponential_noTies_1_4groups"
+  )
+)
+
+combined_combinations_df <- combined_combinations_df %>%
+  select(-any_of(c("tournament_type", "distribution", "group_length"))) %>%
   group_by(dataset) %>%
   mutate(
-    params = list(parse_dataset_name_se(dataset[1]))
+    params = list(parse_dataset_name_combined(dataset[1]))
   ) %>%
@@
   ungroup() %>%
-   filter(third_place == "Third Place" | tournament_type == "stepLadder") %>% 
-  mutate(
-    series = factor(series, levels = c(1, 3, 7)),
+  mutate(
     # Create a label for the additional parameters
-    param_label = paste(tournament_type, reseeded, bad_seeding, sep = "\n")
-  )
-
-# Create a consistent color variable based on the parameter combinations
-single_elimination_plot_df <- single_elimination_plot_df %>%
-  mutate(
-    color_group = paste(tournament_type, reseeded, bad_seeding, sep = " | ")
-  )
-
-# Filter data for No Ties
-single_elim_noties_df <- single_elimination_plot_df %>%
-  filter(ties == "No Ties")
-
-# Filter data for Ties
-single_elim_ties_df <- single_elimination_plot_df %>%
-  filter(ties == "Ties")
-
-# Plot 1: No Ties
-single_elimination_noties_plot <- ggplot(single_elim_noties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+    param_label = paste(tournament_type, sep = "\n")
+  )
+
+combined_combinations_df <- combined_combinations_df %>%
+  mutate(
+    color_group = case_when(
+      tournament_type == "groupStage" ~ "groupStage (4 Groups)",
+      TRUE ~ tournament_type
+    )
+  )
+
+combined_combinations_plot <- ggplot(combined_combinations_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
   geom_line(linewidth = 1) +
@@
   facet_grid(
-    series ~ distribution,
+    ~ distribution,
     labeller = labeller(
-      series = function(x) paste("Series:", x),
       distribution = function(x) x
@@
   labs(
-    title = "Single Elimination: Mutual Information (No Ties)",
+    title = "Mutual Information (No Ties)",
     x = "Top K Teams",
-    y = "Mutual Information (MI)",
-    color = "Parameters"
+    y = "Mutual Information (MI)"
   ) +
@@
 
-# Plot 2: Ties
-single_elimination_ties_plot <- ggplot(single_elim_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+print(combined_plot)
+print(combined_combinations_plot)
+```
+
+## Single Elimination
+
+```{r}
+single_elimination_plot_df <- bind_rows(
+  plot_mi(singleElimination_mi),
+  plot_mi(stepLadder_mi)
+)
+
+# Parse dataset names to extract parameters
+parse_dataset_name_se <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract tournament type (first part)
+  tournament_type <- parts[1]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  if (tournament_type == "singleElimination") {
+    # Extract ties (third part)
+    ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+    # Extract series (fourth part - just the number)
+    series <- as.numeric(parts[4])
+
+    # Check for third_place (could be in position 5+)
+    third_place <- ifelse(any(parts == "thirdplace"), "Third Place", "No Third Place")
+
+    # Check for reseeded
+    reseeded <- ifelse(any(parts == "reseeded"), "Reseeded", "Not Reseeded")
+
+    # Check for badseed
+    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+  } else if (tournament_type == "stepLadder") {
+    # Step ladder doesn't have ties parameter
+    ties <- "No Ties"  # Default for step ladder
+
+    # Extract series (third part for step ladder)
+    series <- as.numeric(parts[3])
+
+    # Step ladder specific defaults
+    third_place <- "No Third Place"
+    reseeded <- "Not Reseeded"
+
+    # Check for badseed
+    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+  }
+
+  return(data.frame(
+    tournament_type = tournament_type,
+    distribution = distribution,
+    ties = ties,
+    series = series,
+    third_place = third_place,
+    reseeded = reseeded,
+    bad_seeding = bad_seeding
+  ))
+}
+
+# Add parameter columns to the plot data
+single_elimination_plot_df <- single_elimination_plot_df %>%
+  select(-any_of(c("tournament_type", "distribution", "ties", "series", "third_place", "reseeded", "bad_seeding", "param_label", "color_group"))) %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_dataset_name_se(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+   filter(third_place == "Third Place" | tournament_type == "stepLadder") %>% 
+  mutate(
+    series = factor(series, levels = c(1, 3, 7)),
+    # Create a label for the additional parameters
+    param_label = paste(tournament_type, reseeded, bad_seeding, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+single_elimination_plot_df <- single_elimination_plot_df %>%
+  mutate(
+    color_group = paste(tournament_type, reseeded, bad_seeding, sep = " | ")
+  )
+
+# Filter data for No Ties
+single_elim_noties_df <- single_elimination_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+single_elim_ties_df <- single_elimination_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties
+single_elimination_noties_plot <- ggplot(single_elim_noties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
   geom_line(linewidth = 1) +
@@
   labs(
-    title = "Single Elimination: Mutual Information (Ties)",
+    title = "Single Elimination: Mutual Information (No Ties)",
     x = "Top K Teams",
@@
 
+# Plot 2: Ties
+single_elimination_ties_plot <- ggplot(single_elim_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Series:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Single Elimination: Mutual Information (Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Parameters"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 7),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 2))
+
+# Combine the combination datasets
+single_elimination_combinations_plot_df <- bind_rows(
+  plot_mi(singleElimination_combinations_mi),
+  plot_mi(stepLadder_combinations_mi)
+)
+
+# Parse dataset names to extract parameters
+parse_dataset_name_se <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract tournament type (first part)
+  tournament_type <- parts[1]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  if (tournament_type == "singleElimination") {
+    # Extract ties (third part)
+    ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+    # Extract series (fourth part - just the number)
+    series <- as.numeric(parts[4])
+
+    # Check for third_place (could be in position 5+)
+    third_place <- ifelse(any(parts == "thirdplace"), "Third Place", "No Third Place")
+
+    # Check for reseeded
+    reseeded <- ifelse(any(parts == "reseeded"), "Reseeded", "Not Reseeded")
+
+    # Check for badseed
+    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+  } else if (tournament_type == "stepLadder") {
+    # Step ladder doesn't have ties parameter
+    ties <- "No Ties"  # Default for step ladder
+
+    # Extract series (third part for step ladder)
+    series <- as.numeric(parts[3])
+
+    # Step ladder specific defaults
+    third_place <- "No Third Place"
+    reseeded <- "Not Reseeded"
+
+    # Check for badseed
+    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+  }
+
+  return(data.frame(
+    tournament_type = tournament_type,
+    distribution = distribution,
+    ties = ties,
+    series = series,
+    third_place = third_place,
+    reseeded = reseeded,
+    bad_seeding = bad_seeding
+  ))
+}
+
+# Add parameter columns to the plot data
+single_elimination_combinations_plot_df <- single_elimination_combinations_plot_df %>%
+  select(-any_of(c("tournament_type", "distribution", "ties", "series", "third_place", "reseeded", "bad_seeding", "param_label", "color_group"))) %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_dataset_name_se(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  filter(third_place == "Third Place" | tournament_type == "stepLadder") %>% 
+  mutate(
+    series = factor(series, levels = c(1, 3, 7)),
+    # Create a label for the additional parameters
+    param_label = paste(tournament_type, reseeded, bad_seeding, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+single_elimination_combinations_plot_df <- single_elimination_combinations_plot_df %>%
+  mutate(
+    color_group = paste(tournament_type, reseeded, bad_seeding, sep = " | ")
+  )
+
+# Filter data for No Ties
+single_elim_combinations_noties_df <- single_elimination_combinations_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+single_elim_combinations_ties_df <- single_elimination_combinations_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties (Combinations)
+single_elimination_combinations_noties_plot <- ggplot(single_elim_combinations_noties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Series:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Single Elimination (Combinations): Mutual Information (No Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Parameters"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 7),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 2))
+
+# Plot 2: Ties (Combinations)
+single_elimination_combinations_ties_plot <- ggplot(single_elim_combinations_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Series:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Single Elimination (Combinations): Mutual Information (Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Parameters"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 7),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 2))
+
 # Display plots
@@
 print(single_elimination_ties_plot)
+
+# Display plots
+print(single_elimination_combinations_noties_plot)
+print(single_elimination_combinations_ties_plot)
+
 ```
@@
 ```{r}
-double_elimination_plot_df <- plot_mi(doubleElimination_mi)
+double_elimination_plot_df <- bind_rows(
+  plot_mi(doubleElimination_mi),
+  plot_mi(Repechage_mi)
+)
 
@@
 
+  # Extract tournament type (first part)
+  tournament_type <- parts[1]
+
   # Extract distribution (second part)
@@
 
+  if (tournament_type == "doubleElimination"){
     # Extract ties (third part)
@@
     bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+  } else if (tournament_type == "Repechage"){
+    # Extract ties (third part)
+    ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+    # Extract series (fourth part - just the number)
+    series <- as.numeric(parts[4])
+
+    # Check for badseed (could be fifth or sixth part)
+    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+  }
 
   return(data.frame(
+    tournament_type = tournament_type,
     distribution = distribution,
@@
 double_elimination_plot_df <- double_elimination_plot_df %>%
+  select(-any_of(c("tournament_type", "distribution", "ties", "series", "bad_seeding", "param_label", "color_group"))) %>%
   group_by(dataset) %>%
@@
     # Create a label for the additional parameters
-    param_label = paste(bad_seeding, sep = "\n")
+    param_label = paste(tournament_type, bad_seeding, sep = "\n")
   )
@@
    mutate(
-    color_group = paste(bad_seeding, sep = " | ")
+    color_group = paste(tournament_type, bad_seeding, sep = " | ")
   )
@@
 
+double_elimination_combinations_plot_df <- bind_rows(
+  plot_mi(doubleElimination_combinations_mi),
+  plot_mi(Repechage_combinations_mi)
+)
+
+# Parse dataset names to extract parameters
+parse_dataset_name <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract tournament type (first part)
+  tournament_type <- parts[1]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  if (tournament_type == "doubleElimination"){
+    # Extract ties (third part)
+    ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+    # Extract series (fourth part - just the number)
+    series <- as.numeric(parts[4])
+
+    # Check for consolation (fifth part might be "consolation")
+    #final <- ifelse(length(parts) >= 5 && parts[5] == "consolation", "Consolation", "Final")
+
+    # Check for badseed (could be fifth or sixth part)
+    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+  } else if (tournament_type == "Repechage"){
+    # Extract ties (third part)
+    ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+    # Extract series (fourth part - just the number)
+    series <- as.numeric(parts[4])
+
+    # Check for badseed (could be fifth or sixth part)
+    bad_seeding <- ifelse(any(parts == "badseed"), "Bad Seed", "Good Seed")
+  }
+
+  return(data.frame(
+    tournament_type = tournament_type,
+    distribution = distribution,
+    ties = ties,
+    series = series,
+    #final = final,
+    bad_seeding = bad_seeding
+  ))
+}
+
+# Add parameter columns to the plot data
+double_elimination_combinations_plot_df <- double_elimination_combinations_plot_df %>%
+  select(-any_of(c("tournament_type", "distribution", "ties", "series", "bad_seeding", "param_label", "color_group"))) %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_dataset_name(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    series = factor(series, levels = c(1, 3, 7)),
+    # Create a label for the additional parameters
+    param_label = paste(tournament_type, bad_seeding, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+double_elimination_combinations_plot_df <- double_elimination_combinations_plot_df %>%
+   mutate(
+    color_group = paste(tournament_type, bad_seeding, sep = " | ")
+  )
+
+# Filter data for No Ties
+double_elimination_combinations_plot_noTies_df <- double_elimination_combinations_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+double_elimination_combinations_plot_ties_df <- double_elimination_combinations_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties (Combinations)
+double_elimination_combinations_noties_plot <- ggplot(double_elimination_combinations_plot_noTies_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Double Elimination (Combinations): Mutual Information (No Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Seeding"
+  ) +
+  scale_y_continuous(limits = c(0.25, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Plot 2: Ties (Combinations)
+double_elimination_combinations_ties_plot <- ggplot(double_elimination_combinations_plot_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Double Elimination (Combinations): Mutual Information (Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Seeding"
+  ) +
+  scale_y_continuous(limits = c(0.25, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
 # Display plots
@@
 print(double_elimination_ties_plot)
+
+# Display plots
+print(double_elimination_combinations_noties_plot)
+print(double_elimination_combinations_ties_plot)
 
@@
 
-# Display plots
-print(roundRobin_noties_plot)
-print(roundRobin_ties_plot)
-
-```
-
-## Group Stage
-
-```{r}
-group_stage_plot_df <- plot_mi(groupStage_mi)
+roundRobin_combinations_plot_df <- plot_mi(roundRobin_combinations_mi)
 
@@
 
-  # Check for group length (fifth part)
-  group_length <- as.numeric(str_extract(parts[5], "\\d+"))
-
   return(data.frame(
@@
     ties = ties,
-    rounds = rounds,
-    group_length = group_length
+    rounds = rounds
   ))
@@
 # Add parameter columns to the plot data
-group_stage_plot_df <- group_stage_plot_df %>%
+roundRobin_combinations_plot_df <- roundRobin_combinations_plot_df %>%
   group_by(dataset) %>%
@@
     series = factor(rounds, levels = c(1, 2, 4)),
-    # Create a label for the additional parameters
-    param_label = paste(group_length, sep = "\n")
-  )
-
-# Create a consistent color variable based on the parameter combinations
-group_stage_plot_df <- group_stage_plot_df %>%
-  mutate(
-    color_group = paste(group_length, sep = " | ")
   )
@@
 # Filter data for No Ties
-group_stage_noties_df <- group_stage_plot_df %>%
+roundRobin_combinations_plot_noties_df <- roundRobin_combinations_plot_df %>%
   filter(ties == "No Ties")
@@
 # Filter data for Ties
-group_stage_ties_df <- group_stage_plot_df %>%
+roundRobin_combinations_plot_ties_df <- roundRobin_combinations_plot_df %>%
   filter(ties == "Ties")
 
-# Plot 1: No Ties
-group_stage_noties_plot <- ggplot(group_stage_noties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+# Plot 1: No Ties (Combinations)
+roundRobin_combinations_noties_plot <- ggplot(roundRobin_combinations_plot_noties_df, aes(x = index, y = mi_values, group = dataset)) +
   geom_line(linewidth = 1) +
@@
   labs(
-    title = "Group Stage: Mutual Information (No Ties)",
+    title = "Round Robin (Combinations): Mutual Information (No Ties)",
     x = "Top K Teams",
-    y = "Mutual Information (MI)",
-    color = "Group Length:"
+    y = "Mutual Information (MI)"
   ) +
@@
   ) +
-  scale_color_d3("category20") +
-  guides(color = guide_legend(nrow = 1))
-
-# Plot 2: Ties
-group_stage_ties_plot <- ggplot(group_stage_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  scale_color_d3("category20")
+
+# Plot 2: Ties (Combinations)
+roundRobin_combinations_ties_plot <- ggplot(roundRobin_combinations_plot_ties_df, aes(x = index, y = mi_values, group = dataset)) +
   geom_line(linewidth = 1) +
@@
   labs(
-    title = "Group Stage: Mutual Information (Ties)",
+    title = "Round Robin (Combinations): Mutual Information (Ties)",
     x = "Top K Teams",
-    y = "Mutual Information (MI)",
-    color = "Group Length:"
+    y = "Mutual Information (MI)"
   ) +
@@
   ) +
+  scale_color_d3("category20")
+
+# Display plots
+print(roundRobin_noties_plot)
+print(roundRobin_ties_plot)
+
+# Display plots
+print(roundRobin_combinations_noties_plot)
+print(roundRobin_combinations_ties_plot)
+
+```
+
+## Group Stage
+
+```{r}
+group_stage_plot_df <- plot_mi(groupStage_mi)
+
+# Parse dataset names to extract parameters
+parse_dataset_name <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  # Extract ties (third part)
+  ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+  # Extract series (fourth part - just the number)
+  rounds <- as.numeric(parts[4])
+
+  # Check for group length (fifth part)
+  group_length <- as.numeric(str_extract(parts[5], "\\d+"))
+
+  return(data.frame(
+    distribution = distribution,
+    ties = ties,
+    rounds = rounds,
+    group_length = group_length
+  ))
+}
+
+# Add parameter columns to the plot data
+group_stage_plot_df <- group_stage_plot_df %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_dataset_name(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    series = factor(rounds, levels = c(1, 2, 4)),
+    # Create a label for the additional parameters
+    param_label = paste(group_length, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+group_stage_plot_df <- group_stage_plot_df %>%
+  mutate(
+    color_group = paste(group_length, sep = " | ")
+  )
+
+# Filter data for No Ties
+group_stage_noties_df <- group_stage_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+group_stage_ties_df <- group_stage_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties
+group_stage_noties_plot <- ggplot(group_stage_noties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Group Stage: Mutual Information (No Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Group Length:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Plot 2: Ties
+group_stage_ties_plot <- ggplot(group_stage_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Group Stage: Mutual Information (Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Group Length:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+group_stage_combinations_plot_df <- plot_mi(groupStage_combinations_mi)
+
+# Parse dataset names to extract parameters
+parse_dataset_name <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  # Extract ties (third part)
+  ties <- ifelse(parts[3] == "ties", "Ties", "No Ties")
+
+  # Extract series (fourth part - just the number)
+  rounds <- as.numeric(parts[4])
+
+  # Check for group length (fifth part)
+  group_length <- as.numeric(str_extract(parts[5], "\\d+"))
+
+  return(data.frame(
+    distribution = distribution,
+    ties = ties,
+    rounds = rounds,
+    group_length = group_length
+  ))
+}
+
+# Add parameter columns to the plot data
+group_stage_combinations_plot_df <- group_stage_combinations_plot_df %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_dataset_name(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    series = factor(rounds, levels = c(1, 2, 4)),
+    # Create a label for the additional parameters
+    param_label = paste(group_length, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+group_stage_combinations_plot_df <- group_stage_combinations_plot_df %>%
+  mutate(
+    color_group = paste(group_length, sep = " | ")
+  )
+
+# Filter data for No Ties
+group_stage_combinations_noties_df <- group_stage_combinations_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+group_stage_combinations_ties_df <- group_stage_combinations_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties (Combinations)
+group_stage_combinations_noties_plot <- ggplot(group_stage_combinations_noties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Group Stage (Combinations): Mutual Information (No Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Group Length:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Plot 2: Ties (Combinations)
+group_stage_combinations_ties_plot <- ggplot(group_stage_combinations_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Group Stage (Combinations): Mutual Information (Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Group Length:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
   scale_color_d3("category20") +
@@
 
+# Display plots
+print(group_stage_combinations_noties_plot)
+print(group_stage_combinations_ties_plot)
+
 ```

### 2025-12-05 repechage.R

@@ -1,27 +1,4 @@
----
-title: "single_elimination"
-format: 
-  html:
-    embed-resources: true
-editor: visual
----
-
-```{r}
 library(tidyverse)
-library(dplyr)
-library(shiny)
-```
-
-### Things To Do
-
--   Add a re-seeded option
-
--   Custom strengths (likely as an input to the function)
-
-# Simulate a Single Match
-
-```{r}
-# Function to simulate a match using Bradley-Terry model
 simulate_match <- function(team1, team2, strength1, strength2) {
   if (is.na(team1)) {
     return(team2)
@@ -32,20 +9,12 @@ simulate_match <- function(team1, team2, strength1, strength2) {
     return(ifelse(runif(1) < p, team1, team2))  # Randomize outcome
   }
 }
-
-```
-
-# Single Elimination Tournament Structures
-
-## Creating the traditional tournament structure for n teams
-
-```{r}
 generate_elimination_tournament_structure <- function(n) {
 
   is_power_of_two <- function(x) {
     x > 0 && (x & (x - 1)) == 0
   }
-  
+
   # If n is not a power of 2, round up to the next power of 2 and add placeholders
   if (!is_power_of_two(n)) {
     next_power_of_two <- 2^ceiling(log2(n))
@@ -58,16 +27,16 @@ generate_elimination_tournament_structure <- function(n) {
 
   rounds <- log(n, base = 2) - 1
   teams <- c(1, 2)  # Initial list of teams
-  
+
   for (i in 1:rounds) {
     teams <- nextLayer(teams)
   }
-  
+
   left_half <- teams[1:(length(teams)/2)]
   right_half <- teams[(length(teams)/2+1):length(teams)]
   right_half <- rev(right_half)
   teams <- c(left_half,right_half)
-  
+
   # Ensure that the lower number team is on the left side of each matchup
   teams <- unlist(lapply(seq(1, length(teams), by = 2), function(i) {
     if (teams[i] > teams[i + 1]) {
@@ -77,48 +46,40 @@ generate_elimination_tournament_structure <- function(n) {
     }
   })
   )
-  
+
   while (extra_teams > 0){
-      teams[which.max(teams)] <- NA
-      extra_teams <- extra_teams - 1
+    teams[which.max(teams)] <- NA
+    extra_teams <- extra_teams - 1
   }
-  
+
   return(teams)
 }
 
 nextLayer <- function(teams) {
   out <- c()  # Initialize an empty vector
[…]

### 2025-12-05 staged_rr.R

@@ -273,3 +273,4 @@ simulate_staged_rr <-  function(num_teams, distribution, group_length = 4, top_k
 }
 
 simulate_staged_rr(16, "Normal", rounds=1)
+

### 2025-12-05 theory4.R
theory for 4 teams and single elimination tournament.

@@ -1,0 +1,75 @@
+
+theta <- c()
+theta <- qnorm((4:1)/5)
+theta <- qunif((4:1)/5,0,sqrt(12))
+theta <- c(0,0,0,0)
+p <- matrix(NA, ncol = 4, nrow = 4)
+for (i in 1:4){
+  for (j in 1:4){
+    p[i,j] <- exp(theta[i] - theta[j])/(1+ exp(theta[i] - theta[j]))
+  }
+}
+
+perms$p <- 0
+perms$p[1] <- p[1,4]*p[2,3]*p[1,2]*.5
+perms$p[2] <- p[1,4]*p[2,3]*p[1,2]*.5
+perms$p[7] <- p[1,4]*p[3,2]*p[1,3]*.5
+perms$p[8] <- p[1,4]*p[3,2]*p[1,3]*.5
+perms$p[9] <- p[1,4]*p[3,2]*p[3,1]*.5
+perms$p[10] <- p[1,4]*p[3,2]*p[3,1]*.5
+
+perms$p[11] <- p[4,1]*p[3,2]*p[3,4]*.5
+perms$p[12] <- p[4,1]*p[3,2]*p[3,4]*.5
+perms$p[13] <- p[4,1]*p[3,2]*p[4,3]*.5
+perms$p[14] <- p[4,1]*p[3,2]*p[4,3]*.5
+
+perms$p[19] <- p[4,1]*p[2,3]*p[2,4]*.5
+perms$p[20] <- p[4,1]*p[2,3]*p[4,2]*.5
+perms$p[21] <- p[4,1]*p[2,3]*p[4,2]*.5
+perms$p[22] <- p[4,1]*p[2,3]*p[2,4]*.5
+
+perms$p[23] <- p[1,4]*p[2,3]*p[2,1]*.5
+perms$p[24] <- p[1,4]*p[2,3]*p[2,1]*.5
+
+#weights for k = 4
+perms$w4 <- c(1,5/6,2/3,1/2, 1/3, 1/2, 2/3, 5/6,
+             2/3,1/2,1/3,1/6,0,1/6,1/3,1/2,2/3,1/2,
+             1/3,1/6,1/3,1/2,2/3,5/6)
+
+perms$w1 <- c(1,1,1,2/3,2/3,1,1,1,
+              2/3,2/3,1/3,1/3,0,0,0,1/3,
+              1/3,0,0,0,1/3,1/3,2/3,2/3)
+
+sub <- perms %>% filter(p > 0)
[…]

### 2025-12-12 simulations.qmd
added more tournaments

@@
 
+```{r}
+# Create separate parameter grids for each configuration
+staged_rr_param_grid_4x4 <- expand.grid( 
+  distribution = c("Normal", "Uniform", "Same", "Exponential"), 
+  ties = c(FALSE, TRUE), 
+  rounds = c(1, 2, 4),
+  group_length = 4,
+  grouping_mode = "by_rank",
+  stringsAsFactors = FALSE
+)
+
+staged_rr_param_grid_2x8 <- expand.grid( 
+  distribution = c("Normal", "Uniform", "Same", "Exponential"), 
+  ties = c(FALSE, TRUE), 
+  rounds = c(1, 2, 4),
+  group_length = 8,
+  grouping_mode = "by_pairs",
+  stringsAsFactors = FALSE
+)
+
+# Combine both grids
+staged_rr_param_grid <- rbind(staged_rr_param_grid_4x4, staged_rr_param_grid_2x8)
+
+staged_rr_run_sim_block <- function(
+  teams, distribution, ties, rounds, group_length, grouping_mode, p_block) {
+
+  # Build the key
+  parts <- c(
+    "stagedRR",
+    tolower(distribution),
+    ifelse(ties, "ties", "noTies"),
+    paste0("r", rounds),
+    paste0("g", group_length)
+  )
+  key <- paste(parts, collapse = "_")
+
+  # Run simulations in parallel
+  sim_list <- future_lapply(
+    X = 1:10000,
+    FUN = function(rep) {
+      simulate_staged_rr(
+        num_teams = teams,
+        distribution = distribution,
+        ties = ties,
+        rounds = rounds,
+        group_length = group_length,
+        grouping_mode = grouping_mode
+      )
+    },
+    future.seed = TRUE    # ensures parallel RNG is valid
+  )
+
+  final_df <- do.call(rbind, sim_list)
+  final_df$teams <- teams
+
+  assign(key, final_df, envir = .GlobalEnv)
+
+  # UPDATE BLOCK-LEVEL PROGRESS BAR
+  p_block()
+
+  message("Finished: ", key)
+}
+
+# Run the simulations with progress bar
+with_progress({
+  p_block <- progressor(steps = nrow(staged_rr_param_grid))
+
+  for (i in seq_len(nrow(staged_rr_param_grid))) {
+    row <- staged_rr_param_grid[i, ]
+
+    staged_rr_run_sim_block(
+      teams        = teams, 
+      distribution = row$distribution,
+      ties         = row$ties,
+      rounds       = row$rounds,
+      group_length = row$group_length,
+      grouping_mode = row$grouping_mode,
+      p_block = p_block
+    )
+  }
+})
+
+```
+
 ## High School Tennis Ladder
+
+```{r}
+tennis_ladder_param_grid <- expand.grid( 
+  distribution = c("Normal","Uniform", "Same", "Exponential"), 
+  r = c(5,10,15),
+  bad_seeding_structure = c(FALSE, TRUE) 
+  )
+
+tennis_ladder_run_sim_block <- function(
+  teams, distribution, ties, r, p_block, bad_seeding_structure) {
+
+  # Apply bad seeding structure override if required
+  seeding_structure <- if (bad_seeding_structure) 16:1 else NULL
+
+  # Build key
+    # Build the key
+  parts <- c(
+    "tennisLadder",
+    tolower(distribution),
+    paste0(r),
+    if (bad_seeding_structure) "badseed"
+  )
+
+  key <- paste(parts, collapse = "_")
+
+  # Run 10,000 simulations in parallel
+  sim_list <- future_lapply(
+    X = 1:10000,
+    FUN = function(rep) {
+      simulate_tennis_ladder(
+        num_teams = teams,
+        distribution = distribution,
+        r = r,
+        seeding_structure = seeding_structure
+      )
+    },
+    future.seed = TRUE    # ensures parallel RNG is valid
+  )
+
+  final_df <- do.call(rbind, sim_list)
+  final_df$teams <- teams
+
+  assign(key, final_df, envir = .GlobalEnv)
+
+  # UPDATE BLOCK-LEVEL PROGRESS BAR
+  p_block()
+
+  message("Finished: ", key)
+}
+
+with_progress({
+  p_block <- progressor(steps = nrow(tennis_ladder_param_grid))
+
+  for (i in seq_len(nrow(tennis_ladder_param_grid))) {
+    row <- tennis_ladder_param_grid[i, ]
+
+    tennis_ladder_run_sim_block(
+      teams        = teams,
+      distribution = row$distribution,
+      r            = row$r,
+      bad_seeding_structure = row$bad_seeding_structure,
+      p_block = p_block
+    )
+  }
+})
+```
 
@@
   # Validate input
-  valid_structures <- c("singleElimination", "doubleElimination", "roundRobin", "groupStage", "stepLadder", "Repechage")
+  valid_structures <- c("singleElimination", "doubleElimination", "roundRobin", 
+                       "groupStage", "stepLadder", "Repechage", "stagedRR", "tennisLadder")
   if (!tournament_structure %in% valid_structures) {
@@
 
+  tennisLadder_key <- function(row, type) {
+    c(
+      type,
+      tolower(row$distribution),
+      paste0(row$r),
+      if (row$bad_seeding_structure) "badseed"
+    )
+  }
+
+  # Key builders for stagedRR (both configurations)
+  staged_rr_4x4_key <- function(row, type) {
+    c(
+      type,
+      tolower(row$distribution),
+      ifelse(row$ties, "ties", "noTies"),
+      paste0("r", row$rounds),
+      paste0("g", row$group_length)
+    )
+  }
+
+  staged_rr_2x8_key <- function(row, type) {
+    c(
+      type,
+      tolower(row$distribution),
+      ifelse(row$ties, "ties", "noTies"),
+      paste0("r", row$rounds),
+      paste0("g", row$group_length)
+    )
+  }
+
   # Function to process a single tournament type
@@
                                       repechege_key)
-    results_list$Repechege <- result
+    results_list$Repechage <- result
   }
@@
   }
+
   if (tournament_structure %in% c("stepLadder")) {
@@
 
+  # Process stagedRR - combined approach using full param grid
+  if (tournament_structure %in% c("stagedRR")) {
+    # Use the combined param grid and a unified key builder
+    result <- process_tournament_type("stagedRR", 
+                                      staged_rr_param_grid,
+                                      staged_rr_4x4_key)  # Works for both since key structure is same
+    results_list$stagedRR <- result
+  }
+
+  if (tournament_structure %in% c("tennisLadder")) {
+    result <- process_tournament_type("tennisLadder", 
+                                      tennis_ladder_param_grid, 
+                                      tennisLadder_key)
+    results_list$tennisLadder <- result
+  }
+
   message("\n✓ MI computation complete!")
@@
 
+# Run computations for all tournament types
 compute_and_store_mi_by_tournament("singleElimination")
@@
 compute_and_store_mi_by_tournament("Repechage")
-
+compute_and_store_mi_by_tournament("stagedRR")
+compute_and_store_mi_by_tournament("tennisLadder")
+
+# Combine all MI results
 all_mi <- bind_rows(
@@
   stepLadder_mi,
-  Repechage_mi
-)
-
+  Repechage_mi,
+  stagedRR_mi,
+  tennisLadder_mi
+)
+
+# Combine all MI combinations results
 all_mi_combinations <- bind_rows(
@@
   stepLadder_combinations_mi,
-  Repechage_combinations_mi
+  Repechage_combinations_mi,
+  stagedRR_combinations_mi,
+  tennisLadder_combinations_mi
 )
@@
   "groupStage_same_noTies_1_4groups",
-  "groupStage_exponential_noTies_1_4groups"
+  "groupStage_exponential_noTies_1_4groups",
+  "stagedRR_normal_noTies_r1_g4",
+  "stagedRR_uniform_noTies_r1_g4",
+  "stagedRR_same_noTies_r1_g4",
+  "stagedRR_exponential_noTies_r1_g4",
+  "tennisLadder_normal_15",
+  "tennisLadder_uniform_15",
+  "tennisLadder_same_15",
+  "tennisLadder_exponential_15"
   )
@@
 
-  # default group length
+  # default group length/rounds
   group_length <- NA
+  rounds <- NA
 
@@
     group_length <- as.numeric(str_extract(parts[5], "\\d+"))
+  }
+
+  if(tournament_type == "stagedRR"){
+    group_length <- as.numeric(str_extract(parts[5], "\\d+"))
+    rounds <- as.numeric(str_extract(parts[4], "\\d+"))
+  }
+
+  if(tournament_type == "tennisLadder"){
+    rounds <- as.numeric(parts[3])
   }
@@
     distribution = distribution,
-    group_length = group_length
+    group_length = group_length,
+    rounds = rounds
   ))
@@
 combined_plot_df <- combined_plot_df %>%
-  select(-any_of(c("tournament_type", "distribution", "group_length"))) %>%
+  select(-any_of(c("tournament_type", "distribution", "group_length", "rounds"))) %>%
   group_by(dataset) %>%
@@
       tournament_type == "groupStage" ~ "groupStage (4 Groups)",
+      tournament_type == "stagedRR" ~ "stagedRR (4 Groups)",
+      tournament_type == "tennisLadder" ~ "tennisLadder (15 Rounds)",
       TRUE ~ tournament_type
@@
 combined_combinations_df <- combined_combinations_df %>%
-  select(-any_of(c("tournament_type", "distribution", "group_length"))) %>%
+  select(-any_of(c("tournament_type", "distribution", "group_length", "rounds"))) %>%
   group_by(dataset) %>%
@@
 ```
+
+## Staged Round Robin
+
+```{r}
+staged_rr_plot_df <- plot_mi(stagedRR_mi)
+
+# Parse dataset names to extract parameters for staged RR
+parse_staged_rr_name <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  # Extract ties (third part)
+  ties <- ifelse(parts[3] == "noTies", "No Ties", "Ties")
+
+  # Extract rounds (fourth part - extract number after 'r')
+  rounds <- as.numeric(str_extract(parts[4], "\\d+"))
+
+  # Extract group length (fifth part - extract number after 'g')
+  group_length <- as.numeric(str_extract(parts[5], "\\d+"))
+
+  return(data.frame(
+    distribution = distribution,
+    ties = ties,
+    rounds = rounds,
+    group_length = group_length
+  ))
+}
+
+# Add parameter columns to the plot data
+staged_rr_plot_df <- staged_rr_plot_df %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_staged_rr_name(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    series = factor(rounds, levels = c(1, 2, 4)),
+    # Create a label for the additional parameters
+    param_label = paste(group_length, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+staged_rr_plot_df <- staged_rr_plot_df %>%
+  mutate(
+    color_group = paste(group_length, sep = " | ")
+  )
+
+# Filter data for No Ties
+staged_rr_noties_df <- staged_rr_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+staged_rr_ties_df <- staged_rr_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties
+staged_rr_noties_plot <- ggplot(staged_rr_noties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Staged RR: Mutual Information (No Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Group Length:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Plot 2: Ties
+staged_rr_ties_plot <- ggplot(staged_rr_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Staged RR: Mutual Information (Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Group Length:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+staged_rr_combinations_plot_df <- plot_mi(stagedRR_combinations_mi)
+
+# Add parameter columns to the plot data
+staged_rr_combinations_plot_df <- staged_rr_combinations_plot_df %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_staged_rr_name(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    series = factor(rounds, levels = c(1, 2, 4)),
+    # Create a label for the additional parameters
+    param_label = paste(group_length, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+staged_rr_combinations_plot_df <- staged_rr_combinations_plot_df %>%
+  mutate(
+    color_group = paste(group_length, sep = " | ")
+  )
+
+# Filter data for No Ties
+staged_rr_combinations_noties_df <- staged_rr_combinations_plot_df %>%
+  filter(ties == "No Ties")
+
+# Filter data for Ties
+staged_rr_combinations_ties_df <- staged_rr_combinations_plot_df %>%
+  filter(ties == "Ties")
+
+# Plot 1: No Ties (Combinations)
+staged_rr_combinations_noties_plot <- ggplot(staged_rr_combinations_noties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Staged RR (Combinations): Mutual Information (No Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Group Length:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Plot 2: Ties (Combinations)
+staged_rr_combinations_ties_plot <- ggplot(staged_rr_combinations_ties_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    series ~ distribution,
+    labeller = labeller(
+      series = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Staged RR (Combinations): Mutual Information (Ties)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Group Length:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Display plots
+print(staged_rr_noties_plot)
+print(staged_rr_ties_plot)
+
+# Display plots
+print(staged_rr_combinations_noties_plot)
+print(staged_rr_combinations_ties_plot)
+```
+
+## Tennis Ladder
+
+```{r}
+tennis_ladder_plot_df <- plot_mi(tennisLadder_mi)
+
+# Parse dataset names to extract parameters for tennis ladder
+parse_tennis_ladder_name <- function(dataset_name) {
+  parts <- str_split(dataset_name, "_")[[1]]
+
+  # Extract distribution (second part)
+  distribution <- parts[2]
+  distribution <- str_to_title(distribution)
+
+  # Extract rounds (third part - the number)
+  rounds <- as.numeric(parts[3])
+
+  # Check if it's a "badseed" variant (fourth part)
+  seed_type <- ifelse(length(parts) >= 4 && parts[4] == "badseed", "Bad Seed", "Good Seed")
+
+  return(data.frame(
+    distribution = distribution,
+    rounds = rounds,
+    seed_type = seed_type
+  ))
+}
+
+# Add parameter columns to the plot data
+tennis_ladder_plot_df <- tennis_ladder_plot_df %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_tennis_ladder_name(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    rounds_factor = factor(rounds, levels = c(5, 10, 15)),
+    # Create a label for the additional parameters
+    param_label = paste(rounds, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+tennis_ladder_plot_df <- tennis_ladder_plot_df %>%
+  mutate(
+    color_group = paste(rounds, sep = " | ")
+  )
+
+# Filter data for Good Seed
+tennis_ladder_goodseed_df <- tennis_ladder_plot_df %>%
+  filter(seed_type == "Good Seed")
+
+# Filter data for Bad Seed
+tennis_ladder_badseed_df <- tennis_ladder_plot_df %>%
+  filter(seed_type == "Bad Seed")
+
+# Plot 1: Good Seed
+tennis_ladder_goodseed_plot <- ggplot(tennis_ladder_goodseed_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    rounds_factor ~ distribution,
+    labeller = labeller(
+      rounds_factor = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Tennis Ladder: Mutual Information (Good Seed)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Rounds:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Plot 2: Bad Seed
+tennis_ladder_badseed_plot <- ggplot(tennis_ladder_badseed_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    rounds_factor ~ distribution,
+    labeller = labeller(
+      rounds_factor = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Tennis Ladder: Mutual Information (Bad Seed)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Rounds:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+tennis_ladder_combinations_plot_df <- plot_mi(tennisLadder_combinations_mi)
+
+# Add parameter columns to the plot data
+tennis_ladder_combinations_plot_df <- tennis_ladder_combinations_plot_df %>%
+  group_by(dataset) %>%
+  mutate(
+    params = list(parse_tennis_ladder_name(dataset[1]))
+  ) %>%
+  unnest(params) %>%
+  ungroup() %>%
+  mutate(
+    rounds_factor = factor(rounds, levels = c(5, 10, 15)),
+    # Create a label for the additional parameters
+    param_label = paste(rounds, sep = "\n")
+  )
+
+# Create a consistent color variable based on the parameter combinations
+tennis_ladder_combinations_plot_df <- tennis_ladder_combinations_plot_df %>%
+  mutate(
+    color_group = paste(rounds, sep = " | ")
+  )
+
+# Filter data for Good Seed
+tennis_ladder_combinations_goodseed_df <- tennis_ladder_combinations_plot_df %>%
+  filter(seed_type == "Good Seed")
+
+# Filter data for Bad Seed
+tennis_ladder_combinations_badseed_df <- tennis_ladder_combinations_plot_df %>%
+  filter(seed_type == "Bad Seed")
+
+# Plot 1: Good Seed (Combinations)
+tennis_ladder_combinations_goodseed_plot <- ggplot(tennis_ladder_combinations_goodseed_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    rounds_factor ~ distribution,
+    labeller = labeller(
+      rounds_factor = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Tennis Ladder (Combinations): Mutual Information (Good Seed)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Rounds:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Plot 2: Bad Seed (Combinations)
+tennis_ladder_combinations_badseed_plot <- ggplot(tennis_ladder_combinations_badseed_df, aes(x = index, y = mi_values, color = color_group, group = dataset)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 1.5) +
+  facet_grid(
+    rounds_factor ~ distribution,
+    labeller = labeller(
+      rounds_factor = function(x) paste("Rounds:", x),
+      distribution = function(x) x
+    )
+  ) +
+  labs(
+    title = "Tennis Ladder (Combinations): Mutual Information (Bad Seed)",
+    x = "Top K Teams",
+    y = "Mutual Information (MI)",
+    color = "Rounds:"
+  ) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_x_continuous(breaks = 1:16) + 
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 11),
+    axis.text = element_text(size = 8),
+    strip.text = element_text(size = 10, face = "bold"),
+    legend.position = "bottom", 
+    legend.title = element_text(size = 10, face = "bold"),
+    legend.text = element_text(size = 8),
+    panel.spacing = unit(1, "lines")
+  ) +
+  scale_color_d3("category20") +
+  guides(color = guide_legend(nrow = 1))
+
+# Display plots
+print(tennis_ladder_goodseed_plot)
+print(tennis_ladder_badseed_plot)
+
+# Display plots
+print(tennis_ladder_combinations_goodseed_plot)
+print(tennis_ladder_combinations_badseed_plot)
+```

### 2025-12-12 staged_rr.R

@@ -1,4 +1,8 @@
-simulate_staged_rr <-  function(num_teams, distribution, group_length = 4, top_k = 2, stages = 2, ties = TRUE, rounds = 1, true_strength_hat = NULL) {
+simulate_staged_rr <- function(num_teams, distribution, group_length = 4, top_k = 2,
+                               stages = 2, ties = TRUE, rounds = 1, theta_hat = NULL,
+                               grouping_mode = "by_rank") {
+  # grouping_mode: "by_rank" for 4x4 (rank 1s together, rank 2s together, etc.)
+  #                "by_pairs" for 2x8 (1st&2nd together, 3rd&4th together, etc.)
 
   if (num_teams <= 3) {
     stop("Number of Teams must be greater than 3.")
@@ -56,7 +60,7 @@ simulate_staged_rr <-  function(num_teams, distribution, group_length = 4, top_k
     df <- arrange(unif_strength, true_rank)
   }
   else if (distribution == "Manual") {
-    strengths <- true_strength_hat
+    strengths <- theta_hat
     manual_strengths <- data.frame(
       true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
       rank_hat = rep(NA, length(teams)),
@@ -74,18 +78,14 @@ simulate_staged_rr <-  function(num_teams, distribution, group_length = 4, top_k
   num_groups <- num_teams / group_length
 
   # Create balanced groups using snake draft seeding
-  # Sort teams by strength (best to worst)
   df <- df %>% arrange(desc(true_strength))
 
-  # Assign groups using snake draft method
   df$groups <- NA
   for (i in 1:nrow(df)) {
     group_round <- ceiling(i / num_groups)
     if (group_round %% 2 == 1) {
-      # Forward direction for odd rounds
       group_id <- ((i - 1) %% num_groups) + 1
     } else {
-      # Reverse direction for even rounds
       group_id <- num_groups - ((i - 1) %% num_groups)
     }
     df$groups[i] <- group_id
@@ -103,19 +103,16 @@ simulate_staged_rr <-  function(num_teams, distribution, group_length = 4, top_k
     group_teams <- df[df$groups == g, ]
     n_group_teams <- nrow(group_teams)
 
-    # Play round robin within the group
     for (r in 1:rounds) {
       for (i in 1:(n_group_teams - 1)) {
         for (j in (i + 1):n_group_teams) {
           team1 <- group_teams$true_rank[i]
[…]

### 2025-12-12 tennis_ladder.R

@@ -1,0 +1,197 @@
+simulate_tennis_ladder <- function(num_teams, distribution, seeding_structure = NULL, r = 3, theta_hat = NULL) {
+
+  if (num_teams <= 1) {
+    stop("Number of Teams must be greater than 1.")
+  }
+
+  if (num_teams %% 2 != 0) {
+    stop("Number of teams must be even for tennis ladder.")
+  }
+
+  teams <- paste(num_teams:1)
+
+  # Create team strengths based on distribution
+  if (distribution == "Normal"){
+    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
+    team_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = strengths,
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- arrange(team_strengths, true_rank)
+  }
+  else if (distribution == "Same"){
+    strengths <- sapply(num_teams, function(n) {rep(0, n)})
+    team_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = strengths,
+      rank_hat = rep(NA, length(teams)),
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams))
+    )
+    df <- arrange(team_strengths, true_rank)
+  }
+  else if (distribution == "Uniform"){
+    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1), 0, sqrt(12)) })
+    team_strengths <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = strengths,
+      game_wins = rep(0, length(teams)),
+      game_losses = rep(0, length(teams)),
+      rank_hat = rep(NA, length(teams))
+    )
+    df <- arrange(team_strengths, true_rank)
+  }
+  else if (distribution == "Exponential"){
+    strengths <- sapply(num_teams, function(n) { qexp(1:n/(n+1)) })
+    unif_strength <- data.frame(
+      true_rank = as.numeric(teams),
+      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
+      game_wins = rep(0,length(teams)),
+      game_losses = rep(0,length(teams)),
+      rank_hat = rep(NA,length(teams))
+    )
+    df <- arrange(unif_strength, true_rank)
+  }
[…]

### 2025-12-12 manuscript/paper.Rmd
updated.

@@
 @Goosens2024, @Appleton1995-zn and @Csato2021.
+
+@Daniels1969
 
@@
 
-Regardless of which method one chooses for defining weights between $\hat{r}$ and $r$, a second set of weights needs to be defined because the number of teams that need to be accurately ordered differs based on the choice of salient ranks.
+In addition to the the weighting function $w$, we defined a second set of weights that we refer to as the "salient weights" and denote as the vector $m$.
+This set of weights is used to indicate which ranks have been determined to salient (as previously described) and consists of a vector of length $n$ with elements that are positive for each of the salient ranks and 0 otherwise.
+As an example, for a four-team tournament, if the organizer decided that only the correctly choosing the best team as the winner was important, then $m = \{1,0,0,0\}$ (i.e.
+measuring effectivity).
+In an event such as the Olympics, the salient weights could be 1's for the first three elements and 0's elsewhere such that $m = \{1,1,1,0\}$.
+Additionally, one can also weight the salient ranks.
+For example, continuing the Olmypics example where getting the top three correct is the main objective, one could argue that it is more important to get the first place correct than the third place.
+This could lead to, for instance, a salient weight vector of $m = \{5,3,1,0, \cdots, 0\}$ giving 5 times the weight to getting the first place team correct as getting the third team correct.
+
+<!-- Regardless of which method one chooses for defining weights between $\hat{r}$ and $r$, a second set of weights needs to be defined because the number of teams that need to be accurately ordered differs based on the choice of salient ranks.
 Therefore, we define a second vector of weights, $m$, which is used to indicate which ranks are salient.
@@
 essentially measuring effectivity)^[7].
+-->
 
-^[7]: One can imagine that in an event like the Olympics, the salient weights would be 1's for the first three elements and 0's elsewhere (i.e. $m = \{1,1,1,0\}$).
+<!-- ^[7]: One can imagine that in an event like the Olympics, the salient weights would be 1's for the first three elements and 0's elsewhere (i.e. $m = \{1,1,1,0\}$).     -->
 
-In addition, one can also weight the salient ranks.
-For example, in the Olmypics where getting the top three correct is the main objective, one could argue that it is more important to get the first place correct than the third place.
-This could lead to, for instance, a salient weight vector of $m = \{5,3,1,0, \cdots, 0\}$ giving 5 times the weight to getting the first place team correct as getting the third team correct.
-
-With the addition of the salient weighting function, $m$, the formula for weighted mutual information becomes:
+With the addition of the salient weighting vector, $m$, the formula for weighted mutual information becomes:
 
 $$
-I^m_w(r,\hat{r})= \sum_{r}\sum_{\hat{r}}w(r, \hat{r}, m)p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
+I_w(r,\hat{r},m)= \sum_{r}\sum_{\hat{r}}w(r, \hat{r}, m)p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
 $$
@@
 
-If we choose to use the weighted Kendall's $\tau$ for our weights, $w(r, \hat{r}, m)$ would be:
+Incorporating $m$ into this function is straightforward for many weighting functions.  However, using weighted Kendall's $\tau$, $w(r, \hat{r}, m)$ is slightly more complex and becomes:
 
@@
 
-Because this mutual information is unitless, the formula can be standardized to be between 0 and 1 as follows: 
+Finally, because this mutual information is unitless, the formula can be standardized to be between 0 and 1 as follows: 
 $$
-\frac{I^m_w(r,\hat{r})}{\text{max}\{H(r), H(\hat{r})\}}
+I^{\star}_w(r,\hat{r},m) = \frac{I_w(r,\hat{r},m)}{\text{max}\{H(r), H(\hat{r})\}}
 $$
 
-where $H(.)$ is the usual Shannon entropy. Competition structures with
-higher levels of efficacy are indicated when our measure is closer to 1.
-Likewise, competition structures that are near 0 by our proposed metric
-indicate low levels of efficacy.
+where $H(.)$ is the usual Shannon entropy.
+Using this framework, values of $I^{\star}_w(r,\hat{r},m)$ near 1 indicated competition structures with high efficacy.
+Values near 0.25 indicate tournament structures where all possible rankings are equally likely and values near 0 indicate rankings that are [what is the right word here "Inverse efficacy"?]
+
+<!-- Likewise, competition structures that are near 0 by our proposed metric indicate low levels of efficacy. -->
 

### 2025-12-12 manuscript/sources_summary.qmd
Rename sources_summary.qmd to manuscript/sources_summary.qmd

--- 2025-11-07 sources_summary.qmd
@@
 editor: visual
+bibliography: references.bib
 ---
@@
 the true ranking of competitors).
+
+# Josie Lit Review
+
+In @Carroll1883, a very early evaluation of tournaments and rankings, Lewis Carroll evaluated paired comparisons and the method of assigning prizes in Lawn Tennis Tournaments.
+He found that competitors were not correctly ranked at the end of the competition, so he proposed a system of rules for conducting tournaments that would take less time and would lead to equitable results.
+He also proposed an equitable system for scoring in matches, which included abolishing the concept of "sets" and to make a "match" consist of "games" instead.
+He concludes that his new methods eliminate the element of luck in tournaments so competitors are fully evaluated on their skills, making it possible to predict who will win.
 

### 2025-12-12 manuscript/paper.Rmd
updated.

@@
 and the number of discordant pairs is equivalent to the number of adjacent swaps needed to get the order of $\hat{r}$ correct.  
-
 As an example, consider $r = (1,2,3,4)$ and $\hat{r} = (1,3,4,2)$.
@@
 $$
+# Tiered Ranking
+
+This framework was developed to assess the efficacy of the full ranking of the top $k$ teams generated by different competition formats.
+However, it may be of interest to have a competition structure that simply selects the top $k$ teams in any order.
+Continuing with the Olympics example, if an organizer only cares about the top 3 teams earning medals *in any order* (i.e.
+$\hat{r} = \{1,2,3,4\}$ and $\hat{r} = \{3,2,1,4\}$ are equivalent), our framework can easily handle this modifying $\hat{r}$ (and $r$?) so that the top $k$ elements are ordered correctly and tied.
+That is, for the Olympic example, if $k$ was 3 and $\hat{r} = \{3,2,1,4\}$, we convert all the elements of $\hat{r} \le k$ to 1 (i.e.
+$\hat{r}$ becomes $\{1,1,1,4\}$), and then we proceed with the exact same formula.
+We refer to this as tiered ranking.
+
+This concept can be extended to multiple tiers, which is applicable to a competition structure such as the premier league where the top 4 teams qualify for the Champions league and the bottom 3 teams are relegated.
+While this straight foward to compute, we do not evaluate scenarios such as this in this article.
 
@@
 
-@Shieh1998: Weighted Kendall's Tau
+<!-- @Shieh1998: Weighted Kendall's Tau -->
 
@@
 
-Incorporating $m$ into this function is straightforward for many weighting functions.  However, using weighted Kendall's $\tau$, $w(r, \hat{r}, m)$ is slightly more complex and becomes:
+Incorporating $m$ into this function is straightforward for many weighting functions.
+We use the weighted Kendall's $\tau$ presented in @Shieh1998 with a slight modification to the sign function.
+Specifically, in the special case that $sign(0) = 1$.
+Then $w(r, \hat{r}, m)$ becomes:
 
@@
 Using this framework, values of $I^{\star}_w(r,\hat{r},m)$ near 1 indicated competition structures with high efficacy.
-Values near 0.25 indicate tournament structures where all possible rankings are equally likely and values near 0 indicate rankings that are [what is the right word here "Inverse efficacy"?]
+Values near 0.25 indicate tournament structures where all possible rankings are equally likely and values near 0 indicate rankings that are inefficacious (e.g.
+complete rank reversal).
 

### 2025-12-12 mutual_information_curves.qmd

@@
 
-weighted_kendall(c(1,1,1,1,5,6), k = 3, r = c(1,1,1,1,5,6))
-# Only works for k = # of top teams tied
-
 make_weights <- function(k) {
@@
 rhat_df <- bind_rows(all_results)
+rhat_df %>% filter(k==1)
 ```
@@
     mutate(w = w)
+  print(weighted_df)
 

### 2025-12-16 manuscript/sources_summary.qmd
lit review

--- 2025-11-07 sources_summary.qmd
@@
 the true ranking of competitors).
+
+# Josie Lit Review
+
+In an early evaluation of tournament design and ranking accuracy, @Carroll1883 evaluated paired comparisons and the method of assigning prizes in Lawn Tennis Tournaments.
+He argued that the standard single-elimination tournament format often failed to rank competitors correctly, as strong players could be eliminated early due to unfavorable pairings or random variation in match outcomes.
+To address this, he proposed a system of rules for conducting tournaments that would take less time while producing more equitable results.
+Carroll also suggested reforms to match scoring, including abolishing the concept of “sets” and instead defining a match as a sequence of individual “games,” with the goal of reducing randomness in outcomes.
+He concluded that his new methods eliminate the element of luck in tournaments so competitors are fully evaluated on their skills, making it possible to predict who will win.
+
+While Carroll focused on improving tournament rules, Kendall and Smith examined rankings as a measurement problem.
+They evaluated the method of paired comparisons, where objects are compared only two at a time rather than against all others simultaneously.
+They argue that the common ranking method of comparison performs poorly when the object quality being measured cannot be represented linearly.
+Instead, they analyzed the method of paired comparisons, in which an observer expresses a preference between two objects at a time, and a final ranking is inferred after many comparisons are made.
+Kendall and Smith also explored how these preferences can be represented in tabular or geometric form, emphasizing that inconsistent or “circular” preferences may occur, such as situations in which A is preferred to B, B to C, and C to A.
+In addition, they examined agreement among multiple judges, distinguishing between disagreement caused by poor judgment and disagreement resulting from indistinguishable differences in object quality.
+They derive a measurement to quantify the agreement between judges, which is based on Kendall’s Tau.
+
+In The World Series Competition, @Mosteller1952 shifted the focus from designing ranking systems to evaluating how well observed competition outcomes reflect underlying team quality.
+Rather than assuming that tournament results accurately rank teams, he estimated the probability that the better team wins an individual contest.
+Using historical data from the World Series, Mosteller modeled game outcomes as events influenced by team strength and random variation.
+His analysis demonstrated that even in a best-of-seven series, the best team does not always win, highlighting the role of chance in tournament outcomes.
+His work provided a foundation for evaluating tournament formats by linking design choices, such as the number of games played, to the probability that final outcomes reflect true team ability.
+
+Building on the paired comparison framework, @BradleyTerry1952 introduced a probabilistic model for paired comparisons that became a cornerstone of modern statistical ranking theory.
+Their model assigns a positive “strength” parameter to each competitor and defines the probability that one competitor beats another as a function of the ratio of their strengths.
+This formulation allows researchers to infer an overall ranking from observed pairwise outcomes by estimating these strength parameters from data, often by using maximum likelihood methods.
+The Bradley-Terry model has become widely used in sports analytics and preference modeling because it translates binary outcomes from games or comparisons into a quantitative ranking of competitors.
+
+@Glenn1960-mn returned to the comparison of tournament formats, offering one of the earliest evaluations of tournament "effectiveness." Focusing on competitions with four participants, Glenn analyzed six tournament designs, including round-robin, variations of single-elimination, and double-elimination formats.
+He compared these designs using two criteria: the probability that the strongest competitor wins and the expected number of games required.
+By modeling match outcomes probabilistically, Glenn found that tournaments with more games are more accurate but less efficient.
+His results showed that round-robin tournaments maximize ranking accuracy, while elimination formats prioritize efficiency.
 
